### 2020-04-25 templates/Abstract_&_Setup.Rmd
Add templates

---
<EN>title:
  Concrete compressive strength prediction with machine learning
date: "April 22, 2020"</EN>
<PT>title:
  Prevendo a resisência à compressão do concreto com técnicas de machine learning
date: "Abril 22, 2020"</PT>
author: "Pedro Bernardino Alves Moreira"
indent: true
<EN>abstract:
 Compressive strength is the main characteristic of concrete. This work used a set of samples widely used to build 6 forecast models of this characteristic, each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers) and 6 relationships between the main ingredients (water/cement, coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate, water / coarse aggregate and water/fine aggregate) to make the predictions. The algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*) . The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31) and the worst for the 100-day model (*RMSE* of 5.85). The 28-day model had *RMSE* of 4.72. The work indicated that the separation in a model for each age is promising and was able to generate compatible results or better than other studies, opening way for new models creation following this proposal, but using different algorithms or the combination of different algorithms.</EN>
<PT>abstract:
  A resistência à compressão é a principal característica do concreto. Esse trabalho utilizou um conjunto de amostras amplamente utilizado para construir 6 modelos de previsão dessa característica, cada um específicamente para uma idade, 3, 7, 14, 28, 56 e 100 dias a partir de 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes) e 6 relações entre os principais ingredientes (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo) para fazer a previsão. Foi utilizado o algorítimo *Parallel Random Forest* (*parRF* do pacote *caret* de linguagem *R*). Os resultados foram satisfatórios sendo o melhor para o modelo de 3 dias (*RMSE* de 3,31) e pior para o modelo de 100 dias (*RMSE* de 5,85). O modelo de 28 dias apresentou *RMSE* de 4,72. O trabalho indicou que a separação em um modelo para cada idade é promissora e foi capaz de gerar resultados compatíveis ou melhores que outros estudos, abrindo espaço para construção de novos modelos seguindo essa proposta, mas utilizando algorítimos diferentes ou a combinação de diversos algorítimos.</PT>
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
bibliography: references.bib
header-includes:
<PT>- \renewcommand{\abstractname}{Resumo}</PT>
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{supertabular}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{caption}
- \usepackage{floatrow}
- \floatsetup[table]{capposition=top}
- \floatsetup[figure]{capposition=top}
- \captionsetup{options=chunk}
- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
- \renewcommand{\thechunk}{\arabic{chunk}}
- \usepackage{indentfirst}
- \usepackage{sectsty} \sectionfont{\centering}
---

```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")

# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)

# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```

<PT>
\renewcommand{\figurename}{Figura}
\renewcommand{\tablename}{Tabela}
\renewcommand{\contentsname}{Sumário}
\renewcommand{\chunkname}{Código}
\renewcommand{\listfigurename}{Lista de figuras}
\renewcommand{\listtablename}{Lista de tabelas}
</PT>

\newpage

\tableofcontents

\newpage

\listoffigures

\listoftables

\newpage

### 2020-04-25 templates/Discussion.Rmd
Add templates

<EN>
# Discussion

The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily. The alternative adopted to create a model for each set of age proved to be a valid method, managing to stratify to obtain specific results for each set. The studies cited in the introduction using the same dataset have similar results, as expected. The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
</EN>

<PT>
# Discussão

Os modelos construidos apresentam resultados satisfatórios e provam que a resistência à compressão do concreto pode ser prevista de forma relativamente fácil. A alternativa adotada de criar um modelo para cada conjunto de idade se mostrou como uma alternativa válida, conseguindo estratificar para obter resultados específicos de cada conjunto. Os estudos citados na introdução utilizando o mesmo conjunto de dados possuem resultados similares, como esperado. A tabela \ref{tab:works-comparison} apresenta os resultados desses trabalhos (\ref{show-works-comparison}), e a tabela \ref{tab:results} apresenta os valores encontrados (\ref{show-results}) para fácil comparação.
</PT>

```{r works-comparison, echo=F}
#<EN># Table - Comparison of other works</EN>
#<PT># Tabela - "Comparação dos estudos de outros autores"</PT>

#<EN>
colNames = c("Author","Year" ,"Algorithm", "RMSE") 
caption <- "Comparison of other works"
#</EN>
#<PT>
colNames = c("Autor","Ano" ,"Algorítimo", "RMSE") 
caption <- "Comparação dos estudos de outros autores"
#</PT>

works <- data.frame(
  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
          "Gradient Boosting Regressor","Random Forest Regressor" ,
          "Regularized Extreme Learning Machine",
          "Support Vector Machines with Radial Basis Function Kernel"),
  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
)

kable(
  works,
  col.names = colNames,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "l"
  )  %>%
    kable_styling(latex_options = c("HOLD_position"))

```

```{r results, echo=F}
#<EN># Table - Final results</EN>
#<PT># Tabela - Resultados finais</PT>

#<EN>
colNames <- c("Model", "RMSE")
caption <- "Final result"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
#</EN>
#<PT>
colNames <- c("Modelo", "RMSE")
caption <- "Resultados finais"
day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
#</PT>

dat_reg_models <- data.frame(
  dia = day,
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)

kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

<EN>
Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results. Another option is to create an *ensemble* of various algorithms, just like @Pierobon2018, but with the separation of age sets proposed here. In addition, it can be performed with a larger dataset, ideally with the same number of samples in each age set, a more homogeneous distribution of compressive strength, and less variance between samples.
</EN>

<PT>
Seguindo a linha de raciocínio desse trabalho, ele pode ser realizado com diferentes algorítimos, os resultados aqui encontrados utilizaram apenas um único (*Parallel Random Forest*), mesmo que tenha sido teoricamente o "melhor" encontrado, outros algorítmos podem apresentar resultados ainda melhores. Outra opção é criar um *ensemble* com diversos algorítimos, como realizado por @Pierobon2018, mas com a separação de conjuntos de idade aqui proposto. Além disso, pode ser realizado com um conjunto maior de dados, idealmente com o mesmo número de amostras em cada conjunto de idade, distribuição mais homgênia da resistência à compressão e menor variância entre as amostras.
</PT>

### 2020-04-25 templates/Materials_and_methods__Data_preparation.Rmd
Add templates

<EN>
## Data preparation

The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies. Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed. In addition, the relevant data has been transformed to better fit the studies in the next steps.
</EN>

<PT>
## Preparação dos dados

A preparação dos dados consistiu em transformar o conjunto de amostra afim de manter apenas dados relevantes para os estudos subsequêntes. Foram retirandos dados considerados irrelevantes ou que com potencial de adicionar um ruído indesejavel nas análises. Além disso, os dados relevântes foram transformados para melhor se enquadrar para os estudos nas próximas etapas.
</PT>

<EN>
### Initial data cleaning

Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).
</EN>

<PT>
### Limpeza inicial dos dados

Inicialmente foram observadas a existência de 25 amostras duplicadas que foram retiradas, resultando em um novo total de 1005 amostras (\ref{show-removing-duplicated-samples}).
</PT>

&nbsp;

```{r removing-duplicated-samples, echo=F}
#<EN># Removing duplicated samples</EN>
#<PT># Removendo amostras duplicadas</PT>

n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```

<EN>
The data show the variables in the columns and samples in the rows. However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
</EN>

<PT>
Os dados apresentam as variaveis nas colunas e amostras nas linhas. Porém foi verificado que algumas amostras são identicas em proporções de ingredientes, alterando apenas o valor da idade e resistência à compressão, por exemplo as amostras 653, 654, 678 e 681, mostradas na tabela \ref{tab:similar-samples}.
</PT>

&nbsp;

```{r similar-samples, echo=FALSE}
#<EN># Tabela - Samples with same composition</EN>
#<PT># Tabela - Amostras com a mesma composição</PT>

same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))

#<EN>
caption <- "Samples with same composition"
#</EN>
#<PT>
caption <- "Amostras com a mesma composição"
#</PT>

kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

<EN>
In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process. This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.
</EN>

<PT>
Além disso, também existem amostras com os mesmos valores e proporções de ingredientes, mas com resistência à compressão diferente, provavelmente devido a diferenças na execução. É o caso por exemplo das amostras 472, 473 e 474, mostradas na tabela \ref{tab:similar-samples-2}.
</PT>

&nbsp;

```{r similar-samples-2, echo=FALSE}
#<EN># Tabela - Same samples with different results</EN>
#<PT># Tabela - Amostras iguais com resultados diferentes</PT>

same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))

#<EN>
caption <- "Same samples with different results"
#</EN>
#<PT>
caption <- "Amostras iguais com resultados diferentes"
#</PT>

kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```

<EN>
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).
</EN>

<PT>
Para facilitar a analise das amostras, todas as amostras que são iguais em relação aos ingredientes, foram atribuidos o mesmo *id*. Além disso, como a resistência à compressão aos 28 dias é o parâmetro para determinar a classe do concreto, foi mantido apenas os elementos que contenham esse dia entre suas amostras. No caso das amostras iguais mas com resultado diferentes de resistência à compressão, foi calculado a média dos valores. Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes (\ref{show-total-de-amostras-apos-processamento}).
</PT>

&nbsp;

```{r initial-data-cleaning, echo=FALSE}
#<EN># Data cleaning</EN>
#<PT># Limpeza inicial das amotras</PT>

dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()

dat <- dat[!duplicated(select(dat, -c(id))),]

dat$id<-factor(dat$id)

n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```

<EN>
The result can be seen in the table \ref{tab:similar-samples-same-id}. All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.
</EN>

<PT>
O resultado pode ser conferido na tabela \ref{tab:similar-samples-same-id}. Todas as amostras com configurações iguais de ingredientes possuem o mesmo id, e quando possuiam resultados diferentes para os mesmos dias, foram transformadas em apenas uma amostra, com a média aritmética na resistência à compressão.
</PT>

&nbsp;

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
#<EN># Tabela - Previous samples after processing</EN>
#<PT># Tabela - Amostras anteriores após processamento</PT>

same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)

#<EN>
caption <- "Previous samples after processing"
#</EN>
#<PT>
caption <- "Amostras anteriores após processamento"
#</PT>

kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))

```

<EN>
### Age selection

As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models. However, it is necessary to verify how relevant the data of these other ages are. Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}.
</EN>

<PT>
### Seleção das idades

Como descrito anteriormente, a principal idade para análise da resistência à compressão é aos 28 dias, mas as outras idades também podem ser utilizadas para construir modelos de previsão. Porém é necessário verificar o quanto relevante os dados dessas outras idades são. Iniciando pela distribuição das amostras em relação a cada idade (\ref{show-boxplot}) mostrado na figura \ref{fig:boxplot}.
</PT>

&nbsp;

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
#<EN># Figure - Compressive strength (MPa) vs age (days)</EN>
#<PT># Figura - Resistência à compressão (MPa) vs idade (dias)</PT>

#<EN>
cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
ylabel <- "Compressive strength (MPa)"
xlabel <- "Age (days)"
#</EN>
#<PT>
cap <- "Boxplot - Resistência à compressão (MPa) vs idade (dias)"
ylabel <- "Resistência à compressão (MPa)"
xlabel <- "Idade (dias)"
#</PT>

dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```

<EN>
It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.

&nbsp;

This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}). The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis. The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.
</EN>

<PT>
Foi observado que as idades de 90, 91 e 100 dias provavelmente representam extremos entre si das configurações de ingredientes, uma vez que são idades relativamente próximas porém com valores muito diferentes, especialmente para 90 e 91.

&nbsp;

Essa hipótese foi verificada utilizando o método de análise de componente principal, aplicado às amostras dessas 3 idades (\ref{show-pca-90-91-100}). A fígura \ref{fig:pca-90-91-100} mostra como as amostras se relacionam umas com as outras (quais são parecidas ou diferentes) e revelou como cada variável contribui para a análise. As duas primeiras dimensões representam $37\%$ e $24\%$ respectivamente da variância.
</PT>

&nbsp;

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
#<EN># Figure - Principal component analysis - 90, 91 e 100 days</EN>
#<PT># Figura - Análise componente principal - 90, 91 e 100 dias</PT>

dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))

#<EN>
cap <- "Principal component analysis - 90, 91 e 100 days"
colnames(dat_90_91_100) <- c(
   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
habillage <- dat_90_91_100$day
#</EN>
#<PT>
cap <- "Análise componente principal - 90, 91 e 100 dias"
colnames(dat_90_91_100) <- c(
   "Cim.","E.G.A.F.","Ci.Vo.","Agu.","Sup.","Ag.G.","Ag.M.","dia"
)
pca <- prcomp(select(dat_90_91_100, -c(dia)), scale = TRUE)
habillage <- dat_90_91_100$dia
#</PT>

fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```

<EN>
Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value. The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).
</EN>

<PT>
Outro ponto importante considerado, da própria natureza do concreto, é o fato da taxa de crescimento de sua resistência à compressão diminuir com o tempo, chegando a um certo valor de establidade. A figura \ref{fig:mpa-on-time} mostra a resistência à compressão ao longo dos dias para amostras com mais de 5 dados, ou seja, dados disponiveis para no mínimo 6 idades distintas (\ref{show-mpa-on-time}). 
</PT>

&nbsp;

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
#<EN># Figure - Compressive strength through time</EN>
#<PT># Figura - Resistência à compressão ao longo do tempo</PT>

#<EN>
cap <- "Compressive strength through time"
ylabel <- "MPa"
xlabel <- "Age (days)"
#</EN>
#<PT>
cap <- "Resistência à compressão ao longo do tempo"
ylabel <- "MPa"
xlabel <- "Idade (dias)"
#</PT>

dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)

dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```

<EN>
For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise. They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}. As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.
</EN>

<PT>
Pelos motivos apresentados nas figuras \ref{fig:boxplot}, \ref{fig:pca-90-91-100} e \ref{fig:mpa-on-time}, foi considerado que as idades de 90, 91 e 100 dias podem ser agrupadas para melhorar a leitura e diminuir o ruído das amostras. Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}, pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
</PT>

&nbsp;

```{r join-90-91-100, echo=F, eval=F}
#<EN># Joining 90, 91 and 100 days data</EN>
#<PT># Juntando amostras de 90, 91 e 100 dias</PT>

ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]

sum(duplicated(c(ind_90, ind_91, ind_100))) # 0

dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```

<EN>
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.
</EN>

<PT>
Mais um tópico analisado na seleção das idades foi a frequência observada de cada valor de idade após essa transformação dos 90, 91 em 100 dias, mostrada na figura \ref{fig:freq-ages}. Alguns valores de dias apresentam concentrações muito baixas de amostras, correndo o risco de prejudicar mais do que ajudar na criação dos modelos, logo elas foram removidas (\ref{show-remove-ages-lower-50}). O critério adotado foi manter apenas idades com frequência maior que 50, ou seja, apenas os valores de 3, 7, 14, 28, 56 e 100 dias.
</PT>

&nbsp;

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
#<EN># Figure - Ages frequency</EN>
#<PT># Figura - Frequência das idades</PT>

#<EN>
cap <- "Ages frequency"
ylabel <- "Frequency"
xlabel <- "Age (days)"
#</EN>
#<PT>
cap <- "Frequência das idades"
ylabel <- "Frequência"
xlabel <- "Idade (dias)"
#</PT>

dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```

```{r remove-ages-lower-50, echo=F}
#<EN># Removing ages with frequency lower than 50</EN>
#<PT># Removendo idades com frequência menor que 50</PT>

dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

<EN>
### Data reorganization

The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}). The result in the first samples after this processing is shown in the table \ref{tab:new-features}.
</EN>

<PT>
### Reorganização dos dados

As amostras foram agrupadas para manter apenas uma amostra distinta de cada conjunto de configuração das proporções dos ingredientes, adicionando novas variáveis/colunas para a resistência em cada idade (\ref{show-reorganizing-dat}). O resultado nas primeiras amostras após esse processamento é mostrado na tabela \ref{tab:new-features}.
</PT>

&nbsp;

```{r reorganizing-dat, echo=F}
#<EN># Data reorganization</EN>
#<PT># Reorganização das amotras</PT>

dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```

```{r new-features, echo=F}
#<EN># Table - First 6 samples after reorganization/EN>
#<PT># Tabela - Primeiras 6 amostras reorganizadas</PT>

#<EN>
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
#</EN>
#<PT>
caption <- "Primeiras 6 amostras reorganizadas"
colNames2 = c("ID", "Cimento", "E.G.A.F.", "C.Volante", "Água",
             "Superp.", "A.Graúdo", "A.Miúdo", "3 dias", "7 dias",
             "14 dias", "28 dias", "56 dias", "100 dias")
#</PT>

dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")

kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

<EN>
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
</EN>

<PT>
O número de amostras e amostras distintas após toda essa manipulação permaneceu o mesmo, um total de 416 (\ref{show-total-samples-2}).
</PT>

```{r total-samples-2, echo=F}
#<EN># Total samples/EN>
#<PT># Total de amostras</PT>

n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

<EN>
### Adding new variables

To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
</EN>

<PT>
### Adicionando novas variáveis

Para finalizar a preparação dos dados, novas colunas foram adicionadas ao conjunto de amostras (\ref{show-new-features-2}). Iniciando pela classe do concreto, por exemplo se a resistência à compressão está entre 25 e 30, recebe a classe *C25*. A inclusão da classe foi importânte pois a resistência em $MPa$ é uma variável contínua, que será utilizada nos modelos de regressão, mas a classe como variável discreta pode fornecer outro ângulo de visualização dos dados. Também foi adicionado o traço aproximado do concreto, que representa as proporções de agregados (miúdo e graúdo) para o cimento. Outras proporções entre os principais ingredientes também foram adicionadas. As novas variáveis são apresentadas na tabela \ref{tab:new-features-table}.
</PT>

```{r new-features-2, echo=F}
#<EN># Adding New features/EN>
#<PT># Adicionando novas variáveis</PT>

concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}

mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}

dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)

lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```

```{r new-features-table, echo=F}
#<EN># Table - New features/EN>
#<PT># Tabela - Novas variáveis</PT>

#<EN>
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
#</EN>
#<PT>
caption <- "Novas variáveis"
colNames7 = c("ID", "Classe","Traço aproximado",
             "Água / Cimento", "A.Miúdo / Cimento", 
             "A.Graúdo / Cimento", "A.Miúdo / A.Graúdo",
             "Água / A.Graúdo", "Água / A.Miúdo")
#</PT>

kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

### 2020-04-25 templates/Materials_and_methods__Data_visualization.Rmd
Add templates

<EN>
## Data visualization

In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.

### Descriptive statistics

The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}). The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data. As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age. The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).
</EN>

<PT>
## Visualização dos dados

Para avaliar a necessidade de mais manipulações antes da construção dos modelos, nesta etapa as 416 amostras já processadas foram visualizadas e analisadas.

### Estatistica descritiva

A tabela \ref{tab:stat-summ} apresenta os dados estatísticos das variáveis contínuas (\ref{show-stat-summ}). A linha de *Null* representa o número de valores zerados para os ingredientes, e a linha *NA* representa o número de dados faltando. Como as amostras foram filtradas para manter apenas conjuntos de amostras com valores conhecidos da resistência à compressão aos 28 dias, o número de *NAs* é zero para essa idade. A figura \ref{fig:stat-summ-categorical} apresenta os dados estatísticos das variáveis discretas (\ref{show-stat-summ-categorical}).

</PT>

&nbsp;

```{r stat-summ, echo=FALSE}
#<EN># Table - Descriptive statistics - continuous variables/EN>
#<PT># Tabela - Estatística descritiva - variáveis contínuas</PT>

summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))

#<EN>
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
#</EN>
#<PT>
caption <- "Estatística descritiva - variáveis contínuas"
colnames(summ) <- c("Amostras", "Null", "NA", "Min", "Max", "Intervalo",
              "Soma", "Mediana", "Média", "Erro padrão  da média", 
              "Intervalo de confiânça da média",
              "Variância", "Desvio Padrão", "Coeficiênte de variação")
rownames(summ) = c("Cimento", "E.G.A.F.", "Cinza Volante", "Água",
              "Superplast.", "A.Graúdo", "A.Miúdo", "3 dias", "7 dias",
              "14 dias", "28 dias", "56 dias", "100 dias",
              "Água/ Cimento", "A.Miúdo/ Cimento", 
              "A.Graúdo/ Cimento", "A.Miúdo / A.Graúdo",
              "Água / A.Graúdo", "Água / A.Miúdo")
#</PT>

kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```

```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
#<EN># Figure - Descriptive statistics - categorical variables/EN>
#<PT># Figura - Estatística descritiva - variáveis discretas</PT>

#<EN>
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
#</EN>
#<PT>
cap <- "Estatística descritiva - variáveis discretas"
name1 <- "Porcentagem"
name2 <- "Porcentagem acumulada"
ylabel <- "Frequencia"
xlabel1 <- "Classe"
xlabel2 <- "Traço aproximado"
#</PT>

format_percent = function(n){
  paste(n, "%", sep = "")
}

format_class <- function(cls){
  str_remove_all(cls, "C")
}

f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))

f_cls_labels = function(n) {
  f_class$class[n]
}

f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}

p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()

format_mix <- function(mix){
  str_remove_all(mix, ":")
}

f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))

f_mix_labels = function(n) {
  f_mix$mix[n]
}

f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}

p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()

grid.arrange(p1, p2, ncol=2)
```

<EN>
### Correlation between ingredients and compressive strength

The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}). The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).
</EN>

<PT>
### Correlação dos ingredientes e resistência à compressão

A figura \ref{fig:correlation} apresenta a correlação das variáveis para cada conjunto de idades (\ref{show-correlation}). A figura \ref{fig:correlation-mpa} apresenta os mesmos dados, mas em vez de correlacionar todos, correlaciona apenas com a resistência à compressão, mostrando os valores mais detalhadamente (\ref{show-correlation-mpa}).
</PT>

&nbsp;

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
#<EN># Figure - Correlation grouped by age/EN>
#<PT># Figura - Correlações em cada idade</PT>

cor_dat <- dat %>% select(-c(id))

#<EN>
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
                "14", "28", "56", "100",
                "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
#</EN>
#<PT>
cap <- "Correlações em cada idade"
f_lvl <- c("3 dias", "7 dias", "14 dias","28 dias", "56 dias", "100 dias")
name <- "Correlação"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                  "Superp.", "A.Graúdo", "A.Miúdo",
                  "3","7","14","28","56","100",
                  "class","mix_app" ,"Ág./Ci.",
                  "A.M./Ci.", "A.G./Ci.", "A.M./A.G.",
                  "Ág./A.G.", "Ág./A.M.")
#</PT>

colnames(cor_dat) <- colnames_dat

cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()

cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}

cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))

melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}

melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))

melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)

melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl

melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}

#<EN>
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
#</EN>
#<PT>
cap <- "Correlação das variáveis com a resistência à compressão no tempo"
name <- "Correlação"
#</PT>

melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

```

<EN>
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.
</EN>

<PT>
A interpretação da figura \ref{fig:correlation-mpa} sugere que a resistência do concreto está relacionada positivamente principalmente com os ingredientes cimento e superplastificante e negativamente com a água e agregado miúdo. Quanto menor a quantidade de cimento para os agregados e para água, mais negativamente estão correlacionados com a resistência à compressão.
</PT>

&nbsp;

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
#<EN># Figure - Relationship between approximated mix, water, MPa and age/EN>
#<PT># Figura - Relação entre o traço aproximado, água, MPa e idade</PT>

#<EN>
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
#</EN>
#<PT>
cap <- "Relação entre o traço aproximado, água, resistência à compressão e idade"
d <- " dias"
xlabel <- "Traço aproximado" 
ylabel <- "Resistência à compressão (MPa)"
label <- "Água /\nCimento"
#</PT>

mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")

mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()

mix_dat$day <- factor(mix_dat$day)
levels(mix_dat$day) <- labs

min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x

mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```

```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
#<EN># Figure - Relationship between concrete main features/EN>
#<PT># Figura - Relação das principais características do concreto</PT>

#<EN>
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
#</EN>
#<PT>
cap <- "Relação das principais proporções do concreto"
d <- " dias"
xlabel <- "Água / Cimento"
ylabel <- "Resistência à compressão (MPa)"
colour <- "Ag. Miúdo /\nCimento"
size <- "Ag. Graúdo /\nCimento"
#</PT>

min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2

mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```

<EN>
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

### Variables distribution

A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias.

The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
</EN>

<PT>
As figuras \ref{fig:mix-app-mpa} e \ref{fig:mix-mpa} mostram a relação entre os principais ingredientes (conhecida como traço) em relação à resistência à compressão (\ref{show-mix-app-mpa} e \ref{show-mix-mpa}). A interpretação dessas figuras mostra que quanto maior a quantidade de cimento em relação aos outros ingredientes, maior será a resistência à compressão.

### Distribuição das variáveis

A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
</PT>

&nbsp;

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
#<EN># Figure - Variables distribution/EN>
#<PT># Figura - Distribuição das variáveis</PT>

#<EN>
cap <- "Variables distribution"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
#</EN>
#<PT>
cap <- "Distribuição das variáveis"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                 "Superplastificante", "A.Graúdo", "A.Miúdo", "MPa",
                 "Água / Cimento", "A.Miúdo / Cimento", "A.Graúdo / Cimento",
                 "A.Miúdo / A.Graúdo","Água / A.Graúdo", "Água / A.Miúdo")
#</PT>

dist_dat <- dat %>%
  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))

colnames(dist_dat) <- colnames_dat

dist_dat <- dist_dat %>%
  gather("Var", "value") %>%
  mutate(value = as.numeric(value))

dist_dat %>%
  ggplot(aes(value)) +
  geom_histogram(aes(y = ..density..), 
                 colour = "black", 
                 fill = "white") +
  geom_density(alpha = .5, fill = "lightseagreen") +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  xlab("") +
  ylab("") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

<EN>
The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}. It shows that the resistance to compression gradually increases over time, as expected. Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
</EN>

<PT>
De outra forma, a figura \ref{fig:vars-distribution-time} mostra a distribuição dos ingredientes e da resistência à compressão para cada conjunto de idades (\ref{show-vars-distribution-time}), ou seja, no caso dos 28 dias apresenta a mesma informação que a figura \ref{fig:vars-distribution}. Através dela é visualizado que como esperado, a resistência à compressão gradualmente aumenta ao longo do tempo. Além disso é visualizado que a concentração dos ingredientes podem variar muito quando estratíficado pelas idades.
</PT>

```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
#<EN># Figure - Variables distribution grouped by age/EN>
#<PT># Figura - Distribuição das variáveis agrupadas por idade</PT>

days_labs <- c("3","7","14","28","56","100")

#<EN>
cap <- "Variables distribution grouped by age"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  days_labs,
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
#</EN>
#<PT>
cap <- "Distribuição das variáveis em relação a idade"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                       "Superplastificante", "A.Graúdo", "A.Miúdo", 
                       days_labs,
                       "Água / Cimento", "A.Miúdo / Cimento", "A.Graúdo / Cimento",
                       "A.Miúdo / A.Graúdo","Água / A.Graúdo", "Água / A.Miúdo")
#</PT>

dist_dat_2 <- dat %>%
  select(-c(id, class, mix_app))

colnames(dist_dat_2) <- colnames_dat

dist_dat_2 <- dist_dat_2 %>%
  gather("day", "mpa", days_labs) %>%
  drop_na() %>%
  gather("Var", "value", -c("day")) %>%
  mutate(day = factor(day, levels = days_labs))

dist_dat_2 %>%
  ggplot(aes(x = day, y = value)) +
  geom_violin(color = NA,
                fill = "lightseagreen",
                alpha = .5,
                na.rm = TRUE,
                scale = "count") +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  xlab("") +
  ylab("")

```

<EN>
### Principal component analysis

Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.

In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}). The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that. It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
</EN>

<PT>
### Análise de componente principal

Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até *20 MPa*, normal até *40 MPa*, médio até *70 MPa* e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
</PT>

```{r pca, echo=F, fig.height=3, fig.cap=cap}
#<EN># Figure - Principal component analysis on ingredients/EN>
#<PT># Figura - Análise componente principal nos ingredientes</PT>

#<EN>
cap <- "Principal component analysis on ingredients"
colnames_dat <- c(
  "Cem.", "B.F.S.", "F.A.", "Wat.", "Sup.", "C.Agg.",
  "F.Agg.","WxC", "F.Agg.xC","C.Agg.xC","F.Agg.xC",
  "WxC.Agg.","WxF.Agg."
)
class_list <- list(low="0", normal="20", medium="40", high="70")
#</EN>
#<PT>
cap <- "Análise componente principal nos ingredientes"
colnames_dat <- c(
  "Cim.", "E.G.A.F.", "Ci.Vo.", "Águ.", "Sup.", "Ag.G.",
  "Ag.M.","AxC", "MxC","GxC","MxC","AxG","AxM"
)
class_list <- list(baixo="0", normal="20", médio="40", alto="70")
#</PT>

dat_pca <- dat %>%
  select(-c(id, class, mix_app,
            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))

colnames(dat_pca) <- colnames_dat

class_2 <- dat$day_28

class_2[class_2 < 20] = "0"
class_2[class_2 >= 20 & class_2 < 40] = "20"
class_2[class_2 >= 40 & class_2 < 70] = "40"
class_2[class_2 >= 70] = "70"

class_2 <- factor(class_2)
levels(class_2) <- class_list

pca <- prcomp(dat_pca, scale = TRUE)

fviz_pca_ind(
  pca,
  geom.ind = "point",
  habillage=class_2,
  addEllipses = T,
  ellipse.level=0.95) +
  ggtitle("") +
  theme_bw()
```

### 2020-04-25 templates/Materials_and_methods__Machine_learning_models.Rmd
Add templates

<EN>
## Machine learning models

The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.

### Pre-processing and data separation

Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado.

As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.
</EN>

<PT>
## Modelos de machine learning

O desenvolvimento dos modelos de machine learning foi realizado com o pacote *caret* [@caret] e baseado em @irizarry2019 e @Kuhn2008.

### Pre processamento e separação dos dados

Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado. 
</PT>

&nbsp;

```{r dummy-var, echo=F, message=F, warning=F}
#<EN># Dummy variables</EN>
#<PT># Variáveis fictícias - dummy vars</PT>

dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))

dummyDat$id <- dat$id

dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```

<EN>
The samples were separated based on age. One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}). For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.
</EN>

<PT>
As amostras foram separadas baseado nas idades. Foram criados um conjunto de dados para cada valor de idade, totalizando 6 conjuntos diferentes (\ref{show-preparation}). Para fins ilustrativos, as primeiras 18 de 45 colunas das primeiras 6 amostras do conjunto de 28 dias são mostradas na tabela \ref{tab:table-preparated-samples}.
</PT>

&nbsp;

```{r preparation, echo=F}
#<EN># Data preparation</EN>
#<PT># Preparação dos dados</PT>

names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")

dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```

```{r table-preparated-samples, echo=F}
#<EN># Table - First 18 columns of the 6 first samples of 28 days</EN>
#<PT># Tabela - Primeiras 18 colunas das primeiras 6 amostras de 28 dias</PT>

#<EN>
colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
caption <- "First 18 columns of the 6 first samples of 28 days"
#</EN>
#<PT>
colNames = c("ID", "Cimento", "E.G.A.F.", "C.Vol.", "Água",
              "Superp.", "A.Graúdo", "A.Miúdo", "MPa", "Classe",
              "Ág./", "A.M./", "A.G./",
              "A.M./","Ág./", "Ág./", "Traço Apox." = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ci.", "Ci.", "A.G.", "A.G.", "Ag.M.", "1:1:2", "1:2:2")
caption <- "Primeiras 18 colunas das primeiras 6 amostras de 28 dias"
#</PT>

dat_tabela <- dat_28[1:18]

kable(
    head(dat_tabela[order(dat_tabela$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

<EN>
For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}). Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact. In addition to them, in the case of the 7-day set, the fly ash variable was also removed. Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}). After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.
</EN>

<PT>
Para cada um dos 6 conjuntos foi verificado a existência ou não de variáveis com variância próxima a zero e sua subsequênte remoção (\ref{show-nzv}). Muitas das 23 variáveis adicionadas referentes ao traço aproximado do concreto foram removidas devido a esse fato. Além delas, no caso do conjunto de 7 dias, a variável cinza volante também foi removida. Depois foi verificado que não existem variáveis com alta correlação, acima de 0,999 em nenhum dos 6 conjuntos de dados (\ref{show-cors}). Após essas etapas, os conjuntos de amostras apresentaram 24, 21, 23, 23, 24 e 25 colunas respectivamente para as idades em sequência crescente.
</PT>

&nbsp;

```{r nzv, echo=F}
#<EN># Removing near zero variance columns</EN>
#<PT># Removendo colunas com variância próxima a zero</PT>

nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)

dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```

```{r cors, echo=F}
#<EN># High correlation variables verification</EN>
#<PT># Verificação de variáveis com alta correlação</PT>

descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))

high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```

<EN>
The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package. If performed at this time, it would be necessary to manually undo these transformations in the predictions. The *caret* allows you to transform before training the models and already transforms the results back.

&nbsp;

Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}). The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
</EN>

<PT>
A etapa de centralização e normalização das variáveis foi realizada mais a frente, junto com a aplicação dos modelos, pois é mais simples fazer dessa forma com o pacote *caret*. Se fosse realizada nesse momento, seria nescessário manualmente desfazer essas transformações nas previsões. O *caret* permite transformar antes do treino dos modelos e já transforma de volta os resultados.

&nbsp;

Cada um dos conjuntos de dados foi separado em conjuntos de teste e treino, *20%* e *80%* respectivamente (\ref{show-split}). A figura \ref{fig:dist-split} mostra como ficou a distribuição dos dados entre os conjuntos em relação a resistência à compressão para cada modelo (\ref{show-dist-split}).
</PT>

```{r split, echo=F, message=F, warning=F}
#<EN># Test and train data split</EN>
#<PT># Separação em conjunto de teste e treino</PT>

reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$MPa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$MPa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$MPa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$MPa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$MPa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$MPa)
)

reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )

split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}

trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)

gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}

dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```

```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
#<EN># Distribution of test and train data</EN>
#<PT># Distribuição dos conjuntos de teste e treino</PT>

#<EN>
cap <- "Distribution of test and train data"
d_lab <- " days"
ylabel <- "Density"
g1 <- "Train"
g2 <- "Test"
#</EN>
#<PT>
cap <- "Distribuição dos conjuntos de teste e treino"
d_lab <- " dias"
ylabel <- "Densidade"
g1 <- "Treino"
g2 <- "Teste"
#</PT>

dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Grup = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Grup = g2, day = str_sub(n, 2))
    )
  dens
}

densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))

labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs

densities %>%
  ggplot(aes(y, fill = Grup)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~dia_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

<EN>
### Performance measures

The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*). The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.

### Naive models

Before creating the real models, for comparison purposes, naive models were created. They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}). In other words, naive models are simply the best guess possible. The results can be checked in the table \ref{tab:table-naive-model-reg}.
</EN>

<PT>
### Medidas de performance

A avaliação da performance dos modelos foi realizada pela Raiz do Erro Quadrático Médio (*RMSE*). O *RMSE* é a medida utilizada em todos os trabalhos citados na introdução e permitirá a comparação dos modelos na discussão.

### Modelos ingênuos

Antes de criar os modelos verdadeiros, para fins de comparação, foram criados modelos ingênuos. Eles simplesmente prevêem que a resistência à compressão do conjunto de teste, é a média da resistência à compressão do conjunto de treino (\ref{show-naive-model-reg}). Em outras palavras, os modelos ingênuos são simplesmente o melhor palpite possível. Os resultados podem ser conferidos na tabela \ref{tab:table-naive-model-reg}.
</PT>

```{r naive-model-reg, echo=F, message=F, warning=F}
#<EN># Naive model</EN>
#<PT># Modelo ingênuo</PT>

res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}

ing_model <- imap(dats_reg, res)

get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}

rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)

ing_model_df <- ing_model %>% reduce(rbind)

df_performance_reg <- full_join(ing_model_df, rmses)

```

```{r table-naive-model-reg, echo=F}
#<EN># Table - Naive models</EN>
#<PT># Tabela - Modelos ingênuo</PT>

#<EN>
colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
caption <- "Naive models"
#</EN>
#<PT>
colNames = c("Idade", "Média $MPa$ (treino)", "RMSE (treino)", "RMSE (teste)")
caption <- "Modelos ingênuos"
#</PT>

kable(
    df_performance_reg,
    col.names = colNames6,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

<EN>
### Choice of algorithms

The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models. The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result. According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).

### Regression models

As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:

1. All *features*;
2. All *features* without the *dummy vars*;
3. Only the original *features*;
4. Only new *features*;
5. Only new *features*, without *dummy vars*;

Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.

For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set. The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
</EN>

<PT>
### Escolha do algorítimo

O pacote *caret* [@caret] expõe mais de 200 algorítimos diferentes para criar modelos de *machine learning*. A documentação do pacote apresenta um código inicial [@modelsClusters] como sugestão para selecionar um portifólio de algóritmos mais distintos possíveis em relação à algum algorítimo pré-selecionado, mas para agilidade e devido a limitações técnicas, foi escolhido utilizar um algorítimo com a maior probabilidade de atingir o melhor resultado possível. Segundo @Fernandez2014, que comparou 179 algorítimos em 121 banco de dados diferentes, o algorítimo mais provável de atingir os melhores resultados possíveis é o *Parallel Random Forest* (denominado *prRF* no *caret*).

### Modelos de regressão

Como ao longo do processamento foram adicionadas novas variáveis (as relações entre os ingredientes e mais algumas *dummy vars* para cada conjunto de idade), foram estudadas 5 possibilidades de configurações das *features* para os modelos:

1. Todas as *features*;
2. Sem as *dummy vars*;
3. Apenas *features* originais;
4. Apenas *features* adicionadas;
5. Apenas *features* adicionadas, sem as *dummy vars*;

Construindo um modelo para cada uma dessas configurações utilizando o conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou que a melhor opção é a configuração 2, ou seja, as *dummy vars* foram completamente descartadas, porém foram mantidas as outras novas varáveis.

Para fins ilustrativos, a tabela \ref{tab:table-done-samples-28} mostra as primeiras 6 amostras do conjunto de treino do modelo de 28 dias. As amostras dos outros modelos, dos conjuntos de teste e treino são similares, sendo a única diferença no modelo de 7 dias, que exlcui a cinza volante devido a variância próxima a zero, realizado anteriormente.
</PT>

```{r test-models-reg, echo=F, warning=F, message=F, include=F}
#<EN># Features selection</EN>
#<PT># Escolha das caracterísitcas (features)</PT>

data1 <- dats_reg$d28$train # full
data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
data4 <- dats_reg$d28$train[c(8:21)] # only new variables
data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars

test1 <- dats_reg$d28$test # full
test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
test4 <- dats_reg$d28$test[c(8:21)] # only new variables
test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars

# parRF
modelLookup("parRF")
control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')

fit_parRF<- function(data, n) {
  set.seed(1, sample.kind = "Rounding")
  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
  train(y ~ ., 
    data = data,
    preProcess = c("center","scale"),
    method='parRF', 
    tuneGrid=tunegrid_parRF,
    trControl=control_parRF)
}

fit_parRF1 <- fit_parRF(data1, 3)
ggplot(fit_parRF1)
min(fit_parRF1$results$RMSE) # 6.26108
p1_parRF1 <- predict(fit_parRF1, newdata = test1)
RMSE(p1_parRF1, test1$y) # 4.887264

fit_parRF2 <- fit_parRF(data2, 1)
ggplot(fit_parRF2)
min(fit_parRF2$results$RMSE) # 6.268789
p_parRF2 <- predict(fit_parRF2, newdata = test2)
RMSE(p_parRF2, test2$y) # 4.812994

fit_parRF3 <- fit_parRF(data3, 1)
ggplot(fit_parRF3)
min(fit_parRF3$results$RMSE) # 6.361905
p_parRF3 <- predict(fit_parRF3, newdata = test3)
RMSE(p_parRF3, test3$y) # 5.083467

fit_parRF4 <- fit_parRF(data4, 1)
ggplot(fit_parRF4)
min(fit_parRF4$results$RMSE) # 8.705984
p_parRF4 <- predict(fit_parRF4, newdata = test4)
RMSE(p_parRF4, test4$y) # 8.083962

fit_parRF5 <- fit_parRF(data5, 1)
ggplot(fit_parRF5)
min(fit_parRF5$results$RMSE) # 8.691755
p_parRF5 <- predict(fit_parRF5, newdata = test5)
RMSE(p_parRF5, test5$y) # 7.991537
```

```{r table-done-samples-28, echo=F}
#<EN># Table - First 6 samples of train data of the 28 days model</EN>
#<PT># Tabela - Primeiras 6 amostras do conjunto de treino do modelo de 28 dias</PT>

#<EN>
caption <- "First 6 samples of train data of the 28 days model"
colNames = c("Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","Ce.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "MPa")
#</EN>
#<PT>
caption <- "Primeiras 6 amostras do conjunto de treino do modelo de 28 dias"
colNames = c("Cimento", "E.G.A.F.", "C.Vol.", "Água",
              "Superp.", "A.Graúdo", "A.Miúdo",
              "Ág./", "A.M./", "A.G./",
              "A.M./","Ág./", "Ág./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$", "Ci.",
              "Ci.", "Ci.", "A.G.", "A.G.", "Ag.M.", "$MPa$")
#</PT>

colNames2 = c("Features"=13, "Outcome"=1)

kable(
    head(data2),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    add_header_above(header = colNames2, line = T, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

<EN>
For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}). For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.
</EN>

<PT>
Para cada conjunto de idade foi criado um modelo utilizando o algorítimo *Parallel Random Forest*, definido anteriormente (\ref{show-reg-models}). Para cada um dos 6 modelos, o parâmetro *mtry* foi otimizado, e foi realizado *repeated cross-validation*, dividindo em 10 ou 30 partes e repetindo 10 vezes.
</PT>

```{r reg-models, echo=F, warning=F, message=F, include=F}
#<EN># Regression models</EN>
#<PT># Modelos de regressão</PT>

# no dummy vars:
data3 <- dats_reg$d3$train[c(1:13, 22)]
data7 <- dats_reg$d7$train[c(1:12, 19)]
data14 <- dats_reg$d14$train[c(1:13, 21)]
data28 <- dats_reg$d28$train[c(1:13, 21)]
data56 <- dats_reg$d56$train[c(1:13, 22)]
data100 <- dats_reg$d100$train[c(1:13, 23)]

test3 <- dats_reg$d3$test[c(1:13, 22)]
test7 <- dats_reg$d7$test[c(1:12, 19)]
test14 <- dats_reg$d14$test[c(1:13, 21)]
test28 <- dats_reg$d28$test[c(1:13, 21)]
test56 <- dats_reg$d56$test[c(1:13, 22)]
test100 <- dats_reg$d100$test[c(1:13, 23)]

get_trControl <- function(n, r){
  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
}

trControl3 <- get_trControl(30, 10)
trControl7 <- get_trControl(10, 10)
trControl14 <- get_trControl(30, 10)
trControl28 <- get_trControl(30, 10)
trControl56 <- get_trControl(30, 10)
trControl100 <- get_trControl(10, 10)

get_tuneGrid <- function(data){
  expand.grid(mtry = seq(1, length(data), 1))
}

tuneGrid3 <- get_tuneGrid(data3)
tuneGrid7 <- get_tuneGrid(data7)
tuneGrid14 <- get_tuneGrid(data14)
tuneGrid28 <- get_tuneGrid(data28)
tuneGrid56 <- get_tuneGrid(data56)
tuneGrid100 <- get_tuneGrid(data100)

set.seed(1, sample.kind = "Rounding")
fit_3 <- train(y ~ ., 
            data = data3,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid3,
            trControl=trControl3)
p_3 <- predict(fit_3, newdata = test3)
RMSE_test_3 <- RMSE(p_3, test3$y)
RMSE_test_3 # 3.31037
fit_3$bestTune # mtry = 6

set.seed(1, sample.kind = "Rounding")
fit_7 <- train(y ~ ., 
            data = data7,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid7,
            trControl=trControl7)
p_7 <- predict(fit_7, newdata = test7)
RMSE_test_7 <- RMSE(p_7, test7$y)
RMSE_test_7 # 4.361987
fit_7$bestTune # mtry = 2

set.seed(1, sample.kind = "Rounding")
fit_14 <- train(y ~ ., 
            data = data14,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid14,
            trControl=trControl14)
p_14 <- predict(fit_14, newdata = test14)
RMSE_test_14 <- RMSE(p_14, test14$y)
RMSE_test_14 # 4.620515
fit_14$bestTune # mtry = 13

set.seed(1, sample.kind = "Rounding")
fit_28 <- train(y ~ ., 
            data = data28,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid28,
            trControl=trControl28)
p_28 <- predict(fit_28, newdata = test28)
RMSE_test_28 <- RMSE(p_28, test28$y)
RMSE_test_28 # 4.716698
fit_28$bestTune # mtry = 11

set.seed(1, sample.kind = "Rounding")
fit_56 <- train(y ~ ., 
            data = data56,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid56,
            trControl=trControl56)
p_56 <- predict(fit_56, newdata = test56)
RMSE_test_56 <- RMSE(p_56, test56$y)
RMSE_test_56 # 5.939163
fit_56$bestTune # mtry = 8

set.seed(1, sample.kind = "Rounding")
fit_100 <- train(y ~ ., 
            data = data100,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid100,
            trControl=trControl100)
p_100 <- predict(fit_100, newdata = test100)
RMSE_test_100 <- RMSE(p_100, test100$y)
RMSE_test_100 # 5.851088
fit_100$bestTune # mtry = 8

```

### 2020-04-25 templates/Materials_and_methods__Materials.Rmd
Add templates

<EN>
# Materials and Methods

## Materials

The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore]. Throughout the process, all code executed was documented in the same order as its execution in [Appendix 3](#appendix3), and reference was always made to codes throughout the text. All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1). In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in [Appendix 2](#appendix2).
</EN>

<PT>
# Metodologia

## Materiais utilizados

A metodologia foi realizada utilizando o software RStudio [@RStudio], ambiente virtual integrado para desenvolvimento de código na linguagem *R* [@RCore]. Ao longo do processo, todo código executado foi documentado na mesma ordem de sua execução no [Apêndice 3](#appendix3), e foi sempre realizada referência aos códigos ao longo do texto. Todas as informações relevantes relacionados ao sistema operacional e pacotes instalados foram apresentados no [Apêndice 1](#appendix1). Além disso foi criado um repositorio online e de código aberto no *Github*, abrigando todo o código utilizado para gerar esse trabalho, o link foi diponibilizado no [Apêndice 2](#appendix2).
</PT>

### 2020-04-25 templates/Materials_and_methods__Obtaining_the_data.Rmd
Add templates

<EN>
## Obtaining the data

The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}). In total there are 1030 samples with 9 columns. The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}). The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}). The first samples can be viewed in the table \ref{tab: first-samples}.
</EN>

<PT>
## Obtenção dos dados

O download dos dados foi realizado no website da Universidade da California Irvine [@downloadData] (\ref{show-download-data}). No total são 1030 amostras com 9 colunas. As amostras foram renomeadas e foi adicionado uma coluna de id para facilitar na manipulação dos dados (\ref{show-rename-dat-cols}). As colunas foram reordenadas para colocar a nova coluna id em primeira posição (\ref{show-reorder-dat}). As primeiras amostras podem ser visualizadas na tabela \ref{tab:first-samples}.
</PT>

&nbsp;

```{r download-data, echo=F, include=F}
#<EN># Data download</EN>
#<PT># Download dos dados</PT>
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")

n_inicial_samples <- nrow(dat)
colnames(dat)
```

```{r rename-dat-cols, echo=F}
#<EN># Ranaming the columns</EN>
#<PT># Renomeando as colunas</PT>
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )

dat$id <- seq.int(nrow(dat))
```

```{r reorder-dat, echo=F}
#<EN># Reordering the data</EN>
#<PT># Reordenando os dados</PT>
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```

```{r col-names-and-units, echo=F}
#<EN># Defining column names and units</EN>
#<PT># Definindo nomes e unidades das colunas</PT>

#<EN>
colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
#</EN>

#<PT>
colNames <- c("ID", "Cimento", "E.G.A.F", "C.Volante", "Água",
             "Superp.", "A.Graúdo", "A.Miúdo", "Dia", "Comp.Str.")
#</PT>

dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```

```{r first-samples, echo=F}
#<EN># Table First samples</EN>
#<PT># Tabela Primeiras amostras</PT>

#<EN>
caption <- "First 6 samples"
#</EN>
#<PT>
caption <- "Primeiras 6 amostras"
#</PT>

kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### 2020-04-25 templates/Results.Rmd
Add templates

<EN>
# Results

The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively. The table \ref{tab:table-reg-models} presents the details and results of each model (\ref{show-table-reg-models}). The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).
</EN>

<PT>
# Results

O *RMSE* de teste de cada modelo em ordem crescente de idade foi respectivamente `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` e `r round(RMSE_test_100,2)`. A tabela \ref{tab:table-reg-models} apresenta os detalhes e resultados de cada modelo (\ref{show-table-reg-models}). A figura \ref{fig:results-comparison} compara os valores reais e previstos (\ref{show-results-comparison}), e as tabelas seguintes mostram os melhores e piores resultados de cada modelo (\ref{show-table-10}).
</PT>

```{r table-reg-models, echo=F}
#<EN># Table - Models details</EN>
#<PT># Tabela - Detalhes dos modelso</PT>

#<EN>
colNames <- c("Model", "mtry", "CV", "Repetitions", 
              "RMSE (train)", "RMSE (test)")
caption <- "Regression models results"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
#</EN>
#<PT>
colNames <- c("Modelo", "mtry", "CV", "Repetições", 
              "RMSE (treino)", "RMSE (teste)")
caption <- "Resultados dos modelos de regressão"
day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
#</PT>

dat_reg_models <- data.frame(
  dia = day,
  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
  number = c(trControl3$number, trControl7$number, trControl14$number,
             trControl28$number,trControl56$number,trControl100$number),
  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
             trControl28$repeats,trControl56$repeats,trControl100$repeats),
  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)

kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
#<EN># Figure - Models comparison</EN>
#<PT># Figura - Comparação dos modelos</PT>

#<EN>
cap <- "Actual vs Predicted values in each model"
models <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
xlabel <- "Actual (MPa)"
ylabel <- "Predicted (MPa)"
#</EN>
#<PT>
cap <- "Comparação dos valores reais e previstos em cada modelo"
models <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
xlabel <- "Real (MPa)"
ylabel <- "Previsto (MPa)"
#</PT>

preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)

gen_res_df <- function(ind){
  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
}

res_df <- lapply(1:6, gen_res_df)
res_df <- bind_rows(res_df)
res_df$model <- factor(res_df$modelo,levels=modelos)

res_df %>%
  ggplot(aes(actual, pred)) +
  facet_wrap(~ modelo, ncol=3) +
  geom_point(alpha=0.5) +
  theme_bw() +
  geom_abline(slope=1, intercept=0) +
  xlab(xlabel) +
  ylab(ylabel)
```

```{r table-10, echo=F}
#<EN># Results tables of 10 best and worst results</EN>
#<PT># Tabelas dos 10 melhores e piores resultados</PT>

#<EN>
colNames1 = c("Actual", "Predicted", "Error", "","Actual", "Predicted", "Error") 
colNames2 = c("Best 10"=3,"", "Worst 10"=3) 
caption_0 <- "Model of "
#</EN>
#<PT>
colNames1 = c("Real", "Previsto", "Erro", "","Real", "Previsto", "Erro") 
colNames2 = c("10 melhores"=3,"", "10 piores"=3)
caption_0 <- "Modelo de "
#</PT>

get_X <- function(pred, actual, X=10){
  diff <- abs(pred - actual)
  diff_2 <- pred - actual
  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
  df_min <- data.frame(
    actual_min=actual[ind_min],
    pred_min=pred[ind_min],
    diff_min=diff[ind_min],
    diff_min_2=diff_2[ind_min]
  )
  df_max <- data.frame(
    actual_max=actual[ind_max],
    pred_max=pred[ind_max],
    diff_max=diff[ind_max],
    diff_max_2=diff_2[ind_max]
  )
  df_max <- df_max[order(-df_max$diff_max),]
  df_min <- df_min[order(df_min$diff_min),]
  df_null <- data.frame(null_col = rep(c(""), X))
  res <- cbind(df_min,df_null, df_max)
  res <- res %>% select(-c(diff_max, diff_min))
  rownames(res) <- c()
  res
}

gen_kable <- function(ind){
  df <- get_X(preds[[ind]], actuals[[ind]])
  caption <- paste0(caption_0,models[ind])
  kable(
    df,
    col.names = colNames1,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  column_spec(4, width = "1cm",) %>%
  add_header_above(header = colNames2, line = T, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position"))
}

gen_kable(1)
gen_kable(2)
gen_kable(3)
gen_kable(4)
gen_kable(5)
gen_kable(6)
```

### 2020-04-25 generated/paper_EN.Rmd
Add generated files

---
title:
  Concrete compressive strength prediction with machine learning
date: "April 22, 2020"
author: "Pedro Bernardino Alves Moreira"
indent: true
abstract:
 Compressive strength is the main characteristic of concrete. This work used a set of samples widely used to build 6 forecast models of this characteristic, each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers) and 6 relationships between the main ingredients (water/cement, coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate, water / coarse aggregate and water/fine aggregate) to make the predictions. The algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*) . The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31) and the worst for the 100-day model (*RMSE* of 5.85). The 28-day model had *RMSE* of 4.72. The work indicated that the separation in a model for each age is promising and was able to generate compatible results or better than other studies, opening way for new models creation following this proposal, but using different algorithms or the combination of different algorithms.
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
bibliography: references.bib
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{supertabular}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{caption}
- \usepackage{floatrow}
- \floatsetup[table]{capposition=top}
- \floatsetup[figure]{capposition=top}
- \captionsetup{options=chunk}
- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
- \renewcommand{\thechunk}{\arabic{chunk}}
- \usepackage{indentfirst}
- \usepackage{sectsty} \sectionfont{\centering}
---
```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```
\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

# Introduction
Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens. Measurement at 28 days is mandatory and represents the grade of the concrete. Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.

&nbsp;

This compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed. @Hasan2011 proposed a mathematical model to predict from the results of tests of 7 and 14 days, and @Kabir2012 from 7 days. However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.

&nbsp;

Many previous studies use the same dataset used by @Yeh1998 to predict the compressive strength of concrete. @Alshamiri2020 got good results with the regularized extreme learning machine (RELM) technique, and @Hameed2020 got even better results with the Artificial Neural Networks and cross-validation technique. This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as @Abban2016, @Raj2018, @Modukuru2020 and @Pierobon2018. At the end of the work, the results found are compared to the works cited here.

&nbsp;

Unlike previous studies with this dataset, this work does data preparation differently. The age of the concrete is the most unique feature that contributes to its compressive strength. For this reason, age is treated separately in the machine learning models, creating models for each age group.

# Materials and Methods

## Materials
The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore]. Throughout the process, all code executed was documented in the same order as its execution in [Appendix 3](#appendix3), and reference was always made to codes throughout the text. All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1). In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in [Appendix 2](#appendix2).

## Reproducibility
In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result. The *seeds* can be checked throughout [Appendix 3](#appendix3) or directly on *Github*.

## Obtaining the data
The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}). In total there are 1030 samples with 9 columns. The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}). The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}). The first samples can be viewed in the table \ref{tab: first-samples}.

&nbsp;

```{r download-data, echo=F, include=F}
# Data download
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```
```{r rename-dat-cols, echo=F}
# Ranaming the columns
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )
dat$id <- seq.int(nrow(dat))
```
```{r reorder-dat, echo=F}
# Reordering the data
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```
```{r col-names-and-units, echo=F}
# Defining column names and units
colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```
```{r first-samples, echo=F}
# Table First samples
caption <- "First 6 samples"
kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

## Data preparation
The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies. Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed. In addition, the relevant data has been transformed to better fit the studies in the next steps.

### Initial data cleaning
Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

&nbsp;

```{r removing-duplicated-samples, echo=F}
# Removing duplicated samples
n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```
The data show the variables in the columns and samples in the rows. However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.

&nbsp;

```{r similar-samples, echo=FALSE}
# Tabela - Samples with same composition
same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))
caption <- "Samples with same composition"
kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process. This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.

&nbsp;

```{r similar-samples-2, echo=FALSE}
# Tabela - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).

&nbsp;

```{r initial-data-cleaning, echo=FALSE}
# Data cleaning
dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()
dat <- dat[!duplicated(select(dat, -c(id))),]
dat$id<-factor(dat$id)
n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```
The result can be seen in the table \ref{tab:similar-samples-same-id}. All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.

&nbsp;

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
# Tabela - Previous samples after processing
same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)
caption <- "Previous samples after processing"
kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### Age selection
As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models. However, it is necessary to verify how relevant the data of these other ages are. Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}.

&nbsp;

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
# Figure - Compressive strength (MPa) vs age (days)
cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
ylabel <- "Compressive strength (MPa)"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```
It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.

&nbsp;

This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}). The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis. The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.

&nbsp;

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
# Figure - Principal component analysis - 90, 91 e 100 days
dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))
cap <- "Principal component analysis - 90, 91 e 100 days"
colnames(dat_90_91_100) <- c(
   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
habillage <- dat_90_91_100$day
fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```
Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value. The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).

&nbsp;

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Compressive strength through time
cap <- "Compressive strength through time"
ylabel <- "MPa"
xlabel <- "Age (days)"
dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)
dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```
For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise. They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}. As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.

&nbsp;

```{r join-90-91-100, echo=F, eval=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

&nbsp;

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Ages frequency
cap <- "Ages frequency"
ylabel <- "Frequency"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r remove-ages-lower-50, echo=F}
# Removing ages with frequency lower than 50
dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

### Data reorganization
The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}). The result in the first samples after this processing is shown in the table \ref{tab:new-features}.

&nbsp;

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
# Table - First 6 samples after reorganization/EN>
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}
# Total samples/EN>
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables
To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
```{r new-features-2, echo=F}
# Adding New features/EN>
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```
```{r new-features-table, echo=F}
# Table - New features/EN>
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

## Data visualization
In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.

### Descriptive statistics
The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}). The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data. As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age. The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).

&nbsp;

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables/EN>
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables/EN>
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

### Correlation between ingredients and compressive strength
The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}). The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).

&nbsp;

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age/EN>
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
                "14", "28", "56", "100",
                "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

&nbsp;

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age/EN>
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
mix_dat$day <- factor(mix_dat$day)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features/EN>
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

### Variables distribution
A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.

&nbsp;

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
# Figure - Variables distribution/EN>

cap <- "Variables distribution"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat <- dat %>%
  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))

colnames(dist_dat) <- colnames_dat
dist_dat <- dist_dat %>%
  gather("Var", "value") %>%
  mutate(value = as.numeric(value))
dist_dat %>%
  ggplot(aes(value)) +
  geom_histogram(aes(y = ..density..), 
                 colour = "black", 
                 fill = "white") +
  geom_density(alpha = .5, fill = "lightseagreen") +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  xlab("") +
  ylab("") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}. It shows that the resistance to compression gradually increases over time, as expected. Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
# Figure - Variables distribution grouped by age/EN>

days_labs <- c("3","7","14","28","56","100")
cap <- "Variables distribution grouped by age"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  days_labs,
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat_2 <- dat %>%
  select(-c(id, class, mix_app))
colnames(dist_dat_2) <- colnames_dat
dist_dat_2 <- dist_dat_2 %>%
  gather("day", "mpa", days_labs) %>%
  drop_na() %>%
  gather("Var", "value", -c("day")) %>%
  mutate(day = factor(day, levels = days_labs))
dist_dat_2 %>%
  ggplot(aes(x = day, y = value)) +
  geom_violin(color = NA,
                fill = "lightseagreen",
                alpha = .5,
                na.rm = TRUE,
                scale = "count") +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  xlab("") +
  ylab("")
```

### Principal component analysis
Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}). The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that. It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
```{r pca, echo=F, fig.height=3, fig.cap=cap}
# Figure - Principal component analysis on ingredients/EN>

cap <- "Principal component analysis on ingredients"
colnames_dat <- c(
  "Cem.", "B.F.S.", "F.A.", "Wat.", "Sup.", "C.Agg.",
  "F.Agg.","WxC", "F.Agg.xC","C.Agg.xC","F.Agg.xC",
  "WxC.Agg.","WxF.Agg."
)
class_list <- list(low="0", normal="20", medium="40", high="70")
dat_pca <- dat %>%
  select(-c(id, class, mix_app,
            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))
colnames(dat_pca) <- colnames_dat
class_2 <- dat$day_28
class_2[class_2 < 20] = "0"
class_2[class_2 >= 20 & class_2 < 40] = "20"
class_2[class_2 >= 40 & class_2 < 70] = "40"
class_2[class_2 >= 70] = "70"
class_2 <- factor(class_2)
levels(class_2) <- class_list
pca <- prcomp(dat_pca, scale = TRUE)
fviz_pca_ind(
  pca,
  geom.ind = "point",
  habillage=class_2,
  addEllipses = T,
  ellipse.level=0.95) +
  ggtitle("") +
  theme_bw()
```

## Machine learning models
The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.

### Pre-processing and data separation
Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado. 
As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.

&nbsp;

```{r dummy-var, echo=F, message=F, warning=F}
# Dummy variables
dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))
dummyDat$id <- dat$id
dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```
The samples were separated based on age. One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}). For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.

&nbsp;

```{r preparation, echo=F}
# Data preparation
names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```
```{r table-preparated-samples, echo=F}
# Table - First 18 columns of the 6 first samples of 28 days
colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
caption <- "First 18 columns of the 6 first samples of 28 days"
dat_tabela <- dat_28[1:18]
kable(
    head(dat_tabela[order(dat_tabela$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}). Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact. In addition to them, in the case of the 7-day set, the fly ash variable was also removed. Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}). After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.

&nbsp;

```{r nzv, echo=F}
# Removing near zero variance columns
nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)
dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```
```{r cors, echo=F}
# High correlation variables verification
descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```
The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package. If performed at this time, it would be necessary to manually undo these transformations in the predictions. The *caret* allows you to transform before training the models and already transforms the results back.

&nbsp;

Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}). The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
```{r split, echo=F, message=F, warning=F}
# Test and train data split
reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$MPa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$MPa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$MPa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$MPa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$MPa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$MPa)
)
reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )
split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}
trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}
dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```
```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
# Distribution of test and train data
cap <- "Distribution of test and train data"
d_lab <- " days"
ylabel <- "Density"
g1 <- "Train"
g2 <- "Test"
dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Grup = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Grup = g2, day = str_sub(n, 2))
    )
  dens
}
densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs
densities %>%
  ggplot(aes(y, fill = Grup)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~dia_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

### Performance measures
The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*). The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.

### Naive models
Before creating the real models, for comparison purposes, naive models were created. They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}). In other words, naive models are simply the best guess possible. The results can be checked in the table \ref{tab:table-naive-model-reg}.
```{r naive-model-reg, echo=F, message=F, warning=F}
# Naive model
res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}
ing_model <- imap(dats_reg, res)
get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}
rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)
ing_model_df <- ing_model %>% reduce(rbind)
df_performance_reg <- full_join(ing_model_df, rmses)
```
```{r table-naive-model-reg, echo=F}
# Table - Naive models
colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
caption <- "Naive models"
kable(
    df_performance_reg,
    col.names = colNames6,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

### Choice of algorithms
The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models. The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result. According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).

### Regression models
As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
1. All *features*;
2. All *features* without the *dummy vars*;
3. Only the original *features*;
4. Only new *features*;
5. Only new *features*, without *dummy vars*;
Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set. The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
```{r test-models-reg, echo=F, warning=F, message=F, include=F}
# Features selection
data1 <- dats_reg$d28$train # full
data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
data4 <- dats_reg$d28$train[c(8:21)] # only new variables
data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
test1 <- dats_reg$d28$test # full
test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
test4 <- dats_reg$d28$test[c(8:21)] # only new variables
test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
# parRF
modelLookup("parRF")
control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
fit_parRF<- function(data, n) {
  set.seed(1, sample.kind = "Rounding")
  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
  train(y ~ ., 
    data = data,
    preProcess = c("center","scale"),
    method='parRF', 
    tuneGrid=tunegrid_parRF,
    trControl=control_parRF)
}
fit_parRF1 <- fit_parRF(data1, 3)
ggplot(fit_parRF1)
min(fit_parRF1$results$RMSE) # 6.26108
p1_parRF1 <- predict(fit_parRF1, newdata = test1)
RMSE(p1_parRF1, test1$y) # 4.887264
fit_parRF2 <- fit_parRF(data2, 1)
ggplot(fit_parRF2)
min(fit_parRF2$results$RMSE) # 6.268789
p_parRF2 <- predict(fit_parRF2, newdata = test2)
RMSE(p_parRF2, test2$y) # 4.812994
fit_parRF3 <- fit_parRF(data3, 1)
ggplot(fit_parRF3)
min(fit_parRF3$results$RMSE) # 6.361905
p_parRF3 <- predict(fit_parRF3, newdata = test3)
RMSE(p_parRF3, test3$y) # 5.083467
fit_parRF4 <- fit_parRF(data4, 1)
ggplot(fit_parRF4)
min(fit_parRF4$results$RMSE) # 8.705984
p_parRF4 <- predict(fit_parRF4, newdata = test4)
RMSE(p_parRF4, test4$y) # 8.083962
fit_parRF5 <- fit_parRF(data5, 1)
ggplot(fit_parRF5)
min(fit_parRF5$results$RMSE) # 8.691755
p_parRF5 <- predict(fit_parRF5, newdata = test5)
RMSE(p_parRF5, test5$y) # 7.991537
```
```{r table-done-samples-28, echo=F}
# Table - First 6 samples of train data of the 28 days model
caption <- "First 6 samples of train data of the 28 days model"
colNames = c("Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","Ce.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "MPa")
colNames2 = c("Features"=13, "Outcome"=1)
kable(
    head(data2),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    add_header_above(header = colNames2, line = T, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}). For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.
```{r reg-models, echo=F, warning=F, message=F, include=F}
# Regression models
# no dummy vars:
data3 <- dats_reg$d3$train[c(1:13, 22)]
data7 <- dats_reg$d7$train[c(1:12, 19)]
data14 <- dats_reg$d14$train[c(1:13, 21)]
data28 <- dats_reg$d28$train[c(1:13, 21)]
data56 <- dats_reg$d56$train[c(1:13, 22)]
data100 <- dats_reg$d100$train[c(1:13, 23)]
test3 <- dats_reg$d3$test[c(1:13, 22)]
test7 <- dats_reg$d7$test[c(1:12, 19)]
test14 <- dats_reg$d14$test[c(1:13, 21)]
test28 <- dats_reg$d28$test[c(1:13, 21)]
test56 <- dats_reg$d56$test[c(1:13, 22)]
test100 <- dats_reg$d100$test[c(1:13, 23)]
get_trControl <- function(n, r){
  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
}
trControl3 <- get_trControl(30, 10)
trControl7 <- get_trControl(10, 10)
trControl14 <- get_trControl(30, 10)
trControl28 <- get_trControl(30, 10)
trControl56 <- get_trControl(30, 10)
trControl100 <- get_trControl(10, 10)
get_tuneGrid <- function(data){
  expand.grid(mtry = seq(1, length(data), 1))
}
tuneGrid3 <- get_tuneGrid(data3)
tuneGrid7 <- get_tuneGrid(data7)
tuneGrid14 <- get_tuneGrid(data14)
tuneGrid28 <- get_tuneGrid(data28)
tuneGrid56 <- get_tuneGrid(data56)
tuneGrid100 <- get_tuneGrid(data100)
set.seed(1, sample.kind = "Rounding")
fit_3 <- train(y ~ ., 
            data = data3,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid3,
            trControl=trControl3)
p_3 <- predict(fit_3, newdata = test3)
RMSE_test_3 <- RMSE(p_3, test3$y)
RMSE_test_3 # 3.31037
fit_3$bestTune # mtry = 6
set.seed(1, sample.kind = "Rounding")
fit_7 <- train(y ~ ., 
            data = data7,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid7,
            trControl=trControl7)
p_7 <- predict(fit_7, newdata = test7)
RMSE_test_7 <- RMSE(p_7, test7$y)
RMSE_test_7 # 4.361987
fit_7$bestTune # mtry = 2
set.seed(1, sample.kind = "Rounding")
fit_14 <- train(y ~ ., 
            data = data14,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid14,
            trControl=trControl14)
p_14 <- predict(fit_14, newdata = test14)
RMSE_test_14 <- RMSE(p_14, test14$y)
RMSE_test_14 # 4.620515
fit_14$bestTune # mtry = 13
set.seed(1, sample.kind = "Rounding")
fit_28 <- train(y ~ ., 
            data = data28,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid28,
            trControl=trControl28)
p_28 <- predict(fit_28, newdata = test28)
RMSE_test_28 <- RMSE(p_28, test28$y)
RMSE_test_28 # 4.716698
fit_28$bestTune # mtry = 11
set.seed(1, sample.kind = "Rounding")
fit_56 <- train(y ~ ., 
            data = data56,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid56,
            trControl=trControl56)
p_56 <- predict(fit_56, newdata = test56)
RMSE_test_56 <- RMSE(p_56, test56$y)
RMSE_test_56 # 5.939163
fit_56$bestTune # mtry = 8
set.seed(1, sample.kind = "Rounding")
fit_100 <- train(y ~ ., 
            data = data100,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid100,
            trControl=trControl100)
p_100 <- predict(fit_100, newdata = test100)
RMSE_test_100 <- RMSE(p_100, test100$y)
RMSE_test_100 # 5.851088
fit_100$bestTune # mtry = 8
```

# Results
The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively. The table \ref{tab:table-reg-models} presents the details and results of each model (\ref{show-table-reg-models}). The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).
```{r table-reg-models, echo=F}
# Table - Models details
colNames <- c("Model", "mtry", "CV", "Repetitions", 
              "RMSE (train)", "RMSE (test)")
caption <- "Regression models results"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
  number = c(trControl3$number, trControl7$number, trControl14$number,
             trControl28$number,trControl56$number,trControl100$number),
  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
             trControl28$repeats,trControl56$repeats,trControl100$repeats),
  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
# Figure - Models comparison
cap <- "Actual vs Predicted values in each model"
models <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
xlabel <- "Actual (MPa)"
ylabel <- "Predicted (MPa)"
preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
gen_res_df <- function(ind){
  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
}
res_df <- lapply(1:6, gen_res_df)
res_df <- bind_rows(res_df)
res_df$model <- factor(res_df$modelo,levels=modelos)
res_df %>%
  ggplot(aes(actual, pred)) +
  facet_wrap(~ modelo, ncol=3) +
  geom_point(alpha=0.5) +
  theme_bw() +
  geom_abline(slope=1, intercept=0) +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r table-10, echo=F}
# Results tables of 10 best and worst results
colNames1 = c("Actual", "Predicted", "Error", "","Actual", "Predicted", "Error") 
colNames2 = c("Best 10"=3,"", "Worst 10"=3) 
caption_0 <- "Model of "
get_X <- function(pred, actual, X=10){
  diff <- abs(pred - actual)
  diff_2 <- pred - actual
  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
  df_min <- data.frame(
    actual_min=actual[ind_min],
    pred_min=pred[ind_min],
    diff_min=diff[ind_min],
    diff_min_2=diff_2[ind_min]
  )
  df_max <- data.frame(
    actual_max=actual[ind_max],
    pred_max=pred[ind_max],
    diff_max=diff[ind_max],
    diff_max_2=diff_2[ind_max]
  )
  df_max <- df_max[order(-df_max$diff_max),]
  df_min <- df_min[order(df_min$diff_min),]
  df_null <- data.frame(null_col = rep(c(""), X))
  res <- cbind(df_min,df_null, df_max)
  res <- res %>% select(-c(diff_max, diff_min))
  rownames(res) <- c()
  res
}
gen_kable <- function(ind){
  df <- get_X(preds[[ind]], actuals[[ind]])
  caption <- paste0(caption_0,models[ind])
  kable(
    df,
    col.names = colNames1,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  column_spec(4, width = "1cm",) %>%
  add_header_above(header = colNames2, line = T, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position"))
}
gen_kable(1)
gen_kable(2)
gen_kable(3)
gen_kable(4)
gen_kable(5)
gen_kable(6)
```

# Discussion
The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily. The alternative adopted to create a model for each set of age proved to be a valid method, managing to stratify to obtain specific results for each set. The studies cited in the introduction using the same dataset have similar results, as expected. The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
```{r works-comparison, echo=F}
# Table - Comparison of other works
colNames = c("Author","Year" ,"Algorithm", "RMSE") 
caption <- "Comparison of other works"
works <- data.frame(
  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
          "Gradient Boosting Regressor","Random Forest Regressor" ,
          "Regularized Extreme Learning Machine",
          "Support Vector Machines with Radial Basis Function Kernel"),
  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
)
kable(
  works,
  col.names = colNames,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "l"
  )  %>%
    kable_styling(latex_options = c("HOLD_position"))
```
```{r results, echo=F}
# Table - Final results
colNames <- c("Model", "RMSE")
caption <- "Final result"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results. Another option is to create an *ensemble* of various algorithms, just like @Pierobon2018, but with the separation of age sets proposed here. In addition, it can be performed with a larger dataset, ideally with the same number of samples in each age set, a more homogeneous distribution of compressive strength, and less variance between samples.

# Literature cited
<div id="refs"></div>
\newpage

# Appendix 1 - Virtual environment {#appendix1}

## Operational system
```{r sistema, echo=FALSE}
v <- as.matrix(R.Version())
kable(v, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```

## Packages
```{r pacotes, echo=FALSE}
p <- matrix(
      c(
        "caret", as.character(packageVersion("caret")),
        "cowplot",as.character(packageVersion("cowplot")),
        "dplyr", as.character(packageVersion("dplyr")),
        "factoextra", as.character(packageVersion("factoextra")),
        "gdata", as.character(packageVersion("gdata")),
        "ggplot2", as.character(packageVersion("ggplot2")),
        "gridExtra",as.character(packageVersion("gridExtra")),
        "kableExtra",as.character(packageVersion("kableExtra")),
        "knitr",as.character(packageVersion("knitr")),
        "pastecs",as.character(packageVersion("pastecs")),
        "purrr", as.character(packageVersion("purrr")),
        "questionr", as.character(packageVersion("questionr")),
        "reshape2", as.character(packageVersion("reshape2")),
        "tidyr", as.character(packageVersion("tidyr")),
        "tidyverse", as.character(packageVersion("tidyverse"))
      ),
      ncol = 2,
      byrow=TRUE
    )
kable(p, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```
\newpage

# Appendix 2 - Online repository {#appendix2}
All the code used to generate this work is available on *GitHub* at the following link:
- https://github.com/pedrobern/concrete-compressive-strength-prediction
\newpage

# Appendix 3 - Code {#appendix3}

## Obtaining the data
### Data download

```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
```

### Ranaming the columns

```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
```

### Reordering the data

```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
```

### Defining column names and units

```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
```

### Table First samples

```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
```

## Data preparation
### Removing duplicated samples

```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
```

### Tabela - Samples with same composition

```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
```

### Tabela - Same samples with different results

```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
```

### Data cleaning

```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
```

### Tabela - Previous samples after processing

```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
```

### Figure - Compressive strength (MPa) vs age (days)

```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
```

### Figure - Principal component analysis - 90, 91 e 100 days

```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
```

### Figure - Compressive strength through time

```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
```

### Joining 90, 91 and 100 days data

```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
```

### Figure - Ages frequency

```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
```

### Removing ages with frequency lower than 50

```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
```

### Data reorganization

```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
```

## Data visualization
## Machine learning models
### Dummy variables

```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
```

### Data preparation

```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
```

### Table - First 18 columns of the 6 first samples of 28 days

```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
```

### Removing near zero variance columns

```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
```

### High correlation variables verification

```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
```

### Test and train data split

```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
```

### Distribution of test and train data

```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
```

### Naive model

```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
```

### Table - Naive models

```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
```

### Features selection

```{r show-test-models-reg, ref.label='test-models-reg', ref='show-test-models-reg', eval=FALSE}
```

### Table - First 6 samples of train data of the 28 days model

```{r show-table-done-samples-28, ref.label='table-done-samples-28', ref='show-table-done-samples-28', eval=FALSE}
```

### Regression models

```{r show-reg-models, ref.label='reg-models', ref='show-reg-models', eval=FALSE}
```

# Results
### Table - Models details

```{r show-table-reg-models, ref.label='table-reg-models', ref='show-table-reg-models', eval=FALSE}
```

### Figure - Models comparison

```{r show-results-comparison, ref.label='results-comparison', ref='show-results-comparison', eval=FALSE}
```

### Results tables of 10 best and worst results

```{r show-table-10, ref.label='table-10', ref='show-table-10', eval=FALSE}
```

# Discussion
### Table - Comparison of other works

```{r show-works-comparison, ref.label='works-comparison', ref='show-works-comparison', eval=FALSE}
```

### Table - Final results

```{r show-results, ref.label='results', ref='show-results', eval=FALSE}
```

### 2020-04-25 generated/paper_PT.Rmd
Add generated files

---
title:
  Prevendo a resisência à compressão do concreto com técnicas de machine learning
date: "Abril 22, 2020"
author: "Pedro Bernardino Alves Moreira"
indent: true
abstract:
  A resistência à compressão é a principal característica do concreto. Esse trabalho utilizou um conjunto de amostras amplamente utilizado para construir 6 modelos de previsão dessa característica, cada um específicamente para uma idade, 3, 7, 14, 28, 56 e 100 dias a partir de 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes) e 6 relações entre os principais ingredientes (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo) para fazer a previsão. Foi utilizado o algorítimo *Parallel Random Forest* (*parRF* do pacote *caret* de linguagem *R*). Os resultados foram satisfatórios sendo o melhor para o modelo de 3 dias (*RMSE* de 3,31) e pior para o modelo de 100 dias (*RMSE* de 5,85). O modelo de 28 dias apresentou *RMSE* de 4,72. O trabalho indicou que a separação em um modelo para cada idade é promissora e foi capaz de gerar resultados compatíveis ou melhores que outros estudos, abrindo espaço para construção de novos modelos seguindo essa proposta, mas utilizando algorítimos diferentes ou a combinação de diversos algorítimos.
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
bibliography: references.bib
header-includes:
- \renewcommand{\abstractname}{Resumo}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{supertabular}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{caption}
- \usepackage{floatrow}
- \floatsetup[table]{capposition=top}
- \floatsetup[figure]{capposition=top}
- \captionsetup{options=chunk}
- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
- \renewcommand{\thechunk}{\arabic{chunk}}
- \usepackage{indentfirst}
- \usepackage{sectsty} \sectionfont{\centering}
---
```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```
\renewcommand{\figurename}{Figura}
\renewcommand{\tablename}{Tabela}
\renewcommand{\contentsname}{Sumário}
\renewcommand{\chunkname}{Código}
\renewcommand{\listfigurename}{Lista de figuras}
\renewcommand{\listtablename}{Lista de tabelas}
\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

# Introdução
A resistência à compressão é a principal característica do concreto, medida por testes de padrões internacionais que consistem na quebra de corpos de prova. A medição aos 28 dias é obrigatória e representa a classe do concreto. Saber com antecêdencia qual o resultado será obtido para uma determinada idade, a partir das proporções de seus ingredientes, é de grande interesse para os fabricantes de concreto, construtoras e engenheiros civis.

&nbsp;

Essa resistência à compressão é uma função não linear de seus ingredientes e idade, tornando difícil o estabelecimento de uma fórmula analítica, apesar de algumas fórmulas já haveram sido propostas. @Hasan2011 propós um modelo matemático para prever a partir dos resultados de testes de 7 e 14 dias, e @Kabir2012 a partir de 7 dias. Porém técnicas de machine learning podem ser utilizadas para modelar essa característica a partir de dados reais de amostras, utilizando apenas os ingredientes.

&nbsp;

Muitos estudos anteriores utilizam o mesmo conjunto de dados utilizado por @Yeh1998 para prever a resistência à compressão do concreto. @Alshamiri2020 obteve bons resultados com a técnica de regularized extreme learning machine (RELM), e @Hameed2020 obteve resultados ainda melhores com a técnica de Artificial Neural Networks e cross-validation. Esse conjunto de amostras é tão conhecido que há ainda páginas na internet de estudos não publicados que o utilizam e possuem bons resultados, como @Abban2016, @Raj2018, @Modukuru2020 e @Pierobon2018. Ao final do trabalho os resultados encontrados são comparados aos trabalhos citados aqui.

&nbsp;

Diferente dos estudos anteriores com esse conjunto de amostras, este trabalho faz a preparação dos dados de forma diferente. A idade do concreto é a variável mais singular que contribui para sua resistência à compressão, por esse motivo, a idade é tratada separadamente nos modelos de machine learning, criando modelos para cada faixa de idade.

# Metodologia

## Materiais utilizados
A metodologia foi realizada utilizando o software RStudio [@RStudio], ambiente virtual integrado para desenvolvimento de código na linguagem *R* [@RCore]. Ao longo do processo, todo código executado foi documentado na mesma ordem de sua execução no [Apêndice 3](#appendix3), e foi sempre realizada referência aos códigos ao longo do texto. Todas as informações relevantes relacionados ao sistema operacional e pacotes instalados foram apresentados no [Apêndice 1](#appendix1). Além disso foi criado um repositorio online e de código aberto no *Github*, abrigando todo o código utilizado para gerar esse trabalho, o link foi diponibilizado no [Apêndice 2](#appendix2).

## Reprodutibilidade
Para garantir a reprodutibilidade, sempre que houve algum código que podesse utilizar operações probabilísticas, foi definido um *seed* antes da sua execução, garantindo que quando rodado em outra máquina, com a mesma versão de *R* e o mesmo *seed*, chegue ao mesmo resultado. Os *seeds* podem ser conferidos ao longo do [apêndice 3](#appendix) ou diretamente no *Github*.

## Obtenção dos dados
O download dos dados foi realizado no website da Universidade da California Irvine [@downloadData] (\ref{show-download-data}). No total são 1030 amostras com 9 colunas. As amostras foram renomeadas e foi adicionado uma coluna de id para facilitar na manipulação dos dados (\ref{show-rename-dat-cols}). As colunas foram reordenadas para colocar a nova coluna id em primeira posição (\ref{show-reorder-dat}). As primeiras amostras podem ser visualizadas na tabela \ref{tab:first-samples}.

&nbsp;

```{r download-data, echo=F, include=F}
# Download dos dados
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```
```{r rename-dat-cols, echo=F}
# Renomeando as colunas
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )
dat$id <- seq.int(nrow(dat))
```
```{r reorder-dat, echo=F}
# Reordenando os dados
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```
```{r col-names-and-units, echo=F}
# Definindo nomes e unidades das colunas
colNames <- c("ID", "Cimento", "E.G.A.F", "C.Volante", "Água",
             "Superp.", "A.Graúdo", "A.Miúdo", "Dia", "Comp.Str.")
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```
```{r first-samples, echo=F}
# Tabela Primeiras amostras
caption <- "Primeiras 6 amostras"
kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

## Preparação dos dados
A preparação dos dados consistiu em transformar o conjunto de amostra afim de manter apenas dados relevantes para os estudos subsequêntes. Foram retirandos dados considerados irrelevantes ou que com potencial de adicionar um ruído indesejavel nas análises. Além disso, os dados relevântes foram transformados para melhor se enquadrar para os estudos nas próximas etapas.

### Limpeza inicial dos dados
Inicialmente foram observadas a existência de 25 amostras duplicadas que foram retiradas, resultando em um novo total de 1005 amostras (\ref{show-removing-duplicated-samples}).

&nbsp;

```{r removing-duplicated-samples, echo=F}
# Removendo amostras duplicadas
n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```
Os dados apresentam as variaveis nas colunas e amostras nas linhas. Porém foi verificado que algumas amostras são identicas em proporções de ingredientes, alterando apenas o valor da idade e resistência à compressão, por exemplo as amostras 653, 654, 678 e 681, mostradas na tabela \ref{tab:similar-samples}.

&nbsp;

```{r similar-samples, echo=FALSE}
# Tabela - Amostras com a mesma composição
same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))
caption <- "Amostras com a mesma composição"
kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
Além disso, também existem amostras com os mesmos valores e proporções de ingredientes, mas com resistência à compressão diferente, provavelmente devido a diferenças na execução. É o caso por exemplo das amostras 472, 473 e 474, mostradas na tabela \ref{tab:similar-samples-2}.

&nbsp;

```{r similar-samples-2, echo=FALSE}
# Tabela - Amostras iguais com resultados diferentes
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Amostras iguais com resultados diferentes"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
Para facilitar a analise das amostras, todas as amostras que são iguais em relação aos ingredientes, foram atribuidos o mesmo *id*. Além disso, como a resistência à compressão aos 28 dias é o parâmetro para determinar a classe do concreto, foi mantido apenas os elementos que contenham esse dia entre suas amostras. No caso das amostras iguais mas com resultado diferentes de resistência à compressão, foi calculado a média dos valores. Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes (\ref{show-total-de-amostras-apos-processamento}).

&nbsp;

```{r initial-data-cleaning, echo=FALSE}
# Limpeza inicial das amotras
dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()
dat <- dat[!duplicated(select(dat, -c(id))),]
dat$id<-factor(dat$id)
n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```
O resultado pode ser conferido na tabela \ref{tab:similar-samples-same-id}. Todas as amostras com configurações iguais de ingredientes possuem o mesmo id, e quando possuiam resultados diferentes para os mesmos dias, foram transformadas em apenas uma amostra, com a média aritmética na resistência à compressão.

&nbsp;

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
# Tabela - Amostras anteriores após processamento
same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)
caption <- "Amostras anteriores após processamento"
kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### Seleção das idades
Como descrito anteriormente, a principal idade para análise da resistência à compressão é aos 28 dias, mas as outras idades também podem ser utilizadas para construir modelos de previsão. Porém é necessário verificar o quanto relevante os dados dessas outras idades são. Iniciando pela distribuição das amostras em relação a cada idade (\ref{show-boxplot}) mostrado na figura \ref{fig:boxplot}.

&nbsp;

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
# Figura - Resistência à compressão (MPa) vs idade (dias)
cap <- "Boxplot - Resistência à compressão (MPa) vs idade (dias)"
ylabel <- "Resistência à compressão (MPa)"
xlabel <- "Idade (dias)"
dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```
Foi observado que as idades de 90, 91 e 100 dias provavelmente representam extremos entre si das configurações de ingredientes, uma vez que são idades relativamente próximas porém com valores muito diferentes, especialmente para 90 e 91.

&nbsp;

Essa hipótese foi verificada utilizando o método de análise de componente principal, aplicado às amostras dessas 3 idades (\ref{show-pca-90-91-100}). A fígura \ref{fig:pca-90-91-100} mostra como as amostras se relacionam umas com as outras (quais são parecidas ou diferentes) e revelou como cada variável contribui para a análise. As duas primeiras dimensões representam $37\%$ e $24\%$ respectivamente da variância.

&nbsp;

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
# Figura - Análise componente principal - 90, 91 e 100 dias
dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))
cap <- "Análise componente principal - 90, 91 e 100 dias"
colnames(dat_90_91_100) <- c(
   "Cim.","E.G.A.F.","Ci.Vo.","Agu.","Sup.","Ag.G.","Ag.M.","dia"
)
pca <- prcomp(select(dat_90_91_100, -c(dia)), scale = TRUE)
habillage <- dat_90_91_100$dia
fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```
Outro ponto importante considerado, da própria natureza do concreto, é o fato da taxa de crescimento de sua resistência à compressão diminuir com o tempo, chegando a um certo valor de establidade. A figura \ref{fig:mpa-on-time} mostra a resistência à compressão ao longo dos dias para amostras com mais de 5 dados, ou seja, dados disponiveis para no mínimo 6 idades distintas (\ref{show-mpa-on-time}).

&nbsp;

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figura - Resistência à compressão ao longo do tempo
cap <- "Resistência à compressão ao longo do tempo"
ylabel <- "MPa"
xlabel <- "Idade (dias)"
dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)
dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```
Pelos motivos apresentados nas figuras \ref{fig:boxplot}, \ref{fig:pca-90-91-100} e \ref{fig:mpa-on-time}, foi considerado que as idades de 90, 91 e 100 dias podem ser agrupadas para melhorar a leitura e diminuir o ruído das amostras. Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}, pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.

&nbsp;

```{r join-90-91-100, echo=F, eval=F}
# Juntando amostras de 90, 91 e 100 dias
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Mais um tópico analisado na seleção das idades foi a frequência observada de cada valor de idade após essa transformação dos 90, 91 em 100 dias, mostrada na figura \ref{fig:freq-ages}. Alguns valores de dias apresentam concentrações muito baixas de amostras, correndo o risco de prejudicar mais do que ajudar na criação dos modelos, logo elas foram removidas (\ref{show-remove-ages-lower-50}). O critério adotado foi manter apenas idades com frequência maior que 50, ou seja, apenas os valores de 3, 7, 14, 28, 56 e 100 dias.

&nbsp;

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figura - Frequência das idades
cap <- "Frequência das idades"
ylabel <- "Frequência"
xlabel <- "Idade (dias)"
dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r remove-ages-lower-50, echo=F}
# Removendo idades com frequência menor que 50
dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

### Reorganização dos dados
As amostras foram agrupadas para manter apenas uma amostra distinta de cada conjunto de configuração das proporções dos ingredientes, adicionando novas variáveis/colunas para a resistência em cada idade (\ref{show-reorganizing-dat}). O resultado nas primeiras amostras após esse processamento é mostrado na tabela \ref{tab:new-features}.

&nbsp;

```{r reorganizing-dat, echo=F}
# Reorganização das amotras
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
caption <- "Primeiras 6 amostras reorganizadas"
colNames2 = c("ID", "Cimento", "E.G.A.F.", "C.Volante", "Água",
             "Superp.", "A.Graúdo", "A.Miúdo", "3 dias", "7 dias",
             "14 dias", "28 dias", "56 dias", "100 dias")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
O número de amostras e amostras distintas após toda essa manipulação permaneceu o mesmo, um total de 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}

### Adicionando novas variáveis
Para finalizar a preparação dos dados, novas colunas foram adicionadas ao conjunto de amostras (\ref{show-new-features-2}). Iniciando pela classe do concreto, por exemplo se a resistência à compressão está entre 25 e 30, recebe a classe *C25*. A inclusão da classe foi importânte pois a resistência em $MPa$ é uma variável contínua, que será utilizada nos modelos de regressão, mas a classe como variável discreta pode fornecer outro ângulo de visualização dos dados. Também foi adicionado o traço aproximado do concreto, que representa as proporções de agregados (miúdo e graúdo) para o cimento. Outras proporções entre os principais ingredientes também foram adicionadas. As novas variáveis são apresentadas na tabela \ref{tab:new-features-table}.
```{r new-features-2, echo=F}
caption <- "Novas variáveis"
colNames7 = c("ID", "Classe","Traço aproximado",
             "Água / Cimento", "A.Miúdo / Cimento", 
             "A.Graúdo / Cimento", "A.Miúdo / A.Graúdo",
             "Água / A.Graúdo", "Água / A.Miúdo")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

## Visualização dos dados
Para avaliar a necessidade de mais manipulações antes da construção dos modelos, nesta etapa as 416 amostras já processadas foram visualizadas e analisadas.

### Estatistica descritiva
A tabela \ref{tab:stat-summ} apresenta os dados estatísticos das variáveis contínuas (\ref{show-stat-summ}). A linha de *Null* representa o número de valores zerados para os ingredientes, e a linha *NA* representa o número de dados faltando. Como as amostras foram filtradas para manter apenas conjuntos de amostras com valores conhecidos da resistência à compressão aos 28 dias, o número de *NAs* é zero para essa idade. A figura \ref{fig:stat-summ-categorical} apresenta os dados estatísticos das variáveis discretas (\ref{show-stat-summ-categorical}).

&nbsp;

```{r stat-summ, echo=FALSE}
caption <- "Estatística descritiva - variáveis contínuas"
colnames(summ) <- c("Amostras", "Null", "NA", "Min", "Max", "Intervalo",
              "Soma", "Mediana", "Média", "Erro padrão  da média", 
              "Intervalo de confiânça da média",
              "Variância", "Desvio Padrão", "Coeficiênte de variação")
rownames(summ) = c("Cimento", "E.G.A.F.", "Cinza Volante", "Água",
              "Superplast.", "A.Graúdo", "A.Miúdo", "3 dias", "7 dias",
              "14 dias", "28 dias", "56 dias", "100 dias",
              "Água/ Cimento", "A.Miúdo/ Cimento", 
              "A.Graúdo/ Cimento", "A.Miúdo / A.Graúdo",
              "Água / A.Graúdo", "Água / A.Miúdo")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
cap <- "Estatística descritiva - variáveis discretas"
name1 <- "Porcentagem"
name2 <- "Porcentagem acumulada"
ylabel <- "Frequencia"
xlabel1 <- "Classe"
xlabel2 <- "Traço aproximado"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

### Correlação dos ingredientes e resistência à compressão
A figura \ref{fig:correlation} apresenta a correlação das variáveis para cada conjunto de idades (\ref{show-correlation}). A figura \ref{fig:correlation-mpa} apresenta os mesmos dados, mas em vez de correlacionar todos, correlaciona apenas com a resistência à compressão, mostrando os valores mais detalhadamente (\ref{show-correlation-mpa}).

&nbsp;

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
cap <- "Correlações em cada idade"
f_lvl <- c("3 dias", "7 dias", "14 dias","28 dias", "56 dias", "100 dias")
name <- "Correlação"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                  "Superp.", "A.Graúdo", "A.Miúdo",
                  "3","7","14","28","56","100",
                  "class","mix_app" ,"Ág./Ci.",
                  "A.M./Ci.", "A.G./Ci.", "A.M./A.G.",
                  "Ág./A.G.", "Ág./A.M.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
cap <- "Correlação das variáveis com a resistência à compressão no tempo"
name <- "Correlação"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
A interpretação da figura \ref{fig:correlation-mpa} sugere que a resistência do concreto está relacionada positivamente principalmente com os ingredientes cimento e superplastificante e negativamente com a água e agregado miúdo. Quanto menor a quantidade de cimento para os agregados e para água, mais negativamente estão correlacionados com a resistência à compressão.

&nbsp;

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
cap <- "Relação entre o traço aproximado, água, resistência à compressão e idade"
d <- " dias"
xlabel <- "Traço aproximado" 
ylabel <- "Resistência à compressão (MPa)"
label <- "Água /\nCimento"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
mix_dat$day <- factor(mix_dat$day)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
cap <- "Relação das principais proporções do concreto"
d <- " dias"
xlabel <- "Água / Cimento"
ylabel <- "Resistência à compressão (MPa)"
colour <- "Ag. Miúdo /\nCimento"
size <- "Ag. Graúdo /\nCimento"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
As figuras \ref{fig:mix-app-mpa} e \ref{fig:mix-mpa} mostram a relação entre os principais ingredientes (conhecida como traço) em relação à resistência à compressão (\ref{show-mix-app-mpa} e \ref{show-mix-mpa}). A interpretação dessas figuras mostra que quanto maior a quantidade de cimento em relação aos outros ingredientes, maior será a resistência à compressão.

### Distribuição das variáveis
A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias.

&nbsp;

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
cap <- "Distribuição das variáveis"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                 "Superplastificante", "A.Graúdo", "A.Miúdo", "MPa",
                 "Água / Cimento", "A.Miúdo / Cimento", "A.Graúdo / Cimento",
                 "A.Miúdo / A.Graúdo","Água / A.Graúdo", "Água / A.Miúdo")
dist_dat <- dat %>%
  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))

colnames(dist_dat) <- colnames_dat
dist_dat <- dist_dat %>%
  gather("Var", "value") %>%
  mutate(value = as.numeric(value))
dist_dat %>%
  ggplot(aes(value)) +
  geom_histogram(aes(y = ..density..), 
                 colour = "black", 
                 fill = "white") +
  geom_density(alpha = .5, fill = "lightseagreen") +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  xlab("") +
  ylab("") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
De outra forma, a figura \ref{fig:vars-distribution-time} mostra a distribuição dos ingredientes e da resistência à compressão para cada conjunto de idades (\ref{show-vars-distribution-time}), ou seja, no caso dos 28 dias apresenta a mesma informação que a figura \ref{fig:vars-distribution}. Através dela é visualizado que como esperado, a resistência à compressão gradualmente aumenta ao longo do tempo. Além disso é visualizado que a concentração dos ingredientes podem variar muito quando estratíficado pelas idades.
```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
cap <- "Distribuição das variáveis em relação a idade"
colnames_dat <- c("Cimento", "E.G.A.F.", "C.Volante", "Água",
                       "Superplastificante", "A.Graúdo", "A.Miúdo", 
                       days_labs,
                       "Água / Cimento", "A.Miúdo / Cimento", "A.Graúdo / Cimento",
                       "A.Miúdo / A.Graúdo","Água / A.Graúdo", "Água / A.Miúdo")
dist_dat_2 <- dat %>%
  select(-c(id, class, mix_app))
colnames(dist_dat_2) <- colnames_dat
dist_dat_2 <- dist_dat_2 %>%
  gather("day", "mpa", days_labs) %>%
  drop_na() %>%
  gather("Var", "value", -c("day")) %>%
  mutate(day = factor(day, levels = days_labs))
dist_dat_2 %>%
  ggplot(aes(x = day, y = value)) +
  geom_violin(color = NA,
                fill = "lightseagreen",
                alpha = .5,
                na.rm = TRUE,
                scale = "count") +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  xlab("") +
  ylab("")
```

### Análise de componente principal
Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até *20 MPa*, normal até *40 MPa*, médio até *70 MPa* e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
```{r pca, echo=F, fig.height=3, fig.cap=cap}
cap <- "Análise componente principal nos ingredientes"
colnames_dat <- c(
  "Cim.", "E.G.A.F.", "Ci.Vo.", "Águ.", "Sup.", "Ag.G.",
  "Ag.M.","AxC", "MxC","GxC","MxC","AxG","AxM"
)
class_list <- list(baixo="0", normal="20", médio="40", alto="70")
dat_pca <- dat %>%
  select(-c(id, class, mix_app,
            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))
colnames(dat_pca) <- colnames_dat
class_2 <- dat$day_28
class_2[class_2 < 20] = "0"
class_2[class_2 >= 20 & class_2 < 40] = "20"
class_2[class_2 >= 40 & class_2 < 70] = "40"
class_2[class_2 >= 70] = "70"
class_2 <- factor(class_2)
levels(class_2) <- class_list
pca <- prcomp(dat_pca, scale = TRUE)
fviz_pca_ind(
  pca,
  geom.ind = "point",
  habillage=class_2,
  addEllipses = T,
  ellipse.level=0.95) +
  ggtitle("") +
  theme_bw()
```

## Modelos de machine learning
O desenvolvimento dos modelos de machine learning foi realizado com o pacote *caret* [@caret] e baseado em @irizarry2019 e @Kuhn2008.

### Pre processamento e separação dos dados
Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado.

&nbsp;

```{r dummy-var, echo=F, message=F, warning=F}
# Variáveis fictícias - dummy vars
dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))
dummyDat$id <- dat$id
dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```
As amostras foram separadas baseado nas idades. Foram criados um conjunto de dados para cada valor de idade, totalizando 6 conjuntos diferentes (\ref{show-preparation}). Para fins ilustrativos, as primeiras 18 de 45 colunas das primeiras 6 amostras do conjunto de 28 dias são mostradas na tabela \ref{tab:table-preparated-samples}.

&nbsp;

```{r preparation, echo=F}
# Preparação dos dados
names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```
```{r table-preparated-samples, echo=F}
# Tabela - Primeiras 18 colunas das primeiras 6 amostras de 28 dias
colNames = c("ID", "Cimento", "E.G.A.F.", "C.Vol.", "Água",
              "Superp.", "A.Graúdo", "A.Miúdo", "MPa", "Classe",
              "Ág./", "A.M./", "A.G./",
              "A.M./","Ág./", "Ág./", "Traço Apox." = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ci.", "Ci.", "A.G.", "A.G.", "Ag.M.", "1:1:2", "1:2:2")
caption <- "Primeiras 18 colunas das primeiras 6 amostras de 28 dias"
dat_tabela <- dat_28[1:18]
kable(
    head(dat_tabela[order(dat_tabela$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
Para cada um dos 6 conjuntos foi verificado a existência ou não de variáveis com variância próxima a zero e sua subsequênte remoção (\ref{show-nzv}). Muitas das 23 variáveis adicionadas referentes ao traço aproximado do concreto foram removidas devido a esse fato. Além delas, no caso do conjunto de 7 dias, a variável cinza volante também foi removida. Depois foi verificado que não existem variáveis com alta correlação, acima de 0,999 em nenhum dos 6 conjuntos de dados (\ref{show-cors}). Após essas etapas, os conjuntos de amostras apresentaram 24, 21, 23, 23, 24 e 25 colunas respectivamente para as idades em sequência crescente.

&nbsp;

```{r nzv, echo=F}
# Removendo colunas com variância próxima a zero
nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)
dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```
```{r cors, echo=F}
# Verificação de variáveis com alta correlação
descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```
A etapa de centralização e normalização das variáveis foi realizada mais a frente, junto com a aplicação dos modelos, pois é mais simples fazer dessa forma com o pacote *caret*. Se fosse realizada nesse momento, seria nescessário manualmente desfazer essas transformações nas previsões. O *caret* permite transformar antes do treino dos modelos e já transforma de volta os resultados.

&nbsp;

Cada um dos conjuntos de dados foi separado em conjuntos de teste e treino, *20%* e *80%* respectivamente (\ref{show-split}). A figura \ref{fig:dist-split} mostra como ficou a distribuição dos dados entre os conjuntos em relação a resistência à compressão para cada modelo (\ref{show-dist-split}).
```{r split, echo=F, message=F, warning=F}
# Separação em conjunto de teste e treino
reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$MPa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$MPa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$MPa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$MPa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$MPa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$MPa)
)
reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )
split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}
trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}
dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```
```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
# Distribuição dos conjuntos de teste e treino
cap <- "Distribuição dos conjuntos de teste e treino"
d_lab <- " dias"
ylabel <- "Densidade"
g1 <- "Treino"
g2 <- "Teste"
dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Grup = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Grup = g2, day = str_sub(n, 2))
    )
  dens
}
densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs
densities %>%
  ggplot(aes(y, fill = Grup)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~dia_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

### Medidas de performance
A avaliação da performance dos modelos foi realizada pela Raiz do Erro Quadrático Médio (*RMSE*). O *RMSE* é a medida utilizada em todos os trabalhos citados na introdução e permitirá a comparação dos modelos na discussão.

### Modelos ingênuos
Antes de criar os modelos verdadeiros, para fins de comparação, foram criados modelos ingênuos. Eles simplesmente prevêem que a resistência à compressão do conjunto de teste, é a média da resistência à compressão do conjunto de treino (\ref{show-naive-model-reg}). Em outras palavras, os modelos ingênuos são simplesmente o melhor palpite possível. Os resultados podem ser conferidos na tabela \ref{tab:table-naive-model-reg}.
```{r naive-model-reg, echo=F, message=F, warning=F}
# Modelo ingênuo
res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}
ing_model <- imap(dats_reg, res)
get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}
rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)
ing_model_df <- ing_model %>% reduce(rbind)
df_performance_reg <- full_join(ing_model_df, rmses)
```
```{r table-naive-model-reg, echo=F}
# Tabela - Modelos ingênuo
colNames = c("Idade", "Média $MPa$ (treino)", "RMSE (treino)", "RMSE (teste)")
caption <- "Modelos ingênuos"
kable(
    df_performance_reg,
    col.names = colNames6,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

### Escolha do algorítimo
O pacote *caret* [@caret] expõe mais de 200 algorítimos diferentes para criar modelos de *machine learning*. A documentação do pacote apresenta um código inicial [@modelsClusters] como sugestão para selecionar um portifólio de algóritmos mais distintos possíveis em relação à algum algorítimo pré-selecionado, mas para agilidade e devido a limitações técnicas, foi escolhido utilizar um algorítimo com a maior probabilidade de atingir o melhor resultado possível. Segundo @Fernandez2014, que comparou 179 algorítimos em 121 banco de dados diferentes, o algorítimo mais provável de atingir os melhores resultados possíveis é o *Parallel Random Forest* (denominado *prRF* no *caret*).

### Modelos de regressão
Como ao longo do processamento foram adicionadas novas variáveis (as relações entre os ingredientes e mais algumas *dummy vars* para cada conjunto de idade), foram estudadas 5 possibilidades de configurações das *features* para os modelos:
1. Todas as *features*;
2. Sem as *dummy vars*;
3. Apenas *features* originais;
4. Apenas *features* adicionadas;
5. Apenas *features* adicionadas, sem as *dummy vars*;
Construindo um modelo para cada uma dessas configurações utilizando o conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou que a melhor opção é a configuração 2, ou seja, as *dummy vars* foram completamente descartadas, porém foram mantidas as outras novas varáveis.
Para fins ilustrativos, a tabela \ref{tab:table-done-samples-28} mostra as primeiras 6 amostras do conjunto de treino do modelo de 28 dias. As amostras dos outros modelos, dos conjuntos de teste e treino são similares, sendo a única diferença no modelo de 7 dias, que exlcui a cinza volante devido a variância próxima a zero, realizado anteriormente.
```{r test-models-reg, echo=F, warning=F, message=F, include=F}
# Escolha das caracterísitcas (features)
data1 <- dats_reg$d28$train # full
data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
data4 <- dats_reg$d28$train[c(8:21)] # only new variables
data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
test1 <- dats_reg$d28$test # full
test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
test4 <- dats_reg$d28$test[c(8:21)] # only new variables
test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
# parRF
modelLookup("parRF")
control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
fit_parRF<- function(data, n) {
  set.seed(1, sample.kind = "Rounding")
  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
  train(y ~ ., 
    data = data,
    preProcess = c("center","scale"),
    method='parRF', 
    tuneGrid=tunegrid_parRF,
    trControl=control_parRF)
}
fit_parRF1 <- fit_parRF(data1, 3)
ggplot(fit_parRF1)
min(fit_parRF1$results$RMSE) # 6.26108
p1_parRF1 <- predict(fit_parRF1, newdata = test1)
RMSE(p1_parRF1, test1$y) # 4.887264
fit_parRF2 <- fit_parRF(data2, 1)
ggplot(fit_parRF2)
min(fit_parRF2$results$RMSE) # 6.268789
p_parRF2 <- predict(fit_parRF2, newdata = test2)
RMSE(p_parRF2, test2$y) # 4.812994
fit_parRF3 <- fit_parRF(data3, 1)
ggplot(fit_parRF3)
min(fit_parRF3$results$RMSE) # 6.361905
p_parRF3 <- predict(fit_parRF3, newdata = test3)
RMSE(p_parRF3, test3$y) # 5.083467
fit_parRF4 <- fit_parRF(data4, 1)
ggplot(fit_parRF4)
min(fit_parRF4$results$RMSE) # 8.705984
p_parRF4 <- predict(fit_parRF4, newdata = test4)
RMSE(p_parRF4, test4$y) # 8.083962
fit_parRF5 <- fit_parRF(data5, 1)
ggplot(fit_parRF5)
min(fit_parRF5$results$RMSE) # 8.691755
p_parRF5 <- predict(fit_parRF5, newdata = test5)
RMSE(p_parRF5, test5$y) # 7.991537
```
```{r table-done-samples-28, echo=F}
# Tabela - Primeiras 6 amostras do conjunto de treino do modelo de 28 dias
caption <- "Primeiras 6 amostras do conjunto de treino do modelo de 28 dias"
colNames = c("Cimento", "E.G.A.F.", "C.Vol.", "Água",
              "Superp.", "A.Graúdo", "A.Miúdo",
              "Ág./", "A.M./", "A.G./",
              "A.M./","Ág./", "Ág./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$", "Ci.",
              "Ci.", "Ci.", "A.G.", "A.G.", "Ag.M.", "$MPa$")
colNames2 = c("Features"=13, "Outcome"=1)
kable(
    head(data2),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    add_header_above(header = colNames2, line = T, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
Para cada conjunto de idade foi criado um modelo utilizando o algorítimo *Parallel Random Forest*, definido anteriormente (\ref{show-reg-models}). Para cada um dos 6 modelos, o parâmetro *mtry* foi otimizado, e foi realizado *repeated cross-validation*, dividindo em 10 ou 30 partes e repetindo 10 vezes.
```{r reg-models, echo=F, warning=F, message=F, include=F}
# Modelos de regressão
# no dummy vars:
data3 <- dats_reg$d3$train[c(1:13, 22)]
data7 <- dats_reg$d7$train[c(1:12, 19)]
data14 <- dats_reg$d14$train[c(1:13, 21)]
data28 <- dats_reg$d28$train[c(1:13, 21)]
data56 <- dats_reg$d56$train[c(1:13, 22)]
data100 <- dats_reg$d100$train[c(1:13, 23)]
test3 <- dats_reg$d3$test[c(1:13, 22)]
test7 <- dats_reg$d7$test[c(1:12, 19)]
test14 <- dats_reg$d14$test[c(1:13, 21)]
test28 <- dats_reg$d28$test[c(1:13, 21)]
test56 <- dats_reg$d56$test[c(1:13, 22)]
test100 <- dats_reg$d100$test[c(1:13, 23)]
get_trControl <- function(n, r){
  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
}
trControl3 <- get_trControl(30, 10)
trControl7 <- get_trControl(10, 10)
trControl14 <- get_trControl(30, 10)
trControl28 <- get_trControl(30, 10)
trControl56 <- get_trControl(30, 10)
trControl100 <- get_trControl(10, 10)
get_tuneGrid <- function(data){
  expand.grid(mtry = seq(1, length(data), 1))
}
tuneGrid3 <- get_tuneGrid(data3)
tuneGrid7 <- get_tuneGrid(data7)
tuneGrid14 <- get_tuneGrid(data14)
tuneGrid28 <- get_tuneGrid(data28)
tuneGrid56 <- get_tuneGrid(data56)
tuneGrid100 <- get_tuneGrid(data100)
set.seed(1, sample.kind = "Rounding")
fit_3 <- train(y ~ ., 
            data = data3,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid3,
            trControl=trControl3)
p_3 <- predict(fit_3, newdata = test3)
RMSE_test_3 <- RMSE(p_3, test3$y)
RMSE_test_3 # 3.31037
fit_3$bestTune # mtry = 6
set.seed(1, sample.kind = "Rounding")
fit_7 <- train(y ~ ., 
            data = data7,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid7,
            trControl=trControl7)
p_7 <- predict(fit_7, newdata = test7)
RMSE_test_7 <- RMSE(p_7, test7$y)
RMSE_test_7 # 4.361987
fit_7$bestTune # mtry = 2
set.seed(1, sample.kind = "Rounding")
fit_14 <- train(y ~ ., 
            data = data14,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid14,
            trControl=trControl14)
p_14 <- predict(fit_14, newdata = test14)
RMSE_test_14 <- RMSE(p_14, test14$y)
RMSE_test_14 # 4.620515
fit_14$bestTune # mtry = 13
set.seed(1, sample.kind = "Rounding")
fit_28 <- train(y ~ ., 
            data = data28,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid28,
            trControl=trControl28)
p_28 <- predict(fit_28, newdata = test28)
RMSE_test_28 <- RMSE(p_28, test28$y)
RMSE_test_28 # 4.716698
fit_28$bestTune # mtry = 11
set.seed(1, sample.kind = "Rounding")
fit_56 <- train(y ~ ., 
            data = data56,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid56,
            trControl=trControl56)
p_56 <- predict(fit_56, newdata = test56)
RMSE_test_56 <- RMSE(p_56, test56$y)
RMSE_test_56 # 5.939163
fit_56$bestTune # mtry = 8
set.seed(1, sample.kind = "Rounding")
fit_100 <- train(y ~ ., 
            data = data100,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid100,
            trControl=trControl100)
p_100 <- predict(fit_100, newdata = test100)
RMSE_test_100 <- RMSE(p_100, test100$y)
RMSE_test_100 # 5.851088
fit_100$bestTune # mtry = 8
```

# Results
O *RMSE* de teste de cada modelo em ordem crescente de idade foi respectivamente `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` e `r round(RMSE_test_100,2)`. A tabela \ref{tab:table-reg-models} apresenta os detalhes e resultados de cada modelo (\ref{show-table-reg-models}). A figura \ref{fig:results-comparison} compara os valores reais e previstos (\ref{show-results-comparison}), e as tabelas seguintes mostram os melhores e piores resultados de cada modelo (\ref{show-table-10}).
```{r table-reg-models, echo=F}
# Tabela - Detalhes dos modelso
colNames <- c("Modelo", "mtry", "CV", "Repetições", 
              "RMSE (treino)", "RMSE (teste)")
caption <- "Resultados dos modelos de regressão"
day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
dat_reg_models <- data.frame(
  dia = day,
  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
  number = c(trControl3$number, trControl7$number, trControl14$number,
             trControl28$number,trControl56$number,trControl100$number),
  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
             trControl28$repeats,trControl56$repeats,trControl100$repeats),
  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
# Figura - Comparação dos modelos
cap <- "Comparação dos valores reais e previstos em cada modelo"
models <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
xlabel <- "Real (MPa)"
ylabel <- "Previsto (MPa)"
preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
gen_res_df <- function(ind){
  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
}
res_df <- lapply(1:6, gen_res_df)
res_df <- bind_rows(res_df)
res_df$model <- factor(res_df$modelo,levels=modelos)
res_df %>%
  ggplot(aes(actual, pred)) +
  facet_wrap(~ modelo, ncol=3) +
  geom_point(alpha=0.5) +
  theme_bw() +
  geom_abline(slope=1, intercept=0) +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r table-10, echo=F}
# Tabelas dos 10 melhores e piores resultados
colNames1 = c("Real", "Previsto", "Erro", "","Real", "Previsto", "Erro") 
colNames2 = c("10 melhores"=3,"", "10 piores"=3)
caption_0 <- "Modelo de "
get_X <- function(pred, actual, X=10){
  diff <- abs(pred - actual)
  diff_2 <- pred - actual
  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
  df_min <- data.frame(
    actual_min=actual[ind_min],
    pred_min=pred[ind_min],
    diff_min=diff[ind_min],
    diff_min_2=diff_2[ind_min]
  )
  df_max <- data.frame(
    actual_max=actual[ind_max],
    pred_max=pred[ind_max],
    diff_max=diff[ind_max],
    diff_max_2=diff_2[ind_max]
  )
  df_max <- df_max[order(-df_max$diff_max),]
  df_min <- df_min[order(df_min$diff_min),]
  df_null <- data.frame(null_col = rep(c(""), X))
  res <- cbind(df_min,df_null, df_max)
  res <- res %>% select(-c(diff_max, diff_min))
  rownames(res) <- c()
  res
}
gen_kable <- function(ind){
  df <- get_X(preds[[ind]], actuals[[ind]])
  caption <- paste0(caption_0,models[ind])
  kable(
    df,
    col.names = colNames1,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  column_spec(4, width = "1cm",) %>%
  add_header_above(header = colNames2, line = T, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position"))
}
gen_kable(1)
gen_kable(2)
gen_kable(3)
gen_kable(4)
gen_kable(5)
gen_kable(6)
```

# Discussão
Os modelos construidos apresentam resultados satisfatórios e provam que a resistência à compressão do concreto pode ser prevista de forma relativamente fácil. A alternativa adotada de criar um modelo para cada conjunto de idade se mostrou como uma alternativa válida, conseguindo estratificar para obter resultados específicos de cada conjunto. Os estudos citados na introdução utilizando o mesmo conjunto de dados possuem resultados similares, como esperado. A tabela \ref{tab:works-comparison} apresenta os resultados desses trabalhos (\ref{show-works-comparison}), e a tabela \ref{tab:results} apresenta os valores encontrados (\ref{show-results}) para fácil comparação.
```{r works-comparison, echo=F}
# Tabela - "Comparação dos estudos de outros autores"
colNames = c("Autor","Ano" ,"Algorítimo", "RMSE") 
caption <- "Comparação dos estudos de outros autores"
works <- data.frame(
  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
          "Gradient Boosting Regressor","Random Forest Regressor" ,
          "Regularized Extreme Learning Machine",
          "Support Vector Machines with Radial Basis Function Kernel"),
  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
)
kable(
  works,
  col.names = colNames,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "l"
  )  %>%
    kable_styling(latex_options = c("HOLD_position"))
```
```{r results, echo=F}
# Tabela - Resultados finais
colNames <- c("Modelo", "RMSE")
caption <- "Resultados finais"
day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
dat_reg_models <- data.frame(
  dia = day,
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
Seguindo a linha de raciocínio desse trabalho, ele pode ser realizado com diferentes algorítimos, os resultados aqui encontrados utilizaram apenas um único (*Parallel Random Forest*), mesmo que tenha sido teoricamente o "melhor" encontrado, outros algorítmos podem apresentar resultados ainda melhores. Outra opção é criar um *ensemble* com diversos algorítimos, como realizado por @Pierobon2018, mas com a separação de conjuntos de idade aqui proposto. Além disso, pode ser realizado com um conjunto maior de dados, idealmente com o mesmo número de amostras em cada conjunto de idade, distribuição mais homgênia da resistência à compressão e menor variância entre as amostras.

# Bibliografia
<div id="refs"></div>
\newpage

# Appendix 1 - Ambiente virtual {#appendix1}

## Sistema operacional
```{r sistema, echo=FALSE}
v <- as.matrix(R.Version())
kable(v, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```

## Pacotes utilizados
```{r pacotes, echo=FALSE}
p <- matrix(
      c(
        "caret", as.character(packageVersion("caret")),
        "cowplot",as.character(packageVersion("cowplot")),
        "dplyr", as.character(packageVersion("dplyr")),
        "factoextra", as.character(packageVersion("factoextra")),
        "gdata", as.character(packageVersion("gdata")),
        "ggplot2", as.character(packageVersion("ggplot2")),
        "gridExtra",as.character(packageVersion("gridExtra")),
        "kableExtra",as.character(packageVersion("kableExtra")),
        "knitr",as.character(packageVersion("knitr")),
        "pastecs",as.character(packageVersion("pastecs")),
        "purrr", as.character(packageVersion("purrr")),
        "questionr", as.character(packageVersion("questionr")),
        "reshape2", as.character(packageVersion("reshape2")),
        "tidyr", as.character(packageVersion("tidyr")),
        "tidyverse", as.character(packageVersion("tidyverse"))
      ),
      ncol = 2,
      byrow=TRUE
    )
kable(p, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```
\newpage

# Appendix 2 - Repositório online {#appendix2}
Todo o código utilizado para gerar esse trabalho está disponível no *GitHub* no link a seguir:
- https://github.com/pedrobern/concrete-compressive-strength-prediction
\newpage

# Appendix 3 - Código {#appendix3}

## Obtenção dos dados
### Download dos dados

```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
```

### Renomeando as colunas

```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
```

### Reordenando os dados

```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
```

### Definindo nomes e unidades das colunas

```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
```

### Tabela Primeiras amostras

```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
```

## Preparação dos dados
### Removendo amostras duplicadas

```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
```

### Tabela - Amostras com a mesma composição

```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
```

### Tabela - Amostras iguais com resultados diferentes

```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
```

### Limpeza inicial das amotras

```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
```

### Tabela - Amostras anteriores após processamento

```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
```

### Figura - Resistência à compressão (MPa) vs idade (dias)

```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
```

### Figura - Análise componente principal - 90, 91 e 100 dias

```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
```

### Figura - Resistência à compressão ao longo do tempo

```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
```

### Juntando amostras de 90, 91 e 100 dias

```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
```

### Figura - Frequência das idades

```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
```

### Removendo idades com frequência menor que 50

```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
```

### Reorganização das amotras

```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
```

### Tabela - Primeiras 6 amostras reorganizadas

```{r show-new-features, ref.label='new-features', ref='show-new-features', eval=FALSE}
```

### Total de amostras

```{r show-total-samples-2, ref.label='total-samples-2', ref='show-total-samples-2', eval=FALSE}
```

### Adicionando novas variáveis

```{r show-new-features-2, ref.label='new-features-2', ref='show-new-features-2', eval=FALSE}
```

### Tabela - Novas variáveis

```{r show-new-features-table, ref.label='new-features-table', ref='show-new-features-table', eval=FALSE}
```

## Visualização dos dados
### Tabela - Estatística descritiva - variáveis contínuas

```{r show-stat-summ, ref.label='stat-summ', ref='show-stat-summ', eval=FALSE}
```

### Figura - Estatística descritiva - variáveis discretas

```{r show-stat-summ-categorical, ref.label='stat-summ-categorical', ref='show-stat-summ-categorical', eval=FALSE}
```

### Figura - Correlações em cada idade

```{r show-correlation, ref.label='correlation', ref='show-correlation', eval=FALSE}
```

### Figura - Relação entre o traço aproximado, água, MPa e idade

```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
```

### Figura - Relação das principais características do concreto

```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
```

### Figura - Distribuição das variáveis

```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
```

### Figura - Distribuição das variáveis agrupadas por idade

```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
```

### Figura - Análise componente principal nos ingredientes

```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
```

## Modelos de machine learning
### Variáveis fictícias - dummy vars

```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
```

### Preparação dos dados

```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
```

### Tabela - Primeiras 18 colunas das primeiras 6 amostras de 28 dias

```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
```

### Removendo colunas com variância próxima a zero

```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
```

### Verificação de variáveis com alta correlação

```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
```

### Separação em conjunto de teste e treino

```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
```

### Distribuição dos conjuntos de teste e treino

```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
```

### Modelo ingênuo

```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
```

### Tabela - Modelos ingênuo

```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
```

### Escolha das caracterísitcas (features)

```{r show-test-models-reg, ref.label='test-models-reg', ref='show-test-models-reg', eval=FALSE}
```

### Tabela - Primeiras 6 amostras do conjunto de treino do modelo de 28 dias

```{r show-table-done-samples-28, ref.label='table-done-samples-28', ref='show-table-done-samples-28', eval=FALSE}
```

### Modelos de regressão

```{r show-reg-models, ref.label='reg-models', ref='show-reg-models', eval=FALSE}
```

# Results
### Tabela - Detalhes dos modelso

```{r show-table-reg-models, ref.label='table-reg-models', ref='show-table-reg-models', eval=FALSE}
```

### Figura - Comparação dos modelos

```{r show-results-comparison, ref.label='results-comparison', ref='show-results-comparison', eval=FALSE}
```

### Tabelas dos 10 melhores e piores resultados

```{r show-table-10, ref.label='table-10', ref='show-table-10', eval=FALSE}
```

# Discussão
### Tabela - "Comparação dos estudos de outros autores"

```{r show-works-comparison, ref.label='works-comparison', ref='show-works-comparison', eval=FALSE}
```

### Tabela - Resultados finais

```{r show-results, ref.label='results', ref='show-results', eval=FALSE}
```

### 2020-04-25 generate_papers.py
Fix appendix hastags

@@ -38,7 +38,10 @@ def generate_appendix_3_subsection(file, L="EN", hastags=2):
         file_title = re.findall("<"+L+">\s*(#{"+str(hastags)+"}.*)\s", f)[0]
         titles = re.findall("^#<"+L+".*?>(#.+?)</"+L+">", f, flags=re.MULTILINE)
         chunk_names = re.findall("```{r\s([\w|-]*),", f, flags=re.DOTALL)
-        s = file_title + "\n"
+        s = ""
+        if hastags == 1:
+            s = "#"	
+        s += file_title + "\n"
         for i in range(len(titles)):
             s += generate_appendix_code_chunk(titles[i], chunk_names[i])
         return(s)

### 2020-04-25 generated/paper_EN.Rmd
Add papers

@@
 In the case of the same samples but with different results of compressive strength, the values were averaged.
-After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).
+After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.
 
@@
 
-```{r join-90-91-100, echo=F, eval=F}
+```{r join-90-91-100, echo=F}
 # Joining 90, 91 and 100 days data
@@
 ```{r new-features, echo=F}
-# Table - First 6 samples after reorganization/EN>
+# Table - First 6 samples after reorganization
 caption <- "First 6 samples after reorganization"
@@
 ```{r total-samples-2, echo=F}
-# Total samples/EN>
+# Total samples
 n_samples <- nrow(dat) # 416
@@
 ```{r new-features-2, echo=F}
-# Adding New features/EN>
+# Adding New features
 concrete_class <- function(mpa){
@@
 ```{r new-features-table, echo=F}
-# Table - New features/EN>
+# Table - New features
 caption <- "New features"
@@
 ```{r stat-summ, echo=FALSE}
-# Table - Descriptive statistics - continuous variables/EN>
+# Table - Descriptive statistics - continuous variables
 summ <- t(
@@
 ```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
-# Figure - Descriptive statistics - categorical variables/EN>
+# Figure - Descriptive statistics - categorical variables
 cap <- "Descriptive statistics - categorical variables"
@@
 ```{r correlation, echo=F, fig.height=4, fig.cap=cap}
-# Figure - Correlation grouped by age/EN>
+# Figure - Correlation grouped by age
 cor_dat <- dat %>% select(-c(id))
@@
 ```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
+# Figure - Correlation over time
 cap <- "Correlation of variables with compressive strength over time"
@@
 ```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
-# Figure - Relationship between approximated mix, water, MPa and age/EN>
+# Figure - Relationship between approximated mix, water, MPa and age
 cap <- "Relationship between approximated mix, water, MPa and age"
@@
   drop_na()
-mix_dat$day <- factor(mix_dat$day)
+lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
+mix_dat$day <- factor(mix_dat$day, levels=lvls)
 levels(mix_dat$day) <- labs
@@
 ```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
-# Figure - Relationship between concrete main features/EN>
+# Figure - Relationship between concrete main features
 cap <- "Relationship between concrete main features"
@@
 ```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
-# Figure - Variables distribution/EN>
+# Figure - Variables distribution
 
@@
 ```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
-# Figure - Variables distribution grouped by age/EN>
+# Figure - Variables distribution grouped by age
 
@@
 ```{r pca, echo=F, fig.height=3, fig.cap=cap}
-# Figure - Principal component analysis on ingredients/EN>
+# Figure - Principal component analysis on ingredients
 
@@
 reg <- list(
-  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$MPa),
-  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$MPa),
-  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$MPa),
-  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$MPa),
-  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$MPa),
-  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$MPa)
+  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$mpa),
+  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$mpa),
+  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$mpa),
+  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$mpa),
+  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$mpa),
+  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$mpa)
 )
@@
         select(y) %>%
-        mutate(Grup = g1, day = str_sub(n, 2)),
+        mutate(Group = g1, day = str_sub(n, 2)),
       d$test %>%
         select(y) %>%
-        mutate(Grup = g2, day = str_sub(n, 2))
+        mutate(Group = g2, day = str_sub(n, 2))
     )
@@
 densities %>%
-  ggplot(aes(y, fill = Grup)) +
+  ggplot(aes(y, fill = Group)) +
   geom_density(alpha = 0.5) +
-  facet_wrap(~dia_f, ncol=3) +
+  facet_wrap(~day_f, ncol=3) +
   xlab("MPa") +
@@
     df_performance_reg,
-    col.names = colNames6,
+    col.names = colNames,
     escape = F,
@@
 res_df <- bind_rows(res_df)
-res_df$model <- factor(res_df$modelo,levels=modelos)
+res_df$model <- factor(res_df$model,levels=models)
 res_df %>%
   ggplot(aes(actual, pred)) +
-  facet_wrap(~ modelo, ncol=3) +
+  facet_wrap(~ model, ncol=3) +
   geom_point(alpha=0.5) +
@@
 # Appendix 2 - Online repository {#appendix2}
-All the code used to generate this work is available on *GitHub* at the following link:
-- https://github.com/pedrobern/concrete-compressive-strength-prediction
+https://github.com/pedrobern/concrete-compressive-strength-prediction
 \newpage
@@
 
+### Table - First 6 samples after reorganization
+
+```{r show-new-features, ref.label='new-features', ref='show-new-features', eval=FALSE}
+```
+
+### Total samples
+
+```{r show-total-samples-2, ref.label='total-samples-2', ref='show-total-samples-2', eval=FALSE}
+```
+
+### Adding New features
+
+```{r show-new-features-2, ref.label='new-features-2', ref='show-new-features-2', eval=FALSE}
+```
+
+### Table - New features
+
+```{r show-new-features-table, ref.label='new-features-table', ref='show-new-features-table', eval=FALSE}
+```
+
 ## Data visualization
+### Table - Descriptive statistics - continuous variables
+
+```{r show-stat-summ, ref.label='stat-summ', ref='show-stat-summ', eval=FALSE}
+```
+
+### Figure - Descriptive statistics - categorical variables
+
+```{r show-stat-summ-categorical, ref.label='stat-summ-categorical', ref='show-stat-summ-categorical', eval=FALSE}
+```
+
+### Figure - Correlation grouped by age
+
+```{r show-correlation, ref.label='correlation', ref='show-correlation', eval=FALSE}
+```
+
+### Figure - Correlation over time
+
+```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
+```
+
+### Figure - Relationship between approximated mix, water, MPa and age
+
+```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
+```
+
+### Figure - Relationship between concrete main features
+
+```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
+```
+
+### Figure - Variables distribution
+
+```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
+```
+
+### Figure - Variables distribution grouped by age
+
+```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
+```
+
+### Figure - Principal component analysis on ingredients
+
+```{r show-pca, ref.label='pca', ref='show-pca', eval=FALSE}
+```
+
 ## Machine learning models
@@
 
-# Results
+## Results
 ### Table - Models details
@@
 
-# Discussion
+## Discussion
 ### Table - Comparison of other works

### 2020-04-25 generated/paper_PT.Rmd
Add papers

@@
 title:
-  Prevendo a resisência à compressão do concreto com técnicas de machine learning
+  Prevendo a resistência à compressão do concreto com técnicas de machine learning
 date: "Abril 22, 2020"
@@
 No caso das amostras iguais mas com resultado diferentes de resistência à compressão, foi calculado a média dos valores.
-Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes (\ref{show-total-de-amostras-apos-processamento}).
+Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes.
 
@@
 
-```{r join-90-91-100, echo=F, eval=F}
+```{r join-90-91-100, echo=F}
 # Juntando amostras de 90, 91 e 100 dias
@@
 ```{r new-features, echo=F}
+# Tabela - Primeiras 6 amostras reorganizadas
 caption <- "Primeiras 6 amostras reorganizadas"
@@
 ```{r total-samples-2, echo=F}
+# Total de amostras
+n_samples <- nrow(dat) # 416
+n_distinct_samples <- n_distinct(dat$id) # 416
+```
 
@@
 ```{r new-features-2, echo=F}
+# Adicionando novas variáveis
+concrete_class <- function(mpa){
+  if (mpa >= 10) {
+    s <- as.character(mpa)
+    first <- substr(s, start = 1, stop = 1)
+    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
+  }
+  else {
+    first <- ""
+    second <- "5"
+  }
+  paste("C", first, second, sep = "")
+}
+mix <- function(c, f_ag, c_ag){
+  paste(
+    1,
+    round(f_ag/c, 0),
+    round(c_ag/c, 0),
+    sep = ":")
+}
+dat <- dat %>%
+  mutate(class = sapply(day_28, concrete_class)) %>%
+  mutate(class = as.factor(class)) %>%
+  mutate(mix_app = factor(
+    mix(cement, fine_aggregate, coarse_aggregate))) %>%
+  mutate(`water_/_cement` = water / cement) %>%
+  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
+  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
+  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
+  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
+  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
+lvl <- levels(dat$class)
+dat$class <- factor(
+  dat$class, 
+  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
+```
+```{r new-features-table, echo=F}
+# Tabela - Novas variáveis
 caption <- "Novas variáveis"
@@
 ```{r stat-summ, echo=FALSE}
+# Tabela - Estatística descritiva - variáveis contínuas
+summ <- t(
+  stat.desc(select(dat, -c(id, class, mix_app))))
 caption <- "Estatística descritiva - variáveis contínuas"
@@
 ```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
+# Figura - Estatística descritiva - variáveis discretas
 cap <- "Estatística descritiva - variáveis discretas"
@@
 ```{r correlation, echo=F, fig.height=4, fig.cap=cap}
+# Figura - Correlações em cada idade
+cor_dat <- dat %>% select(-c(id))
 cap <- "Correlações em cada idade"
@@
 ```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
+# Figura - Correlações no tempo
 cap <- "Correlação das variáveis com a resistência à compressão no tempo"
@@
 ```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
+# Figura - Relação entre o traço aproximado, água, MPa e idade
 cap <- "Relação entre o traço aproximado, água, resistência à compressão e idade"
@@
 ```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
+# Figura - Relação das principais características do concreto
 cap <- "Relação das principais proporções do concreto"
@@
 ```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
+# Figura - Distribuição das variáveis  
 cap <- "Distribuição das variáveis"
@@
 ```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
+# Figura - Distribuição das variáveis agrupadas por idade  
+days_labs <- c("3","7","14","28","56","100")
 cap <- "Distribuição das variáveis em relação a idade"
@@
 ```{r pca, echo=F, fig.height=3, fig.cap=cap}
+# Figura - Análise componente principal nos ingredientes  
 cap <- "Análise componente principal nos ingredientes"
@@
 
-# Results
+# Resultados
 O *RMSE* de teste de cada modelo em ordem crescente de idade foi respectivamente `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` e `r round(RMSE_test_100,2)`.
@@
 ```{r table-reg-models, echo=F}
-# Tabela - Detalhes dos modelso
+# Tabela - Detalhes dos modelos
 colNames <- c("Modelo", "mtry", "CV", "Repetições", 
@@
 # Appendix 2 - Repositório online {#appendix2}
-Todo o código utilizado para gerar esse trabalho está disponível no *GitHub* no link a seguir:
-- https://github.com/pedrobern/concrete-compressive-strength-prediction
+https://github.com/pedrobern/concrete-compressive-strength-prediction
 \newpage
@@
 
+### Figura - Correlações no tempo
+
+```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
+```
+
 ### Figura - Relação entre o traço aproximado, água, MPa e idade
 
-```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
+```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
 ```
@@
 
-```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
+```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
 ```
@@
 
-```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
+```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
 ```
@@
 
-```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
+```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
 ```
@@
 
-```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
+```{r show-pca, ref.label='pca', ref='show-pca', eval=FALSE}
 ```
@@
 
-# Results
-### Tabela - Detalhes dos modelso
+## Resultados
+### Tabela - Detalhes dos modelos
 
@@
 
-# Discussão
+## Discussão
 ### Tabela - "Comparação dos estudos de outros autores"

### 2020-04-25 templates/Abstract_&_Setup.Rmd
Fix typo

@@
 <PT>title:
-  Prevendo a resisência à compressão do concreto com técnicas de machine learning
+  Prevendo a resistência à compressão do concreto com técnicas de machine learning
 date: "Abril 22, 2020"</PT>

### 2020-04-25 templates/Materials_and_methods__Data_preparation.Rmd
Remove wrong ref

@@
 In the case of the same samples but with different results of compressive strength, the values were averaged.
-After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).
+After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.
 </EN>
@@
 No caso das amostras iguais mas com resultado diferentes de resistência à compressão, foi calculado a média dos valores.
-Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes (\ref{show-total-de-amostras-apos-processamento}).
+Após todas essas alterações (\ref{show-initial-data-cleaning}), o novo total de amostras foi reduzido para 970, contendo 416 configurações diferentes das proporções de ingredientes.
 </PT>
@@
 
-```{r join-90-91-100, echo=F, eval=F}
+```{r join-90-91-100, echo=F}
 #<EN># Joining 90, 91 and 100 days data</EN>
@@
 ```{r new-features, echo=F}
-#<EN># Table - First 6 samples after reorganization/EN>
+#<EN># Table - First 6 samples after reorganization</EN>
 #<PT># Tabela - Primeiras 6 amostras reorganizadas</PT>
@@
 ```{r total-samples-2, echo=F}
-#<EN># Total samples/EN>
+#<EN># Total samples</EN>
 #<PT># Total de amostras</PT>
@@
 ```{r new-features-2, echo=F}
-#<EN># Adding New features/EN>
+#<EN># Adding New features</EN>
 #<PT># Adicionando novas variáveis</PT>
@@
 ```{r new-features-table, echo=F}
-#<EN># Table - New features/EN>
+#<EN># Table - New features</EN>
 #<PT># Tabela - Novas variáveis</PT>

### 2020-04-25 templates/Materials_and_methods__Data_visualization.Rmd
Fix typo

@@
 ```{r stat-summ, echo=FALSE}
-#<EN># Table - Descriptive statistics - continuous variables/EN>
+#<EN># Table - Descriptive statistics - continuous variables</EN>
 #<PT># Tabela - Estatística descritiva - variáveis contínuas</PT>
@@
 ```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
-#<EN># Figure - Descriptive statistics - categorical variables/EN>
+#<EN># Figure - Descriptive statistics - categorical variables</EN>
 #<PT># Figura - Estatística descritiva - variáveis discretas</PT>
@@
 ```{r correlation, echo=F, fig.height=4, fig.cap=cap}
-#<EN># Figure - Correlation grouped by age/EN>
+#<EN># Figure - Correlation grouped by age</EN>
 #<PT># Figura - Correlações em cada idade</PT>
@@
 ```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
+#<EN># Figure - Correlation over time</EN>
+#<PT># Figura - Correlações no tempo</PT>
 
@@
 ```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
-#<EN># Figure - Relationship between approximated mix, water, MPa and age/EN>
+#<EN># Figure - Relationship between approximated mix, water, MPa and age</EN>
 #<PT># Figura - Relação entre o traço aproximado, água, MPa e idade</PT>
@@
 
-mix_dat$day <- factor(mix_dat$day)
+lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
+mix_dat$day <- factor(mix_dat$day, levels=lvls)
 levels(mix_dat$day) <- labs
@@
 ```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
-#<EN># Figure - Relationship between concrete main features/EN>
+#<EN># Figure - Relationship between concrete main features</EN>
 #<PT># Figura - Relação das principais características do concreto</PT>
@@
 ```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
-#<EN># Figure - Variables distribution/EN>
+#<EN># Figure - Variables distribution</EN>
 #<PT># Figura - Distribuição das variáveis</PT>
@@
 ```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
-#<EN># Figure - Variables distribution grouped by age/EN>
+#<EN># Figure - Variables distribution grouped by age</EN>
 #<PT># Figura - Distribuição das variáveis agrupadas por idade</PT>
@@
 ```{r pca, echo=F, fig.height=3, fig.cap=cap}
-#<EN># Figure - Principal component analysis on ingredients/EN>
+#<EN># Figure - Principal component analysis on ingredients</EN>
 #<PT># Figura - Análise componente principal nos ingredientes</PT>

### 2020-04-25 generated/paper_EN.Rmd
Fix typo

@@
 ## Operational system
-```{r sistema, echo=FALSE}
+```{r system, echo=FALSE}
 v <- as.matrix(R.Version())
@@
 ## Packages
-```{r pacotes, echo=FALSE}
+```{r packages, echo=FALSE}
 p <- matrix(

### 2020-04-25 generated/paper_PT.Rmd
Fix typo

@@
 ## Sistema operacional
-```{r sistema, echo=FALSE}
+```{r system, echo=FALSE}
 v <- as.matrix(R.Version())
@@
 ## Pacotes utilizados
-```{r pacotes, echo=FALSE}
+```{r packages, echo=FALSE}
 p <- matrix(

### 2020-04-28 templates/Materials_and_methods__Data_visualization.Rmd
Fix typo

@@
 colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
-                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
-                "14", "28", "56", "100",
-                "W./C.", "F.A./C.", 
+                "Superplast.", "Coarse agg.", "Fine agg.",
+                "3", "7", "14", "28", "56", "100",
+                "class", "mix_app", "W./C.", "F.A./C.", 
                 "C.A./C.", "F.A./C.A.",

### 2020-04-29 templates/Abstract_&_Setup.Rmd
Update abstract

--- 2020-04-25 templates/Abstract_&_Setup.Rmd
@@
  Compressive strength is the main characteristic of concrete.
-This work used a set of samples widely used to build 6 forecast models of this characteristic, each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers) and 6 relationships between the main ingredients (water/cement, coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate, water / coarse aggregate and water/fine aggregate) to make the predictions.
-The algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*) .
-The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31) and the worst for the 100-day model (*RMSE* of 5.85).
-The 28-day model had *RMSE* of 4.72.
-The work indicated that the separation in a model for each age is promising and was able to generate compatible results or better than other studies, opening way for new models creation following this proposal, but using different algorithms or the combination of different algorithms.</EN>
+The correct prediction of this parameter means cost and time reduction.
+This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
+A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: resistance to compression, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
+The predictive models were developed in *R* language, using the *caret* package with the *Parallel Random Forest* algorithm and repeated cross-validation technique to optimize the parameters.
+The results were satisfactory and compatible with other studies using the same data set.
+The most important model, 28 days old, obtained *RMSE* of 4.717.
+The 3-day model obtained the best result, *RMSE* of 3.310.
+The worst result was the 56-day model, with *RMSE* of 5.939.
+The work showed that the compressive strength of concrete can be predicted.
+The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age.
+It was a promising alternative since good results were achieved by training with just one algorithm.
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.</EN>
 <PT>abstract:
   A resistência à compressão é a principal característica do concreto.
-Esse trabalho utilizou um conjunto de amostras amplamente utilizado para construir 6 modelos de previsão dessa característica, cada um específicamente para uma idade, 3, 7, 14, 28, 56 e 100 dias a partir de 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes) e 6 relações entre os principais ingredientes (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo) para fazer a previsão.
-Foi utilizado o algorítimo *Parallel Random Forest* (*parRF* do pacote *caret* de linguagem *R*).
-Os resultados foram satisfatórios sendo o melhor para o modelo de 3 dias (*RMSE* de 3,31) e pior para o modelo de 100 dias (*RMSE* de 5,85).
-O modelo de 28 dias apresentou *RMSE* de 4,72.
-O trabalho indicou que a separação em um modelo para cada idade é promissora e foi capaz de gerar resultados compatíveis ou melhores que outros estudos, abrindo espaço para construção de novos modelos seguindo essa proposta, mas utilizando algorítimos diferentes ou a combinação de diversos algorítimos.</PT>
+A previsão correta desse parâmetro significa redução de custo e tempo.
+Esse trabalho construiu modelos de previsão em 6 idades diferentes de amostras de concreto (3, 7, 14, 28, 56, e 100 dias).
+Foi utilizado um conjunto de dados obtido em estudos anteriores, um total de 1030 amostras, com 9 variáveis: resistência à compressão, idade e 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes).
+Outras 6 variáveis foram adicionadas para representar as proporções dos principais ingredientes em cada amostra (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo).
+Os modelos de previsão foram desenvolvidos em linguagem *R*, utilizando o pacote *caret* com o algoritmo *Parallel Random Forest* e técnica de validação cruzada repetida para otimização dos parâmetros.
+Os resultados foram satisfatórios e compatíveis com outros estudos utilizando o mesmo conjunto de dados.
+O modelo mais importante, de 28 dias, obteve *RMSE* de 4,717.
+O modelo de 3 dias obteve o melhor resultado, *RMSE* de 3,310.
+O pior resultado foi o modelo de 56 dias, com *RMSE* de 5,939.
+O trabalho mostrou que a resistência à compressão do concreto pode ser prevista.
+A escolha de criar um modelo para cada idade, ao invés de utilizar a idade como característica para previsão, permitiu chegar a resultados compatíveis com os dados disponíveis de cada idade.
+Foi uma alternativa promissora, visto que bons resultados foram atingidos treinando com apenas um algoritmo.
+Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos.</PT>
 output:

### 2020-05-08 generated/paper_PT.Rmd
Add paper

@@
 indent: true
-abstract:
-  A resistência à compressão é a principal característica do concreto.
-Esse trabalho utilizou um conjunto de amostras amplamente utilizado para construir 6 modelos de previsão dessa característica, cada um específicamente para uma idade, 3, 7, 14, 28, 56 e 100 dias a partir de 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes) e 6 relações entre os principais ingredientes (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo) para fazer a previsão.
-Foi utilizado o algorítimo *Parallel Random Forest* (*parRF* do pacote *caret* de linguagem *R*).
-Os resultados foram satisfatórios sendo o melhor para o modelo de 3 dias (*RMSE* de 3,31) e pior para o modelo de 100 dias (*RMSE* de 5,85).
-O modelo de 28 dias apresentou *RMSE* de 4,72.
-O trabalho indicou que a separação em um modelo para cada idade é promissora e foi capaz de gerar resultados compatíveis ou melhores que outros estudos, abrindo espaço para construção de novos modelos seguindo essa proposta, mas utilizando algorítimos diferentes ou a combinação de diversos algorítimos.
+abstract: "A resistência à compressão é a principal característica do concreto.
+A previsão correta desse parâmetro significa redução de custo e tempo.
+Esse trabalho construiu modelos de previsão em 6 idades diferentes de amostras de concreto (3, 7, 14, 28, 56, e 100 dias).
+Foi utilizado um conjunto de dados obtido em estudos anteriores, um total de 1030 amostras, com 9 variáveis: resistência à compressão, idade e 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes).
+Outras 6 variáveis foram adicionadas para representar as proporções dos principais ingredientes em cada amostra (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo).
+Os modelos de previsão foram desenvolvidos em linguagem *R*, utilizando o pacote *caret* com o algoritmo *Parallel Random Forest* e técnica de validação cruzada repetida para otimização dos parâmetros.
+Os resultados foram satisfatórios e compatíveis com outros estudos utilizando o mesmo conjunto de dados.
+O modelo mais importante, de 28 dias, obteve *RMSE* de 4,717.
+O modelo de 3 dias obteve o melhor resultado, *RMSE* de 3,310.
+O pior resultado foi o modelo de 56 dias, com *RMSE* de 5,939.
+O trabalho mostrou que a resistência à compressão do concreto pode ser prevista.
+A escolha de criar um modelo para cada idade, ao invés de utilizar a idade como característica para previsão, permitiu chegar a resultados compatíveis com os dados disponíveis de cada idade.
+Foi uma alternativa promissora, visto que bons resultados foram atingidos treinando com apenas um algoritmo.
+Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos."
 output:
@@
 - \usepackage{multirow}
-- \usepackage[table]{xcolor}
 - \usepackage{wrapfig}
@@
 ```{r first-samples, echo=F}
-# Tabela Primeiras amostras
+# Tabela - Primeiras amostras
 caption <- "Primeiras 6 amostras"
@@
 A fígura \ref{fig:pca-90-91-100} mostra como as amostras se relacionam umas com as outras (quais são parecidas ou diferentes) e revelou como cada variável contribui para a análise.
-As duas primeiras dimensões representam $37\%$ e $24\%$ respectivamente da variância.
+As duas primeiras dimensões representam 37% e 24% respectivamente da variância.
 
@@
 Pelos motivos apresentados nas figuras \ref{fig:boxplot}, \ref{fig:pca-90-91-100} e \ref{fig:mpa-on-time}, foi considerado que as idades de 90, 91 e 100 dias podem ser agrupadas para melhorar a leitura e diminuir o ruído das amostras.
-Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}, pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
+Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}), pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
 
@@
 ind_100 <- dat$id[which(dat$day == "100")]
-sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
+sum_duplicated <- sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
 dat <- dat %>%
@@
 caption <- "Primeiras 18 colunas das primeiras 6 amostras de 28 dias"
-dat_tabela <- dat_28[1:18]
+dat_table <- dat_28[1:18]
 kable(
-    head(dat_tabela[order(dat_tabela$id),]),
+    head(dat_table[order(dat_table$id),]),
     col.names = dfUnits,
@@
 Como ao longo do processamento foram adicionadas novas variáveis (as relações entre os ingredientes e mais algumas *dummy vars* para cada conjunto de idade), foram estudadas 5 possibilidades de configurações das *features* para os modelos:
+
+&nbsp;
+
 1. Todas as *features*;
@@
 5. Apenas *features* adicionadas, sem as *dummy vars*;
+
+&nbsp;
+
 Construindo um modelo para cada uma dessas configurações utilizando o conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou que a melhor opção é a configuração 2, ou seja, as *dummy vars* foram completamente descartadas, porém foram mantidas as outras novas varáveis.
@@
 
-### Tabela Primeiras amostras
+### Tabela - Primeiras amostras
 

### 2020-05-08 templates/Discussion.Rmd
Add paper

@@
 The studies cited in the introduction using the same dataset have similar results, as expected.
-The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
+The table \ref{tab:works-comparison} presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
 </EN>
@@
   ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
-  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
+  modelo = c("5 Algorithm Ensemble", "Artificial Neural Networks", 
           "Gradient Boosting Regressor","Random Forest Regressor" ,

### 2020-05-08 templates/Materials_and_methods__Data_visualization.Rmd
Add paper

@@
 
-The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}).
+The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}).
 The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data.
@@
 
-A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias.
-
 The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
@@
 
-Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}).
-A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso.
-É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
-
-In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
+In the figure \ref{fig:pca}, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
 The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that.

### 2020-05-08 templates/Materials_and_methods__Obtaining_the_data.Rmd
Add paper

@@
 The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}).
-The first samples can be viewed in the table \ref{tab: first-samples}.
+The first samples can be viewed in the table \ref{tab:first-samples}.
 </EN>
@@
 ```{r rename-dat-cols, echo=F}
-#<EN># Ranaming the columns</EN>
+#<EN># Renaming the columns</EN>
 #<PT># Renomeando as colunas</PT>
@@
 ```{r first-samples, echo=F}
-#<EN># Table First samples</EN>
-#<PT># Tabela Primeiras amostras</PT>
+#<EN># Table - First samples</EN>
+#<PT># Tabela - Primeiras amostras</PT>
 

### 2020-05-14 CILAMCE/paper_EN.Rmd
Add CILAMCE template and paper sketch

---
title: "INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE"
indent: true
output:
  pdf_document:
   keep_tex: yes
   number_sections: yes
   template: CILAMCE_template.tex
   citation_package: natbib
   latex_engine: xelatex
mainfont: Times New Roman
bibliography: references.bib
biblio-style: ./bib-cilamce
---

<!-- \usepackage[figureposition=bottom,tableposition=top]{caption} -->

<!-- spacing before and after equations -->
\setlength{\abovedisplayskip}{18pt}
\setlength{\belowdisplayskip}{18pt}
\setlength{\abovedisplayshortskip}{18pt}
\setlength{\belowdisplayshortskip}{18pt}

<!-- table row spacing -->
\renewcommand{\arraystretch}{1.2}

<!-- table and caption spacing -->
\captionsetup[table]{skip=12pt}

<!-- space before and after figures and tables -->
\setlength\intextsep{13pt}

```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=F, prompt=F, highlight=T )
```

# Lorem ipsum

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

# Lorem ipsum dolor

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

## Lorem ipsum Sub

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

## Lorem ipsum Sub 2

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

## Equations, symbols and units

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

\begin{center}
\begin{equation}
q_r = -4pr^2k\frac{dT}{dr}.
\label{Eq1}
\end{equation}
\end{center}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

\vspace{13pt}

Cite as Eq. \ref{Eq1}.

## Figures and tables

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

```{r table-example, echo=F, eval=T}

df <- data.frame(
   col1 = c("Turbulent tensor", "Turbulent tensor", "Lateral lift", "Virtual mass"),
   col2 = c("C", "C", "C", "C"),
   col2 = c(0.09, 0.69, 0.08, 0.8)
)

kable(
    df,
    "latex",
    col.names = c("Constitutive relation", "Nomenclature", "Value"),
    escape = F,
    booktabs = T,
    caption = "Coefficients in constitutive relations",
    linesep = "",
    align = "c"
    ) %>%
  kable_styling(latex_options = c("HOLD_position"))
```

Cite as Figure \ref{fig:plot-example} and Table \ref{tab:table-example}.

```{r plot-example, echo=F, eval=T, fig.cap="Pressure variation along the nozzle: experimental data", fig.width=4, fig.height=3, fig.pos="!ht"}

df <- data.frame(
   x = 1:10,
   y = (1:10)^2
)

df %>%
  ggplot(aes(x, y)) +
  geom_point()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

# References

Example of reference: "@Yeh1998 proposed… @Hameed2020". Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

### 2020-06-01 HarvardX_Data_Science_Capstone/capstone.Rmd
Update files

--- 2020-04-25 generated/paper_EN.Rmd
@@
 indent: true
-abstract:
- Compressive strength is the main characteristic of concrete.
-This work used a set of samples widely used to build 6 forecast models of this characteristic, each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers) and 6 relationships between the main ingredients (water/cement, coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate, water / coarse aggregate and water/fine aggregate) to make the predictions.
-The algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*) .
-The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31) and the worst for the 100-day model (*RMSE* of 5.85).
-The 28-day model had *RMSE* of 4.72.
-The work indicated that the separation in a model for each age is promising and was able to generate compatible results or better than other studies, opening way for new models creation following this proposal, but using different algorithms or the combination of different algorithms.
+abstract: "Compressive strength is the main characteristic of concrete.
+The correct prediction of this parameter means cost and time reduction.
+This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
+A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
+The predictive models were developed in *R* language, using the *caret* package with the *Parallel Random Forest* algorithm and repeated cross-validation technique to optimize the parameters.
+The results were satisfactory and compatible with other studies using the same data set.
+The most important model, 28 days old, obtained *RMSE* of 4.717.
+The 3-day model obtained the best result, *RMSE* of 3.310.
+The worst result was the 56-day model, with *RMSE* of 5.939.
+The work showed that the compressive strength of concrete can be predicted.
+The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age.
+It was a promising alternative since good results were achieved by training with just one algorithm.
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several."
 output:
@@
 The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}).
-The first samples can be viewed in the table \ref{tab: first-samples}.
+The first samples can be viewed in the table \ref{tab:first-samples}.
 
@@
 ```{r rename-dat-cols, echo=F}
-# Ranaming the columns
+# Renaming the columns
 colnames(dat) <- c(
@@
 ```{r first-samples, echo=F}
-# Table First samples
+# Table - First samples
 caption <- "First 6 samples"
@@
 The data show the variables in the columns and samples in the rows.
-However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
+However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
 
@@
 ```{r similar-samples, echo=FALSE}
-# Tabela - Samples with same composition
+# Table - Samples with same composition
 same_samples <- dat %>%
@@
 ```{r similar-samples-2, echo=FALSE}
-# Tabela - Same samples with different results
+# Table - Same samples with different results
 same_samples_2 <- dat %>%
@@
 ```{r similar-samples-same-id, echo=FALSE, include=TRUE}
-# Tabela - Previous samples after processing
+# Table - Previous samples after processing
 same_samples <- dat %>%
@@
 ### Age selection
-As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models.
+As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models.
 However, it is necessary to verify how relevant the data of these other ages are.
@@
 The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis.
-The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.
+The first two dimensions represent 37% and 24% $ respectively of the variance.
 
@@
 Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value.
-The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).
+The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-time}).
 
@@
 For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise.
-They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}.
+They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}).
 As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.
@@
 ### Data reorganization
-The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
+The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
 The result in the first samples after this processing is shown in the table \ref{tab:new-features}.
@@
 Other proportions between the main ingredients were also added.
-The new variables are presented in the table ref{tab:new-features-table}.
+The new variables are presented in the table \ref{tab:new-features-table}.
 ```{r new-features-2, echo=F}
@@
 ### Descriptive statistics
-The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}).
+The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}).
 The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data.
@@
 ### Variables distribution
-A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
 The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
@@
 ### Principal component analysis
-Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}).
-A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso.
-É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
-In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
+In the figure \ref{fig:pca}, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
 The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that.
@@
 ## Machine learning models
-The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.
+The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.
 
 ### Pre-processing and data separation
-Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado.
 As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.
@@
 One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}).
-For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.
+For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-preparated-samples}.
 
@@
 caption <- "First 18 columns of the 6 first samples of 28 days"
-dat_tabela <- dat_28[1:18]
+dat_table <- dat_28[1:18]
 kable(
-    head(dat_tabela[order(dat_tabela$id),]),
+    head(dat_table[order(dat_table$id),]),
     col.names = dfUnits,
@@
 As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
+
+&nbsp;
+
 1. All *features*;
@@
 5. Only new *features*, without *dummy vars*;
+
+&nbsp;
+
 Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
@@
 
-### Ranaming the columns
+### Renaming the columns
 
@@
 
-### Table First samples
+### Table - First samples
 
@@
 
-### Tabela - Samples with same composition
+### Table - Samples with same composition
 
@@
 
-### Tabela - Same samples with different results
+### Table - Same samples with different results
 
@@
 
-### Tabela - Previous samples after processing
+### Table - Previous samples after processing
 

### 2020-06-01 HarvardX_Data_Science_Capstone/capstone.R

@@ -1,0 +1,1231 @@
+
+# Install libraries, if not already installed
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(ggplot2)) install.packages("ggplot2")
+if(!require(caret)) install.packages("caret")
+if(!require(knitr)) install.packages("knitr")
+if(!require(kableExtra)) install.packages("kableExtra")
+if(!require(gdata)) install.packages("gdata")
+if(!require(dplyr)) install.packages("dplyr")
+if(!require(tidyr)) install.packages("tidyr")
+if(!require(cowplot)) install.packages("cowplot")
+if(!require(pastecs)) install.packages("pastecs")
+if(!require(factoextra)) install.packages("factoextra")
+if(!require(reshape2)) install.packages("reshape2")
+if(!require(purrr)) install.packages("purrr")
+if(!require(gridExtra)) install.packages("gridExtra")
+if(!require(questionr)) install.packages("questionr")
+
+# Load libraries
+library(tidyverse)
+library(caret)
+library(ggplot2)
+library(knitr)
+library(kableExtra)
+library(dplyr)
+library(tidyr)
+library(gdata)
+library(reshape2)
+library(cowplot)
+library(pastecs)
+library(factoextra)
+library(purrr)
+library(gridExtra)
+library(questionr)
+
+
+# Data download
+url_base <- "https://archive.ics.uci.edu"
+url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
+download.file(paste0(url_base, url), "data.xls")
+dat <- read.xls("data.xls")
+n_inicial_samples <- nrow(dat)
+colnames(dat)
+
+
+# Renaming the columns
+colnames(dat) <- c(
+	"cement",
+	"blast_furnace_slag",
+	"fly_ash",
+	"water",
+	"superplasticizers",
+	"coarse_aggregate",
+	"fine_aggregate",
+	"day",
+	"mpa"
+)
+dat$id <- seq.int(nrow(dat))
+
+		
+# Reordering the data
+col_order <- c(
+	"id",
+	"cement",
+	"blast_furnace_slag",
+	"fly_ash",
+	"water",
+	"superplasticizers",
+	"coarse_aggregate",
+	"fine_aggregate",
+	"day",
+	"mpa"
+)
[…]

### 2020-06-01 HarvardX_Data_Science_Capstone/paper_EN.R

@@ -1,1056 +1,0 @@
-# Install libraries, if not already installed
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(ggplot2)) install.packages("ggplot2")
-if(!require(caret)) install.packages("caret")
-if(!require(gdata)) install.packages("gdata")
-if(!require(ggplot2)) install.packages("ggplot2")
-if(!require(dplyr)) install.packages("dplyr")
-if(!require(tidyr)) install.packages("tidyr")
-if(!require(pastecs)) install.packages("pastecs")
-if(!require(factoextra)) install.packages("factoextra")
-if(!require(reshape2)) install.packages("reshape2")
-if(!require(purrr)) install.packages("purrr")
-if(!require(questionr)) install.packages("questionr")
-
-# Load libraries
-library(tidyverse)
-library(caret)
-library(ggplot2)
-library(knitr)
-library(kableExtra)
-library(dplyr)
-library(tidyr)
-library(gdata)
-library(reshape2)
-library(cowplot)
-library(pastecs)
-library(factoextra)
-library(purrr)
-library(gridExtra)
-library(questionr)
-
-
-# Data download
-url_base <- "https://archive.ics.uci.edu"
-url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
-download.file(paste0(url_base, url), "data.xls")
-dat <- read.xls("data.xls")
-n_inicial_samples <- nrow(dat)
-colnames(dat)
-		
-
-# Renaming the columns
-colnames(dat) <- c(
-	"cement",
-	"blast_furnace_slag",
-	"fly_ash",
-	"water",
-	"superplasticizers",
-	"coarse_aggregate",
-	"fine_aggregate",
-	"day",
-	"mpa"
-)
-dat$id <- seq.int(nrow(dat))
-
-
-# Reordering the data
-col_order <- c(
-	"id",
-	"cement",
-	"blast_furnace_slag",
-	"fly_ash",
-	"water",
-	"superplasticizers",
-	"coarse_aggregate",
-	"fine_aggregate",
-	"day",
-	"mpa"
-)
-dat <- dat[, col_order]
-		
-		
-# Removing duplicated samples
[…]

### 2020-06-01 HarvardX_Data_Science_Capstone/capstone.Rmd
Update files

@@
 ```
+
 \newpage
@@
 # Introduction
+
 Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens.
@@
 ## Materials
+
 The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore].
-Throughout the process, all code executed was documented in the same order as its execution in [Appendix 3](#appendix3), and reference was always made to codes throughout the text.
+Throughout the process, all code executed was documented in the same order as its execution in [Appendix 2](#appendix2), and reference was always made to codes throughout the text.
 All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1).
-In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in [Appendix 2](#appendix2).
 
 ## Reproducibility
+
 In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result.
-The *seeds* can be checked throughout [Appendix 3](#appendix3) or directly on *Github*.
+The *seeds* can be checked throughout [Appendix 2](#appendix2).
 
 ## Obtaining the data
+
 The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}).
@@
 ```
+
 ```{r rename-dat-cols, echo=F}
@@
 ```
+
 ```{r reorder-dat, echo=F}
@@
 ```
+
 ```{r col-names-and-units, echo=F}
@@
 ```
+
 ```{r first-samples, echo=F}
@@
 ## Data preparation
+
 The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies.
@@
 ### Initial data cleaning
+
 Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).
@@
 ```
+
 The data show the variables in the columns and samples in the rows.
@@
 ```
+
 In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process.
@@
 ```
+
 To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*.
@@
 ```
+
 The result can be seen in the table \ref{tab:similar-samples-same-id}.
@@
 ### Age selection
+
 As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models.
@@
 ```
+
 It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.
@@
 ```
+
 Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value.
@@
 ```
+
 For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise.
@@
 ```
+
 Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}.
@@
 ```
+
 ```{r remove-ages-lower-50, echo=F}
@@
 ### Data reorganization
+
 The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
@@
 ```
+
 ```{r new-features, echo=F}
@@
 ```
+
 The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
+
 ```{r total-samples-2, echo=F}
@@
 ### Adding new variables
+
 To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}).
@@
 The new variables are presented in the table \ref{tab:new-features-table}.
+
 ```{r new-features-2, echo=F}
@@
 ```
+
 ```{r new-features-table, echo=F}
@@
 ## Data visualization
+
 In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.
@@
 ### Descriptive statistics
+
 The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}).
@@
 ```
+
 ```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
@@
 ### Correlation between ingredients and compressive strength
+
 The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}).
@@
 ```
+
 ```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
@@
 ```
+
 The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate.
@@
 ```
+
 ```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
@@
 ```
+
 The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}).
@@
 ### Variables distribution
+
 The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
@@
 ```
+
 The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}.
@@
 Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
+
 ```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
@@
 ### Principal component analysis
+
 In the figure \ref{fig:pca}, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
@@
 It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
+
 ```{r pca, echo=F, fig.height=3, fig.cap=cap}
@@
 ## Machine learning models
+
 The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.
@@
 ### Pre-processing and data separation
+
 As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.
@@
 ```
+
 The samples were separated based on age.
@@
 ```
+
 ```{r table-preparated-samples, echo=F}
@@
 ```
+
 For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}).
@@
 ```
+
 ```{r cors, echo=F}
@@
 ```
+
 The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package.
@@
 The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
+
 ```{r split, echo=F, message=F, warning=F}
@@
 ```
+
 ```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
@@
 ### Performance measures
+
 The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*).
@@
 ### Naive models
+
 Before creating the real models, for comparison purposes, naive models were created.
@@
 The results can be checked in the table \ref{tab:table-naive-model-reg}.
+
 ```{r naive-model-reg, echo=F, message=F, warning=F}
@@
 ```
+
 ```{r table-naive-model-reg, echo=F}
@@
 ### Choice of algorithms
+
 The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models.
@@
 ### Regression models
+
 As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
@@
 The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
+
 ```{r test-models-reg, echo=F, warning=F, message=F, include=F}
@@
 ```
+
 ```{r table-done-samples-28, echo=F}
@@
 ```
+
 For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}).
 For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.
+
 ```{r reg-models, echo=F, warning=F, message=F, include=F}
@@
 # Results
+
 The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively.
@@
 The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).
+
 ```{r table-reg-models, echo=F}
@@
 ```
+
 ```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
@@
 ```
+
 ```{r table-10, echo=F}
@@
 
-# Discussion
+# Discussion and conclusion
+
 The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.
@@
 The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
+
 ```{r works-comparison, echo=F}
@@
 ```
+
 ```{r results, echo=F}
@@
 ```
+
 Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results.
@@
 # Literature cited
+
 <div id="refs"></div>
+
 \newpage
@@
 ## Operational system
+
 ```{r system, echo=FALSE}
@@
 ## Packages
+
 ```{r packages, echo=FALSE}
@@
 
-# Appendix 2 - Online repository {#appendix2}
-https://github.com/pedrobern/concrete-compressive-strength-prediction
-\newpage
-
-# Appendix 3 - Code {#appendix3}
+# Appendix 2 - Code {#appendix2}
 
@@
 
-## Discussion
+## Discussion and conclusion
 ### Table - Comparison of other works

### 2020-07-13 HarvardX_Data_Science_Capstone/CopyOfcapstone.Rmd
Harvard submission

--- 2020-06-01 HarvardX_Data_Science_Capstone/capstone.Rmd
@@
     toc_depth: 2
-bibliography: references.bib
+# bibliography: references.bib
+references:
+- id: Yeh1998
+  author:
+  - family: Yeh
+    given: I-Cheng
+  year: 1998
+  month: 12
+  pages: 1797-1808
+  title: Modeling of Strength of High-Performance Concrete Using Artificial Neural Networks
+  volume: 28
+  journal: Cement and Concrete Research
+  DOI: 10.1016/S0008-8846(98)00165-3
+- id: Hameed2020
+  author:
+  - family: Hameed
+    given: Mohammed
+  - family: Khalid
+    given: Mohamed
+   year: 2020
+   month: 01
+   pages: 323-335
+   title: "Prediction of Compressive Strength of High-Performance Concrete: Hybrid Artificial Intelligence Technique"
+   # isbn: 978-3-030-38751-8
+   DOI: 10.1007/978-3-030-38752-5_26
 header-includes:
@@
 # Data download
-url_base <- "https://archive.ics.uci.edu"
-url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
-download.file(paste0(url_base, url), "data.xls")
+# url_base <- "https://archive.ics.uci.edu"
+# url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
+# download.file(paste0(url_base, url), "data.xls")
 dat <- read.xls("data.xls")
@@
 
-## Machine learning models
-
-The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.
-
-### Pre-processing and data separation
-
-As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.
-
-&nbsp;
-
-```{r dummy-var, echo=F, message=F, warning=F}
-# Dummy variables
-dummies <- dummyVars( ~ mix_app, data = dat)
-dummyDat <- data.frame(predict(dummies, newdata = dat))
-dummyDat$id <- dat$id
-dat <- dat %>%
-  select(-c(mix_app)) %>%
-  full_join(., dummyDat)
-```
-
-The samples were separated based on age.
-One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}).
-For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-preparated-samples}.
-
-&nbsp;
-
-```{r preparation, echo=F}
-# Data preparation
-names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
-dat_3 <- dat %>%
-  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
-  drop_na() %>%
-  rename_at("day_3",~"mpa")
-dat_7 <- dat %>%
-  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
-  drop_na() %>%
-  rename_at("day_7",~"mpa")
-dat_14 <- dat %>%
-  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
-  drop_na() %>%
-  rename_at("day_14",~"mpa")
-dat_28 <- dat %>%
-  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
-  drop_na() %>%
-  rename_at("day_28",~"mpa")
-dat_56 <- dat %>%
-  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
-  drop_na() %>%
-  rename_at("day_56",~"mpa")
-dat_100 <- dat %>%
-  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
-  drop_na() %>%
-  rename_at("day_100",~"mpa")
-```
-
-```{r table-preparated-samples, echo=F}
-# Table - First 18 columns of the 6 first samples of 28 days
-colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
-              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
-              "Wat./", "F.Agg./", "C.Agg./",
-              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
-dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
-              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
-              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
-caption <- "First 18 columns of the 6 first samples of 28 days"
-dat_table <- dat_28[1:18]
-kable(
-    head(dat_table[order(dat_table$id),]),
-    col.names = dfUnits,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    digits = 2,
-    align = "c"
-    ) %>%
-    add_header_above(header = colNames, line = F, align = "c") %>%
-    kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-
-For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}).
-Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact.
-In addition to them, in the case of the 7-day set, the fly ash variable was also removed.
-Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}).
-After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.
-
-&nbsp;
-
-```{r nzv, echo=F}
-# Removing near zero variance columns
-nzv_3 <- nearZeroVar(dat_3)
-nzv_7 <- nearZeroVar(dat_7)
-nzv_14 <- nearZeroVar(dat_14)
-nzv_28 <- nearZeroVar(dat_28)
-nzv_56 <- nearZeroVar(dat_56)
-nzv_100 <- nearZeroVar(dat_100)
-dat_3 <- dat_3[,-nzv_3]
-dat_7 <- dat_7[,-nzv_7]
-dat_14 <- dat_14[,-nzv_14]
-dat_28 <- dat_28[,-nzv_28]
-dat_56 <- dat_56[,-nzv_56]
-dat_100 <- dat_100[,-nzv_100]
-```
-
-```{r cors, echo=F}
-# High correlation variables verification
-descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
-descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
-descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
-descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
-descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
-descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
-high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
-high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
-high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
-high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
-high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
-high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
-```
-
-The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package.
-If performed at this time, it would be necessary to manually undo these transformations in the predictions.
-The *caret* allows you to transform before training the models and already transforms the results back.
-
-&nbsp;
-
-Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}).
-The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
-
-```{r split, echo=F, message=F, warning=F}
-# Test and train data split
-reg <- list(
-  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$mpa),
-  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$mpa),
-  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$mpa),
-  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$mpa),
-  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$mpa),
-  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$mpa)
-)
-reg_seed <- c(
-  1111, # 3
-  1, # 7
-  22, # 14
-  11111, # 28
-  111, # 56
-  11 # 100
-  )
-split_reg <- function(n){
-  set.seed(reg_seed[[n]], sample.kind="Rounding")
-  createDataPartition(reg[[n]]$y, p = .8, list = F)
-}
-trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
-gen_dat <- function(n){
-  regIndex <- trainIndex_reg[[n]]
-  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
-}
-dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
-names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
-```
-
-```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
-# Distribution of test and train data
-cap <- "Distribution of test and train data"
-d_lab <- " days"
-ylabel <- "Density"
-g1 <- "Train"
-g2 <- "Test"
-dens <- function(d, n){
-  dens <- full_join(
-      d$train %>%
-        select(y) %>%
-        mutate(Group = g1, day = str_sub(n, 2)),
-      d$test %>%
-        select(y) %>%
-        mutate(Group = g2, day = str_sub(n, 2))
-    )
-  dens
-}
-densities <- imap(dats_reg, dens) %>%
-  reduce(rbind) %>%
-  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
-labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
-levels(densities$day_f) <- labs
-densities %>%
-  ggplot(aes(y, fill = Group)) +
-  geom_density(alpha = 0.5) +
-  facet_wrap(~day_f, ncol=3) +
-  xlab("MPa") +
-  ylab(ylabel) +
-  theme_bw() +
-  theme_minimal() +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(), 
-        axis.line = element_line(colour = "black"))
-```
-
-### Performance measures
-
-The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*).
-The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.
-
-### Naive models
-
-Before creating the real models, for comparison purposes, naive models were created.
-They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}).
-In other words, naive models are simply the best guess possible.
-The results can be checked in the table \ref{tab:table-naive-model-reg}.
-
-```{r naive-model-reg, echo=F, message=F, warning=F}
-# Naive model
-res <- function(d, n){
-  data.frame(
-    day = str_sub(n, 2),
-    mean_mpa = mean(d$train$y)   
-    )
-}
-ing_model <- imap(dats_reg, res)
-get_rmse <- function(d, n){
-   data.frame(
-    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
-    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
-    day = str_sub(n, 2) 
-    )
-}
-rmses <- imap(dats_reg, get_rmse) %>%
-  reduce(rbind)
-ing_model_df <- ing_model %>% reduce(rbind)
-df_performance_reg <- full_join(ing_model_df, rmses)
-```
-
-```{r table-naive-model-reg, echo=F}
-# Table - Naive models
-colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
-caption <- "Naive models"
-kable(
-    df_performance_reg,
-    col.names = colNames,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  kable_styling(latex_options = c("HOLD_position"))
-```
-
-### Choice of algorithms
-
-The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models.
-The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result.
-According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).
-
-### Regression models
-
-As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
-
-&nbsp;
-
-1. All *features*;
-2. All *features* without the *dummy vars*;
-3. Only the original *features*;
-4. Only new *features*;
-5. Only new *features*, without *dummy vars*;
-
-&nbsp;
-
-Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
-For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set.
-The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
-
-```{r test-models-reg, echo=F, warning=F, message=F, include=F}
-# Features selection
-data1 <- dats_reg$d28$train # full
-data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
-data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
-data4 <- dats_reg$d28$train[c(8:21)] # only new variables
-data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
-test1 <- dats_reg$d28$test # full
-test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
-test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
-test4 <- dats_reg$d28$test[c(8:21)] # only new variables
-test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
-# parRF
-modelLookup("parRF")
-control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
-fit_parRF<- function(data, n) {
-  set.seed(1, sample.kind = "Rounding")
-  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
-  train(y ~ ., 
-    data = data,
-    preProcess = c("center","scale"),
-    method='parRF', 
-    tuneGrid=tunegrid_parRF,
-    trControl=control_parRF)
-}
-fit_parRF1 <- fit_parRF(data1, 3)
-ggplot(fit_parRF1)
-min(fit_parRF1$results$RMSE) # 6.26108
-p1_parRF1 <- predict(fit_parRF1, newdata = test1)
-RMSE(p1_parRF1, test1$y) # 4.887264
-fit_parRF2 <- fit_parRF(data2, 1)
-ggplot(fit_parRF2)
-min(fit_parRF2$results$RMSE) # 6.268789
-p_parRF2 <- predict(fit_parRF2, newdata = test2)
-RMSE(p_parRF2, test2$y) # 4.812994
-fit_parRF3 <- fit_parRF(data3, 1)
-ggplot(fit_parRF3)
-min(fit_parRF3$results$RMSE) # 6.361905
-p_parRF3 <- predict(fit_parRF3, newdata = test3)
-RMSE(p_parRF3, test3$y) # 5.083467
-fit_parRF4 <- fit_parRF(data4, 1)
-ggplot(fit_parRF4)
-min(fit_parRF4$results$RMSE) # 8.705984
-p_parRF4 <- predict(fit_parRF4, newdata = test4)
-RMSE(p_parRF4, test4$y) # 8.083962
-fit_parRF5 <- fit_parRF(data5, 1)
-ggplot(fit_parRF5)
-min(fit_parRF5$results$RMSE) # 8.691755
-p_parRF5 <- predict(fit_parRF5, newdata = test5)
-RMSE(p_parRF5, test5$y) # 7.991537
-```
-
-```{r table-done-samples-28, echo=F}
-# Table - First 6 samples of train data of the 28 days model
-caption <- "First 6 samples of train data of the 28 days model"
-colNames = c("Cement", "B.F.S.", "Fly ash", "water",
-              "Superp.", "Coarse Agg.", "Fine Agg.",
-              "Wat./", "F.Agg./", "C.Agg./",
-              "F.Agg./","Wat./", "Wat./", "y")
-dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
-              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","Ce.",
-              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "MPa")
-colNames2 = c("Features"=13, "Outcome"=1)
-kable(
-    head(data2),
-    col.names = dfUnits,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    digits = 2,
-    align = "c"
-    ) %>%
-    add_header_above(header = colNames, line = F, align = "c") %>%
-    add_header_above(header = colNames2, line = T, align = "c") %>%
-    kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-
-For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}).
-For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.
-
-```{r reg-models, echo=F, warning=F, message=F, include=F}
-# Regression models
-# no dummy vars:
-data3 <- dats_reg$d3$train[c(1:13, 22)]
-data7 <- dats_reg$d7$train[c(1:12, 19)]
-data14 <- dats_reg$d14$train[c(1:13, 21)]
-data28 <- dats_reg$d28$train[c(1:13, 21)]
-data56 <- dats_reg$d56$train[c(1:13, 22)]
-data100 <- dats_reg$d100$train[c(1:13, 23)]
-test3 <- dats_reg$d3$test[c(1:13, 22)]
-test7 <- dats_reg$d7$test[c(1:12, 19)]
-test14 <- dats_reg$d14$test[c(1:13, 21)]
-test28 <- dats_reg$d28$test[c(1:13, 21)]
-test56 <- dats_reg$d56$test[c(1:13, 22)]
-test100 <- dats_reg$d100$test[c(1:13, 23)]
-get_trControl <- function(n, r){
-  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
-}
-trControl3 <- get_trControl(30, 10)
-trControl7 <- get_trControl(10, 10)
-trControl14 <- get_trControl(30, 10)
-trControl28 <- get_trControl(30, 10)
-trControl56 <- get_trControl(30, 10)
-trControl100 <- get_trControl(10, 10)
-get_tuneGrid <- function(data){
-  expand.grid(mtry = seq(1, length(data), 1))
-}
-tuneGrid3 <- get_tuneGrid(data3)
-tuneGrid7 <- get_tuneGrid(data7)
-tuneGrid14 <- get_tuneGrid(data14)
-tuneGrid28 <- get_tuneGrid(data28)
-tuneGrid56 <- get_tuneGrid(data56)
-tuneGrid100 <- get_tuneGrid(data100)
-set.seed(1, sample.kind = "Rounding")
-fit_3 <- train(y ~ ., 
-            data = data3,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid3,
-            trControl=trControl3)
-p_3 <- predict(fit_3, newdata = test3)
-RMSE_test_3 <- RMSE(p_3, test3$y)
-RMSE_test_3 # 3.31037
-fit_3$bestTune # mtry = 6
-set.seed(1, sample.kind = "Rounding")
-fit_7 <- train(y ~ ., 
-            data = data7,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid7,
-            trControl=trControl7)
-p_7 <- predict(fit_7, newdata = test7)
-RMSE_test_7 <- RMSE(p_7, test7$y)
-RMSE_test_7 # 4.361987
-fit_7$bestTune # mtry = 2
-set.seed(1, sample.kind = "Rounding")
-fit_14 <- train(y ~ ., 
-            data = data14,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid14,
-            trControl=trControl14)
-p_14 <- predict(fit_14, newdata = test14)
-RMSE_test_14 <- RMSE(p_14, test14$y)
-RMSE_test_14 # 4.620515
-fit_14$bestTune # mtry = 13
-set.seed(1, sample.kind = "Rounding")
-fit_28 <- train(y ~ ., 
-            data = data28,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid28,
-            trControl=trControl28)
-p_28 <- predict(fit_28, newdata = test28)
-RMSE_test_28 <- RMSE(p_28, test28$y)
-RMSE_test_28 # 4.716698
-fit_28$bestTune # mtry = 11
-set.seed(1, sample.kind = "Rounding")
-fit_56 <- train(y ~ ., 
-            data = data56,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid56,
-            trControl=trControl56)
-p_56 <- predict(fit_56, newdata = test56)
-RMSE_test_56 <- RMSE(p_56, test56$y)
-RMSE_test_56 # 5.939163
-fit_56$bestTune # mtry = 8
-set.seed(1, sample.kind = "Rounding")
-fit_100 <- train(y ~ ., 
-            data = data100,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid100,
-            trControl=trControl100)
-p_100 <- predict(fit_100, newdata = test100)
-RMSE_test_100 <- RMSE(p_100, test100$y)
-RMSE_test_100 # 5.851088
-fit_100$bestTune # mtry = 8
-```
-
-# Results
-
-The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively.
-The table \ref{tab:table-reg-models} presents the details and results of each model (\ref{show-table-reg-models}).
-The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).
-
-```{r table-reg-models, echo=F}
-# Table - Models details
-colNames <- c("Model", "mtry", "CV", "Repetitions", 
-              "RMSE (train)", "RMSE (test)")
-caption <- "Regression models results"
-day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
-dat_reg_models <- data.frame(
-  dia = day,
-  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
-           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
-  number = c(trControl3$number, trControl7$number, trControl14$number,
-             trControl28$number,trControl56$number,trControl100$number),
-  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
-             trControl28$repeats,trControl56$repeats,trControl100$repeats),
-  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
-                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
-                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
-  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
-                RMSE_test_28, RMSE_test_56, RMSE_test_100)
-)
-kable(
-    dat_reg_models,
-    col.names = colNames,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  kable_styling(latex_options = c("HOLD_position"))
-```
-
-```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
-# Figure - Models comparison
-cap <- "Actual vs Predicted values in each model"
-models <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
-xlabel <- "Actual (MPa)"
-ylabel <- "Predicted (MPa)"
-preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
-actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
-gen_res_df <- function(ind){
-  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
-}
-res_df <- lapply(1:6, gen_res_df)
-res_df <- bind_rows(res_df)
-res_df$model <- factor(res_df$model,levels=models)
-res_df %>%
-  ggplot(aes(actual, pred)) +
-  facet_wrap(~ model, ncol=3) +
-  geom_point(alpha=0.5) +
-  theme_bw() +
-  geom_abline(slope=1, intercept=0) +
-  xlab(xlabel) +
-  ylab(ylabel)
-```
-
-```{r table-10, echo=F}
-# Results tables of 10 best and worst results
-colNames1 = c("Actual", "Predicted", "Error", "","Actual", "Predicted", "Error") 
-colNames2 = c("Best 10"=3,"", "Worst 10"=3) 
-caption_0 <- "Model of "
-get_X <- function(pred, actual, X=10){
-  diff <- abs(pred - actual)
-  diff_2 <- pred - actual
-  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
-  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
-  df_min <- data.frame(
-    actual_min=actual[ind_min],
-    pred_min=pred[ind_min],
-    diff_min=diff[ind_min],
-    diff_min_2=diff_2[ind_min]
-  )
-  df_max <- data.frame(
-    actual_max=actual[ind_max],
-    pred_max=pred[ind_max],
-    diff_max=diff[ind_max],
-    diff_max_2=diff_2[ind_max]
-  )
-  df_max <- df_max[order(-df_max$diff_max),]
-  df_min <- df_min[order(df_min$diff_min),]
-  df_null <- data.frame(null_col = rep(c(""), X))
-  res <- cbind(df_min,df_null, df_max)
-  res <- res %>% select(-c(diff_max, diff_min))
-  rownames(res) <- c()
-  res
-}
-gen_kable <- function(ind){
-  df <- get_X(preds[[ind]], actuals[[ind]])
-  caption <- paste0(caption_0,models[ind])
-  kable(
-    df,
-    col.names = colNames1,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  column_spec(4, width = "1cm",) %>%
-  add_header_above(header = colNames2, line = T, align = "c") %>%
-  kable_styling(latex_options = c("HOLD_position"))
-}
-gen_kable(1)
-gen_kable(2)
-gen_kable(3)
-gen_kable(4)
-gen_kable(5)
-gen_kable(6)
-```
-
-# Discussion and conclusion
-
-The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.
-The alternative adopted to create a model for each set of age proved to be a valid method, managing to stratify to obtain specific results for each set.
-The studies cited in the introduction using the same dataset have similar results, as expected.
-The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
-
-```{r works-comparison, echo=F}
-# Table - Comparison of other works
-colNames = c("Author","Year" ,"Algorithm", "RMSE") 
-caption <- "Comparison of other works"
-works <- data.frame(
-  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
-  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
-  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
-          "Gradient Boosting Regressor","Random Forest Regressor" ,
-          "Regularized Extreme Learning Machine",
-          "Support Vector Machines with Radial Basis Function Kernel"),
-  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
-)
-kable(
-  works,
-  col.names = colNames,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "l"
-  )  %>%
-    kable_styling(latex_options = c("HOLD_position"))
-```
-
-```{r results, echo=F}
-# Table - Final results
-colNames <- c("Model", "RMSE")
-caption <- "Final result"
-day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
-dat_reg_models <- data.frame(
-  dia = day,
-  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
-                RMSE_test_28, RMSE_test_56, RMSE_test_100)
-)
-kable(
-    dat_reg_models,
-    col.names = colNames,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  kable_styling(latex_options = c("HOLD_position"))
-```
-
-Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results.
-Another option is to create an *ensemble* of various algorithms, just like @Pierobon2018, but with the separation of age sets proposed here.
-In addition, it can be performed with a larger dataset, ideally with the same number of samples in each age set, a more homogeneous distribution of compressive strength, and less variance between samples.
-
 # Literature cited
@@
 <div id="refs"></div>
-
-\newpage
-
-# Appendix 1 - Virtual environment {#appendix1}
-
-## Operational system
-
-```{r system, echo=FALSE}
-v <- as.matrix(R.Version())
-kable(v, booktabs = T) %>%
-    kable_styling(latex_options = c("HOLD_position"))
-```
-
-## Packages
-
-```{r packages, echo=FALSE}
-p <- matrix(
-      c(
-        "caret", as.character(packageVersion("caret")),
-        "cowplot",as.character(packageVersion("cowplot")),
-        "dplyr", as.character(packageVersion("dplyr")),
-        "factoextra", as.character(packageVersion("factoextra")),
-        "gdata", as.character(packageVersion("gdata")),
-        "ggplot2", as.character(packageVersion("ggplot2")),
-        "gridExtra",as.character(packageVersion("gridExtra")),
-        "kableExtra",as.character(packageVersion("kableExtra")),
-        "knitr",as.character(packageVersion("knitr")),
-        "pastecs",as.character(packageVersion("pastecs")),
-        "purrr", as.character(packageVersion("purrr")),
-        "questionr", as.character(packageVersion("questionr")),
-        "reshape2", as.character(packageVersion("reshape2")),
-        "tidyr", as.character(packageVersion("tidyr")),
-        "tidyverse", as.character(packageVersion("tidyverse"))
-      ),
-      ncol = 2,
-      byrow=TRUE
-    )
-kable(p, booktabs = T) %>%
-    kable_styling(latex_options = c("HOLD_position"))
-```
-\newpage
-
-# Appendix 2 - Code {#appendix2}
-
-## Obtaining the data
-### Data download
-
-```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
-```
-
-### Renaming the columns
-
-```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
-```
-
-### Reordering the data
-
-```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
-```
-
-### Defining column names and units
-
-```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
-```
-
-### Table - First samples
-
-```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
-```
-
-## Data preparation
-### Removing duplicated samples
-
-```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
-```
-
-### Table - Samples with same composition
-
-```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
-```
-
-### Table - Same samples with different results
-
-```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
-```
-
-### Data cleaning
-
-```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
-```
-
-### Table - Previous samples after processing
-
-```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
-```
-
-### Figure - Compressive strength (MPa) vs age (days)
-
-```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
-```
-
-### Figure - Principal component analysis - 90, 91 e 100 days
-
-```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
-```
-
-### Figure - Compressive strength through time
-
-```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
-```
-
-### Joining 90, 91 and 100 days data
-
-```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
-```
-
-### Figure - Ages frequency
-
-```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
-```
-
-### Removing ages with frequency lower than 50
-
-```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
-```
-
-### Data reorganization
-
-```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
-```
-
-### Table - First 6 samples after reorganization
-
-```{r show-new-features, ref.label='new-features', ref='show-new-features', eval=FALSE}
-```
-
-### Total samples
-
-```{r show-total-samples-2, ref.label='total-samples-2', ref='show-total-samples-2', eval=FALSE}
-```
-
-### Adding New features
-
-```{r show-new-features-2, ref.label='new-features-2', ref='show-new-features-2', eval=FALSE}
-```
-
-### Table - New features
-
-```{r show-new-features-table, ref.label='new-features-table', ref='show-new-features-table', eval=FALSE}
-```
-
-## Data visualization
-### Table - Descriptive statistics - continuous variables
-
-```{r show-stat-summ, ref.label='stat-summ', ref='show-stat-summ', eval=FALSE}
-```
-
-### Figure - Descriptive statistics - categorical variables
-
-```{r show-stat-summ-categorical, ref.label='stat-summ-categorical', ref='show-stat-summ-categorical', eval=FALSE}
-```
-
-### Figure - Correlation grouped by age
-
-```{r show-correlation, ref.label='correlation', ref='show-correlation', eval=FALSE}
-```
-
-### Figure - Correlation over time
-
-```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
-```
-
-### Figure - Relationship between approximated mix, water, MPa and age
-
-```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
-```
-
-### Figure - Relationship between concrete main features
-
-```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
-```
-
-### Figure - Variables distribution
-
-```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
-```
-
-### Figure - Variables distribution grouped by age
-
-```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
-```
-
-### Figure - Principal component analysis on ingredients
-
-```{r show-pca, ref.label='pca', ref='show-pca', eval=FALSE}
-```
-
-## Machine learning models
-### Dummy variables
-
-```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
-```
-
-### Data preparation
-
-```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
-```
-
-### Table - First 18 columns of the 6 first samples of 28 days
-
-```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
-```
-
-### Removing near zero variance columns
-
-```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
-```
-
-### High correlation variables verification
-
-```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
-```
-
-### Test and train data split
-
-```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
-```
-
-### Distribution of test and train data
-
-```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
-```
-
-### Naive model
-
-```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
-```
-
-### Table - Naive models
-
-```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
-```
-
-### Features selection
-
-```{r show-test-models-reg, ref.label='test-models-reg', ref='show-test-models-reg', eval=FALSE}
-```
-
-### Table - First 6 samples of train data of the 28 days model
-
-```{r show-table-done-samples-28, ref.label='table-done-samples-28', ref='show-table-done-samples-28', eval=FALSE}
-```
-
-### Regression models
-
-```{r show-reg-models, ref.label='reg-models', ref='show-reg-models', eval=FALSE}
-```
-
-## Results
-### Table - Models details
-
-```{r show-table-reg-models, ref.label='table-reg-models', ref='show-table-reg-models', eval=FALSE}
-```
-
-### Figure - Models comparison
-
-```{r show-results-comparison, ref.label='results-comparison', ref='show-results-comparison', eval=FALSE}
-```
-
-### Results tables of 10 best and worst results
-
-```{r show-table-10, ref.label='table-10', ref='show-table-10', eval=FALSE}
-```
-
-## Discussion and conclusion
-### Table - Comparison of other works
-
-```{r show-works-comparison, ref.label='works-comparison', ref='show-works-comparison', eval=FALSE}
-```
-
-### Table - Final results
-
-```{r show-results, ref.label='results', ref='show-results', eval=FALSE}
-```

### 2020-07-13 HarvardX_Data_Science_Capstone/CopyOfcapstone.tex
Harvard submission

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Concrete compressive strength prediction with machine learning},
            pdfauthor={Pedro Bernardino Alves Moreira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{supertabular}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage[normalem]{ulem}
\usepackage{caption}
\usepackage{floatrow}
\floatsetup[table]{capposition=top}
\floatsetup[figure]{capposition=top}
\captionsetup{options=chunk}
\DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
\renewcommand{\thechunk}{\arabic{chunk}}
\usepackage{indentfirst}
\usepackage{sectsty} \sectionfont{\centering}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Concrete compressive strength prediction with machine learning}
\author{Pedro Bernardino Alves Moreira}
\date{April 22, 2020}

\begin{document}
\maketitle
\begin{abstract}
Compressive strength is the main characteristic of concrete. The correct
prediction of this parameter means cost and time reduction. This work
built predictive models for 6 different ages of concrete samples (3, 7,
14, 28, 56, and 100 days). A set of data obtained in previous studies
was used, a total of 1030 samples, with 9 variables: compressive
strength, age, and 7 ingredients (water, cement, fine aggregate, coarse
aggregate, fly ash, blast furnace slag, and superplasticizers). Another
6 variables were added to represent the proportions of the main
ingredients in each sample (water/cement, fine aggregate/cement, coarse
aggregate/cement, fine aggregate/coarse aggregate, water/coarse
aggregate, and water/fine aggregate). The predictive models were
developed in \emph{R} language, using the \emph{caret} package with the
\emph{Parallel Random Forest} algorithm and repeated cross-validation
technique to optimize the parameters. The results were satisfactory and
compatible with other studies using the same data set. The most
important model, 28 days old, obtained \emph{RMSE} of 4.717. The 3-day
model obtained the best result, \emph{RMSE} of 3.310. The worst result
was the 56-day model, with \emph{RMSE} of 5.939. The work showed that
the compressive strength of concrete can be predicted. The choice of
creating a model for each age, instead of using age as a predictor,
allowed to get compatible results with the available data at each age.
It was a promising alternative since good results were achieved by
training with just one algorithm. This work facilitates exploration and
new efforts to predict the compressive strength of concrete, it can be
replicated using different algorithms or the combination of several.
\end{abstract}

\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Compressive strength is the main characteristic of concrete, measured by
tests of international standards that consist of the breaking of
specimens. Measurement at 28 days is mandatory and represents the grade
of the concrete. Knowing in advance what the result will be obtained for
a given age, based on the proportions of its ingredients, is of great
interest to concrete manufacturers, construction companies, and civil
engineers.

~

This compressive strength is a nonlinear function of its ingredients and
age, making it difficult to establish an analytical formula, although
some formulas have already been proposed. ({\textbf{???}}) proposed a
mathematical model to predict from the results of tests of 7 and 14
days, and ({\textbf{???}}) from 7 days. However, machine learning
techniques can be used to model this characteristic from real sample
data, using only the ingredients.

~

Many previous studies use the same dataset used by Yeh (n.d.) to predict
the compressive strength of concrete. ({\textbf{???}}) got good results
with the regularized extreme learning machine (RELM) technique, and
({\textbf{???}}) got even better results with the Artificial Neural
Networks and cross-validation technique. This set of samples is so well
known that there are many pages on the internet of unpublished studies
that use it and have good results, such as ({\textbf{???}}),
({\textbf{???}}), ({\textbf{???}}) and ({\textbf{???}}). At the end of
the work, the results found are compared to the works cited here.

~

Unlike previous studies with this dataset, this work does data
preparation differently. The age of the concrete is the most unique
feature that contributes to its compressive strength. For this reason,
age is treated separately in the machine learning models, creating
models for each age group.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

The methodology was carried out using RStudio software ({\textbf{???}}),
an integrated virtual environment for code development in \emph{R}
({\textbf{???}}). Throughout the process, all code executed was
documented in the same order as its execution in
\protect\hyperlink{appendix2}{Appendix 2}, and reference was always made
to codes throughout the text. All relevant information related to the
operating system and installed packages has been presented in
\protect\hyperlink{appendix1}{Appendix 1}.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

In order to guarantee reproducibility, whenever there was some code that
could use probabilistic operations, a \emph{seed} was defined before its
execution, ensuring that when run on another machine, with the same
version of \emph{R} and the same \emph{seed}, get the same result. The
\emph{seeds} can be checked throughout
\protect\hyperlink{appendix2}{Appendix 2}.

\hypertarget{obtaining-the-data}{%
\subsection{Obtaining the data}\label{obtaining-the-data}}

The data was downloaded from the University of California Irvine website
({\textbf{???}}) (\ref{show-download-data}). In total there are 1030
samples with 9 columns. The samples were renamed and an id column was
added to facilitate data manipulation (\ref{show-rename-dat-cols}). The
columns were reordered to put the new id column in the first position
(\ref{show-reorder-dat}). The first samples can be viewed in the table
\ref{tab:first-samples}.

~

\begin{table}[H]

\caption{\label{tab:first-samples}First 6 samples}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 28 & 79.99\\
\addlinespace
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 28 & 61.89\\
\addlinespace
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 270 & 40.27\\
\addlinespace
4 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 365 & 41.05\\
\addlinespace
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 360 & 44.30\\
\addlinespace
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 90 & 47.03\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{data-preparation}{%
\subsection{Data preparation}\label{data-preparation}}

The preparation of the data consisted of transforming the sample set in
order to maintain only relevant data for the subsequent studies. Data
that were considered irrelevant or that had the potential to add
undesirable noise to the analysis were removed. In addition, the
relevant data has been transformed to better fit the studies in the next
steps.

\hypertarget{initial-data-cleaning}{%
\subsubsection{Initial data cleaning}\label{initial-data-cleaning}}

Initially, there were 25 duplicate samples that were removed, resulting
in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

~

The data show the variables in the columns and samples in the rows.
However it was found that some samples are identical in proportions of
ingredients, changing only the value of age and compressive strength,
for example, samples 653, 654, 678 and 681, shown in the table
\ref{tab:similar-samples}.

~

\begin{table}[H]

\caption{\label{tab:similar-samples}Samples with same composition}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
653 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 3 & 4.57\\
\addlinespace
678 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 7 & 7.68\\
\addlinespace
681 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 28 & 17.28\\
\addlinespace
654 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 90 & 25.46\\
\bottomrule
\end{tabular}}
\end{table}

In addition, there are also samples with the same values and proportions
of ingredients, but with different compressive strength, probably due to
differences in the building process. This is the case, for example, of
samples 472, 473 and 474, shown in the table
\ref{tab:similar-samples-2}.

~

\begin{table}

\caption{\label{tab:similar-samples-2}Same samples with different results}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 57.03\\
\addlinespace
473 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 44.42\\
\addlinespace
474 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 51.02\\
\bottomrule
\end{tabular}}
\end{table}

To facilitate the analysis of the samples, all samples that are the same
in relation to the ingredients, have been assigned the same \emph{id}.
In addition, as the compressive strength at 28 days is the parameter to
determine the grade of the concrete, only the elements containing that
day among its samples were maintained. In the case of the same samples
but with different results of compressive strength, the values were
averaged. After all these changes (\ref{show-initial-data-cleaning}),
the new total of samples was reduced to 970, containing 416 different
settings for the proportions of ingredients.

~

The result can be seen in the table \ref{tab:similar-samples-same-id}.
All samples with equal ingredient settings have the same id, and when
they had different results for the same days, they were transformed into
just one sample, with the arithmetic mean in the compressive strength.

~

\begin{table}[H]

\caption{\label{tab:similar-samples-same-id}Previous samples after processing}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 50.82\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 3 & 4.57\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 7 & 7.68\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 28 & 17.28\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 90 & 25.46\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{age-selection}{%
\subsubsection{Age selection}\label{age-selection}}

As previously described, the main age for analysis of compressive
strength is 28 days, but other ages can also be used to build predictive
models. However, it is necessary to verify how relevant the data of
these other ages are. Starting with the distribution of the samples in
relation to each age (\ref{show-boxplot}) shown in the figure
\ref{fig:boxplot}.

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/boxplot-1}

}

\caption{Boxplot - Compressive strength (MPa) vs age (days)}\label{fig:boxplot}
\end{figure}

It was observed that the ages of 90, 91 and 100 days probably represent
extremes to each other in the ingredient configurations, since they are
relatively close ages but with very different values, especially for 90
and 91.

~

This hypothesis was verified using the principal component analysis
method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}).
The figure \ref{fig:pca-90-91-100} shows how the samples relate to each
other (which are similar or different) and revealed how each variable
contributes to the analysis. The first two dimensions represent 37\% and
24\% \$ respectively of the variance.

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/pca-90-91-100-1}

}

\caption{Principal component analysis - 90, 91 e 100 days}\label{fig:pca-90-91-100}
\end{figure}

Another important point considered, of the concrete nature itself, is
the fact that the growth rate of its compressive strength decreases with
time, reaching a certain stability value. The figure
\ref{fig:mpa-on-time} shows the compressive strength over the days for
samples with more than 5 data, that is, data available for at least 6
different ages (\ref{show-mpa-on-time}).

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/mpa-on-time-1}

}

\caption{Compressive strength through time}\label{fig:mpa-on-time}
\end{figure}

For the reasons presented in the figures \ref{fig:boxplot},
\ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered
that the ages of 90, 91 and 100 days can be grouped to improve reading
and decrease sample noise. They were converted to the same value, the
age of 100 days was chosen (\ref{show-join-90-91-100}). As shown in the
figure \ref{fig:mpa-on-time}, the resistance only increases, after 100
days the resistance to compression will be greater than or equal to the
value of 90 or 91 days.

~

Another topic analyzed in the selection of ages was the observed
frequency of each age value after this transformation from 90, 91 in 100
days, shown in the figure \ref{fig:freq-ages}. Some values of days have
very low concentrations of samples, at the risk of damaging more than
helping to create the models, so they were removed
(\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain
only ages with a frequency greater than 50, only the values of 3, 7, 14,
28, 56 and 100 days.

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/freq-ages-1}

}

\caption{Ages frequency}\label{fig:freq-ages}
\end{figure}

\hypertarget{data-reorganization}{%
\subsubsection{Data reorganization}\label{data-reorganization}}

The samples were grouped to maintain only one sample from each set of
configuration of the proportions of the ingredients, adding new
variables/columns for resistance at each age
(\ref{show-reorganizing-dat}). The result in the first samples after
this processing is shown in the table \ref{tab:new-features}.

~

\begin{table}[H]

\caption{\label{tab:new-features}First 6 samples after reorganization}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Ag.} & \multicolumn{1}{c}{Fine Ag.} & \multicolumn{1}{c}{3 days} & \multicolumn{1}{c}{7 days} & \multicolumn{1}{c}{14 days} & \multicolumn{1}{c}{28 days} & \multicolumn{1}{c}{56 days} & \multicolumn{1}{c}{100 days} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 &  &  &  & 79.99 &  & \\
\addlinespace
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 &  &  &  & 61.89 &  & \\
\addlinespace
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 &  & 30.28 &  & 33.02 &  & 37.72\\
\addlinespace
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 9.13 & 14.64 &  & 28.02 &  & 38.07\\
\addlinespace
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 &  &  &  & 45.85 &  & 47.03\\
\addlinespace
7 & 380.0 & 95.0 & 0 & 228 & 0.0 & 932.0 & 594.0 &  & 32.82 &  & 36.45 &  & 40.56\\
\bottomrule
\end{tabular}}
\end{table}

The number of samples and distinct samples after all this manipulation
remained the same, a total of 416 (\ref{show-total-samples-2}).

\hypertarget{adding-new-variables}{%
\subsubsection{Adding new variables}\label{adding-new-variables}}

To finish the data preparation, new columns were added to the dataset
(\ref{show-new-features-2}). Starting with the concrete class, for
example if the compressive strength is between 25 and 30, it receives
the class \emph{C25}. The inclusion of the class was important because
the compressive strength in \emph{MPa} is a continuous variable, which
will be used in the regression models, but the class as a discrete
variable can provide another visualization of the data. The approximate
mix of concrete was also added, which represents the proportions of
aggregates (fine and coarse) for cement. Other proportions between the
main ingredients were also added. The new variables are presented in the
table \ref{tab:new-features-table}.

\begin{table}[H]

\caption{\label{tab:new-features-table}New features}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{c>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}}
\toprule
ID & Class & Approximated Mix & Water / Cement & Fine Ag. / Cement & Coarse Ag. / Cement & Fine Ag. / Coarse Ag. & Water / Coarse Ag. & Water / Fine Ag.\\
\midrule
1 & C75 & 1:1:2 & 0.3000 & 1.2519 & 1.9259 & 0.6500 & 0.1558 & 0.2396\\
\addlinespace
2 & C60 & 1:1:2 & 0.3000 & 1.2519 & 1.9537 & 0.6408 & 0.1536 & 0.2396\\
\addlinespace
3 & C30 & 1:2:3 & 0.6857 & 1.7865 & 2.8030 & 0.6373 & 0.2446 & 0.3838\\
\addlinespace
5 & C25 & 1:4:5 & 0.9668 & 4.1566 & 4.9265 & 0.8437 & 0.1962 & 0.2326\\
\addlinespace
6 & C45 & 1:3:4 & 0.8571 & 2.5188 & 3.5038 & 0.7189 & 0.2446 & 0.3403\\
\addlinespace
7 & C35 & 1:2:2 & 0.6000 & 1.5632 & 2.4526 & 0.6373 & 0.2446 & 0.3838\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{data-visualization}{%
\subsection{Data visualization}\label{data-visualization}}

In order to assess the need for further manipulation before building the
models, in this step the 416 samples already processed were visualized
and analyzed.

\hypertarget{descriptive-statistics}{%
\subsubsection{Descriptive statistics}\label{descriptive-statistics}}

The table \ref{tab:stat-summ} presents the statistical data of the
continuous variables (\ref {show-stat-summ}). The \emph{Null} line
represents the number of zeroed values for the ingredients, and the
\emph{NA} line represents the number of missing data. As the samples
were filtered to maintain only sets of samples with known values of
compressive strength at 28 days, the number of \emph{NAs} is zero for
that age. The figure \ref{fig:stat-summ-categorical} presents the
statistical data of the discrete variables
(\ref{show-stat-summ-categorical}).

~

\begin{table}[H]

\caption{\label{tab:stat-summ}Descriptive statistics - continuous variables}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{1.5cm}cccccccc>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
  & Samples & Null & NA & Min & Max & Range & Sum & Median & Mean & SE mean & CI mean & Variance & Std.Dev. & Coef.Var\\
\midrule
Cement & 416 & 0 & 0 & 102.00 & 540.00 & 438.00 & 109373.10 & 257.70 & 262.92 & 5.10 & 10.02 & 10817.50 & 104.01 & 0.40\\
\addlinespace
B.F.S. & 416 & 174 & 0 & 0.00 & 359.40 & 359.40 & 35824.60 & 94.25 & 86.12 & 4.32 & 8.49 & 7755.00 & 88.06 & 1.02\\
\addlinespace
Fly ash & 416 & 202 & 0 & 0.00 & 200.10 & 200.10 & 26389.00 & 71.25 & 63.44 & 3.26 & 6.40 & 4407.81 & 66.39 & 1.05\\
\addlinespace
Water & 416 & 0 & 0 & 121.80 & 247.00 & 125.20 & 76335.60 & 185.00 & 183.50 & 0.94 & 1.86 & 370.73 & 19.25 & 0.10\\
\addlinespace
Superplast. & 416 & 107 & 0 & 0.00 & 32.20 & 32.20 & 2871.30 & 7.60 & 6.90 & 0.26 & 0.52 & 28.85 & 5.37 & 0.78\\
\addlinespace
Coarse agg. & 416 & 0 & 0 & 801.00 & 1145.00 & 344.00 & 397799.90 & 953.35 & 956.25 & 4.12 & 8.10 & 7063.06 & 84.04 & 0.09\\
\addlinespace
Fine agg. & 416 & 0 & 0 & 594.00 & 992.60 & 398.60 & 317809.80 & 769.65 & 763.97 & 3.59 & 7.06 & 5371.89 & 73.29 & 0.10\\
\addlinespace
3 days & 121 & 0 & 295 & 2.33 & 41.64 & 39.31 & 2210.82 & 15.52 & 18.27 & 0.87 & 1.72 & 91.64 & 9.57 & 0.52\\
\addlinespace
7 days & 114 & 0 & 302 & 7.51 & 59.09 & 51.58 & 2845.52 & 21.06 & 24.96 & 1.29 & 2.55 & 188.81 & 13.74 & 0.55\\
\addlinespace
14 days & 62 & 0 & 354 & 12.84 & 59.76 & 46.92 & 1782.56 & 26.54 & 28.75 & 1.10 & 2.19 & 74.62 & 8.64 & 0.30\\
\addlinespace
28 days & 416 & 0 & 0 & 8.54 & 81.75 & 73.21 & 15101.13 & 33.72 & 36.30 & 0.70 & 1.38 & 206.30 & 14.36 & 0.40\\
\addlinespace
56 days & 83 & 0 & 333 & 23.25 & 80.20 & 56.95 & 4178.77 & 50.77 & 50.35 & 1.52 & 3.02 & 190.82 & 13.81 & 0.27\\
\addlinespace
100 days & 120 & 0 & 296 & 21.86 & 82.60 & 60.74 & 5701.90 & 45.61 & 47.52 & 1.17 & 2.31 & 163.40 & 12.78 & 0.27\\
\addlinespace
Water / Cement & 416 & 0 & 0 & 0.27 & 1.88 & 1.62 & 340.60 & 0.73 & 0.82 & 0.02 & 0.03 & 0.11 & 0.34 & 0.41\\
\addlinespace
Fine agg. / Cement & 416 & 0 & 0 & 1.14 & 9.24 & 8.10 & 1415.82 & 2.94 & 3.40 & 0.07 & 0.14 & 1.96 & 1.40 & 0.41\\
\addlinespace
Coarse agg. / Cement & 416 & 0 & 0 & 1.55 & 8.70 & 7.14 & 1761.06 & 3.67 & 4.23 & 0.08 & 0.16 & 2.70 & 1.64 & 0.39\\
\addlinespace
Fine agg. / Coarse agg. & 416 & 0 & 0 & 0.53 & 1.16 & 0.63 & 335.42 & 0.80 & 0.81 & 0.01 & 0.01 & 0.01 & 0.11 & 0.14\\
\addlinespace
Water / Coarse agg. & 416 & 0 & 0 & 0.12 & 0.29 & 0.17 & 80.66 & 0.19 & 0.19 & 0.00 & 0.00 & 0.00 & 0.03 & 0.16\\
\addlinespace
Water / Fine agg. & 416 & 0 & 0 & 0.13 & 0.38 & 0.26 & 101.26 & 0.24 & 0.24 & 0.00 & 0.00 & 0.00 & 0.04 & 0.17\\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/stat-summ-categorical-1}

}

\caption{Descriptive statistics - categorical variables}\label{fig:stat-summ-categorical}
\end{figure}

\hypertarget{correlation-between-ingredients-and-compressive-strength}{%
\subsubsection{Correlation between ingredients and compressive
strength}\label{correlation-between-ingredients-and-compressive-strength}}

The figure \ref{fig:correlation} shows the correlation of variables for
each set of ages (\ref{show-correlation}). The figure
\ref{fig:correlation-mpa} presents the same data, but instead of
correlating them all, it only correlates with the compressive strength,
showing the values in more detail (\ref{show-correlation-mpa}).

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/correlation-1}

}

\caption{Correlations at each age}\label{fig:correlation}
\end{figure}

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/correlation-mpa-1}

}

\caption{Correlation of variables with compressive strength over time}\label{fig:correlation-mpa}
\end{figure}

The interpretation of the figure \ref{fig:correlation-mpa} suggests that
the strength of the concrete is positively related mainly to the cement
and superplasticizer ingredients and negatively to the water and fine
aggregate. The smaller the amount of cement for aggregates and water,
the more negatively they are correlated with compressive strength.

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/mix-app-mpa-1}

}

\caption{Relationship between approximated mix, water, MPa and age}\label{fig:mix-app-mpa}
\end{figure}

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/mix-mpa-1}

}

\caption{Relationship between concrete main features}\label{fig:mix-mpa}
\end{figure}

The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the
relationship between the main ingredients (known as mix) in relation to
the compressive strength (\ref{show-mix-app-mpa} and
\ref{show-mix-mpa}). The interpretation of these figures shows that the
greater the amount of cement in relation to the other ingredients, the
greater the resistance to compression.

\hypertarget{variables-distribution}{%
\subsubsection{Variables distribution}\label{variables-distribution}}

The figure\ref{fig:vars-distribution} shows the distribution of
variables in the samples (\ref{show-vars-distribution}). It was
calculated using data only at 28 days.

~

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/vars-distribution-1}

}

\caption{Variables distribution}\label{fig:vars-distribution}
\end{figure}

The figure \ref{fig:vars-distribution-time} shows the distribution of
ingredients and compressive strength for each set of ages
(\ref{show-vars-distribution-time}), in case of the 28 days it presents
the same information as the figure \ref{fig:vars-distribution}. It shows
that the resistance to compression gradually increases over time, as
expected. Furthermore it is seen that the concentration of the
ingredients can vary a lot when stratified by ages.

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/vars-distribution-time-1}

}

\caption{Variables distribution grouped by age}\label{fig:vars-distribution-time}
\end{figure}

\hypertarget{principal-component-analysis}{%
\subsubsection{Principal component
analysis}\label{principal-component-analysis}}

In the figure \ref{fig:pca}, using an alternative classification, the
principal component analysis was performed on the ingredients
(\ref{show-pca}). The classification separates concrete into 4 different
compressive strength groups, low up to \emph{20 MPa}, normal up to
\emph{40 MPa}, medium up to \emph{70 MPa} and high above that. It is
possible to notice that the groups overlap, but there is a
differentiation between the high and low group.

\begin{figure}

{\centering \includegraphics{CopyOfcapstone_files/figure-latex/pca-1}

}

\caption{Principal component analysis on ingredients}\label{fig:pca}
\end{figure}

\hypertarget{literature-cited}{%
\section{Literature cited}\label{literature-cited}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Yeh1998}{}%
Yeh, I-Cheng. n.d. ``Modeling of Strength of High-Performance Concrete
Using Artificial Neural Networks.''
\url{https://doi.org/10.1016/S0008-8846(98)00165-3}.

\end{document}

### 2020-07-13 generated/paper_PT_latex_template.Rmd
Update abstract

--- 2020-05-08 generated/paper_PT.Rmd
@@
 ```
+
 \renewcommand{\figurename}{Figura}
@@
 
-### Escolha do algorítimo
-O pacote *caret* [@caret] expõe mais de 200 algorítimos diferentes para criar modelos de *machine learning*.
-A documentação do pacote apresenta um código inicial [@modelsClusters] como sugestão para selecionar um portifólio de algóritmos mais distintos possíveis em relação à algum algorítimo pré-selecionado, mas para agilidade e devido a limitações técnicas, foi escolhido utilizar um algorítimo com a maior probabilidade de atingir o melhor resultado possível.
-Segundo @Fernandez2014, que comparou 179 algorítimos em 121 banco de dados diferentes, o algorítimo mais provável de atingir os melhores resultados possíveis é o *Parallel Random Forest* (denominado *prRF* no *caret*).
-
-### Modelos de regressão
-Como ao longo do processamento foram adicionadas novas variáveis (as relações entre os ingredientes e mais algumas *dummy vars* para cada conjunto de idade), foram estudadas 5 possibilidades de configurações das *features* para os modelos:
-
-&nbsp;
-
-1. Todas as *features*;
-2. Sem as *dummy vars*;
-3. Apenas *features* originais;
-4. Apenas *features* adicionadas;
-5. Apenas *features* adicionadas, sem as *dummy vars*;
-
-&nbsp;
-
-Construindo um modelo para cada uma dessas configurações utilizando o conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou que a melhor opção é a configuração 2, ou seja, as *dummy vars* foram completamente descartadas, porém foram mantidas as outras novas varáveis.
-Para fins ilustrativos, a tabela \ref{tab:table-done-samples-28} mostra as primeiras 6 amostras do conjunto de treino do modelo de 28 dias.
-As amostras dos outros modelos, dos conjuntos de teste e treino são similares, sendo a única diferença no modelo de 7 dias, que exlcui a cinza volante devido a variância próxima a zero, realizado anteriormente.
-```{r test-models-reg, echo=F, warning=F, message=F, include=F}
-# Escolha das caracterísitcas (features)
-data1 <- dats_reg$d28$train # full
-data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
-data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
-data4 <- dats_reg$d28$train[c(8:21)] # only new variables
-data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
-test1 <- dats_reg$d28$test # full
-test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
-test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
-test4 <- dats_reg$d28$test[c(8:21)] # only new variables
-test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
-# parRF
-modelLookup("parRF")
-control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
-fit_parRF<- function(data, n) {
-  set.seed(1, sample.kind = "Rounding")
-  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
-  train(y ~ ., 
-    data = data,
-    preProcess = c("center","scale"),
-    method='parRF', 
-    tuneGrid=tunegrid_parRF,
-    trControl=control_parRF)
-}
-fit_parRF1 <- fit_parRF(data1, 3)
-ggplot(fit_parRF1)
-min(fit_parRF1$results$RMSE) # 6.26108
-p1_parRF1 <- predict(fit_parRF1, newdata = test1)
-RMSE(p1_parRF1, test1$y) # 4.887264
-fit_parRF2 <- fit_parRF(data2, 1)
-ggplot(fit_parRF2)
-min(fit_parRF2$results$RMSE) # 6.268789
-p_parRF2 <- predict(fit_parRF2, newdata = test2)
-RMSE(p_parRF2, test2$y) # 4.812994
-fit_parRF3 <- fit_parRF(data3, 1)
-ggplot(fit_parRF3)
-min(fit_parRF3$results$RMSE) # 6.361905
-p_parRF3 <- predict(fit_parRF3, newdata = test3)
-RMSE(p_parRF3, test3$y) # 5.083467
-fit_parRF4 <- fit_parRF(data4, 1)
-ggplot(fit_parRF4)
-min(fit_parRF4$results$RMSE) # 8.705984
-p_parRF4 <- predict(fit_parRF4, newdata = test4)
-RMSE(p_parRF4, test4$y) # 8.083962
-fit_parRF5 <- fit_parRF(data5, 1)
-ggplot(fit_parRF5)
-min(fit_parRF5$results$RMSE) # 8.691755
-p_parRF5 <- predict(fit_parRF5, newdata = test5)
-RMSE(p_parRF5, test5$y) # 7.991537
-```
-```{r table-done-samples-28, echo=F}
-# Tabela - Primeiras 6 amostras do conjunto de treino do modelo de 28 dias
-caption <- "Primeiras 6 amostras do conjunto de treino do modelo de 28 dias"
-colNames = c("Cimento", "E.G.A.F.", "C.Vol.", "Água",
-              "Superp.", "A.Graúdo", "A.Miúdo",
-              "Ág./", "A.M./", "A.G./",
-              "A.M./","Ág./", "Ág./", "y")
-dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
-              "$kg/m^3$", "$kg/m^3$","$kg/m^3$", "Ci.",
-              "Ci.", "Ci.", "A.G.", "A.G.", "Ag.M.", "$MPa$")
-colNames2 = c("Features"=13, "Outcome"=1)
-kable(
-    head(data2),
-    col.names = dfUnits,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    digits = 2,
-    align = "c"
-    ) %>%
-    add_header_above(header = colNames, line = F, align = "c") %>%
-    add_header_above(header = colNames2, line = T, align = "c") %>%
-    kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-Para cada conjunto de idade foi criado um modelo utilizando o algorítimo *Parallel Random Forest*, definido anteriormente (\ref{show-reg-models}).
-Para cada um dos 6 modelos, o parâmetro *mtry* foi otimizado, e foi realizado *repeated cross-validation*, dividindo em 10 ou 30 partes e repetindo 10 vezes.
-```{r reg-models, echo=F, warning=F, message=F, include=F}
-# Modelos de regressão
-# no dummy vars:
-data3 <- dats_reg$d3$train[c(1:13, 22)]
-data7 <- dats_reg$d7$train[c(1:12, 19)]
-data14 <- dats_reg$d14$train[c(1:13, 21)]
-data28 <- dats_reg$d28$train[c(1:13, 21)]
-data56 <- dats_reg$d56$train[c(1:13, 22)]
-data100 <- dats_reg$d100$train[c(1:13, 23)]
-test3 <- dats_reg$d3$test[c(1:13, 22)]
-test7 <- dats_reg$d7$test[c(1:12, 19)]
-test14 <- dats_reg$d14$test[c(1:13, 21)]
-test28 <- dats_reg$d28$test[c(1:13, 21)]
-test56 <- dats_reg$d56$test[c(1:13, 22)]
-test100 <- dats_reg$d100$test[c(1:13, 23)]
-get_trControl <- function(n, r){
-  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
-}
-trControl3 <- get_trControl(30, 10)
-trControl7 <- get_trControl(10, 10)
-trControl14 <- get_trControl(30, 10)
-trControl28 <- get_trControl(30, 10)
-trControl56 <- get_trControl(30, 10)
-trControl100 <- get_trControl(10, 10)
-get_tuneGrid <- function(data){
-  expand.grid(mtry = seq(1, length(data), 1))
-}
-tuneGrid3 <- get_tuneGrid(data3)
-tuneGrid7 <- get_tuneGrid(data7)
-tuneGrid14 <- get_tuneGrid(data14)
-tuneGrid28 <- get_tuneGrid(data28)
-tuneGrid56 <- get_tuneGrid(data56)
-tuneGrid100 <- get_tuneGrid(data100)
-set.seed(1, sample.kind = "Rounding")
-fit_3 <- train(y ~ ., 
-            data = data3,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid3,
-            trControl=trControl3)
-p_3 <- predict(fit_3, newdata = test3)
-RMSE_test_3 <- RMSE(p_3, test3$y)
-RMSE_test_3 # 3.31037
-fit_3$bestTune # mtry = 6
-set.seed(1, sample.kind = "Rounding")
-fit_7 <- train(y ~ ., 
-            data = data7,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid7,
-            trControl=trControl7)
-p_7 <- predict(fit_7, newdata = test7)
-RMSE_test_7 <- RMSE(p_7, test7$y)
-RMSE_test_7 # 4.361987
-fit_7$bestTune # mtry = 2
-set.seed(1, sample.kind = "Rounding")
-fit_14 <- train(y ~ ., 
-            data = data14,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid14,
-            trControl=trControl14)
-p_14 <- predict(fit_14, newdata = test14)
-RMSE_test_14 <- RMSE(p_14, test14$y)
-RMSE_test_14 # 4.620515
-fit_14$bestTune # mtry = 13
-set.seed(1, sample.kind = "Rounding")
-fit_28 <- train(y ~ ., 
-            data = data28,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid28,
-            trControl=trControl28)
-p_28 <- predict(fit_28, newdata = test28)
-RMSE_test_28 <- RMSE(p_28, test28$y)
-RMSE_test_28 # 4.716698
-fit_28$bestTune # mtry = 11
-set.seed(1, sample.kind = "Rounding")
-fit_56 <- train(y ~ ., 
-            data = data56,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid56,
-            trControl=trControl56)
-p_56 <- predict(fit_56, newdata = test56)
-RMSE_test_56 <- RMSE(p_56, test56$y)
-RMSE_test_56 # 5.939163
-fit_56$bestTune # mtry = 8
-set.seed(1, sample.kind = "Rounding")
-fit_100 <- train(y ~ ., 
-            data = data100,
-            preProcess = c("center","scale"),
-            method='parRF', 
-            tuneGrid=tuneGrid100,
-            trControl=trControl100)
-p_100 <- predict(fit_100, newdata = test100)
-RMSE_test_100 <- RMSE(p_100, test100$y)
-RMSE_test_100 # 5.851088
-fit_100$bestTune # mtry = 8
-```
-
-# Resultados
-O *RMSE* de teste de cada modelo em ordem crescente de idade foi respectivamente `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` e `r round(RMSE_test_100,2)`.
-A tabela \ref{tab:table-reg-models} apresenta os detalhes e resultados de cada modelo (\ref{show-table-reg-models}).
-A figura \ref{fig:results-comparison} compara os valores reais e previstos (\ref{show-results-comparison}), e as tabelas seguintes mostram os melhores e piores resultados de cada modelo (\ref{show-table-10}).
-```{r table-reg-models, echo=F}
-# Tabela - Detalhes dos modelos
-colNames <- c("Modelo", "mtry", "CV", "Repetições", 
-              "RMSE (treino)", "RMSE (teste)")
-caption <- "Resultados dos modelos de regressão"
-day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
-dat_reg_models <- data.frame(
-  dia = day,
-  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
-           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
-  number = c(trControl3$number, trControl7$number, trControl14$number,
-             trControl28$number,trControl56$number,trControl100$number),
-  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
-             trControl28$repeats,trControl56$repeats,trControl100$repeats),
-  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
-                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
-                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
-  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
-                RMSE_test_28, RMSE_test_56, RMSE_test_100)
-)
-kable(
-    dat_reg_models,
-    col.names = colNames,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  kable_styling(latex_options = c("HOLD_position"))
-```
-```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
-# Figura - Comparação dos modelos
-cap <- "Comparação dos valores reais e previstos em cada modelo"
-models <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
-xlabel <- "Real (MPa)"
-ylabel <- "Previsto (MPa)"
-preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
-actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
-gen_res_df <- function(ind){
-  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
-}
-res_df <- lapply(1:6, gen_res_df)
-res_df <- bind_rows(res_df)
-res_df$model <- factor(res_df$model,levels=models)
-res_df %>%
-  ggplot(aes(actual, pred)) +
-  facet_wrap(~ model, ncol=3) +
-  geom_point(alpha=0.5) +
-  theme_bw() +
-  geom_abline(slope=1, intercept=0) +
-  xlab(xlabel) +
-  ylab(ylabel)
-```
-```{r table-10, echo=F}
-# Tabelas dos 10 melhores e piores resultados
-colNames1 = c("Real", "Previsto", "Erro", "","Real", "Previsto", "Erro") 
-colNames2 = c("10 melhores"=3,"", "10 piores"=3)
-caption_0 <- "Modelo de "
-get_X <- function(pred, actual, X=10){
-  diff <- abs(pred - actual)
-  diff_2 <- pred - actual
-  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
-  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
-  df_min <- data.frame(
-    actual_min=actual[ind_min],
-    pred_min=pred[ind_min],
-    diff_min=diff[ind_min],
-    diff_min_2=diff_2[ind_min]
-  )
-  df_max <- data.frame(
-    actual_max=actual[ind_max],
-    pred_max=pred[ind_max],
-    diff_max=diff[ind_max],
-    diff_max_2=diff_2[ind_max]
-  )
-  df_max <- df_max[order(-df_max$diff_max),]
-  df_min <- df_min[order(df_min$diff_min),]
-  df_null <- data.frame(null_col = rep(c(""), X))
-  res <- cbind(df_min,df_null, df_max)
-  res <- res %>% select(-c(diff_max, diff_min))
-  rownames(res) <- c()
-  res
-}
-gen_kable <- function(ind){
-  df <- get_X(preds[[ind]], actuals[[ind]])
-  caption <- paste0(caption_0,models[ind])
-  kable(
-    df,
-    col.names = colNames1,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  column_spec(4, width = "1cm",) %>%
-  add_header_above(header = colNames2, line = T, align = "c") %>%
-  kable_styling(latex_options = c("HOLD_position"))
-}
-gen_kable(1)
-gen_kable(2)
-gen_kable(3)
-gen_kable(4)
-gen_kable(5)
-gen_kable(6)
-```
-
-# Discussão
-Os modelos construidos apresentam resultados satisfatórios e provam que a resistência à compressão do concreto pode ser prevista de forma relativamente fácil.
-A alternativa adotada de criar um modelo para cada conjunto de idade se mostrou como uma alternativa válida, conseguindo estratificar para obter resultados específicos de cada conjunto.
-Os estudos citados na introdução utilizando o mesmo conjunto de dados possuem resultados similares, como esperado.
-A tabela \ref{tab:works-comparison} apresenta os resultados desses trabalhos (\ref{show-works-comparison}), e a tabela \ref{tab:results} apresenta os valores encontrados (\ref{show-results}) para fácil comparação.
-```{r works-comparison, echo=F}
-# Tabela - "Comparação dos estudos de outros autores"
-colNames = c("Autor","Ano" ,"Algorítimo", "RMSE") 
-caption <- "Comparação dos estudos de outros autores"
-works <- data.frame(
-  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
-  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
-  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
-          "Gradient Boosting Regressor","Random Forest Regressor" ,
-          "Regularized Extreme Learning Machine",
-          "Support Vector Machines with Radial Basis Function Kernel"),
-  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
-)
-kable(
-  works,
-  col.names = colNames,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "l"
-  )  %>%
-    kable_styling(latex_options = c("HOLD_position"))
-```
-```{r results, echo=F}
-# Tabela - Resultados finais
-colNames <- c("Modelo", "RMSE")
-caption <- "Resultados finais"
-day <- c("3 dias", "7 dias", "14 dias", "28 dias", "56 dias", "100 dias")
-dat_reg_models <- data.frame(
-  dia = day,
-  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
-                RMSE_test_28, RMSE_test_56, RMSE_test_100)
-)
-kable(
-    dat_reg_models,
-    col.names = colNames,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    )  %>%
-  kable_styling(latex_options = c("HOLD_position"))
-```
-Seguindo a linha de raciocínio desse trabalho, ele pode ser realizado com diferentes algorítimos, os resultados aqui encontrados utilizaram apenas um único (*Parallel Random Forest*), mesmo que tenha sido teoricamente o "melhor" encontrado, outros algorítmos podem apresentar resultados ainda melhores.
-Outra opção é criar um *ensemble* com diversos algorítimos, como realizado por @Pierobon2018, mas com a separação de conjuntos de idade aqui proposto.
-Além disso, pode ser realizado com um conjunto maior de dados, idealmente com o mesmo número de amostras em cada conjunto de idade, distribuição mais homgênia da resistência à compressão e menor variância entre as amostras.
-
 # Bibliografia

### 2020-07-13 templates/Abstract.Rmd
Update abstract

--- 2020-04-29 templates/Abstract_&_Setup.Rmd
@@
 ---
-<EN>title:
-  Concrete compressive strength prediction with machine learning
-date: "April 22, 2020"</EN>
-<PT>title:
-  Prevendo a resistência à compressão do concreto com técnicas de machine learning
-date: "Abril 22, 2020"</PT>
-author: "Pedro Bernardino Alves Moreira"
-indent: true
-<EN>abstract:
+output:
+  pdf_document: default
+---
+# Novo Resumo - Inglês
 Compressive strength is the main characteristic of concrete.
@@
 It was a promising alternative since good results were achieved by training with just one algorithm.
-This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.</EN>
-<PT>abstract:
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.
+
+Compressive strength is the main characteristic of concrete.
+The correct prediction of this parameter results in cost and time reduction.
+This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
+Was used a dataset with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
+The predictive models were developed in R language, using the caret package with the Parallel Random Forest algorithm and repeated cross-validation technique to optimize the parameters.
+The results were satisfactory and compatible with other studies using the same data set.
+The most important model, 28 days old, obtained RMSE of 4.717.
+The 3-day model obtained the best result, RMSE of 3.310.
+The worst result was the 56-day model, with RMSE of 5.939.
+The work showed that the compressive strength of concrete can be predicted.
+The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age.
+It was a promising alternative since good results were achieved by training with just one algorithm.
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.
+
+# Novo Resumo - Português
 A resistência à compressão é a principal característica do concreto.
@@
 Foi uma alternativa promissora, visto que bons resultados foram atingidos treinando com apenas um algoritmo.
-Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos.</PT>
-output:
-  pdf_document:
-    keep_tex: yes
-    number_sections: yes
-    toc_depth: 2
-bibliography: references.bib
-header-includes:
-<PT>- \renewcommand{\abstractname}{Resumo}</PT>
-- \usepackage{booktabs}
-- \usepackage{longtable}
-- \usepackage{supertabular}
-- \usepackage{array}
-- \usepackage{multirow}
-- \usepackage[table]{xcolor}
-- \usepackage{wrapfig}
-- \usepackage{colortbl}
-- \usepackage{pdflscape}
-- \usepackage{tabu}
-- \usepackage{threeparttable}
-- \usepackage[normalem]{ulem}
-- \usepackage{caption}
-- \usepackage{floatrow}
-- \floatsetup[table]{capposition=top}
-- \floatsetup[figure]{capposition=top}
-- \captionsetup{options=chunk}
-- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
-- \renewcommand{\thechunk}{\arabic{chunk}}
-- \usepackage{indentfirst}
-- \usepackage{sectsty} \sectionfont{\centering}
----
+Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos.
 
-```{r setup, include=FALSE, cache=F}
-# Install libraries, if not already installed
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(ggplot2)) install.packages("ggplot2")
-if(!require(caret)) install.packages("caret")
-if(!require(knitr)) install.packages("knitr")
-if(!require(kableExtra)) install.packages("kableExtra")
-if(!require(gdata)) install.packages("gdata")
-if(!require(dplyr)) install.packages("dplyr")
-if(!require(tidyr)) install.packages("tidyr")
-if(!require(cowplot)) install.packages("cowplot")
-if(!require(pastecs)) install.packages("pastecs")
-if(!require(factoextra)) install.packages("factoextra")
-if(!require(reshape2)) install.packages("reshape2")
-if(!require(purrr)) install.packages("purrr")
-if(!require(gridExtra)) install.packages("gridExtra")
-if(!require(questionr)) install.packages("questionr")
-
-# Load libraries
-library(tidyverse)
-library(caret)
-library(ggplot2)
-library(knitr)
-library(kableExtra)
-library(dplyr)
-library(tidyr)
-library(gdata)
-library(reshape2)
-library(cowplot)
-library(pastecs)
-library(factoextra)
-library(purrr)
-library(gridExtra)
-library(questionr)
-
-# Set global options
-options(pillar.sigfig = 5)
-options(knitr.kable.NA = '')
-opts_knit$set(eval.after = "fig.cap")
-opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
-oldSource <- knit_hooks$get("source")
-knit_hooks$set(source = function(x, options) {
-  x <- oldSource(x, options)
-  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
-  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
-})
-```
-
-<PT>
-\renewcommand{\figurename}{Figura}
-\renewcommand{\tablename}{Tabela}
-\renewcommand{\contentsname}{Sumário}
-\renewcommand{\chunkname}{Código}
-\renewcommand{\listfigurename}{Lista de figuras}
-\renewcommand{\listtablename}{Lista de tabelas}
+# Resumo antigo
+A resistência à compressão é a principal característica do concreto.
+Esse trabalho utilizou um conjunto de amostras amplamente utilizado para construir 6 modelos de previsão dessa característica, cada um específicamente para uma idade, 3, 7, 14, 28, 56 e 100 dias a partir de 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes) e 6 relações entre os principais ingredientes (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo) para fazer a previsão.
+Foi utilizado o algoritmo *Parallel Random Forest* (*parRF* do pacote *caret* de linguagem *R*).
+Os resultados foram satisfatórios sendo o melhor para o modelo de 3 dias (*RMSE* de 3,31) e pior para o modelo de 100 dias (*RMSE* de 5,85).
+O modelo de 28 dias apresentou *RMSE* de 4,72.
+O trabalho indicou que a separação em um modelo para cada idade é promissora e foi capaz de gerar resultados compatíveis ou melhores que outros estudos, abrindo espaço para construção de novos modelos seguindo essa proposta, mas utilizando algoritmos diferentes ou a combinação de diversos algoritmos.
 </PT>
 
-\newpage
+# Resumo explodido (o Novo, mesma coisa, não precisa ler, é apenas pra eu me organizar)
 
-\tableofcontents
+## Introdução
+A resistência à compressão é a principal característica do concreto. A previsão correta desse parâmetro significa redução de custo e tempo.
 
-\newpage
+## Objetivo
+Esse trabalho construiu modelos de previsão em 6 idades diferentes de amostras de concreto (3, 7, 14, 28, 56, e 100 dias).
 
-\listoffigures
+## Métodos
+Foi utilizado um conjunto de amostras obtido em estudos anteriores, um total de 1030 amostras, com 9 variáveis: resistência à compressão, idade e 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória granulada de alto forno e superplastificantes).
+Outras 6 variáveis foram adicionadas para representar as proporções dos principais ingredientes em cada amostra (água/cimento, agregado miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo, água/agregado graúdo e água/agregado miúdo).
+Os modelos de previsão foram desenvolvidos em linguagem *R*, utilizando o pacote *caret* com o algoritmo *Parallel Random Forest* e técnica de validação cruzada repetida para otimização dos parâmetros.
 
-\listoftables
+## Resultados
+Os resultados foram satisfatórios e compatíveis com outros estudos utilizando o mesmo conjunto de dados.
+O modelo mais importante, de 28 dias, obteve *RMSE* de 4,717.
+O modelo de 3 dias obteve o melhor resultado, *RMSE* de 3,310.
+O pior resultado foi o modelo de 56 dias, com *RMSE* de 5,939.
 
-\newpage
+## Discussão
+O trabalho mostrou que a resistência à compressão do concreto pode ser prevista.
+A escolha de criar um modelo para cada idade, ao invés de utilizar a idade como característica para previsão, permitiu chegar a resultados compatíveis com os dados disponíveis de cada idade.
+Foi uma alternativa promissora, visto que bons resultados foram atingidos treinando com apenas um algoritmo.
+Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos.

### 2020-07-13 CILAMCE/paper_EN.Rmd
Update CILAMCE folder

@@
 opts_knit$set(eval.after = "fig.cap")
-opts_chunk$set(echo = TRUE, fig.align = 'center', cache=F, prompt=F, highlight=T )
+opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
 ```
 
-# Lorem ipsum
+# Introduction
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens.
+Measurement at 28 days is mandatory and represents the grade of the concrete.
+Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.
 
-# Lorem ipsum dolor
+This compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed.
+@Hasan2011 proposed a mathematical model to predict from the results of tests of 7 and 14 days, and @Kabir2012 from 7 days.
+However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+```{r comentario-1, echo=T}
+# Aqui vale incluir um parágrafo para explicar o que é machine learning,
+# e por que é possível obter a resistência a partir apenas dos ingredientes.
+```
 
+Many previous studies use the same dataset used by @Yeh1998 to predict the compressive strength of concrete.
+@Alshamiri2020 got good results with the regularized extreme learning machine (RELM) technique, and @Hameed2020 got even better results with the Artificial Neural Networks and cross-validation technique.
+This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as @Abban2016, @Raj2018, @Modukuru2020 and @Pierobon2018.
+At the end of the work, the results found are compared to the works cited here.
+
+Unlike previous studies with this dataset, this work does data preparation differently.
+The age of the concrete is the most unique feature that contributes to its compressive strength.
+For this reason, age is treated separately in the machine learning models, creating models for each age group.
+
+# Materials and Methods
+
+## Materials
+
+```{r comentario-2, echo=T}
+# Editar, porque não vai ter mais os apêndices
+```
+
+The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore].
+Throughout the process, all code executed was documented in the same order as its execution in Appendix 3, and reference was always made to codes throughout the text.
+All relevant information related to the operating system and installed packages has been presented in Appendix 1.
+In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in[Appendix 2.
+
+## Reproducibility
+
+```{r comentario-3, echo=T}
+# Editar, porque não vai ter mais os apêndices
+# Definir o que é seed
+```
+
+In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result.
+The *seeds* can be checked throughout Appendix 3 or directly on *Github*.
+
+## Obtaining the data
+The data was downloaded from the University of California Irvine website [@downloadData].
+In total there are 1030 samples with 9 columns.
+The samples were renamed and an id column was added to facilitate data manipulation.
+The columns were reordered to put the new id column in the first position.
+The first samples can be viewed in the Table \ref{tab:first-samples}.
+
+```{r download-data, echo=F, include=F}
+# Data download
+url_base <- "https://archive.ics.uci.edu"
+url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
+download.file(paste0(url_base, url), "data.xls")
+dat <- read.xls("data.xls")
+n_inicial_samples <- nrow(dat)
+colnames(dat)
+```
+```{r rename-dat-cols, echo=F}
+# Renaming the columns
+colnames(dat) <- c(
+  "cement",
+  "blast_furnace_slag",
+  "fly_ash",
+  "water",
+  "superplasticizers",
+  "coarse_aggregate",
+  "fine_aggregate",
+  "day",
+  "mpa"
+  )
+dat$id <- seq.int(nrow(dat))
+```
+```{r reorder-dat, echo=F}
+# Reordering the data
+col_order <- c(
+  "id",
+  "cement",
+  "blast_furnace_slag",
+  "fly_ash",
+  "water",
+  "superplasticizers",
+  "coarse_aggregate",
+  "fine_aggregate",
+  "day",
+  "mpa"
+)
+dat <- dat[, col_order]
+```
+```{r col-names-and-units, echo=F}
+# Defining column names and units
+colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
+             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
+dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
+             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
+```
+```{r first-samples, echo=F}
+# Table - First samples
+caption <- "First 6 samples"
+kable(
+    dat[1:6,],
+    col.names = dfUnits,
+    escape = F,
+    booktabs = T,
+    caption = caption,
+    linesep = "",
+    align = "c"
+    ) %>%
+    add_header_above(header = colNames, line = F, align = "c") %>%
+    kable_styling(latex_options = c("HOLD_position", "scale_down"))
+```
+
+## Data preparation
+
+The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies.
+Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed.
+In addition, the relevant data has been transformed to better fit the studies in the next steps.
+
+### Initial data cleaning
+
+Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples.
+
+```{r removing-duplicated-samples, echo=F}
+# Removing duplicated samples
+n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
+n_duplicated_samples <- n_inicial_samples - n_distinct_samples
+dat <- dat[!duplicated(select(dat, -c(id))),]
+n_samples <- nrow(dat)
+```
+
+The data show the variables in the columns and samples in the rows.
+However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the Table \ref{tab:similar-samples}.
+
+```{r similar-samples, echo=FALSE}
+# Table - Samples with same composition
+same_samples <- dat %>%
+  filter(id %in% c(653, 678, 654, 681))
+caption <- "Samples with same composition"
+kable(
+  same_samples[order(same_samples$day),],
+  col.names = dfUnits,
+  escape = F,
+  booktabs = T,
+  caption = caption,
+  linesep = "\\addlinespace",
+  align = "c",
+  row.names = FALSE
+  ) %>%
+  add_header_above(header = colNames, line = F, align = "c") %>%
+  kable_styling(latex_options = c("HOLD_position", "scale_down"))
+```
+
+In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process.
+This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.
+
+--- 
 ## Lorem ipsum Sub
@@
 
-# References
+# tes
 
-Example of reference: "@Yeh1998 proposed… @Hameed2020".
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Example of reference: "@Yeh1998 proposed… @Hameed2020.
+@Alshamiri2020,  ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

### 2020-07-13 CILAMCE/paper_EN.tex
Update CILAMCE folder

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

%% --------------------------------------------------------------------------
% LaTeX template for the XL CILAMCE.
%
% This latex document tries to copy the Microsoft Word template.
% --------------------------------------------------------------------------
\documentclass[a4paper,11pt]{book}

% PACKAGES USED - packages that need to be previously installed on your computer
\usepackage[lmargin=2.5cm, rmargin=2.5cm, tmargin=2.5cm, bmargin=2.5cm ]{geometry}
\usepackage{graphicx}
\usepackage{times}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[english]{babel}
\usepackage{parskip} 
\usepackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% My Additional Packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
%\usepackage{cleveref}
% \usepackage{./pkg-crefNames}
\usepackage[labelsep=period]{caption}

% % %BibTeX compatible with the CILAMCE format
% \usepackage[numbers,sort&compress]{natbib}
% \makeatletter
% \renewcommand\bibsection
% {
%   \section*{References}
% }
% \makeatother
% \bibliographystyle{./bib-cilamce}
% \setlength{\bibhang}{0pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% CONFIGURATION
\renewcommand*\arraystretch{1.5}
\renewcommand*\thesection{\arabic{section}}
%\hyphenpenalty=10000 % You can uncomment this to avoid hyphenization
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\bfseries}
\titlespacing\section{0pt}{18pt plus 2pt minus 2pt}{18pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{13pt plus 0pt minus 2pt}{13pt plus 0pt minus 2pt}
\setlength{\parskip}{0pt} % Spacing between paragraphs
\setlength{\parindent}{0.75cm} % Paragraph identation
\setlength\abovecaptionskip{6pt}

% --------------------------------------------------------------------------
% DO NOT EDIT - SPECIAL HEADINGS OF XLI CILAMCE
% --------------------------------------------------------------------------
\fancypagestyle{first}
{
\fancyhf{}
\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}
\renewcommand{\headrulewidth}{.0pt}
\renewcommand{\footrulewidth}{.5pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[RO,LE]{}

\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\fancyfoot[LE]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\renewcommand{\headrulewidth}{.5pt}
\renewcommand{\footrulewidth}{.5pt}

% --------------------------------------------------------------------------
% PLEASE, EDIT THIS!
\fancyhead[LE]{\footnotesize \textit{Concrete compressive strength prediction with machine learning}}
\fancyhead[RO]{\footnotesize \textit{P. Moreira, V. da Silva}}
% --------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
% \usepackage[]{natbib}
\usepackage[numbers,sort&compress,sectionbib]{natbib}
\bibliographystyle{./bib-cilamce}

\title{INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR
PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE}
\author{}
\date{}

\begin{document}\thispagestyle{first}

\begin{figure}[ht!]
\vspace{-50pt}
\flushright
\includegraphics[width=5.5cm]{logoCILAMCE2019.png}
%scale=0.25
\end{figure}

% \maketitle
% 
% manually adding title
\noindent
\textbf{\large
INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR
PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE}
\vspace{18pt}

\noindent  \textbf{Pedro B. A. Moreira}

\noindent  \textit{pedrobermoreira@gmail.com}

\noindent \textit{Student, Dept. of Engineering, University Veiga de Almeida}

\noindent \textit{Address, Zip-Code, State/Province, Country}

\noindent  \textbf{Victor M. da Silva}

\noindent  \textit{somebody2@somewhere.com}

\noindent \textit{Assistant Professor, Dept. of Engineering, IBMEC/RJ}

\noindent \textit{Address, Zip-Code, State/Province, Country}

\vspace{18pt}

\noindent \textbf{Abstract.}
Compressive strength is the main characteristic of concrete. The correct prediction of this parameter results in cost and time reduction. This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days). Was used a dataset with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers). Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate). The predictive models were developed in R language, using the caret package with the Parallel Random Forest algorithm and repeated cross-validation technique to optimize the parameters. The results were satisfactory and compatible with other studies using the same data set. The most important model, 28 days old, obtained RMSE of 4.717. The 3-day model obtained the best result, RMSE of 3.310. The worst result was the 56-day model, with RMSE of 5.939. The work showed that the compressive strength of concrete can be predicted. The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age. It was a promising alternative since good results were achieved by training with just one algorithm. This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.

\vspace{18pt} % <- keep this vertical space!

\noindent \textbf{Keywords:} Concrete, Compressive Strength, Machine Learning, Prediction

\pagenumbering{gobble}

\clearpage

\renewcommand{\bibname}{References}

\setlength{\abovedisplayskip}{18pt}
\setlength{\belowdisplayskip}{18pt}
\setlength{\abovedisplayshortskip}{18pt}
\setlength{\belowdisplayshortskip}{18pt}

\renewcommand{\arraystretch}{1.2}

\captionsetup[table]{skip=12pt}

\setlength\intextsep{13pt}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Compressive strength is the main characteristic of concrete, measured by
tests of international standards that consist of the breaking of
specimens. Measurement at 28 days is mandatory and represents the grade
of the concrete. Knowing in advance what the result will be obtained for
a given age, based on the proportions of its ingredients, is of great
interest to concrete manufacturers, construction companies, and civil
engineers.

This compressive strength is a nonlinear function of its ingredients and
age, making it difficult to establish an analytical formula, although
some formulas have already been proposed. \citet{Hasan2011} proposed a
mathematical model to predict from the results of tests of 7 and 14
days, and \citet{Kabir2012} from 7 days. However, machine learning
techniques can be used to model this characteristic from real sample
data, using only the ingredients.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Aqui vale incluir um parágrafo para explicar o que é machine learning,}
\CommentTok{# e por que é possível obter a resistência a partir apenas dos ingredientes.}
\end{Highlighting}
\end{Shaded}

Many previous studies use the same dataset used by \citet{Yeh1998} to
predict the compressive strength of concrete. \citet{Alshamiri2020} got
good results with the regularized extreme learning machine (RELM)
technique, and \citet{Hameed2020} got even better results with the
Artificial Neural Networks and cross-validation technique. This set of
samples is so well known that there are many pages on the internet of
unpublished studies that use it and have good results, such as
\citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and
\citet{Pierobon2018}. At the end of the work, the results found are
compared to the works cited here.

Unlike previous studies with this dataset, this work does data
preparation differently. The age of the concrete is the most unique
feature that contributes to its compressive strength. For this reason,
age is treated separately in the machine learning models, creating
models for each age group.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Editar, porque não vai ter mais os apêndices}
\end{Highlighting}
\end{Shaded}

The methodology was carried out using RStudio software \citep{RStudio},
an integrated virtual environment for code development in \emph{R}
\citep{RCore}. Throughout the process, all code executed was documented
in the same order as its execution in Appendix 3, and reference was
always made to codes throughout the text. All relevant information
related to the operating system and installed packages has been
presented in Appendix 1. In addition, an online and open-source
repository was created in \emph{Github}, housing all the code used to
generate this work, the link was made available in{[}Appendix 2.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Editar, porque não vai ter mais os apêndices}
\CommentTok{# Definir o que é seed}
\end{Highlighting}
\end{Shaded}

In order to guarantee reproducibility, whenever there was some code that
could use probabilistic operations, a \emph{seed} was defined before its
execution, ensuring that when run on another machine, with the same
version of \emph{R} and the same \emph{seed}, get the same result. The
\emph{seeds} can be checked throughout Appendix 3 or directly on
\emph{Github}.

\hypertarget{obtaining-the-data}{%
\subsection{Obtaining the data}\label{obtaining-the-data}}

The data was downloaded from the University of California Irvine website
\citep{downloadData}. In total there are 1030 samples with 9 columns.
The samples were renamed and an id column was added to facilitate data
manipulation. The columns were reordered to put the new id column in the
first position. The first samples can be viewed in the Table
\ref{tab:first-samples}.

\begin{table}[H]

\caption{\label{tab:first-samples}First 6 samples}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 28 & 79.99\\
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 28 & 61.89\\
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 270 & 40.27\\
4 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 365 & 41.05\\
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 360 & 44.30\\
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 90 & 47.03\\
\bottomrule
\end{tabular}}
\end{table}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{lorem-ipsum-sub}{%
\subsection{Lorem ipsum Sub}\label{lorem-ipsum-sub}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\hypertarget{lorem-ipsum-sub-2}{%
\subsection{Lorem ipsum Sub 2}\label{lorem-ipsum-sub-2}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\hypertarget{equations-symbols-and-units}{%
\subsection{Equations, symbols and
units}\label{equations-symbols-and-units}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\begin{center}
\begin{equation}
q_r = -4pr^2k\frac{dT}{dr}.
\label{Eq1}
\end{equation}
\end{center}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\vspace{13pt}

Cite as Eq. \ref{Eq1}.

\hypertarget{figures-and-tables}{%
\subsection{Figures and tables}\label{figures-and-tables}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\begin{table}[H]

\caption{\label{tab:table-example}Coefficients in constitutive relations}
\centering
\begin{tabular}[t]{ccc}
\toprule
Constitutive relation & Nomenclature & Value\\
\midrule
Turbulent tensor & C & 0.09\\
Turbulent tensor & C & 0.69\\
Lateral lift & C & 0.08\\
Virtual mass & C & 0.80\\
\bottomrule
\end{tabular}
\end{table}

Cite as Figure \ref{fig:plot-example} and Table \ref{tab:table-example}.

\begin{figure}[!ht]

{\centering \includegraphics{paper_EN_files/figure-latex/plot-example-1}

}

\caption{Pressure variation along the nozzle: experimental data}\label{fig:plot-example}
\end{figure}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur.

\hypertarget{tes}{%
\section{tes}\label{tes}}

Example of reference: "\citet{Yeh1998} proposed\ldots{}
\citet{Hameed2020}. \citet{Alshamiri2020}, ipsum dolor sit amet,
consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore
et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud
exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
Duis aute irure dolor in reprehenderit in voluptate velit esse cillum
dolore eu fugiat nulla pariatur. Lorem ipsum dolor sit amet, consectetur
adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco
laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur.

% 
  \bibliography{references.bib}

% 
\end{document}

### 2020-07-13 CILAMCE_2020/paper.tex
Add new CILAMCE folder

--- 2020-07-13 CILAMCE/paper_EN.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-
 %% --------------------------------------------------------------------------
-% LaTeX template for the XL CILAMCE.
+% LaTeX template for the XLI CILAMCE.
 %
@@
 % --------------------------------------------------------------------------
-\documentclass[a4paper,11pt]{book}
+\documentclass[a4paper,10pt]{book}
 
@@
 \usepackage[utf8]{inputenc}
+%\usepackage{amssymb} %Mathematics
+%\usepackage{amsfonts}%Mathematics
+%\usepackage{amsmath,amscd}%Mathematics
+%\usepackage{amsthm}%Mathematics
+%\usepackage{mathrsfs}%Mathematics font
+%\usepackage{xspace}
+%\usepackage{booktabs}
+%\usepackage{stmaryrd}%Particular Brackets
+%\usepackage{graphicx} %Tables and Figures
+%\usepackage{subfigure}
+%\usepackage{url}
 \usepackage{hyperref}
-%\usepackage{cleveref}
-% \usepackage{./pkg-crefNames}
+\usepackage{cleveref}
+\usepackage{./pkg-crefNames}
 \usepackage[labelsep=period]{caption}
 
-% % %BibTeX compatible with the CILAMCE format
-% \usepackage[numbers,sort&compress]{natbib}
-% \makeatletter
-% \renewcommand\bibsection
-% {
-%   \section*{References}
-% }
-% \makeatother
-% \bibliographystyle{./bib-cilamce}
-% \setlength{\bibhang}{0pt}
+%BibTeX compatible with the CILAMCE format
+\usepackage[numbers,sort&compress]{natbib}
+
+\setlength{\bibsep}{0pt plus 0.3ex}
+
+\renewcommand*{\bibfont}{\small}
+
+\makeatletter
+\renewcommand\bibsection
+{
+  \section*{References}
+}
+
+\renewenvironment{thebibliography}[1]
+      {\section*{\refname}%
+       \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
+       \list{\@biblabel{\@arabic\c@enumiv}}%
+            {\settowidth\labelwidth{\@biblabel{#1}}%
+             \leftmargin\labelwidth
+             \advance\leftmargin-10pt% change 20 pt according to your needs
+             \advance\leftmargin\labelsep
+             \setlength\itemindent{10pt}% change using the inverse of the length used before
+             \@openbib@code
+             \usecounter{enumiv}%
+             \let\p@enumiv\@empty
+             \renewcommand\theenumiv{\@arabic\c@enumiv}}%
+       \sloppy
+       \clubpenalty4000
+       \@clubpenalty \clubpenalty
+       \widowpenalty4000%
+       \sfcode`\.\@m}
+      {\def\@noitemerr
+        {\@latex@warning{Empty `thebibliography' environment}}%
+       \endlist}
+\renewcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
+\makeatother
+
+\makeatother
+\bibliographystyle{./bib-cilamce}
+%\bibliographystyle{plain}
 
@@
 \titleformat*{\subsection}{\bfseries}
-\titlespacing\section{0pt}{18pt plus 2pt minus 2pt}{18pt plus 2pt minus 2pt}
-\titlespacing\subsection{0pt}{13pt plus 0pt minus 2pt}{13pt plus 0pt minus 2pt}
+\titlespacing\section{0pt}{20pt plus 2pt minus 2pt}{12pt plus 2pt minus 2pt}
+\titlespacing\subsection{0pt}{20pt plus 0pt minus 0pt}{12pt plus 0pt minus 0pt}
 \setlength{\parskip}{0pt} % Spacing between paragraphs
@@
 % PLEASE, EDIT THIS!
-\fancyhead[LE]{\footnotesize \textit{Concrete compressive strength prediction with machine learning}}
-\fancyhead[RO]{\footnotesize \textit{P. Moreira, V. da Silva}}
+\fancyhead[LE]{\footnotesize \textit{Template for CILAMCE 2020 Concrete compressive strength prediction with machine learning}}
+\fancyhead[RO]{\footnotesize \textit{P. Moreira, V. Silva}}
 % --------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-  \setmainfont[]{Times New Roman}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-% \usepackage[]{natbib}
-\usepackage[numbers,sort&compress,sectionbib]{natbib}
-\bibliographystyle{./bib-cilamce}
-
-\title{INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR
-PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE}
-\author{}
-\date{}
 
@@
 
+% --------------------------------------------------------------------------
+% DO NOT EDIT - LOGO OF XLI CILAMCE
+% --------------------------------------------------------------------------
+
 \begin{figure}[ht!]
-\vspace{-50pt}
+\vspace{-30pt}
 \flushright
-\includegraphics[width=5.5cm]{logoCILAMCE2019.png}
+\includegraphics[width=5.5cm]{Figures/logoCILAMCE2020.png}
 %scale=0.25
@@
 
-% \maketitle
-% 
-% manually adding title
+% --------------------------------------------------------------------------
+% TITLE OF PAPER
+% --------------------------------------------------------------------------
+
 \noindent
-\textbf{\large
-INSTRUCTIONS FOR PREPARATION AND SUBMISSION OF FULL-PAPERS FOR
-PUBLICATION IN THE PROCEEDINGS OF XL CILAMCE}
-\vspace{18pt}
+\textbf{\Large
+Concrete compressive strength prediction with machine learning} 
+\vspace{18pt} % <- keep this vertical space!
 
-\noindent  \textbf{Pedro B. A. Moreira}
+% --------------------------------------------------------------------------
+% AUTHORS
+% --------------------------------------------------------------------------
+
+\noindent Pedro B. A. Moreira$^1$, Victor M. Silva$^1$
+
+\vspace{18pt} % <- keep this vertical space!
+
+\noindent $^1$\textit{Student, Dept. of Engineering, University Veiga de Almeida}
+
+\noindent \textit{Avenida das Américas, 22631-004, Rio de Janeiro, Brazil}
 
@@
 
-\noindent \textit{Student, Dept. of Engineering, University Veiga de Almeida}
+\noindent $^2$\textit{Assistant Professor, Dept. of Engineering, IBMEC/RJ}
 
-\noindent \textit{Address, Zip-Code, State/Province, Country}
+\noindent \textit{TODO: Address, Zip-Code, State/Province, Country}
 
-\noindent  \textbf{Victor M. da Silva}
+\noindent \textit{TODO: somebody3@somewhere.com}
 
-\noindent  \textit{somebody2@somewhere.com}
+\vspace{18pt} % <- keep this vertical space!
 
-\noindent \textit{Assistant Professor, Dept. of Engineering, IBMEC/RJ}
-
-\noindent \textit{Address, Zip-Code, State/Province, Country}
-
-\vspace{18pt}
+% --------------------------------------------------------------------------
+% ABSTRACT
+% --------------------------------------------------------------------------
 
 \noindent \textbf{Abstract.}
+TODO: REDUCE TO 200 WORDS, CURRENTRLY IT HAS 270.
 Compressive strength is the main characteristic of concrete.
-The correct prediction of this parameter results in cost and time reduction.
+The correct prediction of this parameter means cost and time reduction.
 This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
-Was used a dataset with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
 Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
@@
 It was a promising alternative since good results were achieved by training with just one algorithm.
-This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be used as a baseline to predict with different algorithms or the combination of several.
 
@@
 
-\pagenumbering{gobble}
+% --------------------------------------------------------------------------
+\section{Introduction}\label{sec:introduction}
+% --------------------------------------------------------------------------
 
-\clearpage
+% --------------------------------------------------------------------------
+\section{Materials and methods}\label{sec:MaterialsAndMethods}
+% --------------------------------------------------------------------------
 
-\renewcommand{\bibname}{References}
+% --------------------------------------------------------------------------
+\subsection{Materials} \label{subsec:materials}
+% --------------------------------------------------------------------------
 
-\setlength{\abovedisplayskip}{18pt}
-\setlength{\belowdisplayskip}{18pt}
-\setlength{\abovedisplayshortskip}{18pt}
-\setlength{\belowdisplayshortskip}{18pt}
+% --------------------------------------------------------------------------
+\subsection{Reproducibility} \label{subsec:reproducibility}
+% --------------------------------------------------------------------------
 
-\renewcommand{\arraystretch}{1.2}
+% --------------------------------------------------------------------------
+\subsection{Obtaining the data} \label{subsec:Obtainingthedata}
+% --------------------------------------------------------------------------
 
-\captionsetup[table]{skip=12pt}
+% --------------------------------------------------------------------------
+\subsection{Data preparation} \label{subsec:datapreparation}
+% --------------------------------------------------------------------------
 
-\setlength\intextsep{13pt}
+% --------------------------------------------------------------------------
+\subsection{Data visualization} \label{subsec:datavisualization}
+% --------------------------------------------------------------------------
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+% --------------------------------------------------------------------------
+\subsection{Machine learning models} \label{subsec:machinelearningmodels}
+% --------------------------------------------------------------------------
 
-Compressive strength is the main characteristic of concrete, measured by
-tests of international standards that consist of the breaking of
-specimens. Measurement at 28 days is mandatory and represents the grade
-of the concrete. Knowing in advance what the result will be obtained for
-a given age, based on the proportions of its ingredients, is of great
-interest to concrete manufacturers, construction companies, and civil
-engineers.
+% --------------------------------------------------------------------------
+\section{Results}\label{sec:Results}
+% --------------------------------------------------------------------------
 
-This compressive strength is a nonlinear function of its ingredients and
-age, making it difficult to establish an analytical formula, although
-some formulas have already been proposed. \citet{Hasan2011} proposed a
-mathematical model to predict from the results of tests of 7 and 14
-days, and \citet{Kabir2012} from 7 days. However, machine learning
-techniques can be used to model this characteristic from real sample
-data, using only the ingredients.
+% --------------------------------------------------------------------------
+\section{Discussion and conclusion}\label{sec:DiscussionAndConclusion}
+% --------------------------------------------------------------------------
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Aqui vale incluir um parágrafo para explicar o que é machine learning,}
-\CommentTok{# e por que é possível obter a resistência a partir apenas dos ingredientes.}
-\end{Highlighting}
-\end{Shaded}
+%--------------------------------------------------------------------------
+\noindent \textbf{Authorship statement.} The authors hereby confirm that they are the sole liable persons responsible for the authorship of this work, and that all material that has been herein included as part of the present paper is either the property (and authorship) of the authors, or has the permission of the owners to be included here.
 
-Many previous studies use the same dataset used by \citet{Yeh1998} to
-predict the compressive strength of concrete. \citet{Alshamiri2020} got
-good results with the regularized extreme learning machine (RELM)
-technique, and \citet{Hameed2020} got even better results with the
-Artificial Neural Networks and cross-validation technique. This set of
-samples is so well known that there are many pages on the internet of
-unpublished studies that use it and have good results, such as
-\citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and
-\citet{Pierobon2018}. At the end of the work, the results found are
-compared to the works cited here.
+\bibliography{bibliography}
 
-Unlike previous studies with this dataset, this work does data
-preparation differently. The age of the concrete is the most unique
-feature that contributes to its compressive strength. For this reason,
-age is treated separately in the machine learning models, creating
-models for each age group.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{materials}{%
-\subsection{Materials}\label{materials}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Editar, porque não vai ter mais os apêndices}
-\end{Highlighting}
-\end{Shaded}
-
-The methodology was carried out using RStudio software \citep{RStudio},
-an integrated virtual environment for code development in \emph{R}
-\citep{RCore}. Throughout the process, all code executed was documented
-in the same order as its execution in Appendix 3, and reference was
-always made to codes throughout the text. All relevant information
-related to the operating system and installed packages has been
-presented in Appendix 1. In addition, an online and open-source
-repository was created in \emph{Github}, housing all the code used to
-generate this work, the link was made available in{[}Appendix 2.
-
-\hypertarget{reproducibility}{%
-\subsection{Reproducibility}\label{reproducibility}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Editar, porque não vai ter mais os apêndices}
-\CommentTok{# Definir o que é seed}
-\end{Highlighting}
-\end{Shaded}
-
-In order to guarantee reproducibility, whenever there was some code that
-could use probabilistic operations, a \emph{seed} was defined before its
-execution, ensuring that when run on another machine, with the same
-version of \emph{R} and the same \emph{seed}, get the same result. The
-\emph{seeds} can be checked throughout Appendix 3 or directly on
-\emph{Github}.
-
-\hypertarget{obtaining-the-data}{%
-\subsection{Obtaining the data}\label{obtaining-the-data}}
-
-The data was downloaded from the University of California Irvine website
-\citep{downloadData}. In total there are 1030 samples with 9 columns.
-The samples were renamed and an id column was added to facilitate data
-manipulation. The columns were reordered to put the new id column in the
-first position. The first samples can be viewed in the Table
-\ref{tab:first-samples}.
-
-\begin{table}[H]
-
-\caption{\label{tab:first-samples}First 6 samples}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccccccc}
-\toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
- & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
-\midrule
-1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 28 & 79.99\\
-2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 28 & 61.89\\
-3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 270 & 40.27\\
-4 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 365 & 41.05\\
-5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 360 & 44.30\\
-6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 90 & 47.03\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-\hypertarget{lorem-ipsum-sub}{%
-\subsection{Lorem ipsum Sub}\label{lorem-ipsum-sub}}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\hypertarget{lorem-ipsum-sub-2}{%
-\subsection{Lorem ipsum Sub 2}\label{lorem-ipsum-sub-2}}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\hypertarget{equations-symbols-and-units}{%
-\subsection{Equations, symbols and
-units}\label{equations-symbols-and-units}}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\begin{center}
-\begin{equation}
-q_r = -4pr^2k\frac{dT}{dr}.
-\label{Eq1}
-\end{equation}
-\end{center}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\vspace{13pt}
-
-Cite as Eq. \ref{Eq1}.
-
-\hypertarget{figures-and-tables}{%
-\subsection{Figures and tables}\label{figures-and-tables}}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\begin{table}[H]
-
-\caption{\label{tab:table-example}Coefficients in constitutive relations}
-\centering
-\begin{tabular}[t]{ccc}
-\toprule
-Constitutive relation & Nomenclature & Value\\
-\midrule
-Turbulent tensor & C & 0.09\\
-Turbulent tensor & C & 0.69\\
-Lateral lift & C & 0.08\\
-Virtual mass & C & 0.80\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Cite as Figure \ref{fig:plot-example} and Table \ref{tab:table-example}.
-
-\begin{figure}[!ht]
-
-{\centering \includegraphics{paper_EN_files/figure-latex/plot-example-1}
-
-}
-
-\caption{Pressure variation along the nozzle: experimental data}\label{fig:plot-example}
-\end{figure}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur.
-
-\hypertarget{tes}{%
-\section{tes}\label{tes}}
-
-Example of reference: "\citet{Yeh1998} proposed\ldots{}
-\citet{Hameed2020}. \citet{Alshamiri2020}, ipsum dolor sit amet,
-consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore
-et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud
-exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum
-dolore eu fugiat nulla pariatur. Lorem ipsum dolor sit amet, consectetur
-adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco
-laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
-in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur.
-
-% 
-  \bibliography{references.bib}
-
-% 
 \end{document}
+% --------------------------------------------------------------------------

### 2020-07-13 CILAMCE_2020/template.tex
Add new CILAMCE folder

%% --------------------------------------------------------------------------
% LaTeX template for the XLI CILAMCE.
%
% This latex document tries to copy the Microsoft Word template.
% --------------------------------------------------------------------------
\documentclass[a4paper,10pt]{book}

% PACKAGES USED - packages that need to be previously installed on your computer
\usepackage[lmargin=2.5cm, rmargin=2.5cm, tmargin=2.5cm, bmargin=2.5cm ]{geometry}
\usepackage{graphicx}
\usepackage{times}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[english]{babel}
\usepackage{parskip} 
\usepackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% My Additional Packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
%\usepackage{amssymb} %Mathematics
%\usepackage{amsfonts}%Mathematics
%\usepackage{amsmath,amscd}%Mathematics
%\usepackage{amsthm}%Mathematics
%\usepackage{mathrsfs}%Mathematics font
%\usepackage{xspace}
%\usepackage{booktabs}
%\usepackage{stmaryrd}%Particular Brackets
%\usepackage{graphicx} %Tables and Figures
%\usepackage{subfigure}
%\usepackage{url}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{./pkg-crefNames}
\usepackage[labelsep=period]{caption}

%BibTeX compatible with the CILAMCE format
\usepackage[numbers,sort&compress]{natbib}

\setlength{\bibsep}{0pt plus 0.3ex}

\renewcommand*{\bibfont}{\small}

\makeatletter
\renewcommand\bibsection
{
  \section*{References}
}

\renewenvironment{thebibliography}[1]
      {\section*{\refname}%
       \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
       \list{\@biblabel{\@arabic\c@enumiv}}%
            {\settowidth\labelwidth{\@biblabel{#1}}%
             \leftmargin\labelwidth
             \advance\leftmargin-10pt% change 20 pt according to your needs
             \advance\leftmargin\labelsep
             \setlength\itemindent{10pt}% change using the inverse of the length used before
             \@openbib@code
             \usecounter{enumiv}%
             \let\p@enumiv\@empty
             \renewcommand\theenumiv{\@arabic\c@enumiv}}%
       \sloppy
       \clubpenalty4000
       \@clubpenalty \clubpenalty
       \widowpenalty4000%
       \sfcode`\.\@m}
      {\def\@noitemerr
        {\@latex@warning{Empty `thebibliography' environment}}%
       \endlist}
\renewcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother

\makeatother
\bibliographystyle{./bib-cilamce}
%\bibliographystyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% CONFIGURATION
\renewcommand*\arraystretch{1.5}
\renewcommand*\thesection{\arabic{section}}
%\hyphenpenalty=10000 % You can uncomment this to avoid hyphenization
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\bfseries}
\titlespacing\section{0pt}{20pt plus 2pt minus 2pt}{12pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{20pt plus 0pt minus 0pt}{12pt plus 0pt minus 0pt}
\setlength{\parskip}{0pt} % Spacing between paragraphs
\setlength{\parindent}{0.75cm} % Paragraph identation
\setlength\abovecaptionskip{6pt}

% --------------------------------------------------------------------------
% DO NOT EDIT - SPECIAL HEADINGS OF XLI CILAMCE
% --------------------------------------------------------------------------
\fancypagestyle{first}
{
\fancyhf{}
\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}
\renewcommand{\headrulewidth}{.0pt}
\renewcommand{\footrulewidth}{.5pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[RO,LE]{}

\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\fancyfoot[LE]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\renewcommand{\headrulewidth}{.5pt}
\renewcommand{\footrulewidth}{.5pt}

% --------------------------------------------------------------------------
% PLEASE, EDIT THIS!
\fancyhead[LE]{\footnotesize \textit{Template for CILAMCE 2020 (enter here the short title of your paper)}}
\fancyhead[RO]{\footnotesize \textit{F. Author, S. Author, T. Author}}
% --------------------------------------------------------------------------

\begin{document}\thispagestyle{first}

% --------------------------------------------------------------------------
% DO NOT EDIT - LOGO OF XLI CILAMCE
% --------------------------------------------------------------------------

\begin{figure}[ht!]
\vspace{-30pt}
\flushright
\includegraphics[width=5.5cm]{Figures/logoCILAMCE2020.png}
%scale=0.25
\end{figure}

% --------------------------------------------------------------------------
% TITLE OF PAPER
% --------------------------------------------------------------------------

\noindent
\textbf{\Large
Instructions for preparation and submission of full-papers for publication in the Proceedings of CILAMCE-2020} 
\vspace{18pt} % <- keep this vertical space!

% --------------------------------------------------------------------------
% AUTHORS
% --------------------------------------------------------------------------

\noindent First A. Author$^1$, Second B. Author$^1$, Third C. Author$^2$

\vspace{18pt} % <- keep this vertical space!

\noindent $^1$\textit{Dept. of Something, University of Somewhere}

\noindent \textit{Address, Zip-Code, State/Province, Country}

\noindent \textit{somebody1@somewhere.com, somebody2@somewhere.com}

\noindent $^2$\textit{Dept. of Something Else, University of Somewhere Else}

\noindent \textit{Address, Zip-Code, State/Province, Country}

\noindent \textit{somebody3@somewhere.com}

\vspace{18pt} % <- keep this vertical space!

% --------------------------------------------------------------------------
% ABSTRACT
% --------------------------------------------------------------------------

\noindent \textbf{Abstract.}
This document-template provides detailed text formatting for preparing your full-length paper to the XLI Ibero-Latin American Congress on Computational Methods in Engineering (CILAMCE-2020). It is strongly recommended that you use the pre-defined styles of this template file. Also, it is strongly encouraged that you write your full-length paper in English.

\vspace{18pt} % <- keep this vertical space!

\noindent \textbf{Keywords:} First keyword, Second keyword, Third keyword (up to 5 keywords)

% --------------------------------------------------------------------------
\section{Introduction}\label{sec:introduction}
% --------------------------------------------------------------------------
Only full-length papers that have been orally presented will be published in the congress proceedings. It is extremely important that you prepare your full-length paper in strict accordance with the text formatting of this document, which can be enforced either using the pre-defined styles of this template file or manually setting the specifications described in the next section. After the preparation of your paper, you should generate a PDF file for submission. Only PDF files will be accepted by the online submission system.

% --------------------------------------------------------------------------
\section{Format instructions}\label{sec:format}
% --------------------------------------------------------------------------

Please follow these general instructions carefully: (a) type the body of the paper in 
single column; (b) use no more than 7 pages (5 pages for papers at the Research Beginners 
Mini-symposium), A4-sized, with 2.5 cm margins on all sides, and do not insert page numbers;
(c) use 10pt Times New Roman throughout the body of the text, with 15pt for the paper's title
13pt for first-level headings and 10pt for second-level headings; (d) always use either
exactly 13pt-spaced lines (preferred) or single-spaced lines, with justified alignment; 
(e) type no more than 200 words in the abstract; (f) cite references by Name  [number],
and list them consecutively in the reference list by the order of citation in the text 
(the list should only include works that have been cited in the text); (g) provide good
quality figures; (h) define all quantities, variables and symbols as soon as they first appear in the
text; (i) use only SI units. We strongly encourage you to use the pre-defined styles of this template file. Also, we strongly encourage you to write your
full-length paper in English. The general appearance of your paper should look like this
document.

% --------------------------------------------------------------------------
\subsection{Writing style} \label{subsec:writestyle}
% --------------------------------------------------------------------------

Please limit your paper by writing concisely, rather than by reducing figures or tables to a size at which symbols or labels would become difficult to read.

% --------------------------------------------------------------------------
\subsection{Additional specifications} \label{subsec:additionalspecifications}
% --------------------------------------------------------------------------

The first page must include the paper's title (in sentence case), authors, affiliations, abstract, keywords, as well as the beginning of the text itself (i.e., the first section, which is usually the introduction section). Do not insert a page break in between the keywords and the beginning of the text.

\vspace{13pt}

\noindent \textbf{\textit{Remark 1: Line spacing.}}. Please follow strictly the line spacing specifications defined next: either exactly 13pt spaced lines (preferred) or single-spaced lines for the body of the text, and 20pt space before and 12pt space after first-level and second-level headings.

\vspace{13pt}

\noindent \textbf{\textit{Remark 2: Author(s) and affiliation(s).}} Type authors' names in regular(plain) type, flush left, one per line, including first name, middle initial(s) and last name. Group the authors by their affiliations with a superscript number. Each name or group of names must be followed by the corresponding affiliations and emails, which should be both in italics. A 12pt space must be left between the authors’ names and their affiliations.

\vspace{13pt}

\noindent \textbf{\textit{Remark 3: Abstract and keywords.}} Type {``Abstract''} in boldface, flush left, followed by a period. On the same line, type the abstract in regular (plain) format, justified alignment. The abstract should be no longer than 200 words. Please pay attention to the line spacing between the authors' affiliations and the abstract, as well as between the abstract and the keywords. For the keywords, type the heading {``Keywords''}, followed by a colon, in boldface, flush left and type 3 to 5 keywords, separated by commas.

\vspace{13pt}

\noindent \textbf{\textit{Remark 4: Headings.}} First-level headings must be typed in sentence letters, 13 pt boldface type, flush left, as in Section 2 above. Use Arabic numbering (if you use this template's styles, the automatic numbering is already assigned to the first-level heading style). Notice the line spacing before and after first-level headings. For second-level headings, use sentence letters, 10 pt boldface type, flush left, with Arabic double numbering (if you use this template's styles, the automatic double numbering is already assigned to the second-level heading style. Notice the line spacing before and after second-level headings. Do not use third-level headings. Instead, at most, ``Remarks'' (or the like) are allowed, as shown here. These shall start with the word {``Remark''} (or the like) in boldface italics, followed by a number and title. The text that follows must be in regular (plain) format. Notice the line spacing before and after theses paragraphs.

\vspace{13pt}

\noindent \textbf{\textit{Remark 5: Body of text.}} As said before, the body of the text should be typed in 10pt Times New Roman, using  exactly 13 pt spaced lines (preferred) or single-spaced, and justified alignment. Start each paragraph with an indentation of 0.75 cm from the left margin, and allow no space between paragraphs.

% --------------------------------------------------------------------------
\subsection{Equations, symbols and units}\label{subsec:equationsSymbolsUnits}
% --------------------------------------------------------------------------

Equations must be centered, right-numbered, with numbers enclosed in parentheses and placed flush right. Allow 6 pt line spacing before and after an equation. For example:

\vspace{6pt}
\begin{center}
\begin{equation}
q_r = -4pr^2k\frac{dT}{dr}.
\label{Eq1}
\end{equation}
\end{center}
\vspace{6pt}

Please pay attention to the punctuation after the equations, as equations are part of the text and must be punctuated accordingly.
When referring to an equation in the text, write \cref{Eq1}, except at the beginning of a sentence, wherein Equation \ref{Eq1} should be used. Please describe the notation adopted and be careful as to define all quantities, variables and symbols as soon as they first appear in the text. A nomenclature section is not necessary.

All data, including those shown in tables and figures, must be reported in SI units. Decimal points rather than commas should always indicate decimals.

% --------------------------------------------------------------------------
\subsection{Figures and tables}\label{subsec:figuresAndTables}
% --------------------------------------------------------------------------

Figures and tables should be inserted as close as possible to their mentioning in the text. Enclosed text and symbols must be clearly readable; avoid small symbols thus. Supply good quality pictures and illustrations. Figures and tables and their captions should be centered in the text. Place figure captions below the figures and allow 12 pt line spacing between both, and 20 pt line spacing after the caption (before the adjacent text). Place table title above the table, leaving 20 pt line spacing before the title and 6 pt between the title and the table. Leave 20pt line spacing after the table (before the subsequent text)
\vspace{8pt}
\begin{table}[!ht]
\centering
\caption{Coefficients in constitutive relations}
\label{Table1}
\vspace{6pt}
\begin{tabular}{ccc}
\hline
Constitutive relation & Nomenclature & Value \\ \hline
Turbulent tensor & C$_{\mu}$ & 0.09 \\
Turbulent tensor & C$_{\mu b}$ & 0.69 \\
Lateral lift & C$_{L}$ & 0.08 \\
Virtual mass & C$_{VM}$ & 0.8 \\ \hline
\end{tabular}
\end{table}
\vspace{8pt}
Arabic numerals should be used in figures and tables (e.g., Figure \ref{Fig1}, Figure 2, Table \ref{Table1}, Table 2, etc). Refer to them in the text as ``Table \ref{Table1}'' and ``\cref{Fig1}'' (no quotation marks), except at the beginning of a sentence, wherein ``Figure \ref{Fig1}'' should be used (again, no quotation marks).
\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=1.0]{Figures/Figure1.png}
\vspace{12pt}
\caption{Pressure variation along the nozzle: experimental data}
\label{Fig1}
\end{center}
\end{figure}
\vspace{8pt}
When constructing graphs or plots, do not forget to label coordinates and state the corresponding units. Similarly, label all columns/rows in tables and state corresponding units whenever applicable.

% --------------------------------------------------------------------------
\subsection{Permission}\label{subsec:permission}
% --------------------------------------------------------------------------

You are the sole responsible for making sure that you have the right to publish everything in your paper. If you use material from a copyrighted source, or from other authors, you may need to obtain permission from the copyright holder or the respective authors. An authorship statement must be placed at the end of your paper, immediately before the list of references, as shown further below.

% --------------------------------------------------------------------------
\subsection{References}\label{subsec:references}
% --------------------------------------------------------------------------

References should be cited in the text by Name [number]. For example: ``In a recent work, \citet{Drucker1951} proposed… \citet{Bazant1988}''. Numbers must be Arabic, enclosed in square brackets and used consecutively. References must be listed at the end of the paper, in a separate section entitled ``References''. They must be listed consecutively by the order of citation in the text. Type the word {``References''} in boldface, 13 pt Times New Roman type from the left margin, leaving 18 pt line spacing before and after. Type the reference list below it, in 11 pt Times New Roman (please do not leave blank spaces between references). See examples below. The list should only include works that are cited in the text.

% --------------------------------------------------------------------------
\section{Conclusions}\label{sec:conclusion}
% --------------------------------------------------------------------------

Type your conclusions or closing remarks here. Please be as concise and objective as possible. Do not make a ``summary'' of the paper, but instead list the main findings and results, even if these are only partial conclusions so far.

%-------------------------------------------------------------------------
\vspace{20pt}
\noindent \textbf{Acknowledgements.} This section should be positioned immediately after the Conclusion section. Type {Acknowledgements} in boldface, 10 pt Times New Roman type from left margin, leaving 20 pt line spacing before and 12pt after.
\vspace{12pt}

%--------------------------------------------------------------------------
\noindent \textbf{Authorship statement.} This section is mandatory and should be positioned immediately before the References section. The text should be exactly as follows: The authors hereby confirm that they are the sole liable persons responsible for the authorship of this work, and that all material that has been herein included as part of the present paper is either the property (and authorship) of the authors, or has the permission of the owners to be included here.

\bibliography{bibliography_example}

\end{document}
% --------------------------------------------------------------------------

### 2020-07-13 CILAMCE/paper_EN.tex
Update CILAMCE LOGO

@@
 \flushright
-\includegraphics[width=5.5cm]{logoCILAMCE2019.png}
+\includegraphics[width=5.5cm]{logoCILAMCE2020.png}
 %scale=0.25
@@
 
+\hypertarget{data-preparation}{%
+\subsection{Data preparation}\label{data-preparation}}
+
+The preparation of the data consisted of transforming the sample set in
+order to maintain only relevant data for the subsequent studies. Data
+that were considered irrelevant or that had the potential to add
+undesirable noise to the analysis were removed. In addition, the
+relevant data has been transformed to better fit the studies in the next
+steps.
+
+\hypertarget{initial-data-cleaning}{%
+\subsubsection{Initial data cleaning}\label{initial-data-cleaning}}
+
+Initially, there were 25 duplicate samples that were removed, resulting
+in a new total of 1005 samples.
+
+The data show the variables in the columns and samples in the rows.
+However it was found that some samples are identical in proportions of
+ingredients, changing only the value of age and compressive strength,
+for example, samples 653, 654, 678 and 681, shown in the Table
+\ref{tab:similar-samples}.
+
+\begin{table}[H]
+
+\caption{\label{tab:similar-samples}Samples with same composition}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{cccccccccc}
+\toprule
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
+ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
+\midrule
+653 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 3 & 4.57\\
+\addlinespace
+678 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 7 & 7.68\\
+\addlinespace
+681 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 28 & 17.28\\
+\addlinespace
+654 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 90 & 25.46\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+In addition, there are also samples with the same values and proportions
+of ingredients, but with different compressive strength, probably due to
+differences in the building process. This is the case, for example, of
+samples 472, 473 and 474, shown in the table
+\ref{tab:similar-samples-2}.
+
 \begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

### 2020-07-13 CILAMCE_2020/paper.tex
Update TODO tag
Add intro and materials

@@
 
+Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens.
+Measurement at 28 days is mandatory and represents the grade of the concrete.
+Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.
+
+The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed.
+\citet{Hasan2011} proposed a mathematical model to predict from the results of tests of 7 and 14 days, and \citet{Kabir2012} from 7 days.
+However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.
+
+Many previous studies use the same dataset used by \citet{Yeh1998} to predict the compressive strength of concrete.
+\citet{Alshamiri2020} got good results with the regularized extreme learning machine (RELM) technique, and \citet{Hameed2020} got even better results with the Artificial Neural Networks and cross-validation technique.
+This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as \citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and \citet{Pierobon2018}.
+At the end of the work, the results found are compared to the works cited here.
+
+Unlike previous studies with this dataset, this work does data preparation differently.
+The age of the concrete is the most unique feature that contributes to its compressive strength.
+For this reason, age is treated separately in the machine learning models, creating models for each age group.
+
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
-\subsection{Materials} \label{subsec:materials}
+\subsection{Materials and reproducibility} \label{subsec:materialsandreproducibility}
 % --------------------------------------------------------------------------
 
-% --------------------------------------------------------------------------
-\subsection{Reproducibility} \label{subsec:reproducibility}
-% --------------------------------------------------------------------------
+The methodology was carried out using RStudio software (\citet{RStudio}), an integrated virtual environment for code development in R language (\citet{RCore}).
+Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: CREATE GITHUB REPO BIBLIO AND ADD HERE).
+The repository contains a longer version of this paper, including all the code, required packages and versions.
+In order to guarantee reproducibility, whenever there was code that uses probabilistic operations, a seed was defined before its execution, ensuring results consistency when running on another machine.
 

### 2020-07-14 CILAMCE_2020/paper.tex
Add results and conclusion sections

@@
 % --------------------------------------------------------------------------
-\subsection{Obtaining the data} \label{subsec:Obtainingthedata}
+% \subsection{Obtaining the data} \label{subsec:Obtainingthedata}
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
-\subsection{Data preparation} \label{subsec:datapreparation}
+% \subsection{Data preparation} \label{subsec:datapreparation}
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
-\subsection{Data visualization} \label{subsec:datavisualization}
+% \subsection{Data visualization} \label{subsec:datavisualization}
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
-\subsection{Machine learning models} \label{subsec:machinelearningmodels}
+% \subsection{Machine learning models} \label{subsec:machinelearningmodels}
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
+The test RMSE for each model in ascending order of age was 3.31, 4.36, 4.62, 4.72, 5.94 and 5.85 respectively.
+The table \ref{tab:results} presents the details and results of each model, including the naive one.
+The figure \ref{fig:results} compares the actual and predicted values for the final models.
+
+\vspace{8pt}
+\begin{table}[!ht]
+\centering
+\caption{Final models}
+\label{tab:results}
+\vspace{6pt}
+\begin{tabular}{ccccccc}
+\hline
+Model & mtry & CV & Repetitions & Naive RMSE (test) & Final RMSE (train) & Final RMSE (test)\\ \hline
+3 days & 6 & 30 & 10 & 9.303229 & 3.905196 & 3.310370\\
+7 days & 2 & 10 & 10 & 13.443646 & 4.475981 & 4.361987\\
+14 days & 13 & 30 & 10 & 7.593319 & 5.136687 & 4.620515\\
+28 days & 11 & 30 & 10 & 14.283824 & 5.847334 & 4.716698\\
+56 days & 8 & 30 & 10 & 12.702112 & 6.702565 & 5.939163\\
+100 days & 8 & 10 & 10 & 12.614652 & 6.381940 & 5.851088\\ \hline
+\end{tabular}
+\end{table}
+\vspace{8pt}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.8]{Figures/results.jpg}
+\vspace{12pt}
+\caption{Actual vs Predicted values for each model}
+\label{fig:results}
+\end{center}
+\end{figure}
+\vspace{8pt}
 
@@
 
+The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.The alternative adopted to create a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like all studies cited in the introduction with the same dataset.
+This stratification managed to obtain different results for each age group.
+Although the RMSE of these studies and the found here were close.
+The table \ref{tab:comparison} shows the comparison between these studies and the 28 days model developed here.
+
+\vspace{8pt}
+\begin{table}[!ht]
+\centering
+\caption{Comparison to other works with same dataset}
+\label{tab:comparison}
+\vspace{6pt}
+\begin{tabular}{cccl}
+\hline
+Author & Year & Algorithm & RMSE \\ \hline
+\citet{Pierobon2018} & 2018 & 5 algorithms Ensemble & 4.150\\
+This work & 2020 & Parallel Random Forest & 4.717 (28 day model)\\
+\citet{Hameed2020} & 2020 & Artificial Neural Networks & 4.736\\
+\citet{Raj2018} & 2018 & Gradient Boosting Regressor & 4.957\\
+\citet{Modukuru2020} & 2020 & Random Forest Regressor & 5.080\\
+\citet{Alshamiri2020} & 2020 & Regularized Extreme Learning Machine & 5.508\\
+\citet{Abban2016} & 2016 & Support Vector Machines with Radial Basis Function Kernel & 6.105\\ \hline
+\end{tabular}
+\end{table}
+\vspace{8pt}
+
+Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (Parallel Random Forest), other algorithms can present better results.
+Another option is to create an ensemble of various algorithms, just like \citet{Pierobon2018}, but with the separation of age sets proposed here.
+In addition, it can be performed with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength.
+
 %--------------------------------------------------------------------------
+\vspace{20pt}
 \noindent \textbf{Authorship statement.} The authors hereby confirm that they are the sole liable persons responsible for the authorship of this work, and that all material that has been herein included as part of the present paper is either the property (and authorship) of the authors, or has the permission of the owners to be included here.

### 2020-07-14 CILAMCE_2020/paper.tex
Update table and figure to uppercase

@@
 The test RMSE for each model in ascending order of age was 3.31, 4.36, 4.62, 4.72, 5.94 and 5.85 respectively.
-The table \ref{tab:results} presents the details and results of each model, including the naive one.
-The figure \ref{fig:results} compares the actual and predicted values for the final models.
+The Table \ref{tab:results} presents the details and results of each model, including the naive one.
+The Fig.
+\ref{fig:results} compares the actual and predicted values for the final models.
 
@@
 Although the RMSE of these studies and the found here were close.
-The table \ref{tab:comparison} shows the comparison between these studies and the 28 days model developed here.
+The Table \ref{tab:comparison} shows the comparison between these studies and the 28 days model developed here.
 

### 2020-07-21 CILAMCE_2020/paper.tex
Update results comparison figure
Update paper

--- 2020-07-13 CILAMCE_2020/paper.tex
@@
 \makeatother
-\bibliographystyle{./bib-cilamce}
+\bibliographystyle{bib-cilamce}
 %\bibliographystyle{plain}
@@
 % PLEASE, EDIT THIS!
-\fancyhead[LE]{\footnotesize \textit{Template for CILAMCE 2020 Concrete compressive strength prediction with machine learning}}
+\fancyhead[LE]{\footnotesize \textit{CILAMCE 2020 Concrete compressive strength prediction with machine learning}}
 \fancyhead[RO]{\footnotesize \textit{P. Moreira, V. Silva}}
@@
 
-\noindent \textit{TODO: Address, Zip-Code, State/Province, Country}
+\noindent \textit{Avenida Armando Lombardi 940, 22640-000, Rio de Janeiro, Brazil}
 
-\noindent \textit{TODO: somebody3@somewhere.com}
+\noindent \textit{victor.silva@professores.ibmec.edu.br}
 
@@
 The correct prediction of this parameter means cost and time reduction.
-This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
-A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+This work built predictive models for 6 different ages of concrete samples.
+A set of 1030 samples from previous studies was used, with 9 variables: compressive strength, age, and 7 ingredients.
 Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
@@
 The results were satisfactory and compatible with other studies using the same data set.
-The most important model, 28 days old, obtained RMSE of 4.717.
+The most important model, 28 days old, obtained a root mean square error (RMSE) of 4.717.
 The 3-day model obtained the best result, RMSE of 3.310.
@@
 
-Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens.
+Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens (\citet{Gambhir1990}).
 Measurement at 28 days is mandatory and represents the grade of the concrete.
@@
 
-The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed.
-\citet{Hasan2011} proposed a mathematical model to predict from the results of tests of 7 and 14 days, and \citet{Kabir2012} from 7 days.
-However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.
+The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed (\cite{Hasan2011}, \cite{Kabir2012}).
+However, most of the studies have build models including the age as a feature along with the ingredients, but because of the non linearity between the compressive strength and age, we have find the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
 
-Many previous studies use the same dataset used by \citet{Yeh1998} to predict the compressive strength of concrete.
+Therefore, the present study aims at building predictive models for the concrete compressive strength at different ages using only it's ingredients as features.
+
+% --------------------------------------------------------------------------
+\section{Related work}\label{sec:introduction}
+% --------------------------------------------------------------------------
+
+TODO: Estou com dúvida, se coloco os resultados dos outros trabalhos aqui, porque são apresentados na seção de conclusão e discussão.
+
+TODO: Essa seção ainda tem que ser mais desenvolvida.
+
+TODO: Focar nos que foram publicados/peer review.
+
+% \citet{Hasan2011} proposed a mathematical model to predict 28 day compressive strength from the results of tests of 7 and 14 days, and \citet{Kabir2012} from 7 days.
+These are analytical formulas, but use
+
+The data set used in this study was the one used originally by \citet{Yeh1998}, and then by many othe works.
 \citet{Alshamiri2020} got good results with the regularized extreme learning machine (RELM) technique, and \citet{Hameed2020} got even better results with the Artificial Neural Networks and cross-validation technique.
 This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as \citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and \citet{Pierobon2018}.
-At the end of the work, the results found are compared to the works cited here.
-
-Unlike previous studies with this dataset, this work does data preparation differently.
-The age of the concrete is the most unique feature that contributes to its compressive strength.
-For this reason, age is treated separately in the machine learning models, creating models for each age group.
+Later in this paper all studies are compared with the results found here.
 
@@
 Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: CREATE GITHUB REPO BIBLIO AND ADD HERE).
-The repository contains a longer version of this paper, including all the code, required packages and versions.
+The repository contains an extended version of this paper, including all the code, required packages and versions.
 In order to guarantee reproducibility, whenever there was code that uses probabilistic operations, a seed was defined before its execution, ensuring results consistency when running on another machine.
 
+The main package used to build the machine learning models was the Caret Package (\citet{caret}).
+It provides all functionalities and utilities to build prediction models for any data set, has a straight and clear documentation that guides the process and provide around 200 different algorithms to build models.
+We used only one algorithm in this work, chosen by the highest probability to achieve the best possible result, according to \citet{Fernandez2014}, who compared 179 algorithms across 121 different databases, and find out that the most likely to achieve the best possible results is the Parallel Random Forest (called prRF in the caret).
+
 % --------------------------------------------------------------------------
-\subsection{Obtaining the data} \label{subsec:Obtainingthedata}
+\subsection{TODO ...} \label{subsec:todo}
+% --------------------------------------------------------------------------
+
+...
+
+% --------------------------------------------------------------------------
+% \subsection{Obtaining the data} \label{subsec:Obtainingthedata}
 % --------------------------------------------------------------------------
@@
 % --------------------------------------------------------------------------
+The test RMSE for each model in ascending order of age was 3.31, 4.36, 4.62, 4.72, 5.94 and 5.85 respectively.
+Table \ref{tab:results} presents the details and results of each model, including the naive one.
+Fig.
+\ref{fig:results} compares the actual and predicted values for the final models.
+
+\vspace{8pt}
+\begin{table}[!ht]
+\centering
+\caption{Final models}
+\label{tab:results}
+\vspace{6pt}
+\begin{tabular}{ccccccc}
+\hline
+Model & mtry & CV & Repetitions & Naive RMSE (test) & Final RMSE (train) & Final RMSE (test)\\ \hline
+3 days & 6 & 30 & 10 & 9.303229 & 3.905196 & 3.310370\\
+7 days & 2 & 10 & 10 & 13.443646 & 4.475981 & 4.361987\\
+14 days & 13 & 30 & 10 & 7.593319 & 5.136687 & 4.620515\\
+28 days & 11 & 30 & 10 & 14.283824 & 5.847334 & 4.716698\\
+56 days & 8 & 30 & 10 & 12.702112 & 6.702565 & 5.939163\\
+100 days & 8 & 10 & 10 & 12.614652 & 6.381940 & 5.851088\\ \hline
+\end{tabular}
+\end{table}
+\vspace{8pt}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.8]{Figures/results-comparison.pdf}
+\vspace{12pt}
+\caption{TODO:(O que tem que atualizar aqui?) Actual vs predicted values for each model}
+\label{fig:results}
+\end{center}
+\end{figure}
+\vspace{8pt}
 
@@
 
+The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.The alternative adopted, to create a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like the related studies with the same dataset.
+The adoption of this stratification achieved different results for each age group.
+However, the RMSE calculated in our work and the one obtained in the related works were close.
+Table \ref{tab:comparison} shows the comparison between these studies and the 28 days model developed here.
+
+\vspace{8pt}
+\begin{table}[!ht]
+\centering
+\caption{Comparison to other works with same dataset}
+\label{tab:comparison}
+\vspace{6pt}
+\begin{tabular}{lclcc}
+\hline
+Author & Year & Algorithm & RMSE & Difference ({\%}) \\ \hline
+\citet{Pierobon2018} & 2018 & 5 algorithms Ensemble & 4.150 & -12\\
+\textbf{This work (28 day)} & \textbf{2020} & \textbf{Parallel Random Forest} & \textbf{4.717}  & \textbf{0}\\
+\citet{Hameed2020} & 2020 & Artificial Neural Networks & 4.736 & 0\\
+\citet{Raj2018} & 2018 & Gradient Boosting Regressor & 4.957  & +5\\
+\citet{Modukuru2020} & 2020 & Random Forest Regressor & 5.080  & +8\\
+\citet{Alshamiri2020} & 2020 & Regularized Extreme Learning Machine & 5.508 & +17\\
+\citet{Abban2016} & 2016 & SVM with Radial Basis Function Kernel & 6.105 & +29\\ \hline
+\end{tabular}
+\end{table}
+\vspace{8pt}
+
+Following the line of reasoning of this work, the same hypothesis can be evaluated using other algorithms besides the one used here (Parallel Random Forest), as they can present better results.
+Another option is to create an ensemble of various algorithms, just like \citet{Pierobon2018}, but with the separation of age sets proposed here.
+In addition, this study can be reproduced with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength.
+
 %--------------------------------------------------------------------------
+\vspace{20pt}
 \noindent \textbf{Authorship statement.} The authors hereby confirm that they are the sole liable persons responsible for the authorship of this work, and that all material that has been herein included as part of the present paper is either the property (and authorship) of the authors, or has the permission of the owners to be included here.
+
+\vspace{20pt}
+TODO: Na bibliografia, os ROI não entram por conta do modelo deles, que é inspirado no apa, o mais próximo que posso usar pra colocar o ROI é a chave "howpublished", mas ela não é usada em todas a bibliografias, depende do tipo, por exemplo @article, @inproceedings e @inbook não usam ela, acho que só a @misc usa (para publicacoes online), mas também não adianta porque as publicações online não tem ROI.
+Mas no geral atulizei as referencias e agora estão mais completas
 

### 2020-07-27 CILAMCE_2020/paper.tex
Add related work section

@@
 \noindent \textbf{Abstract.}
-TODO: REDUCE TO 200 WORDS, CURRENTRLY IT HAS 270.
 Compressive strength is the main characteristic of concrete.
@@
 This work built predictive models for 6 different ages of concrete samples.
-A set of 1030 samples from previous studies was used, with 9 variables: compressive strength, age, and 7 ingredients.
-Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
-The predictive models were developed in R language, using the caret package with the Parallel Random Forest algorithm and repeated cross-validation technique to optimize the parameters.
-The results were satisfactory and compatible with other studies using the same data set.
+A set of 1030 samples from previous studies was used, with 9 variables.
+Another 6 variables were added to represent the proportions of the main ingredients in each sample.
+The predictive models were developed in R language, using the Parallel Random Forest algorithm and repeated cross-validation technique to optimize the parameters.
+The results were compatible with other studies using the same data set.
 The most important model, 28 days old, obtained a root mean square error (RMSE) of 4.717.
 The 3-day model obtained the best result, RMSE of 3.310.
-The worst result was the 56-day model, with RMSE of 5.939.
 The work showed that the compressive strength of concrete can be predicted.
-The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age.
+The choice of creating a model for each age allowed to get compatible results with the available data at each age.
 It was a promising alternative since good results were achieved by training with just one algorithm.
@@
 
-\noindent \textbf{Keywords:} Concrete, Compressive Strength, Machine Learning, Prediction
+\noindent \textbf{Keywords:} Concrete, Compressive Strength, Machine Learning, Prediction, Parallel Random Forest
 
@@
 The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed (\cite{Hasan2011}, \cite{Kabir2012}).
-However, most of the studies have build models including the age as a feature along with the ingredients, but because of the non linearity between the compressive strength and age, we have find the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
+However, most of the studies have build models including the age as a feature along with the ingredients, but due to the non linearity between the compressive strength and age, we have find the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
 
@@
 % --------------------------------------------------------------------------
-\section{Related work}\label{sec:introduction}
+\section{Related work}\label{sec:RelatedWork}
 % --------------------------------------------------------------------------
-
-TODO: Estou com dúvida, se coloco os resultados dos outros trabalhos aqui, porque são apresentados na seção de conclusão e discussão.
-
-TODO: Essa seção ainda tem que ser mais desenvolvida.
-
-TODO: Focar nos que foram publicados/peer review.
 
@@
 
-The data set used in this study was the one used originally by \citet{Yeh1998}, and then by many othe works.
-\citet{Alshamiri2020} got good results with the regularized extreme learning machine (RELM) technique, and \citet{Hameed2020} got even better results with the Artificial Neural Networks and cross-validation technique.
-This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as \citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and \citet{Pierobon2018}.
+%  This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as \citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and \citet{Pierobon2018}.
 Later in this paper all studies are compared with the results found here.
+
+\citet{Yeh1998} demonstrated the possibility of using Artificial Neural Networks to predict the compressive strength of concrete, concluding that it is a more accurate method than regression models.
+In this study, more than 1000 concrete samples were collected from 17 different sources.
+This data set was later used in several studies about concrete, some of which are mentioned below.
+
+\citet{Alshamiri2020} proposed a new Regularized Extreme Learning Machine (RELM) method to train Artificial Neural Networks models to predict the compressive strenght.
+The results were compared with several known algorithms running on the same dataset, including individual and essembles, and the proposed model had the best result by far.
+
+\citet{Hameed2020} compares Artificial Neural Network models with Multiple Linear Regression to predict a compressive strength force and have found that Artificial Neural Network models obtain much more accuracy than the Multiple Lenear Regression.
+
+In addition to these published studies, it is now very common to publish side projects on web pages.
+For easy access to this database and the growing interest in data science and machine learning, some unpublished studies using this same database include \citet{Modukuru2020}, \citet{Raj2018}, \citet{Abban2016} and  \citet{Pierobon2018}.
+Overall, they all followed standard steps in the development of machine learning models, the first two using the scikit-learn package in python language developed by \citet{pedregosa2011scikit} and both
+the latter used the caret package developed by \citet{caret} in R language.
+
+At the end of this work, in the discussion and conclusion section, the results found in this work are compared with all these related studies.
 
@@
 \vspace{12pt}
-\caption{TODO:(O que tem que atualizar aqui?) Actual vs predicted values for each model}
+\caption{Actual vs predicted values for each model}
 \label{fig:results}
@@
 
-\vspace{20pt}
-TODO: Na bibliografia, os ROI não entram por conta do modelo deles, que é inspirado no apa, o mais próximo que posso usar pra colocar o ROI é a chave "howpublished", mas ela não é usada em todas a bibliografias, depende do tipo, por exemplo @article, @inproceedings e @inbook não usam ela, acho que só a @misc usa (para publicacoes online), mas também não adianta porque as publicações online não tem ROI.
-Mas no geral atulizei as referencias e agora estão mais completas
-
 \bibliography{bibliography}

### 2020-07-27 CILAMCE_2020/paper.tex
Add data preparation section

@@
 Overall, they all followed standard steps in the development of machine learning models, the first two using the scikit-learn package in python language developed by \citet{pedregosa2011scikit} and both
-the latter used the caret package developed by \citet{caret} in R language.
+the latter used the caret package developed by \citet{caret} in R language \cite{RCore}.
 
@@
 
-The methodology was carried out using RStudio software (\citet{RStudio}), an integrated virtual environment for code development in R language (\citet{RCore}).
+The methodology was carried out using RStudio software \cite{RStudio}, an integrated virtual environment for code development in R language \cite{RCore}.
 Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: CREATE GITHUB REPO BIBLIO AND ADD HERE).
@@
 
-The main package used to build the machine learning models was the Caret Package (\citet{caret}).
+The main package used to build the machine learning models was the Caret Package \cite{caret}.
 It provides all functionalities and utilities to build prediction models for any data set, has a straight and clear documentation that guides the process and provide around 200 different algorithms to build models.
@@
 % --------------------------------------------------------------------------
-\subsection{TODO ...} \label{subsec:todo}
+\subsection{Dataset} \label{subsec:dataset}
+% --------------------------------------------------------------------------
+The data was downloaded from the University of California Irvine website \cite{downloadData}.
+The same dataset used by \citet{Yeh1998} and the related works.
+In total there are 1030 rows with 9 columns.
+Each row represents a sample with the variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+
+TODO: Possivel tabela das primeiras amostras aqui
+
+% --------------------------------------------------------------------------
+\subsection{Data preparation} \label{subsec:datapreparation}
+% --------------------------------------------------------------------------
+
+The related works used the data set in its entirety or performed a minimum of preparation.
+In a different way, in this work a specific step was dedicated just to clean the samples and prepare them for the next steps.
+The major steps executed in this section are listed below:
+
+\begin{enumerate}
+\item Duplicate samples were removed;
+
+\item Samples aggregated and identified (with a new ID column) by the proportion settings of ingredients, independent of its age;
+
+\item Ages of 90, 91, and 100 days were joined, transformed into 100 days.
+Three criteria were used to evaluate this step: First the plot of samples against a boxplot, that showed for example that most samples of 91 days have compressive strength between 60 and 80 MPa, while at the of age 90 days they are mostly between 30 and 50 MPa, but the difference of 1 day can't have such influence in the compressive strength.
+Then, because of the boxplot, a principal component analysis (PCA) of the ingredients were made, showing that the this fact is because the combination of ingredients among samples o 90 and 91 days are very distinct, same happens for 100 days.
+So it is very fair that we can join these ages without any prejudice to the predictions, in fact it can enhance the predictions by reducing noise.
+Finally, it was chosen to be 100 days because a plot of the compressive strength through time for IDs that have at least 5 different ages show that the compressive tends to increase through time, never decreases, as expected.
+
+\item For samples with same ID and the same age, but the different value of compressive strength, the compressive strength was averaged, unifying into only one row for each combination of ID and age;
+
+\item Only the IDs that include the age of 28 days among the ages for that ID were maintained;
+
+\item After joining the ages of 90, 91 and 100 days, the ages with a frequency less than 50 were removed, leaving only the ages of 3, 7, 14, 28, 56 and 100 days;
+
+\item The final number of samples has been reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs);
+
+\item Addition of 6 new continuous variables representing the proportions between the main ingredients that were used in the prediction models (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate and water/fine aggregate);
+
+\item Addition of 2 new categorical variables used to visualize the distribution of the samples (concrete class and approximate mix).
+\end{enumerate}
+
+% --------------------------------------------------------------------------
+\subsection{Data visualization} \label{subsec:datavisualization}
+% --------------------------------------------------------------------------
+
+...
+
+% --------------------------------------------------------------------------
+\subsection{Machine learning models} \label{subsec:machinelearningmodels}
 % --------------------------------------------------------------------------

### 2020-07-27 generated/paper_EN.Rmd
Add preparated data xls

@@
 if(!require(questionr)) install.packages("questionr")
+if(!require(xlsx)) install.packages("xlsx")
+
 # Load libraries
@@
 library(questionr)
+library("xlsx")
+
 # Set global options
@@
 ```
+
 ```{r rename-dat-cols, echo=F}
-# Ranaming the columns
+# Renaming the columns
 colnames(dat) <- c(
@@
 ```
+
 ```{r first-samples, echo=F}
-# Table First samples
+# Table - First samples
 caption <- "First 6 samples"
@@
 ```
+
 The data show the variables in the columns and samples in the rows.
-However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
+However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
 
@@
 ### Age selection
-As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models.
+
+As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models.
 However, it is necessary to verify how relevant the data of these other ages are.
@@
 ```
+
 Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value.
-The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).
+The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-time}).
 
@@
 ```
+
 For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise.
-They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}.
+They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}).
 As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.
@@
 
+```{r dat-preparated, echo=F}
+dat_preparated <- dat
+```
+
 ### Data reorganization
-The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
+
+The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
 The result in the first samples after this processing is shown in the table \ref{tab:new-features}.
@@
 Other proportions between the main ingredients were also added.
-The new variables are presented in the table ref{tab:new-features-table}.
+The new variables are presented in the table \ref{tab:new-features-table}.
+
 ```{r new-features-2, echo=F}
@@
 
+```{r dat-preparated-2, echo=F}
+dat_preparated_2 <- dat
+```
+
+```{r dat-preparated-to-xml, echo=F}
+dat_preparated_2 <- dat
+names(dat_preparated_2) <-  gsub("day_", "", names(dat_preparated_2))
+dat_preparated_2 <- dat_preparated_2 %>%
+  gather("age", "Mpa", "3":"100", na.rm = TRUE, convert = TRUE)
+
+```
+
 ## Data visualization
+
 In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.
@@
 ### Descriptive statistics
-The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}).
+
+The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}).
 The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data.
@@
 ### Variables distribution
-A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
+
 The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
@@
 ### Principal component analysis
-Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}).
-A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso.
-É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
-In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
+
+In the figure \ref{fig:pca}, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
 The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that.
 It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
+
 ```{r pca, echo=F, fig.height=3, fig.cap=cap}
@@
 ## Machine learning models
-The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.
+
+The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.
 
 ### Pre-processing and data separation
-Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado.
+
 As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.
@@
 ```
+
 The samples were separated based on age.
 One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}).
-For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.
+For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-preparated-samples}.
 
@@
 ### Regression models
+
 As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
+
+&nbsp;
+
 1. All *features*;
@@
 The studies cited in the introduction using the same dataset have similar results, as expected.
-The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
+The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}).
+
 ```{r works-comparison, echo=F}

### 2020-07-27 CILAMCE_2020/paper.tex
Update data preparation section

@@
 
-\item The final number of samples has been reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs);
-
 \item Addition of 6 new continuous variables representing the proportions between the main ingredients that were used in the prediction models (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate and water/fine aggregate);
@@
 \end{enumerate}
+
+After these manipulations, the final number of samples has been reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs).
+A xls file named "TODO" of the data at this point is available at the github repo (TODO: referencia bibliografica do link do arquivo no github, que tem o nome de "data\_2.xls", preciso fazer uma pasta com os arquivos e nomear de forma melhor no github - ainda vou fazer uma limpa no github e excluir muita coisa lá).
 
@@
 ...
-
-% --------------------------------------------------------------------------
-% \subsection{Obtaining the data} \label{subsec:Obtainingthedata}
-% --------------------------------------------------------------------------
-
-% --------------------------------------------------------------------------
-% \subsection{Data preparation} \label{subsec:datapreparation}
-% --------------------------------------------------------------------------
-
-% --------------------------------------------------------------------------
-% \subsection{Data visualization} \label{subsec:datavisualization}
-% --------------------------------------------------------------------------
-
-% --------------------------------------------------------------------------
-% \subsection{Machine learning models} \label{subsec:machinelearningmodels}
-% --------------------------------------------------------------------------
 

### 2020-07-27 CILAMCE_2020/paper.tex
New todo
Add visualization section

@@
 Each row represents a sample with the variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
-
-TODO: Possivel tabela das primeiras amostras aqui
+TODO: referencia link do github com o arquivo xls utilizado (os dados originais)
 
@@
 
-...
+Several plots and tables were built to perform the visualization of the samples prepared in the previous step.
+Including analysis of the descriptive statistics of the continuous and categorical variables, the distribution of the variables in relation to the compressive strength, the correlation between the variables grouped by age, the relationship between the approximate mix and the compressive strength, the relationships between the main concrete ingredients and the compressive strength and principal component analysis (PCA) of the ingredients.
+All of these plots/tables are part of the extended version available on github.
+(TODO: reference).
+However, to keep the article short, here only the plots of the statistical analysis of the categorical variables are presented, which are used for an insight in the conclusion of this work.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.8]{Figures/stat-summ-categorical.pdf}
+\vspace{12pt}
+\caption{Descriptive statistics of the categorical variables}
+\label{fig:categorical}
+\end{center}
+\end{figure}
+\vspace{8pt}
 

### 2020-07-28 CILAMCE_2020/paper.tex
Machine learning section and more figures

@@
 The methodology was carried out using RStudio software \cite{RStudio}, an integrated virtual environment for code development in R language \cite{RCore}.
-Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: CREATE GITHUB REPO BIBLIO AND ADD HERE).
+Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: reference).
 The repository contains an extended version of this paper, including all the code, required packages and versions.
 In order to guarantee reproducibility, whenever there was code that uses probabilistic operations, a seed was defined before its execution, ensuring results consistency when running on another machine.
-
-The main package used to build the machine learning models was the Caret Package \cite{caret}.
-It provides all functionalities and utilities to build prediction models for any data set, has a straight and clear documentation that guides the process and provide around 200 different algorithms to build models.
-We used only one algorithm in this work, chosen by the highest probability to achieve the best possible result, according to \citet{Fernandez2014}, who compared 179 algorithms across 121 different databases, and find out that the most likely to achieve the best possible results is the Parallel Random Forest (called prRF in the caret).
 
@@
 Each row represents a sample with the variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
-TODO: referencia link do github com o arquivo xls utilizado (os dados originais)
+TODO: reference
 
@@
 \item Ages of 90, 91, and 100 days were joined, transformed into 100 days.
-Three criteria were used to evaluate this step: First the plot of samples against a boxplot, that showed for example that most samples of 91 days have compressive strength between 60 and 80 MPa, while at the of age 90 days they are mostly between 30 and 50 MPa, but the difference of 1 day can't have such influence in the compressive strength.
-Then, because of the boxplot, a principal component analysis (PCA) of the ingredients were made, showing that the this fact is because the combination of ingredients among samples o 90 and 91 days are very distinct, same happens for 100 days.
+Three criteria were used to evaluate this step: First the plot of samples against a boxplot grouped by age, shown in Fig.
+\ref{fig:boxplot}, that showed for example that most samples of 91 days have compressive strength between 60 and 80 MPa, while at the of age 90 days they are mostly between 30 and 50 MPa, but the difference of 1 day can't have such influence in the compressive strength.
+Then, because of the boxplot, a principal component analysis (PCA) of the ingredients were made in Fig.
+\ref{fig:pca}, showing that the this fact is because the combination of ingredients among samples o 90 and 91 days are very distinct, same happens for 100 days.
 So it is very fair that we can join these ages without any prejudice to the predictions, in fact it can enhance the predictions by reducing noise.
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.7]{Figures/boxplot.pdf}
+\vspace{12pt}
+\caption{Boxplot - compressive strength grouped by age}
+\label{fig:boxplot}
+\end{center}
+\end{figure}
+\vspace{8pt}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.7]{Figures/pca-90-91-100.pdf}
+\vspace{12pt}
+\caption{Principal component analysis - 90, 91 and 100 days}
+\label{fig:pca}
+\end{center}
+\end{figure}
+\vspace{8pt}
+
 After these manipulations, the final number of samples has been reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs).
-A xls file named "TODO" of the data at this point is available at the github repo (TODO: referencia bibliografica do link do arquivo no github, que tem o nome de "data\_2.xls", preciso fazer uma pasta com os arquivos e nomear de forma melhor no github - ainda vou fazer uma limpa no github e excluir muita coisa lá).
+A xls file named "TODO" of the data at this point is available at the github repo (TODO: referencia).
 
@@
 
-Several plots and tables were built to perform the visualization of the samples prepared in the previous step.
+Several plots and tables were built to perform the exploration and visualization of the samples prepared in the previous step.
 Including analysis of the descriptive statistics of the continuous and categorical variables, the distribution of the variables in relation to the compressive strength, the correlation between the variables grouped by age, the relationship between the approximate mix and the compressive strength, the relationships between the main concrete ingredients and the compressive strength and principal component analysis (PCA) of the ingredients.
@@
 (TODO: reference).
-However, to keep the article short, here only the plots of the statistical analysis of the categorical variables are presented, which are used for an insight in the conclusion of this work.
+However, to keep the article short, here only the plots of the statistical analysis of the categorical variables are presented, in the Fig.
+\ref{fig:categorical}, which are used for an insight in the conclusion of this work.
 
@@
 \begin{center}
-\includegraphics[scale=0.8]{Figures/stat-summ-categorical.pdf}
+\includegraphics[scale=0.7]{Figures/stat-summ-categorical.pdf}
 \vspace{12pt}
@@
 % --------------------------------------------------------------------------
+\subsection{Pre-processing and data split} \label{subsec:preprocessing}
+% --------------------------------------------------------------------------
+
+The main package used to build the machine learning models was the Caret Package \cite{caret}.
+It provides all functionalities and utilities to build prediction models for any data set, has a straight and clear documentation that guides the process and provide around 200 different algorithms to build models.
+In this work, it was done key steps described by \citet{irizarry2019} and \citet{Kuhn2008}.
+Starting by some pre-processing steps described below:
+
+\begin{enumerate}
+
+\item Removal of the categorical variables;
+
+\item Separation of the dataset by age, resulting in 6 smaller datasets;
+
+\item Removal of variables with near zero variance (only the fly ash of the 7-day set was removed);
+
+\item Verification of variables with a correlation above 0.999, which did not occur;
+
+\item Each data set was split into test and training sets, 20\% and 80\% respectively, shown in Table \ref{tab:split-config} and the distribution in the Fig. \ref{fig:split-dist}.
+
+\end{enumerate}
+
+\vspace{8pt}
+\begin{table}[!ht]
+\centering
+\caption{Dataset split configurations}
+\label{tab:split-config}
+\vspace{6pt}
+\begin{tabular}{cccc}
+\hline
+Model & Total samples & Train (80\%) & Test (20\%)\\ \hline
+3 days & 121 & 97 & 24\\
+7 days & 114 & 94 & 20\\
+14 days & 62 & 50 & 12\\
+28 days & 416 & 335 & 81\\
+56 days & 83 & 67 & 16\\
+100 days & 120 & 96 & 24\\ \hline
+\end{tabular}
+\end{table}
+\vspace{8pt}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[scale=0.7]{Figures/dist-split.pdf}
+\vspace{12pt}
+\caption{Distribution of the test and train data in relation to compressive strength}
+\label{fig:split-dist}
+\end{center}
+\end{figure}
+\vspace{8pt}
+
+% --------------------------------------------------------------------------
+\subsection{Naive models} \label{subsec:naivemodels}
+% --------------------------------------------------------------------------
+
+Before building the real models, for comparison purposes, naive models were created.
+They simply predict that the compressive strength of the test set is the average compressive strength of the training set.
+In other words, naive models are simply the best guess possible to evaluate how close/far the real model is from a guess.
+
+% --------------------------------------------------------------------------
 \subsection{Machine learning models} \label{subsec:machinelearningmodels}
@@
 
-...
+We used only one algorithm in this work, chosen by the highest probability to achieve the best possible result, according to \citet{Fernandez2014}, who compared 179 algorithms across 121 different databases, and find out that the most likely to achieve the best possible results is the Parallel Random Forest (called prRF in the caret \cite{caret}).
+Six different models were built, one for each age-set, the following steps were made for each one:
+
+\begin{enumerate}
+
+\item Define the resampling scheme, with method of repeated cross validation;
+
+\item Define a tuning grid for the "mtry" tuning parameter, which is a sequence from 1 to the number of columns of each dataset.
+All but the 7-day are equal since only the 7-day set have a column removed in the pre-processing;
+
+\item Set seed equal to "1", chosen arbitrarily to guarantee reproducibility. This seed can be manipulated to obtain more satisfactory results, but it was chosen not to.
+
+\item Do pre-processing transformation of the data with "center" and "scale" methods;
+
+\item Run the caret "train" function with the above configurations and model "parRF";
+
+\end{enumerate}
 
@@
 % --------------------------------------------------------------------------
+The performance evaluation of the models was performed by the Root Mean Square Error (RMSE).
+The RMSE is the measure used in all the works mentioned in the introduction allowing the comparison of the models in the discussion.
+
 The test RMSE for each model in ascending order of age was 3.31, 4.36, 4.62, 4.72, 5.94 and 5.85 respectively.
@@
 \centering
-\caption{Final models}
+\caption{Final models results}
 \label{tab:results}

### 2020-07-28 CILAMCE_2020/paper.tex
Update conclusion

@@
 Another option is to create an ensemble of various algorithms, just like \citet{Pierobon2018}, but with the separation of age sets proposed here.
-In addition, this study can be reproduced with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength.
+In addition, this study can be reproduced with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength and concrete class.
 

### 2020-08-04 CILAMCE_2020/paper.tex
Update

@@
 The methodology was carried out using RStudio software \cite{RStudio}, an integrated virtual environment for code development in R language \cite{RCore}.
-Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository (TODO: reference).
+Throughout the process, all code executed was documented in the same order as its execution and pushed to the github repository \cite{github}.
 The repository contains an extended version of this paper, including all the code, required packages and versions.
@@
 The data was downloaded from the University of California Irvine website \cite{downloadData}.
-The same dataset used by \citet{Yeh1998} and the related works.
 In total there are 1030 rows with 9 columns.
 Each row represents a sample with the variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
-TODO: reference
 
@@
 
-\item Ages of 90, 91, and 100 days were joined, transformed into 100 days.
-Three criteria were used to evaluate this step: First the plot of samples against a boxplot grouped by age, shown in Fig.
-\ref{fig:boxplot}, that showed for example that most samples of 91 days have compressive strength between 60 and 80 MPa, while at the of age 90 days they are mostly between 30 and 50 MPa, but the difference of 1 day can't have such influence in the compressive strength.
-Then, because of the boxplot, a principal component analysis (PCA) of the ingredients were made in Fig.
-\ref{fig:pca}, showing that the this fact is because the combination of ingredients among samples o 90 and 91 days are very distinct, same happens for 100 days.
-So it is very fair that we can join these ages without any prejudice to the predictions, in fact it can enhance the predictions by reducing noise.
-Finally, it was chosen to be 100 days because a plot of the compressive strength through time for IDs that have at least 5 different ages show that the compressive tends to increase through time, never decreases, as expected.
-
-\item For samples with same ID and the same age, but the different value of compressive strength, the compressive strength was averaged, unifying into only one row for each combination of ID and age;
-
-\item Only the IDs that include the age of 28 days among the ages for that ID were maintained;
+\item Ages of 90, 91, and 100 days were merged into a single 100-day category.
+This step was taken with the following development: First, plotting and analysing the boxplot grouped by age in Fig.
+\ref{fig:boxplot} showed that samples of 90, 91 and 100 days have distinct concentrations of compressive strength values.
+Then a principal component analysis (PCA) of the ingredients were made in Fig.
+\ref{fig:pca}, showing that the ingredients combination of these ages are very distinct.
+Due to they are very close ages it is reasonable that we can join these ages without any prejudice to the predictions.
+Finally, the 100-day mark was chosen because the analysis of samples that had data on at least five different ages showed that the compressive strength tends to increase through time.
+Therefore, an older age tends to provide more conservative results.
 
 \item After joining the ages of 90, 91 and 100 days, the ages with a frequency less than 50 were removed, leaving only the ages of 3, 7, 14, 28, 56 and 100 days;
+
+\item For samples with the same ID and the same age, but with different values of compressive strength, the data was combined into a single row for each combination of ID and age, containing an average of the compressive strengths;
+
+\item Only the IDs that have data on the 28-day mark were kept;
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/boxplot.pdf}
+\includegraphics[scale=0.9]{Figures/boxplot.pdf}
 \vspace{12pt}
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/pca-90-91-100.pdf}
+\includegraphics[scale=0.9]{Figures/pca-90-91-100.pdf}
 \vspace{12pt}
@@
 
-After these manipulations, the final number of samples has been reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs).
-A xls file named "TODO" of the data at this point is available at the github repo (TODO: referencia).
+After these manipulations, the final number of samples was reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs).
+A xls file named "TODO" of the data at this point is available at the github repository \cite{github}.
 
@@
 Including analysis of the descriptive statistics of the continuous and categorical variables, the distribution of the variables in relation to the compressive strength, the correlation between the variables grouped by age, the relationship between the approximate mix and the compressive strength, the relationships between the main concrete ingredients and the compressive strength and principal component analysis (PCA) of the ingredients.
-All of these plots/tables are part of the extended version available on github.
-(TODO: reference).
-However, to keep the article short, here only the plots of the statistical analysis of the categorical variables are presented, in the Fig.
+The plots for the statistical analysis of the categorical variables are presented, in the Fig.
 \ref{fig:categorical}, which are used for an insight in the conclusion of this work.
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/stat-summ-categorical.pdf}
+\includegraphics[scale=0.8]{Figures/stat-summ-categorical.pdf}
 \vspace{12pt}
@@
 \begin{center}
-\includegraphics[scale=0.7]{Figures/dist-split.pdf}
+\includegraphics[scale=0.9]{Figures/dist-split.pdf}
 \vspace{12pt}
@@
 \begin{center}
-\includegraphics[scale=0.8]{Figures/results-comparison.pdf}
+\includegraphics[scale=0.9]{Figures/results-comparison.pdf}
 \vspace{12pt}
@@
 
+\pagebreak
+
 % --------------------------------------------------------------------------
@@
 
-The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.The alternative adopted, to create a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like the related studies with the same dataset.
+The built models present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.The alternative adopted, creating a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like the related studies with the same dataset.
 The adoption of this stratification achieved different results for each age group.
@@
 
+\pagebreak
+
 %--------------------------------------------------------------------------

### 2020-08-04 CILAMCE_2020/paper.tex
Update

@@
 
-Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens (\citet{Gambhir1990}).
+Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens \cite{Gambhir1990}.
 Measurement at 28 days is mandatory and represents the grade of the concrete.
@@
 
-The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed (\cite{Hasan2011}, \cite{Kabir2012}).
+The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed by \citet{Hasan2011} and  \citet{Kabir2012}.
 However, most of the studies have build models including the age as a feature along with the ingredients, but due to the non linearity between the compressive strength and age, we have find the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
@@
 
-\citet{Hameed2020} compares Artificial Neural Network models with Multiple Linear Regression to predict a compressive strength force and have found that Artificial Neural Network models obtain much more accuracy than the Multiple Lenear Regression.
+\citet{Hameed2020} compares Artificial Neural Network models with Multiple Linear Regression to predict the compressive strength force and have found that Artificial Neural Network models obtain much more accuracy than the Multiple Linear Regression.
 
@@
 \item Ages of 90, 91, and 100 days were merged into a single 100-day category.
-This step was taken with the following development: First, plotting and analysing the boxplot grouped by age in Fig.
+This step was taken with the following development: first, plotting and analysing the boxplot grouped by age in Fig.
 \ref{fig:boxplot} showed that samples of 90, 91 and 100 days have distinct concentrations of compressive strength values.
@@
 \ref{fig:pca}, showing that the ingredients combination of these ages are very distinct.
-Due to they are very close ages it is reasonable that we can join these ages without any prejudice to the predictions.
+As they are very close ages, it is reasonable that we can join these ages without any prejudice to the predictions.
 Finally, the 100-day mark was chosen because the analysis of samples that had data on at least five different ages showed that the compressive strength tends to increase through time.
@@
 After these manipulations, the final number of samples was reduced from 1030 rows to 916 rows, with 416 different ingredient configurations (IDs).
-A xls file named "TODO" of the data at this point is available at the github repository \cite{github}.
+A xls file of the data at this point is available for download at the github repository \cite{github}.
 
@@
 Including analysis of the descriptive statistics of the continuous and categorical variables, the distribution of the variables in relation to the compressive strength, the correlation between the variables grouped by age, the relationship between the approximate mix and the compressive strength, the relationships between the main concrete ingredients and the compressive strength and principal component analysis (PCA) of the ingredients.
-The plots for the statistical analysis of the categorical variables are presented, in the Fig.
-\ref{fig:categorical}, which are used for an insight in the conclusion of this work.
+
+The plots for the statistical analysis of the categorical variables are presented, in the Fig. \ref{fig:categorical}.
 
@@
 
-We used only one algorithm in this work, chosen by the highest probability to achieve the best possible result, according to \citet{Fernandez2014}, who compared 179 algorithms across 121 different databases, and find out that the most likely to achieve the best possible results is the Parallel Random Forest (called prRF in the caret \cite{caret}).
+We used only one algorithm in this work, chosen by the highest probability to achieve the best possible result, according to \citet{Fernandez2014}, who compared 179 algorithms across 121 different databases, and find out that the most likely to achieve the best possible results is the Parallel Random Forest, called prRF in the Caret Package \cite{caret}.
+
 Six different models were built, one for each age-set, the following steps were made for each one:
@@
 
-The built models present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.The alternative adopted, creating a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like the related studies with the same dataset.
+The built models present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily.
+The alternative adopted, creating a model for each set of age proved to be a valid method, instead of using the age as a predictor along with the ingredients like the related studies with the same dataset.
 The adoption of this stratification achieved different results for each age group.
-However, the RMSE calculated in our work and the one obtained in the related works were close.
+The RMSE calculated in our work and the one obtained in the related works were close.
 Table \ref{tab:comparison} shows the comparison between these studies and the 28 days model developed here.
+
+Following the line of reasoning of this work, the same hypothesis can be evaluated using other algorithms besides the one used here (Parallel Random Forest), as they can present better results.
+Another option is to create an ensemble of various algorithms, just like \citet{Pierobon2018}, but with the separation of age sets proposed here.
+In addition, this study can be reproduced with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength and concrete class, as seen in Fig.
+\ref{fig:categorical} that this dataset is very biased to concrete class between C25 and C35.
 
@@
 
-Following the line of reasoning of this work, the same hypothesis can be evaluated using other algorithms besides the one used here (Parallel Random Forest), as they can present better results.
-Another option is to create an ensemble of various algorithms, just like \citet{Pierobon2018}, but with the separation of age sets proposed here.
-In addition, this study can be reproduced with a larger dataset, ideally with a similar number of samples in each age group and a more homogeneous distribution of compressive strength and concrete class.
-
 \pagebreak

### 2020-08-10 CILAMCE_2020/paper.tex
Minor fixes and make visualiation plots to list

@@
 
-Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens \cite{Gambhir1990}.
-Measurement at 28 days is mandatory and represents the grade of the concrete.
-Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.
+Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of breaking of specimens \cite{Gambhir1990}.
+Measurement when the concrete is 28 days old is mandatory and represents the grade of the concrete.
+Knowing in advance the result for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.
 
 The compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed by \citet{Hasan2011} and  \citet{Kabir2012}.
-However, most of the studies have build models including the age as a feature along with the ingredients, but due to the non linearity between the compressive strength and age, we have find the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
+However, most of the studies have built models including the age as a feature along with the ingredients, but due to the non linearity between the compressive strength and age, we have found the need of further investigation of models that separate the age and analyse only the ingredients, then specify for each age.
 
@@
 
-% \citet{Hasan2011} proposed a mathematical model to predict 28 day compressive strength from the results of tests of 7 and 14 days, and \citet{Kabir2012} from 7 days.
-These are analytical formulas, but use
-
-%  This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as \citet{Abban2016}, \citet{Raj2018}, \citet{Modukuru2020} and \citet{Pierobon2018}.
-Later in this paper all studies are compared with the results found here.
-
 \citet{Yeh1998} demonstrated the possibility of using Artificial Neural Networks to predict the compressive strength of concrete, concluding that it is a more accurate method than regression models.
@@
 \citet{Alshamiri2020} proposed a new Regularized Extreme Learning Machine (RELM) method to train Artificial Neural Networks models to predict the compressive strenght.
-The results were compared with several known algorithms running on the same dataset, including individual and essembles, and the proposed model had the best result by far.
+The results were compared with several known algorithms running on the same dataset, including individual and ensembles, and the proposed model had the best result by far.
 
@@
 
-Several plots and tables were built to perform the exploration and visualization of the samples prepared in the previous step.
-Including analysis of the descriptive statistics of the continuous and categorical variables, the distribution of the variables in relation to the compressive strength, the correlation between the variables grouped by age, the relationship between the approximate mix and the compressive strength, the relationships between the main concrete ingredients and the compressive strength and principal component analysis (PCA) of the ingredients.
+In addition to the already presented plots, several other plots and tables were built to perform the exploration and visualization of the samples prepared in the previous step, including:
+
+\begin{itemize}
+
+\item Analysis of the descriptive statistics of the continuous and categorical variables;
+
+\item Distribution of the variables in relation to the compressive strength;
+
+\item Correlation between the variables grouped by age;
+
+\item Relationship between the approximate mix and the compressive strength;
+
+\item Relationship between the main concrete ingredients and the compressive strength;
+
+\item Principal component analysis (PCA) of the ingredients.
+
+\end{itemize}
 
@@
 
-Six different models were built, one for each age-set, the following steps were made for each one:
+Six different models were built, one for each age-set, and the following steps were made for each one:
 

### 2020-08-14 _meta/generated/paper_EN.Rmd
refactor

--- 2020-07-27 generated/paper_EN.Rmd
@@
 
-```{r dat-preparated-to-xml, echo=F}
+```{r dat-preparated-to-xls, echo=F}
 dat_preparated_2 <- dat
@@
   gather("age", "Mpa", "3":"100", na.rm = TRUE, convert = TRUE)
+
+write.xlsx(dat_preparated_2, "data_2.xls", sheetName = "Sheet1", 
+  col.names = TRUE, row.names = TRUE, append = FALSE)
 

### 2020-08-14 _meta/generated/paper_EN.tex
refactor

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Concrete compressive strength prediction with machine learning},
  pdfauthor={Pedro Bernardino Alves Moreira},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{supertabular}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage[normalem]{ulem}
\usepackage{caption}
\usepackage{floatrow}
\floatsetup[table]{capposition=top}
\floatsetup[figure]{capposition=top}
\captionsetup{options=chunk}
\DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
\renewcommand{\thechunk}{\arabic{chunk}}
\usepackage{indentfirst}
\usepackage{sectsty} \sectionfont{\centering}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Concrete compressive strength prediction with machine learning}
\author{Pedro Bernardino Alves Moreira}
\date{April 22, 2020}

\begin{document}
\maketitle
\begin{abstract}
Compressive strength is the main characteristic of concrete. The correct
prediction of this parameter means cost and time reduction. This work
built predictive models for 6 different ages of concrete samples (3, 7,
14, 28, 56, and 100 days). A set of data obtained in previous studies
was used, a total of 1030 samples, with 9 variables: compressive
strength, age, and 7 ingredients (water, cement, fine aggregate, coarse
aggregate, fly ash, blast furnace slag, and superplasticizers). Another
6 variables were added to represent the proportions of the main
ingredients in each sample (water/cement, fine aggregate/cement, coarse
aggregate/cement, fine aggregate/coarse aggregate, water/coarse
aggregate, and water/fine aggregate). The predictive models were
developed in \emph{R} language, using the \emph{caret} package with the
\emph{Parallel Random Forest} algorithm and repeated cross-validation
technique to optimize the parameters. The results were satisfactory and
compatible with other studies using the same data set. The most
important model, 28 days old, obtained \emph{RMSE} of 4.717. The 3-day
model obtained the best result, \emph{RMSE} of 3.310. The worst result
was the 56-day model, with \emph{RMSE} of 5.939. The work showed that
the compressive strength of concrete can be predicted. The choice of
creating a model for each age, instead of using age as a predictor,
allowed to get compatible results with the available data at each age.
It was a promising alternative since good results were achieved by
training with just one algorithm. This work facilitates exploration and
new efforts to predict the compressive strength of concrete, it can be
replicated using different algorithms or the combination of several.
\end{abstract}

\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Compressive strength is the main characteristic of concrete, measured by
tests of international standards that consist of the breaking of
specimens. Measurement at 28 days is mandatory and represents the grade
of the concrete. Knowing in advance what the result will be obtained for
a given age, based on the proportions of its ingredients, is of great
interest to concrete manufacturers, construction companies, and civil
engineers.

~

This compressive strength is a nonlinear function of its ingredients and
age, making it difficult to establish an analytical formula, although
some formulas have already been proposed. Hasan (2011) proposed a
mathematical model to predict from the results of tests of 7 and 14
days, and Kabir (2012) from 7 days. However, machine learning techniques
can be used to model this characteristic from real sample data, using
only the ingredients.

~

Many previous studies use the same dataset used by Yeh (1998) to predict
the compressive strength of concrete. Alshamiri (2020) got good results
with the regularized extreme learning machine (RELM) technique, and
Hameed (2020) got even better results with the Artificial Neural
Networks and cross-validation technique. This set of samples is so well
known that there are many pages on the internet of unpublished studies
that use it and have good results, such as Abban (2016), Raj (2018),
Modukuru (2020) and Pierobon (2018). At the end of the work, the results
found are compared to the works cited here.

~

Unlike previous studies with this dataset, this work does data
preparation differently. The age of the concrete is the most unique
feature that contributes to its compressive strength. For this reason,
age is treated separately in the machine learning models, creating
models for each age group.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

The methodology was carried out using RStudio software (RStudio Team
2020), an integrated virtual environment for code development in
\emph{R} (R Core Team 2020). Throughout the process, all code executed
was documented in the same order as its execution in
\protect\hyperlink{appendix2}{Appendix 2}, and reference was always made
to codes throughout the text. All relevant information related to the
operating system and installed packages has been presented in
\protect\hyperlink{appendix1}{Appendix 1}.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

In order to guarantee reproducibility, whenever there was some code that
could use probabilistic operations, a \emph{seed} was defined before its
execution, ensuring that when run on another machine, with the same
version of \emph{R} and the same \emph{seed}, get the same result. The
\emph{seeds} can be checked throughout
\protect\hyperlink{appendix2}{Appendix 2}.

\hypertarget{obtaining-the-data}{%
\subsection{Obtaining the data}\label{obtaining-the-data}}

The data was downloaded from the University of California Irvine website
(``Concrete Compressive Strength Data Set'' 2008)
(\ref{show-download-data}). In total there are 1030 samples with 9
columns. The samples were renamed and an id column was added to
facilitate data manipulation (\ref{show-rename-dat-cols}). The columns
were reordered to put the new id column in the first position
(\ref{show-reorder-dat}). The first samples can be viewed in the table
\ref{tab:first-samples}.

~

\begin{table}[H]

\caption{\label{tab:first-samples}First 6 samples}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 28 & 79.99\\
\addlinespace
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 28 & 61.89\\
\addlinespace
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 270 & 40.27\\
\addlinespace
4 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 365 & 41.05\\
\addlinespace
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 360 & 44.30\\
\addlinespace
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 90 & 47.03\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{data-preparation}{%
\subsection{Data preparation}\label{data-preparation}}

The preparation of the data consisted of transforming the sample set in
order to maintain only relevant data for the subsequent studies. Data
that were considered irrelevant or that had the potential to add
undesirable noise to the analysis were removed. In addition, the
relevant data has been transformed to better fit the studies in the next
steps.

\hypertarget{initial-data-cleaning}{%
\subsubsection{Initial data cleaning}\label{initial-data-cleaning}}

Initially, there were 25 duplicate samples that were removed, resulting
in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

~

The data show the variables in the columns and samples in the rows.
However it was found that some samples are identical in proportions of
ingredients, changing only the value of age and compressive strength,
for example, samples 653, 654, 678 and 681, shown in the table
\ref{tab:similar-samples}.

~

\begin{table}[H]

\caption{\label{tab:similar-samples}Samples with same composition}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
653 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 3 & 4.57\\
\addlinespace
678 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 7 & 7.68\\
\addlinespace
681 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 28 & 17.28\\
\addlinespace
654 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 90 & 25.46\\
\bottomrule
\end{tabular}}
\end{table}

In addition, there are also samples with the same values and proportions
of ingredients, but with different compressive strength, probably due to
differences in the building process. This is the case, for example, of
samples 472, 473 and 474, shown in the table
\ref{tab:similar-samples-2}.

~

\begin{table}

\caption{\label{tab:similar-samples-2}Same samples with different results}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 57.03\\
\addlinespace
473 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 44.42\\
\addlinespace
474 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 51.02\\
\bottomrule
\end{tabular}}
\end{table}

To facilitate the analysis of the samples, all samples that are the same
in relation to the ingredients, have been assigned the same \emph{id}.
In addition, as the compressive strength at 28 days is the parameter to
determine the grade of the concrete, only the elements containing that
day among its samples were maintained. In the case of the same samples
but with different results of compressive strength, the values were
averaged. After all these changes (\ref{show-initial-data-cleaning}),
the new total of samples was reduced to 970, containing 416 different
settings for the proportions of ingredients.

~

The result can be seen in the table \ref{tab:similar-samples-same-id}.
All samples with equal ingredient settings have the same id, and when
they had different results for the same days, they were transformed into
just one sample, with the arithmetic mean in the compressive strength.

~

\begin{table}[H]

\caption{\label{tab:similar-samples-same-id}Previous samples after processing}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
\midrule
472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 50.82\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 3 & 4.57\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 7 & 7.68\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 28 & 17.28\\
\addlinespace
653 & 102 & 153 & 0 & 192 & 0.0 & 887 & 942 & 90 & 25.46\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{age-selection}{%
\subsubsection{Age selection}\label{age-selection}}

As previously described, the main age for analysis of compressive
strength is 28 days, but other ages can also be used to build predictive
models. However, it is necessary to verify how relevant the data of
these other ages are. Starting with the distribution of the samples in
relation to each age (\ref{show-boxplot}) shown in the figure
\ref{fig:boxplot}.

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/boxplot-1}

}

\caption{Boxplot - Compressive strength (MPa) vs age (days)}\label{fig:boxplot}
\end{figure}

It was observed that the ages of 90, 91 and 100 days probably represent
extremes to each other in the ingredient configurations, since they are
relatively close ages but with very different values, especially for 90
and 91.

~

This hypothesis was verified using the principal component analysis
method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}).
The figure \ref{fig:pca-90-91-100} shows how the samples relate to each
other (which are similar or different) and revealed how each variable
contributes to the analysis. The first two dimensions represent 37\% and
24\% \$ respectively of the variance.

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/pca-90-91-100-1}

}

\caption{Principal component analysis - 90, 91 e 100 days}\label{fig:pca-90-91-100}
\end{figure}

Another important point considered, of the concrete nature itself, is
the fact that the growth rate of its compressive strength decreases with
time, reaching a certain stability value. The figure
\ref{fig:mpa-on-time} shows the compressive strength over the days for
samples with more than 5 data, that is, data available for at least 6
different ages (\ref{show-mpa-on-time}).

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/mpa-on-time-1}

}

\caption{Compressive strength through time}\label{fig:mpa-on-time}
\end{figure}

For the reasons presented in the figures \ref{fig:boxplot},
\ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered
that the ages of 90, 91 and 100 days can be grouped to improve reading
and decrease sample noise. They were converted to the same value, the
age of 100 days was chosen (\ref{show-join-90-91-100}). As shown in the
figure \ref{fig:mpa-on-time}, the resistance only increases, after 100
days the resistance to compression will be greater than or equal to the
value of 90 or 91 days.

~

Another topic analyzed in the selection of ages was the observed
frequency of each age value after this transformation from 90, 91 in 100
days, shown in the figure \ref{fig:freq-ages}. Some values of days have
very low concentrations of samples, at the risk of damaging more than
helping to create the models, so they were removed
(\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain
only ages with a frequency greater than 50, only the values of 3, 7, 14,
28, 56 and 100 days.

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/freq-ages-1}

}

\caption{Ages frequency}\label{fig:freq-ages}
\end{figure}

\hypertarget{data-reorganization}{%
\subsubsection{Data reorganization}\label{data-reorganization}}

The samples were grouped to maintain only one sample from each set of
configuration of the proportions of the ingredients, adding new
variables/columns for resistance at each age
(\ref{show-reorganizing-dat}). The result in the first samples after
this processing is shown in the table \ref{tab:new-features}.

~

\begin{table}[H]

\caption{\label{tab:new-features}First 6 samples after reorganization}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Ag.} & \multicolumn{1}{c}{Fine Ag.} & \multicolumn{1}{c}{3 days} & \multicolumn{1}{c}{7 days} & \multicolumn{1}{c}{14 days} & \multicolumn{1}{c}{28 days} & \multicolumn{1}{c}{56 days} & \multicolumn{1}{c}{100 days} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 &  &  &  & 79.99 &  & \\
\addlinespace
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 &  &  &  & 61.89 &  & \\
\addlinespace
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 &  & 30.28 &  & 33.02 &  & 37.72\\
\addlinespace
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 9.13 & 14.64 &  & 28.02 &  & 38.07\\
\addlinespace
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 &  &  &  & 45.85 &  & 47.03\\
\addlinespace
7 & 380.0 & 95.0 & 0 & 228 & 0.0 & 932.0 & 594.0 &  & 32.82 &  & 36.45 &  & 40.56\\
\bottomrule
\end{tabular}}
\end{table}

The number of samples and distinct samples after all this manipulation
remained the same, a total of 416 (\ref{show-total-samples-2}).

\hypertarget{adding-new-variables}{%
\subsubsection{Adding new variables}\label{adding-new-variables}}

To finish the data preparation, new columns were added to the dataset
(\ref{show-new-features-2}). Starting with the concrete class, for
example if the compressive strength is between 25 and 30, it receives
the class \emph{C25}. The inclusion of the class was important because
the compressive strength in \emph{MPa} is a continuous variable, which
will be used in the regression models, but the class as a discrete
variable can provide another visualization of the data. The approximate
mix of concrete was also added, which represents the proportions of
aggregates (fine and coarse) for cement. Other proportions between the
main ingredients were also added. The new variables are presented in the
table \ref{tab:new-features-table}.

\begin{table}[H]

\caption{\label{tab:new-features-table}New features}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{c>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}}
\toprule
ID & Class & Approximated Mix & Water / Cement & Fine Ag. / Cement & Coarse Ag. / Cement & Fine Ag. / Coarse Ag. & Water / Coarse Ag. & Water / Fine Ag.\\
\midrule
1 & C75 & 1:1:2 & 0.3000 & 1.2519 & 1.9259 & 0.6500 & 0.1558 & 0.2396\\
\addlinespace
2 & C60 & 1:1:2 & 0.3000 & 1.2519 & 1.9537 & 0.6408 & 0.1536 & 0.2396\\
\addlinespace
3 & C30 & 1:2:3 & 0.6857 & 1.7865 & 2.8030 & 0.6373 & 0.2446 & 0.3838\\
\addlinespace
5 & C25 & 1:4:5 & 0.9668 & 4.1566 & 4.9265 & 0.8437 & 0.1962 & 0.2326\\
\addlinespace
6 & C45 & 1:3:4 & 0.8571 & 2.5188 & 3.5038 & 0.7189 & 0.2446 & 0.3403\\
\addlinespace
7 & C35 & 1:2:2 & 0.6000 & 1.5632 & 2.4526 & 0.6373 & 0.2446 & 0.3838\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{data-visualization}{%
\subsection{Data visualization}\label{data-visualization}}

In order to assess the need for further manipulation before building the
models, in this step the 416 samples already processed were visualized
and analyzed.

\hypertarget{descriptive-statistics}{%
\subsubsection{Descriptive statistics}\label{descriptive-statistics}}

The table \ref{tab:stat-summ} presents the statistical data of the
continuous variables (\ref {show-stat-summ}). The \emph{Null} line
represents the number of zeroed values for the ingredients, and the
\emph{NA} line represents the number of missing data. As the samples
were filtered to maintain only sets of samples with known values of
compressive strength at 28 days, the number of \emph{NAs} is zero for
that age. The figure \ref{fig:stat-summ-categorical} presents the
statistical data of the discrete variables
(\ref{show-stat-summ-categorical}).

~

\begin{table}[H]

\caption{\label{tab:stat-summ}Descriptive statistics - continuous variables}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{1.5cm}cccccccc>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
  & Samples & Null & NA & Min & Max & Range & Sum & Median & Mean & SE mean & CI mean & Variance & Std.Dev. & Coef.Var\\
\midrule
Cement & 416 & 0 & 0 & 102.00 & 540.00 & 438.00 & 109373.10 & 257.70 & 262.92 & 5.10 & 10.02 & 10817.50 & 104.01 & 0.40\\
\addlinespace
B.F.S. & 416 & 174 & 0 & 0.00 & 359.40 & 359.40 & 35824.60 & 94.25 & 86.12 & 4.32 & 8.49 & 7755.00 & 88.06 & 1.02\\
\addlinespace
Fly ash & 416 & 202 & 0 & 0.00 & 200.10 & 200.10 & 26389.00 & 71.25 & 63.44 & 3.26 & 6.40 & 4407.81 & 66.39 & 1.05\\
\addlinespace
Water & 416 & 0 & 0 & 121.80 & 247.00 & 125.20 & 76335.60 & 185.00 & 183.50 & 0.94 & 1.86 & 370.73 & 19.25 & 0.10\\
\addlinespace
Superplast. & 416 & 107 & 0 & 0.00 & 32.20 & 32.20 & 2871.30 & 7.60 & 6.90 & 0.26 & 0.52 & 28.85 & 5.37 & 0.78\\
\addlinespace
Coarse agg. & 416 & 0 & 0 & 801.00 & 1145.00 & 344.00 & 397799.90 & 953.35 & 956.25 & 4.12 & 8.10 & 7063.06 & 84.04 & 0.09\\
\addlinespace
Fine agg. & 416 & 0 & 0 & 594.00 & 992.60 & 398.60 & 317809.80 & 769.65 & 763.97 & 3.59 & 7.06 & 5371.89 & 73.29 & 0.10\\
\addlinespace
3 days & 121 & 0 & 295 & 2.33 & 41.64 & 39.31 & 2210.82 & 15.52 & 18.27 & 0.87 & 1.72 & 91.64 & 9.57 & 0.52\\
\addlinespace
7 days & 114 & 0 & 302 & 7.51 & 59.09 & 51.58 & 2845.52 & 21.06 & 24.96 & 1.29 & 2.55 & 188.81 & 13.74 & 0.55\\
\addlinespace
14 days & 62 & 0 & 354 & 12.84 & 59.76 & 46.92 & 1782.56 & 26.54 & 28.75 & 1.10 & 2.19 & 74.62 & 8.64 & 0.30\\
\addlinespace
28 days & 416 & 0 & 0 & 8.54 & 81.75 & 73.21 & 15101.13 & 33.72 & 36.30 & 0.70 & 1.38 & 206.30 & 14.36 & 0.40\\
\addlinespace
56 days & 83 & 0 & 333 & 23.25 & 80.20 & 56.95 & 4178.77 & 50.77 & 50.35 & 1.52 & 3.02 & 190.82 & 13.81 & 0.27\\
\addlinespace
100 days & 120 & 0 & 296 & 21.86 & 82.60 & 60.74 & 5701.90 & 45.61 & 47.52 & 1.17 & 2.31 & 163.40 & 12.78 & 0.27\\
\addlinespace
Water / Cement & 416 & 0 & 0 & 0.27 & 1.88 & 1.62 & 340.60 & 0.73 & 0.82 & 0.02 & 0.03 & 0.11 & 0.34 & 0.41\\
\addlinespace
Fine agg. / Cement & 416 & 0 & 0 & 1.14 & 9.24 & 8.10 & 1415.82 & 2.94 & 3.40 & 0.07 & 0.14 & 1.96 & 1.40 & 0.41\\
\addlinespace
Coarse agg. / Cement & 416 & 0 & 0 & 1.55 & 8.70 & 7.14 & 1761.06 & 3.67 & 4.23 & 0.08 & 0.16 & 2.70 & 1.64 & 0.39\\
\addlinespace
Fine agg. / Coarse agg. & 416 & 0 & 0 & 0.53 & 1.16 & 0.63 & 335.42 & 0.80 & 0.81 & 0.01 & 0.01 & 0.01 & 0.11 & 0.14\\
\addlinespace
Water / Coarse agg. & 416 & 0 & 0 & 0.12 & 0.29 & 0.17 & 80.66 & 0.19 & 0.19 & 0.00 & 0.00 & 0.00 & 0.03 & 0.16\\
\addlinespace
Water / Fine agg. & 416 & 0 & 0 & 0.13 & 0.38 & 0.26 & 101.26 & 0.24 & 0.24 & 0.00 & 0.00 & 0.00 & 0.04 & 0.17\\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/stat-summ-categorical-1}

}

\caption{Descriptive statistics - categorical variables}\label{fig:stat-summ-categorical}
\end{figure}

\hypertarget{correlation-between-ingredients-and-compressive-strength}{%
\subsubsection{Correlation between ingredients and compressive
strength}\label{correlation-between-ingredients-and-compressive-strength}}

The figure \ref{fig:correlation} shows the correlation of variables for
each set of ages (\ref{show-correlation}). The figure
\ref{fig:correlation-mpa} presents the same data, but instead of
correlating them all, it only correlates with the compressive strength,
showing the values in more detail (\ref{show-correlation-mpa}).

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/correlation-1}

}

\caption{Correlations at each age}\label{fig:correlation}
\end{figure}

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/correlation-mpa-1}

}

\caption{Correlation of variables with compressive strength over time}\label{fig:correlation-mpa}
\end{figure}

The interpretation of the figure \ref{fig:correlation-mpa} suggests that
the strength of the concrete is positively related mainly to the cement
and superplasticizer ingredients and negatively to the water and fine
aggregate. The smaller the amount of cement for aggregates and water,
the more negatively they are correlated with compressive strength.

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/mix-app-mpa-1}

}

\caption{Relationship between approximated mix, water, MPa and age}\label{fig:mix-app-mpa}
\end{figure}

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/mix-mpa-1}

}

\caption{Relationship between concrete main features}\label{fig:mix-mpa}
\end{figure}

The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the
relationship between the main ingredients (known as mix) in relation to
the compressive strength (\ref{show-mix-app-mpa} and
\ref{show-mix-mpa}). The interpretation of these figures shows that the
greater the amount of cement in relation to the other ingredients, the
greater the resistance to compression.

\hypertarget{variables-distribution}{%
\subsubsection{Variables distribution}\label{variables-distribution}}

The figure\ref{fig:vars-distribution} shows the distribution of
variables in the samples (\ref{show-vars-distribution}). It was
calculated using data only at 28 days.

~

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/vars-distribution-1}

}

\caption{Variables distribution}\label{fig:vars-distribution}
\end{figure}

The figure \ref{fig:vars-distribution-time} shows the distribution of
ingredients and compressive strength for each set of ages
(\ref{show-vars-distribution-time}), in case of the 28 days it presents
the same information as the figure \ref{fig:vars-distribution}. It shows
that the resistance to compression gradually increases over time, as
expected. Furthermore it is seen that the concentration of the
ingredients can vary a lot when stratified by ages.

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/vars-distribution-time-1}

}

\caption{Variables distribution grouped by age}\label{fig:vars-distribution-time}
\end{figure}

\hypertarget{principal-component-analysis}{%
\subsubsection{Principal component
analysis}\label{principal-component-analysis}}

In the figure \ref{fig:pca}, using an alternative classification, the
principal component analysis was performed on the ingredients
(\ref{show-pca}). The classification separates concrete into 4 different
compressive strength groups, low up to \emph{20 MPa}, normal up to
\emph{40 MPa}, medium up to \emph{70 MPa} and high above that. It is
possible to notice that the groups overlap, but there is a
differentiation between the high and low group.

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/pca-1}

}

\caption{Principal component analysis on ingredients}\label{fig:pca}
\end{figure}

\hypertarget{machine-learning-models}{%
\subsection{Machine learning models}\label{machine-learning-models}}

The development of machine learning models was carried out with the
\emph{caret} package (Kuhn 2020) and based on Irizarry (2019) and Kuhn
(2008).

\hypertarget{pre-processing-and-data-separation}{%
\subsubsection{Pre-processing and data
separation}\label{pre-processing-and-data-separation}}

As the approximate mix is a categorical variable, it was converted into
dummy variables (\ref{show-dummy-var}), going from 22 columns (id,
class, compressive strength and 19 more \emph{features}) to 45 columns,
an addition of 23 variables, one for each approximate mix.

~

The samples were separated based on age. One dataset was created for
each age value, totaling 6 different sets (\ref{show-preparation}). For
illustrative purposes, the first 18 of 45 columns of the first 6 samples
from the 28-day set are shown in the table
\ref{tab:table-preparated-samples}.

~

\begin{table}[H]

\caption{\label{tab:table-preparated-samples}First 18 columns of the 6 first samples of 28 days}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccccccccccc}
\toprule
\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Agg.} & \multicolumn{1}{c}{Fine Agg.} & \multicolumn{1}{c}{MPa} & \multicolumn{1}{c}{Class} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{C.Agg./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{Wat./} & \multicolumn{2}{c}{App Mix} \\
 & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ &  & Ci. & Ce. & Ce. & C.Agg. & C.Agg. & F.Agg. & 1:1:2 & 1:2:2\\
\midrule
1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 79.99 & C75 & 0.30 & 1.25 & 1.93 & 0.65 & 0.16 & 0.24 & 1 & 0\\
\addlinespace
2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 61.89 & C60 & 0.30 & 1.25 & 1.95 & 0.64 & 0.15 & 0.24 & 1 & 0\\
\addlinespace
3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 33.02 & C30 & 0.69 & 1.79 & 2.80 & 0.64 & 0.24 & 0.38 & 0 & 0\\
\addlinespace
5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 28.02 & C25 & 0.97 & 4.16 & 4.93 & 0.84 & 0.20 & 0.23 & 0 & 0\\
\addlinespace
6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 45.85 & C45 & 0.86 & 2.52 & 3.50 & 0.72 & 0.24 & 0.34 & 0 & 0\\
\addlinespace
7 & 380.0 & 95.0 & 0 & 228 & 0.0 & 932.0 & 594.0 & 36.45 & C35 & 0.60 & 1.56 & 2.45 & 0.64 & 0.24 & 0.38 & 0 & 1\\
\bottomrule
\end{tabular}}
\end{table}

For each of the 6 sets, the existence or not of variables with variance
close to zero and their subsequent removal was verified
(\ref{show-nzv}). Many of the 23 variables added referring to the
approximate concrete mix were removed due to this fact. In addition to
them, in the case of the 7-day set, the fly ash variable was also
removed. Then it was verified that there are no variables with high
correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}).
After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25
columns respectively for ages in increasing sequence.

~

The stage of centralization and normalization of the variables was
carried out later, together with the application of the models, as it is
simpler to do this with the \emph{caret} package. If performed at this
time, it would be necessary to manually undo these transformations in
the predictions. The \emph{caret} allows you to transform before
training the models and already transforms the results back.

~

Each data set was separated into test and training sets, \emph{20\%} and
\emph{80\%} respectively (\ref{show-split}). The figure
\ref{fig:dist-split} shows the distribution of data between the sets in
relation to the compressive strength for each model
(\ref{show-dist-split}).

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/dist-split-1}

}

\caption{Distribution of test and train data}\label{fig:dist-split}
\end{figure}

\hypertarget{performance-measures}{%
\subsubsection{Performance measures}\label{performance-measures}}

The performance evaluation of the models was performed by the Root Mean
Square Error (\emph{RMSE}). The \emph{RMSE} is the measure used in all
the works mentioned in the introduction and will allow the comparison of
the models in the discussion.

\hypertarget{naive-models}{%
\subsubsection{Naive models}\label{naive-models}}

Before creating the real models, for comparison purposes, naive models
were created. They simply predict that the compressive strength of the
test set is the average compressive strength of the training set
(\ref{show-naive-model-reg}). In other words, naive models are simply
the best guess possible. The results can be checked in the table
\ref{tab:table-naive-model-reg}.

\begin{table}[H]

\caption{\label{tab:table-naive-model-reg}Naive models}
\centering
\begin{tabular}[t]{cccc}
\toprule
Age & Mean $MPa$ (train) & RMSE (train) & RMSE (test)\\
\midrule
3 & 18.08887 & 9.591344 & 9.303229\\
\addlinespace
7 & 25.12383 & 13.731569 & 13.443646\\
\addlinespace
14 & 28.63980 & 8.786823 & 7.593319\\
\addlinespace
28 & 36.33605 & 14.361021 & 14.283824\\
\addlinespace
56 & 50.06555 & 13.968077 & 12.702112\\
\addlinespace
100 & 47.57000 & 12.758042 & 12.614652\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{choice-of-algorithms}{%
\subsubsection{Choice of algorithms}\label{choice-of-algorithms}}

The \emph{caret} (Kuhn 2020) package exposes more than 200 different
algorithms for building \emph{machine learning} models. The package
documentation presents an initial code (``Models Clustered by Tag
Similarity'' 2020) as a suggestion to select a portfolio of the most
distinct algorithms possible in relation to some pre-selected algorithm,
but for agility and due to technical limitations, it was chosen to use
an algorithm with the highest probability to achieve the best possible
result. According to Fernandez-Delgado et al. (2014), who compared 179
algorithms in 121 different databases, the algorithm most likely to
achieve the best possible results is the \emph{Parallel Random Forest}
(called \emph{prRF } in \emph{caret}).

\hypertarget{regression-models}{%
\subsubsection{Regression models}\label{regression-models}}

As new variables were added throughout the processing (the relationships
between the ingredients and a few more \emph{dummy vars} for each age
set), 5 possibilities for configuring the \emph{features} for the models
were studied:

~

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  All \emph{features};
\item
  All \emph{features} without the \emph{dummy vars};
\item
  Only the original \emph{features};
\item
  Only new \emph{features};
\item
  Only new \emph{features}, without \emph{dummy vars};
\end{enumerate}

~

Building a model for each of these configurations using the set at 28
days (\ref{show-test-models-reg}), showed that the best option is
configuration 2, the \emph{dummy vars} were completely discarded, but
the other new variables were kept. For illustrative purposes, the
\ref{tab:table-done-samples-28} table shows the first 6 samples from the
28-day model training set. The samples of the other models, of the test
and training sets are similar, the only difference being in the 7-day
model, which excludes fly ash due to the variance close to zero,
performed previously.

\begin{table}[H]

\caption{\label{tab:table-done-samples-28}First 6 samples of train data of the 28 days model}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{cccccccccccccc}
\toprule
\multicolumn{13}{c}{Features} & \multicolumn{1}{c}{Outcome} \\
\cmidrule(l{3pt}r{3pt}){1-13} \cmidrule(l{3pt}r{3pt}){14-14}
\multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Agg.} & \multicolumn{1}{c}{Fine Agg.} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{C.Agg./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{y} \\
$kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & Ce. & Ce. & Ce. & C.Agg. & C.Agg. & F.Agg. & MPa\\
\midrule
540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 0.30 & 1.25 & 1.93 & 0.65 & 0.16 & 0.24 & 79.99\\
\addlinespace
540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 0.30 & 1.25 & 1.95 & 0.64 & 0.15 & 0.24 & 61.89\\
\addlinespace
266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 0.86 & 2.52 & 3.50 & 0.72 & 0.24 & 0.34 & 45.85\\
\addlinespace
380.0 & 95.0 & 0 & 228 & 0.0 & 932.0 & 594.0 & 0.60 & 1.56 & 2.45 & 0.64 & 0.24 & 0.38 & 36.45\\
\addlinespace
475.0 & 0.0 & 0 & 228 & 0.0 & 932.0 & 594.0 & 0.48 & 1.25 & 1.96 & 0.64 & 0.24 & 0.38 & 39.29\\
\addlinespace
198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 0.97 & 4.16 & 4.93 & 0.84 & 0.20 & 0.23 & 28.02\\
\bottomrule
\end{tabular}}
\end{table}

For each age set, a model was created using the \emph{Parallel Random
Forest} algorithm, previously defined (\ref{show-reg-models}). For each
of the 6 models, the parameter \emph{mtry} was optimized, and
\emph{repeated cross-validation } was performed, dividing into 10 or 30
parts and repeating 10 times.

\hypertarget{results}{%
\section{Results}\label{results}}

The test \emph{RMSE} for each model in ascending order of age was 3.31,
4.36, 4.62, 4.72, 5.94 and 5.85 respectively. The table
\ref{tab:table-reg-models} presents the details and results of each
model (\ref{show-table-reg-models}). The figure
\ref{fig:results-comparison} compares the actual and predicted values
(\ref{show-results-comparison}), and the following tables show the best
and worst results for each model (\ref{show-table-10}).

\begin{table}[H]

\caption{\label{tab:table-reg-models}Regression models results}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Model & mtry & CV & Repetitions & RMSE (train) & RMSE (test)\\
\midrule
3 days & 6 & 30 & 10 & 3.905196 & 3.310370\\
\addlinespace
7 days & 2 & 10 & 10 & 4.475981 & 4.361987\\
\addlinespace
14 days & 13 & 30 & 10 & 5.136687 & 4.620515\\
\addlinespace
28 days & 11 & 30 & 10 & 5.847334 & 4.716698\\
\addlinespace
56 days & 8 & 30 & 10 & 6.702565 & 5.939163\\
\addlinespace
100 days & 8 & 10 & 10 & 6.381940 & 5.851088\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics{paper_EN_files/figure-latex/results-comparison-1}

}

\caption{Actual vs Predicted values in each model}\label{fig:results-comparison}
\end{figure}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 3 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
26.06 & 26.297746 & 0.2377457 &  & 41.10 & 32.758262 & -8.341738\\
\addlinespace
23.80 & 23.432367 & -0.3676333 &  & 19.93 & 12.732172 & -7.197828\\
\addlinespace
15.52 & 16.436292 & 0.9162920 &  & 35.30 & 29.781989 & -5.518011\\
\addlinespace
10.76 & 11.830651 & 1.0706507 &  & 9.45 & 14.316742 & 4.866742\\
\addlinespace
32.11 & 30.977619 & -1.1323807 &  & 4.90 & 8.154245 & 3.254245\\
\addlinespace
11.36 & 12.660044 & 1.3000440 &  & 13.57 & 10.464572 & -3.105428\\
\addlinespace
15.62 & 17.079698 & 1.4596980 &  & 24.10 & 27.016321 & 2.916321\\
\addlinespace
24.39 & 26.018261 & 1.6282610 &  & 33.80 & 31.045796 & -2.754204\\
\addlinespace
11.41 & 9.688879 & -1.7211213 &  & 8.49 & 11.196910 & 2.706910\\
\addlinespace
11.85 & 13.788864 & 1.9388640 &  & 14.99 & 17.677379 & 2.687379\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 7 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
17.24 & 17.61947 & 0.3794733 &  & 55.60 & 43.71082 & -11.889177\\
\addlinespace
15.75 & 16.29318 & 0.5431847 &  & 9.01 & 18.32999 & 9.319987\\
\addlinespace
10.09 & 10.67438 & 0.5843837 &  & 45.90 & 39.45277 & -6.447230\\
\addlinespace
15.07 & 14.10010 & -0.9698955 &  & 42.80 & 36.78526 & -6.014737\\
\addlinespace
43.11 & 44.35190 & 1.2419040 &  & 20.42 & 16.03647 & -4.383533\\
\addlinespace
35.10 & 33.69951 & -1.4004926 &  & 12.37 & 15.58829 & 3.218292\\
\addlinespace
11.17 & 12.71080 & 1.5408013 &  & 17.17 & 14.38220 & -2.787795\\
\addlinespace
21.86 & 20.23260 & -1.6274033 &  & 34.57 & 37.26745 & 2.697450\\
\addlinespace
10.79 & 12.85092 & 2.0609233 &  & 21.16 & 18.79017 & -2.369827\\
\addlinespace
23.52 & 21.42631 & -2.0936880 &  & 21.18 & 18.84231 & -2.337694\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 14 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
22.14 & 22.69248 & 0.5524823 &  & 47.71 & 38.87416 & -8.835845\\
\addlinespace
20.77 & 22.36561 & 1.5956093 &  & 34.24 & 25.88502 & -8.354984\\
\addlinespace
24.45 & 22.74412 & -1.7058757 &  & 21.50 & 26.91740 & 5.417403\\
\addlinespace
25.37 & 27.62261 & 2.2526130 &  & 29.75 & 25.02926 & -4.720744\\
\addlinespace
31.81 & 29.53499 & -2.2750130 &  & 33.36 & 37.88168 & 4.521675\\
\addlinespace
23.51 & 26.19615 & 2.6861493 &  & 35.96 & 32.35358 & -3.606417\\
\addlinespace
35.96 & 32.35358 & -3.6064167 &  & 23.51 & 26.19615 & 2.686149\\
\addlinespace
33.36 & 37.88168 & 4.5216750 &  & 31.81 & 29.53499 & -2.275013\\
\addlinespace
29.75 & 25.02926 & -4.7207443 &  & 25.37 & 27.62261 & 2.252613\\
\addlinespace
21.50 & 26.91740 & 5.4174030 &  & 24.45 & 22.74412 & -1.705876\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 28 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
27.83 & 27.86776 & 0.0377573 &  & 78.80 & 61.57401 & -17.225986\\
\addlinespace
43.80 & 43.84951 & 0.0495087 &  & 39.38 & 52.47227 & 13.092267\\
\addlinespace
26.92 & 26.82812 & -0.0918800 &  & 33.04 & 43.81819 & 10.778193\\
\addlinespace
31.87 & 32.06794 & 0.1979360 &  & 50.60 & 39.85258 & -10.747415\\
\addlinespace
19.77 & 19.97517 & 0.2051693 &  & 47.40 & 37.11150 & -10.288501\\
\addlinespace
31.88 & 32.13467 & 0.2546680 &  & 69.84 & 60.88636 & -8.953635\\
\addlinespace
59.00 & 59.36857 & 0.3685727 &  & 31.38 & 40.09934 & 8.719345\\
\addlinespace
23.79 & 24.17376 & 0.3837583 &  & 49.77 & 42.71779 & -7.052213\\
\addlinespace
28.99 & 28.55592 & -0.4340790 &  & 49.77 & 42.72394 & -7.046057\\
\addlinespace
32.24 & 31.79166 & -0.4483390 &  & 25.10 & 32.07616 & 6.976158\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 56 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
35.85 & 36.06413 & 0.2141317 &  & 47.97 & 59.74774 & 11.777736\\
\addlinespace
38.33 & 37.86955 & -0.4604540 &  & 74.36 & 64.08547 & -10.274529\\
\addlinespace
64.90 & 65.81326 & 0.9132646 &  & 55.20 & 64.00472 & 8.804718\\
\addlinespace
47.13 & 48.41224 & 1.2822413 &  & 45.08 & 52.62057 & 7.540568\\
\addlinespace
61.86 & 59.49031 & -2.3696857 &  & 42.03 & 48.63652 & 6.606522\\
\addlinespace
34.20 & 36.72639 & 2.5263850 &  & 73.70 & 67.19523 & -6.504768\\
\addlinespace
53.46 & 50.04835 & -3.4116453 &  & 54.77 & 60.36449 & 5.594488\\
\addlinespace
35.34 & 38.82656 & 3.4865560 &  & 60.20 & 65.67068 & 5.470680\\
\addlinespace
60.20 & 65.67068 & 5.4706798 &  & 35.34 & 38.82656 & 3.486556\\
\addlinespace
54.77 & 60.36449 & 5.5944884 &  & 53.46 & 50.04835 & -3.411645\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table-10}Model of 100 days}
\centering
\begin{tabular}[t]{ccc>{\centering\arraybackslash}p{1cm}ccc}
\toprule
\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
Actual & Predicted & Error &  & Actual & Predicted & Error\\
\midrule
67.80 & 67.53752 & -0.2624793 &  & 66.42 & 50.11266 & -16.307345\\
\addlinespace
55.64 & 55.13289 & -0.5071103 &  & 82.60 & 70.59383 & -12.006168\\
\addlinespace
43.06 & 42.33515 & -0.7248493 &  & 31.35 & 40.69152 & 9.341519\\
\addlinespace
45.84 & 46.89673 & 1.0567320 &  & 59.30 & 51.07318 & -8.226824\\
\addlinespace
47.74 & 46.24704 & -1.4929567 &  & 51.86 & 44.27586 & -7.584144\\
\addlinespace
58.78 & 57.23738 & -1.5426163 &  & 47.22 & 52.72019 & 5.500191\\
\addlinespace
48.85 & 46.67286 & -2.1771403 &  & 32.92 & 38.02416 & 5.104161\\
\addlinespace
44.21 & 46.50377 & 2.2937710 &  & 32.53 & 37.56541 & 5.035406\\
\addlinespace
36.59 & 34.26929 & -2.3207093 &  & 35.17 & 40.18011 & 5.010112\\
\addlinespace
52.96 & 50.52831 & -2.4316947 &  & 32.07 & 27.31722 & -4.752779\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion}}

The models built present satisfactory results and prove that the
compressive strength of concrete can be predicted relatively easily. The
alternative adopted to create a model for each set of age proved to be a
valid method, managing to stratify to obtain specific results for each
set. The studies cited in the introduction using the same dataset have
similar results, as expected. The \ref{tab:works-comparison} table
presents the results of these works (\ref{show-works-comparison}), and
the table \ref{tab:results} presents the values found for easy
comparison (\ref{show-results}).

\begin{table}[H]

\caption{\label{tab:works-comparison}Comparison of other works}
\centering
\begin{tabular}[t]{llll}
\toprule
Author & Year & Algorithm & RMSE\\
\midrule
Pierobon & 2018 & Ensemble com 5 algorítimos & 4.150\\
\addlinespace
Hameed & 2020 & Artificial Neural Networks & 4.736\\
\addlinespace
Raj & 2018 & Gradient Boosting Regressor & 4.957\\
\addlinespace
Modukuru & 2020 & Random Forest Regressor & 5.080\\
\addlinespace
Alshamiri & 2020 & Regularized Extreme Learning Machine & 5.508\\
\addlinespace
Abban & 2016 & Support Vector Machines with Radial Basis Function Kernel & 6.105\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:results}Final result}
\centering
\begin{tabular}[t]{cc}
\toprule
Model & RMSE\\
\midrule
3 days & 3.310370\\
\addlinespace
7 days & 4.361987\\
\addlinespace
14 days & 4.620515\\
\addlinespace
28 days & 4.716698\\
\addlinespace
56 days & 5.939163\\
\addlinespace
100 days & 5.851088\\
\bottomrule
\end{tabular}
\end{table}

Following the line of reasoning of this work, it can be performed with
different algorithms, the results found here used only one
(\emph{Parallel Random Forest}), even though it was theoretically the
``best'' found, other algorithms can present even better results.
Another option is to create an \emph{ensemble} of various algorithms,
just like Pierobon (2018), but with the separation of age sets proposed
here. In addition, it can be performed with a larger dataset, ideally
with the same number of samples in each age set, a more homogeneous
distribution of compressive strength, and less variance between samples.

\hypertarget{literature-cited}{%
\section{Literature cited}\label{literature-cited}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abban2016}{}%
Abban, Daniel. 2016. ``Concrete Compresive Strength.'' October 2016.
\url{https://rpubs.com/brother_abban/220101}.

\leavevmode\hypertarget{ref-Alshamiri2020}{}%
Alshamiri, Tian-Feng e Kim, Abobakr e Yuan. 2020. ``Non-Tuned Machine
Learning Approach for Predicting the Compressive Strength of
High-Performance Concrete.'' \emph{Materials} 13 (February): 1023.
\url{https://doi.org/10.3390/ma13051023}.

\leavevmode\hypertarget{ref-downloadData}{}%
``Concrete Compressive Strength Data Set.'' 2008. University of
California Irvine. March 2008.
\url{https://archive.ics.uci.edu/ml/datasets/concrete+compressive+strength}.

\leavevmode\hypertarget{ref-Fernandez2014}{}%
Fernandez-Delgado, Manuel, E. Cernadas, S. Barro, and Dinani Amorim.
2014. ``Do We Need Hundreds of Classifiers to Solve Real World
Classification Problems?'' \emph{Journal of Machine Learning Research}
15 (October): 3133--81.

\leavevmode\hypertarget{ref-Hameed2020}{}%
Hameed, Mohamed, Mohammed e Khalid. 2020. ``Prediction of Compressive
Strength of High-Performance Concrete: Hybrid Artificial Intelligence
Technique.'' In, 323--35.
\url{https://doi.org/10.1007/978-3-030-38752-5_26}.

\leavevmode\hypertarget{ref-Hasan2011}{}%
Hasan, Ahsanul, Md e Kabir. 2011. ``Prediction of Compressive Strength
of Concrete from Early Age Test Result.'' In.
\url{https://doi.org/10.13140/RG.2.1.3270.7684}.

\leavevmode\hypertarget{ref-irizarry2019}{}%
Irizarry, R. A. 2019. \emph{Introduction to Data Science: Data Analysis
and Prediction Algorithms with R}. Chapman \& Hall/Crc Data Science
Series. CRC Press.
\url{https://books.google.com.br/books?id=xb29DwAAQBAJ}.

\leavevmode\hypertarget{ref-Kabir2012}{}%
Kabir, Md e Miah, Ahsanul e Hasan. 2012. ``Predicting 28 Days
Compressive Strength of Concrete from 7 Days Test Result,'' January,
18--22.
\url{https://www.researchgate.net/publication/258255513_Predicting_28_Days_Compressive_Strength_of_Concrete_from_7_Days_Test_Result}.

\leavevmode\hypertarget{ref-Kuhn2008}{}%
Kuhn, Max. 2008. ``Building Predictive Models in R Using the Caret
Package.'' \emph{Journal of Statistical Software, Articles} 28 (5).
\url{https://doi.org/10.18637/jss.v028.i05}.

\leavevmode\hypertarget{ref-caret}{}%
Kuhn, Max et al. 2020. \emph{Caret: Classification E Regression
Training}.
\url{https://cran.r-project.org/web/packages/caret/index.html}.

\leavevmode\hypertarget{ref-modelsClusters}{}%
``Models Clustered by Tag Similarity.'' 2020. 2020.
\url{http://topepo.github.io/caret/models-clustered-by-tag-similarity.html}.

\leavevmode\hypertarget{ref-Modukuru2020}{}%
Modukuru, Pranay. 2020. ``Concrete Compressive Strength Prediction Using
Machine Learning.'' 2020.
\url{https://towardsdatascience.com/concrete-compressive-strength-prediction-using-machine-learning-4a531b3c43f3}.

\leavevmode\hypertarget{ref-Pierobon2018}{}%
Pierobon, Gabriel. 2018. ``A Comprehensive Machine Learning Workflow
with Multiple Modelling Using Caret and caretEnsemble in R.'' September
2018.
\url{https://towardsdatascience.com/a-comprehensive-machine-learning-workflow-with-multiple-modelling-using-caret-and-caretensemble-in-fcbf6d80b5f2}.

\leavevmode\hypertarget{ref-Raj2018}{}%
Raj, Pavan. 2018. ``Predicting Compressive Strength of Concrete.'' June
2018.
\url{https://www.kaggle.com/pavanraj159/predicting-compressive-strength-of-concrete}.

\leavevmode\hypertarget{ref-RCore}{}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-RStudio}{}%
RStudio Team. 2020. \emph{RStudio: Integrated Development Environment
for R}. Boston, MA: RStudio, Inc. \url{http://www.rstudio.com/}.

\leavevmode\hypertarget{ref-Yeh1998}{}%
Yeh, I-Cheng. 1998. ``Modeling of Strength of High-Performance Concrete
Using Artificial Neural Networks.'' Cement and Concrete Research,
28(12), 1797-1808.'' \emph{Cement and Concrete Research} 28 (December):
1797--1808. \url{https://doi.org/10.1016/S0008-8846(98)00165-3}.

\newpage

\hypertarget{appendix1}{%
\section{Appendix 1 - Virtual environment}\label{appendix1}}

\hypertarget{operational-system}{%
\subsection{Operational system}\label{operational-system}}

\begin{table}[H]
\centering
\begin{tabular}{ll}
\toprule
platform & x86\_64-apple-darwin15.6.0\\
arch & x86\_64\\
os & darwin15.6.0\\
system & x86\_64, darwin15.6.0\\
status & \\
\addlinespace
major & 3\\
minor & 6.2\\
year & 2019\\
month & 12\\
day & 12\\
\addlinespace
svn rev & 77560\\
language & R\\
version.string & R version 3.6.2 (2019-12-12)\\
nickname & Dark and Stormy Night\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{packages}{%
\subsection{Packages}\label{packages}}

\begin{table}[H]
\centering
\begin{tabular}{ll}
\toprule
caret & 6.0.85\\
cowplot & 1.0.0\\
dplyr & 0.8.4\\
factoextra & 1.0.6\\
gdata & 2.18.0\\
\addlinespace
ggplot2 & 3.2.1\\
gridExtra & 2.3\\
kableExtra & 1.1.0\\
knitr & 1.28\\
pastecs & 1.3.21\\
\addlinespace
purrr & 0.3.3\\
questionr & 0.7.0\\
reshape2 & 1.4.3\\
tidyr & 1.0.2\\
tidyverse & 1.3.0\\
\bottomrule
\end{tabular}
\end{table}
\newpage

\hypertarget{appendix2}{%
\section{Appendix 2 - Code}\label{appendix2}}

\hypertarget{obtaining-the-data-1}{%
\subsection{Obtaining the data}\label{obtaining-the-data-1}}

\hypertarget{data-download}{%
\subsubsection{Data download}\label{data-download}}

\label{show-download-data}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Data download}
\NormalTok{url_base <-}\StringTok{ "https://archive.ics.uci.edu"}
\NormalTok{url <-}\StringTok{  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"}
\KeywordTok{download.file}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(url_base, url), }\StringTok{"data.xls"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.xls}\NormalTok{(}\StringTok{"data.xls"}\NormalTok{)}
\NormalTok{n_inicial_samples <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(dat)}
\KeywordTok{colnames}\NormalTok{(dat)}
\end{Highlighting}
\end{Shaded}

\hypertarget{renaming-the-columns}{%
\subsubsection{Renaming the columns}\label{renaming-the-columns}}

\label{show-rename-dat-cols}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Renaming the columns}
\KeywordTok{colnames}\NormalTok{(dat) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
  \StringTok{"cement"}\NormalTok{,}
  \StringTok{"blast_furnace_slag"}\NormalTok{,}
  \StringTok{"fly_ash"}\NormalTok{,}
  \StringTok{"water"}\NormalTok{,}
  \StringTok{"superplasticizers"}\NormalTok{,}
  \StringTok{"coarse_aggregate"}\NormalTok{,}
  \StringTok{"fine_aggregate"}\NormalTok{,}
  \StringTok{"day"}\NormalTok{,}
  \StringTok{"mpa"}
\NormalTok{  )}
\NormalTok{dat}\OperatorTok{$}\NormalTok{id <-}\StringTok{ }\KeywordTok{seq.int}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(dat))}
\end{Highlighting}
\end{Shaded}

\hypertarget{reordering-the-data}{%
\subsubsection{Reordering the data}\label{reordering-the-data}}

\label{show-reorder-dat}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Reordering the data}
\NormalTok{col_order <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
  \StringTok{"id"}\NormalTok{,}
  \StringTok{"cement"}\NormalTok{,}
  \StringTok{"blast_furnace_slag"}\NormalTok{,}
  \StringTok{"fly_ash"}\NormalTok{,}
  \StringTok{"water"}\NormalTok{,}
  \StringTok{"superplasticizers"}\NormalTok{,}
  \StringTok{"coarse_aggregate"}\NormalTok{,}
  \StringTok{"fine_aggregate"}\NormalTok{,}
  \StringTok{"day"}\NormalTok{,}
  \StringTok{"mpa"}
\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[, col_order]}
\end{Highlighting}
\end{Shaded}

\hypertarget{defining-column-names-and-units}{%
\subsubsection{Defining column names and
units}\label{defining-column-names-and-units}}

\label{show-col-names-and-units}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Defining column names and units}
\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
             \StringTok{"Superp."}\NormalTok{, }\StringTok{"C.Aggregate"}\NormalTok{, }\StringTok{"F.Aggregate"}\NormalTok{, }\StringTok{"Day"}\NormalTok{, }\StringTok{"Comp.Str."}\NormalTok{)}
\NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
             \StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{""}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---first-samples}{%
\subsubsection{Table - First samples}\label{table---first-samples}}

\label{show-first-samples}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - First samples}
\NormalTok{caption <-}\StringTok{ "First 6 samples"}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{    dat[}\DecValTok{1}\OperatorTok{:}\DecValTok{6}\NormalTok{,],}
    \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-preparation-1}{%
\subsection{Data preparation}\label{data-preparation-1}}

\hypertarget{removing-duplicated-samples}{%
\subsubsection{Removing duplicated
samples}\label{removing-duplicated-samples}}

\label{show-removing-duplicated-samples}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Removing duplicated samples}
\NormalTok{n_distinct_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{n_distinct}\NormalTok{()}
\NormalTok{n_duplicated_samples <-}\StringTok{ }\NormalTok{n_inicial_samples }\OperatorTok{-}\StringTok{ }\NormalTok{n_distinct_samples}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[}\OperatorTok{!}\KeywordTok{duplicated}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(id))),]}
\NormalTok{n_samples <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(dat)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---samples-with-same-composition}{%
\subsubsection{Table - Samples with same
composition}\label{table---samples-with-same-composition}}

\label{show-similar-samples}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Samples with same composition}
\NormalTok{same_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{653}\NormalTok{, }\DecValTok{678}\NormalTok{, }\DecValTok{654}\NormalTok{, }\DecValTok{681}\NormalTok{))}
\NormalTok{caption <-}\StringTok{ "Samples with same composition"}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{  same_samples[}\KeywordTok{order}\NormalTok{(same_samples}\OperatorTok{$}\NormalTok{day),],}
  \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
  \DataTypeTok{escape =}\NormalTok{ F,}
  \DataTypeTok{booktabs =}\NormalTok{ T,}
  \DataTypeTok{caption =}\NormalTok{ caption,}
  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
  \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
  \DataTypeTok{row.names =} \OtherTok{FALSE}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---same-samples-with-different-results}{%
\subsubsection{Table - Same samples with different
results}\label{table---same-samples-with-different-results}}

\label{show-similar-samples-2}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Same samples with different results}
\NormalTok{same_samples_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{472}\NormalTok{, }\DecValTok{473}\NormalTok{, }\DecValTok{474}\NormalTok{))}
\NormalTok{caption <-}\StringTok{ "Same samples with different results"}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{  same_samples_}\DecValTok{2}\NormalTok{[}\KeywordTok{order}\NormalTok{(same_samples_}\DecValTok{2}\OperatorTok{$}\NormalTok{day),],}
  \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
  \DataTypeTok{escape =}\NormalTok{ F,}
  \DataTypeTok{booktabs =}\NormalTok{ T,}
  \DataTypeTok{caption =}\NormalTok{ caption,}
  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
  \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
  \DataTypeTok{row.names =} \OtherTok{FALSE}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-cleaning}{%
\subsubsection{Data cleaning}\label{data-cleaning}}

\label{show-initial-data-cleaning}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Data cleaning}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(}
\NormalTok{    cement,}
\NormalTok{    blast_furnace_slag,}
\NormalTok{    fly_ash,}
\NormalTok{    water,}
\NormalTok{    superplasticizers,}
\NormalTok{    coarse_aggregate,}
\NormalTok{    fine_aggregate,}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\StringTok{"28"} \OperatorTok{%in%}\StringTok{ }\NormalTok{day) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ id[}\KeywordTok{which.min}\NormalTok{(id)]) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(id, day) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{mpa =} \KeywordTok{mean}\NormalTok{(mpa)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{()}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[}\OperatorTok{!}\KeywordTok{duplicated}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(id))),]}
\NormalTok{dat}\OperatorTok{$}\NormalTok{id<-}\KeywordTok{factor}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id)}
\NormalTok{n_samples <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(dat)}
\NormalTok{n_distinct_samples <-}\StringTok{ }\KeywordTok{n_distinct}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---previous-samples-after-processing}{%
\subsubsection{Table - Previous samples after
processing}\label{table---previous-samples-after-processing}}

\label{show-similar-samples-same-id}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Previous samples after processing}
\NormalTok{same_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{==}\StringTok{ }\DecValTok{653} \OperatorTok{|}\StringTok{ }\NormalTok{id }\OperatorTok{==}\StringTok{ }\DecValTok{472} \OperatorTok{&}\StringTok{ }\NormalTok{day }\OperatorTok{==}\StringTok{ }\DecValTok{28}\NormalTok{)}
\NormalTok{caption <-}\StringTok{ "Previous samples after processing"}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{  same_samples[}\KeywordTok{order}\NormalTok{(same_samples}\OperatorTok{$}\NormalTok{id, same_samples}\OperatorTok{$}\NormalTok{day),],}
  \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
  \DataTypeTok{escape =}\NormalTok{ F,}
  \DataTypeTok{booktabs =}\NormalTok{ T,}
  \DataTypeTok{caption =}\NormalTok{ caption,}
  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
  \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
  \DataTypeTok{row.names =} \OtherTok{FALSE}\NormalTok{,}
  \DataTypeTok{digits =} \DecValTok{2}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---compressive-strength-mpa-vs-age-days}{%
\subsubsection{Figure - Compressive strength (MPa) vs age
(days)}\label{figure---compressive-strength-mpa-vs-age-days}}

\label{show-boxplot}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Compressive strength (MPa) vs age (days)}
\NormalTok{cap <-}\StringTok{ "Boxplot - Compressive strength (MPa) vs age (days)"}
\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{factor}\NormalTok{(day), }\DataTypeTok{y=}\NormalTok{mpa)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha=}\FloatTok{0.2}\NormalTok{)  }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel)}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---principal-component-analysis---90-91-e-100-days}{%
\subsubsection{Figure - Principal component analysis - 90, 91 e 100
days}\label{figure---principal-component-analysis---90-91-e-100-days}}

\label{show-pca-90-91-100}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Principal component analysis - 90, 91 e 100 days}
\NormalTok{dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(day }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{90}\NormalTok{, }\DecValTok{91}\NormalTok{, }\DecValTok{100}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, mpa))}
\NormalTok{cap <-}\StringTok{ "Principal component analysis - 90, 91 e 100 days"}
\KeywordTok{colnames}\NormalTok{(dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
   \StringTok{"Cem."}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly.A."}\NormalTok{,}\StringTok{"Water"}\NormalTok{,}\StringTok{"Sup."}\NormalTok{,}\StringTok{"C.Ag."}\NormalTok{,}\StringTok{"F.Ag."}\NormalTok{,}\StringTok{"day"}\NormalTok{)}
\NormalTok{pca <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(day)), }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{habillage <-}\StringTok{ }\NormalTok{dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\OperatorTok{$}\NormalTok{day}
\KeywordTok{fviz_pca_biplot}\NormalTok{(}
\NormalTok{  pca,}
  \DataTypeTok{geom.ind =} \StringTok{"point"}\NormalTok{,}
  \DataTypeTok{habillage=}\NormalTok{habillage,}
  \DataTypeTok{addEllipses =} \OtherTok{TRUE}\NormalTok{,}
  \DataTypeTok{ellipse.level=}\FloatTok{0.75}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{3}\NormalTok{, }\FloatTok{3.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{-5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---compressive-strength-through-time}{%
\subsubsection{Figure - Compressive strength through
time}\label{figure---compressive-strength-through-time}}

\label{show-mpa-on-time}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Compressive strength through time}
\NormalTok{cap <-}\StringTok{ "Compressive strength through time"}
\NormalTok{ylabel <-}\StringTok{ "MPa"}
\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
\NormalTok{dat_duplicated_only <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(id) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{n}\NormalTok{()}\OperatorTok{>}\DecValTok{5}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(id, mpa, day)}
\NormalTok{dat_duplicated_only }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(day, mpa, }\DataTypeTok{fill=}\NormalTok{id, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_line}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{joining-90-91-and-100-days-data}{%
\subsubsection{Joining 90, 91 and 100 days
data}\label{joining-90-91-and-100-days-data}}

\label{show-join-90-91-100}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Joining 90, 91 and 100 days data}
\NormalTok{ind_}\DecValTok{90}\NormalTok{ <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{id[}\KeywordTok{which}\NormalTok{(dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{==}\StringTok{ "90"}\NormalTok{)]}
\NormalTok{ind_}\DecValTok{91}\NormalTok{ <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{id[}\KeywordTok{which}\NormalTok{(dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{==}\StringTok{ "91"}\NormalTok{)]}
\NormalTok{ind_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{id[}\KeywordTok{which}\NormalTok{(dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{==}\StringTok{ "100"}\NormalTok{)]}
\NormalTok{sum_duplicated <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{duplicated}\NormalTok{(}\KeywordTok{c}\NormalTok{(ind_}\DecValTok{90}\NormalTok{, ind_}\DecValTok{91}\NormalTok{, ind_}\DecValTok{100}\NormalTok{))) }\CommentTok{# 0}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{day =} \KeywordTok{ifelse}\NormalTok{(day }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{91}\NormalTok{, }\DecValTok{90}\NormalTok{), }\DecValTok{100}\NormalTok{, day))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---ages-frequency}{%
\subsubsection{Figure - Ages frequency}\label{figure---ages-frequency}}

\label{show-freq-ages}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Ages frequency}
\NormalTok{cap <-}\StringTok{ "Ages frequency"}
\NormalTok{ylabel <-}\StringTok{ "Frequency"}
\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{factor}\NormalTok{(day))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_bar}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel)}
\end{Highlighting}
\end{Shaded}

\hypertarget{removing-ages-with-frequency-lower-than-50}{%
\subsubsection{Removing ages with frequency lower than
50}\label{removing-ages-with-frequency-lower-than-50}}

\label{show-remove-ages-lower-50}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Removing ages with frequency lower than 50}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{28}\NormalTok{, }\DecValTok{56}\NormalTok{, }\DecValTok{100}\NormalTok{),]}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-reorganization-1}{%
\subsubsection{Data reorganization}\label{data-reorganization-1}}

\label{show-reorganizing-dat}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Data reorganization}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by_at}\NormalTok{(}\KeywordTok{vars}\NormalTok{(}\OperatorTok{-}\NormalTok{mpa)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{row_id =} \DecValTok{1}\OperatorTok{:}\KeywordTok{n}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{spread}\NormalTok{(day, mpa, }\DataTypeTok{sep =} \StringTok{"_"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{row_id)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---first-6-samples-after-reorganization}{%
\subsubsection{Table - First 6 samples after
reorganization}\label{table---first-6-samples-after-reorganization}}

\label{show-new-features}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - First 6 samples after reorganization}
\NormalTok{caption <-}\StringTok{ "First 6 samples after reorganization"}
\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S"}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
             \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Ag."}\NormalTok{, }\StringTok{"Fine Ag."}\NormalTok{, }\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{,}
             \StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
\NormalTok{dfUnits2 <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
              \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}
              \StringTok{"$MPa$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{head}\NormalTok{(dat[}\KeywordTok{order}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id),]),}
    \DataTypeTok{col.names =}\NormalTok{ dfUnits2,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames2, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{total-samples}{%
\subsubsection{Total samples}\label{total-samples}}

\label{show-total-samples-2}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Total samples}
\NormalTok{n_samples <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(dat) }\CommentTok{# 416}
\NormalTok{n_distinct_samples <-}\StringTok{ }\KeywordTok{n_distinct}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id) }\CommentTok{# 416}
\end{Highlighting}
\end{Shaded}

\hypertarget{adding-new-features}{%
\subsubsection{Adding New features}\label{adding-new-features}}

\label{show-new-features-2}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Adding New features}
\NormalTok{concrete_class <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mpa)\{}
  \ControlFlowTok{if}\NormalTok{ (mpa }\OperatorTok{>=}\StringTok{ }\DecValTok{10}\NormalTok{) \{}
\NormalTok{    s <-}\StringTok{ }\KeywordTok{as.character}\NormalTok{(mpa)}
\NormalTok{    first <-}\StringTok{ }\KeywordTok{substr}\NormalTok{(s, }\DataTypeTok{start =} \DecValTok{1}\NormalTok{, }\DataTypeTok{stop =} \DecValTok{1}\NormalTok{)}
\NormalTok{    second <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{substr}\NormalTok{(s, }\DataTypeTok{start =} \DecValTok{2}\NormalTok{, }\DataTypeTok{stop =} \DecValTok{2}\NormalTok{) }\OperatorTok{>=}\StringTok{ }\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{  \}}
  \ControlFlowTok{else}\NormalTok{ \{}
\NormalTok{    first <-}\StringTok{ ""}
\NormalTok{    second <-}\StringTok{ "5"}
\NormalTok{  \}}
  \KeywordTok{paste}\NormalTok{(}\StringTok{"C"}\NormalTok{, first, second, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}
\NormalTok{\}}
\NormalTok{mix <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(c, f_ag, c_ag)\{}
  \KeywordTok{paste}\NormalTok{(}
    \DecValTok{1}\NormalTok{,}
    \KeywordTok{round}\NormalTok{(f_ag}\OperatorTok{/}\NormalTok{c, }\DecValTok{0}\NormalTok{),}
    \KeywordTok{round}\NormalTok{(c_ag}\OperatorTok{/}\NormalTok{c, }\DecValTok{0}\NormalTok{),}
    \DataTypeTok{sep =} \StringTok{":"}\NormalTok{)}
\NormalTok{\}}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{class =} \KeywordTok{sapply}\NormalTok{(day_}\DecValTok{28}\NormalTok{, concrete_class)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{class =} \KeywordTok{as.factor}\NormalTok{(class)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{mix_app =} \KeywordTok{factor}\NormalTok{(}
    \KeywordTok{mix}\NormalTok{(cement, fine_aggregate, coarse_aggregate))) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{water }\OperatorTok{/}\StringTok{ }\NormalTok{cement) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{fine_aggregate}\OperatorTok{/}\NormalTok{cement) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{coarse_aggregate_/_cement}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{coarse_aggregate}\OperatorTok{/}\NormalTok{cement) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{fine_aggregate_/_coarse_aggregate}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{fine_aggregate}\OperatorTok{/}\NormalTok{coarse_aggregate) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{water_/_coarse_aggregate}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{water}\OperatorTok{/}\NormalTok{coarse_aggregate) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{`}\DataTypeTok{water_/_fine_aggregate}\StringTok{`}\NormalTok{ =}\StringTok{ }\NormalTok{water}\OperatorTok{/}\NormalTok{fine_aggregate)}
\NormalTok{lvl <-}\StringTok{ }\KeywordTok{levels}\NormalTok{(dat}\OperatorTok{$}\NormalTok{class)}
\NormalTok{dat}\OperatorTok{$}\NormalTok{class <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(}
\NormalTok{  dat}\OperatorTok{$}\NormalTok{class, }
  \DataTypeTok{levels=}\KeywordTok{c}\NormalTok{( }\StringTok{"C5"}\NormalTok{, }\KeywordTok{sort}\NormalTok{(lvl[lvl}\OperatorTok{!=}\StringTok{"C5"}\NormalTok{], }\DataTypeTok{decreasing=}\NormalTok{F)))}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---new-features}{%
\subsubsection{Table - New features}\label{table---new-features}}

\label{show-new-features-table}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - New features}
\NormalTok{caption <-}\StringTok{ "New features"}
\NormalTok{colNames7 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Class"}\NormalTok{,}\StringTok{"Approximated Mix"}\NormalTok{,}
             \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine Ag. / Cement"}\NormalTok{, }
             \StringTok{"Coarse Ag. / Cement"}\NormalTok{, }\StringTok{"Fine Ag. / Coarse Ag."}\NormalTok{,}
             \StringTok{"Water / Coarse Ag."}\NormalTok{, }\StringTok{"Water / Fine Ag."}\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{head}\NormalTok{(dat[}\KeywordTok{order}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id),][,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{15}\OperatorTok{:}\DecValTok{22}\NormalTok{)]),}
    \DataTypeTok{col.names =}\NormalTok{ colNames7,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
    \DataTypeTok{digits =} \DecValTok{4}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DataTypeTok{width =} \StringTok{"1.5cm"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DataTypeTok{width =} \StringTok{"2cm"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{4}\OperatorTok{:}\DecValTok{9}\NormalTok{, }\DataTypeTok{width =} \StringTok{"1.7cm"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-visualization-1}{%
\subsection{Data visualization}\label{data-visualization-1}}

\hypertarget{table---descriptive-statistics---continuous-variables}{%
\subsubsection{Table - Descriptive statistics - continuous
variables}\label{table---descriptive-statistics---continuous-variables}}

\label{show-stat-summ}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Descriptive statistics - continuous variables}
\NormalTok{summ <-}\StringTok{ }\KeywordTok{t}\NormalTok{(}
  \KeywordTok{stat.desc}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, mix_app))))}
\NormalTok{caption <-}\StringTok{ "Descriptive statistics - continuous variables"}
\KeywordTok{colnames}\NormalTok{(summ) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Samples"}\NormalTok{, }\StringTok{"Null"}\NormalTok{, }\StringTok{"NA"}\NormalTok{, }\StringTok{"Min"}\NormalTok{, }\StringTok{"Max"}\NormalTok{, }\StringTok{"Range"}\NormalTok{,}
              \StringTok{"Sum"}\NormalTok{, }\StringTok{"Median"}\NormalTok{, }\StringTok{"Mean"}\NormalTok{, }\StringTok{"SE mean"}\NormalTok{, }
              \StringTok{"CI mean"}\NormalTok{,}\StringTok{"Variance"}\NormalTok{, }\StringTok{"Std.Dev."}\NormalTok{, }\StringTok{"Coef.Var"}\NormalTok{)}
\KeywordTok{rownames}\NormalTok{(summ) =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
              \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{, }\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{,}
              \StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{,}
              \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
              \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
              \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{  summ,}
  \DataTypeTok{escape =}\NormalTok{ F,}
  \DataTypeTok{booktabs =}\NormalTok{ T,}
  \DataTypeTok{caption =}\NormalTok{ caption,}
  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
  \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
  \DataTypeTok{digits =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{10}\OperatorTok{:}\DecValTok{15}\NormalTok{), }\DataTypeTok{width =} \StringTok{"1.5cm"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---descriptive-statistics---categorical-variables}{%
\subsubsection{Figure - Descriptive statistics - categorical
variables}\label{figure---descriptive-statistics---categorical-variables}}

\label{show-stat-summ-categorical}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Descriptive statistics - categorical variables}
\NormalTok{cap <-}\StringTok{ "Descriptive statistics - categorical variables"}
\NormalTok{name1 <-}\StringTok{ "Percentage"}
\NormalTok{name2 <-}\StringTok{ "Accumulated percentage"}
\NormalTok{ylabel <-}\StringTok{ "Frequency"}
\NormalTok{xlabel1 <-}\StringTok{ "Class"}
\NormalTok{xlabel2 <-}\StringTok{ "Approximate Mix"}
\NormalTok{format_percent =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
  \KeywordTok{paste}\NormalTok{(n, }\StringTok{"%"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}
\NormalTok{\}}
\NormalTok{format_class <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(cls)\{}
  \KeywordTok{str_remove_all}\NormalTok{(cls, }\StringTok{"C"}\NormalTok{)}
\NormalTok{\}}
\NormalTok{f_class <-}\StringTok{ }\KeywordTok{freq}\NormalTok{(dat}\OperatorTok{$}\NormalTok{class, }\DataTypeTok{cum =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{sort =} \StringTok{"dec"}\NormalTok{, }\DataTypeTok{total =}\NormalTok{ F) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(n, }\StringTok{"%"}\NormalTok{, }\StringTok{"%cum"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{class =} \KeywordTok{row.names}\NormalTok{(.)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{class_n =} \KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{format_class}\NormalTok{(class)))}
\NormalTok{f_cls_labels =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n) \{}
\NormalTok{  f_class}\OperatorTok{$}\NormalTok{class[n]}
\NormalTok{\}}
\NormalTok{f_cls_acc_labels =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
  \KeywordTok{paste}\NormalTok{(f_class}\OperatorTok{$}\StringTok{`}\DataTypeTok{%cum}\StringTok{`}\NormalTok{[n], }\StringTok{"%"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}
\NormalTok{\}}
\NormalTok{p1 <-}\StringTok{ }\NormalTok{f_class }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{as.integer}\NormalTok{(}\KeywordTok{reorder}\NormalTok{(class_n, }\OperatorTok{-}\NormalTok{n)), }\DataTypeTok{y =}\NormalTok{ n)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{'identity'}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_y_continuous}\NormalTok{(}
      \DataTypeTok{sec.axis =} \KeywordTok{sec_axis}\NormalTok{(}\OperatorTok{~}\NormalTok{.}\OperatorTok{/}\KeywordTok{length}\NormalTok{(dat}\OperatorTok{$}\NormalTok{class) }\OperatorTok{*}\StringTok{ }\DecValTok{100}\NormalTok{,}
                          \DataTypeTok{name =}\NormalTok{ name1,}
                          \DataTypeTok{labels =}\NormalTok{ format_percent)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ f_cls_labels, }\DataTypeTok{breaks =} \DecValTok{1}\OperatorTok{:}\DecValTok{16}\NormalTok{, }\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{16.5}\NormalTok{),}
                     \DataTypeTok{sec.axis =} \KeywordTok{sec_axis}\NormalTok{(}\OperatorTok{~}\NormalTok{., }\DataTypeTok{breaks =} \DecValTok{1}\OperatorTok{:}\DecValTok{16}\NormalTok{,}
                              \DataTypeTok{name =}\NormalTok{ name2,}
                              \DataTypeTok{labels =}\NormalTok{ f_cls_acc_labels)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.grid.minor.y =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel1) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{()}
\NormalTok{format_mix <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mix)\{}
  \KeywordTok{str_remove_all}\NormalTok{(mix, }\StringTok{":"}\NormalTok{)}
\NormalTok{\}}
\NormalTok{f_mix <-}\StringTok{ }\KeywordTok{freq}\NormalTok{(dat}\OperatorTok{$}\NormalTok{mix_app, }\DataTypeTok{cum =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{sort =} \StringTok{"dec"}\NormalTok{, }\DataTypeTok{total =}\NormalTok{ F) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(n, }\StringTok{"%"}\NormalTok{, }\StringTok{"%cum"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{mix =} \KeywordTok{row.names}\NormalTok{(.)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{mix_n =} \KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{format_mix}\NormalTok{(mix)))}
\NormalTok{f_mix_labels =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n) \{}
\NormalTok{  f_mix}\OperatorTok{$}\NormalTok{mix[n]}
\NormalTok{\}}
\NormalTok{f_mix_acc_labels =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
  \KeywordTok{paste}\NormalTok{(f_mix}\OperatorTok{$}\StringTok{`}\DataTypeTok{%cum}\StringTok{`}\NormalTok{[n], }\StringTok{"%"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}
\NormalTok{\}}
\NormalTok{p2 <-}\StringTok{ }\NormalTok{f_mix }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{as.integer}\NormalTok{(}\KeywordTok{reorder}\NormalTok{(mix_n, }\OperatorTok{-}\NormalTok{n)), }\DataTypeTok{y =}\NormalTok{ n)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{'identity'}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_y_continuous}\NormalTok{(}
      \DataTypeTok{sec.axis =} \KeywordTok{sec_axis}\NormalTok{(}\OperatorTok{~}\NormalTok{.}\OperatorTok{/}\KeywordTok{length}\NormalTok{(dat}\OperatorTok{$}\NormalTok{mix_app) }\OperatorTok{*}\StringTok{ }\DecValTok{100}\NormalTok{,}
                          \DataTypeTok{name =}\NormalTok{ name1,}
                          \DataTypeTok{labels =}\NormalTok{ format_percent)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ f_mix_labels, }\DataTypeTok{breaks =} \DecValTok{1}\OperatorTok{:}\DecValTok{24}\NormalTok{, }\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{24.5}\NormalTok{),}
                     \DataTypeTok{sec.axis =} \KeywordTok{sec_axis}\NormalTok{(}\OperatorTok{~}\NormalTok{., }\DataTypeTok{breaks =} \DecValTok{1}\OperatorTok{:}\DecValTok{24}\NormalTok{,}
                              \DataTypeTok{name =}\NormalTok{ name2,}
                              \DataTypeTok{labels =}\NormalTok{ f_mix_acc_labels)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.grid.minor.y =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel2) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{()}
\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---correlation-grouped-by-age}{%
\subsubsection{Figure - Correlation grouped by
age}\label{figure---correlation-grouped-by-age}}

\label{show-correlation}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Correlation grouped by age}
\NormalTok{cor_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id))}
\NormalTok{cap <-}\StringTok{ "Correlations at each age"}
\NormalTok{f_lvl <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{,}\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
\NormalTok{name <-}\StringTok{ "Correlation"}
\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
                \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
                \StringTok{"3"}\NormalTok{, }\StringTok{"7"}\NormalTok{, }\StringTok{"14"}\NormalTok{, }\StringTok{"28"}\NormalTok{, }\StringTok{"56"}\NormalTok{, }\StringTok{"100"}\NormalTok{,}
                \StringTok{"class"}\NormalTok{, }\StringTok{"mix_app"}\NormalTok{, }\StringTok{"W./C."}\NormalTok{, }\StringTok{"F.A./C."}\NormalTok{, }
                \StringTok{"C.A./C."}\NormalTok{, }\StringTok{"F.A./C.A."}\NormalTok{,}
                \StringTok{"W./C.A."}\NormalTok{, }\StringTok{"W./F.A."}\NormalTok{)}
\KeywordTok{colnames}\NormalTok{(cor_dat) <-}\StringTok{ }\NormalTok{colnames_dat}
\NormalTok{cor_dat <-}\StringTok{ }\NormalTok{cor_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"day"}\NormalTok{, }\StringTok{"mpa"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{, }\StringTok{"7"}\NormalTok{, }\StringTok{"14"}\NormalTok{, }\StringTok{"28"}\NormalTok{, }\StringTok{"56"}\NormalTok{, }\StringTok{"100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{()}
\NormalTok{cor_day <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d)\{}
\NormalTok{  res <-}\StringTok{ }\NormalTok{cor_dat }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{filter}\NormalTok{(day }\OperatorTok{==}\StringTok{ }\NormalTok{d) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(day, class, mix_app)) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{cor}\NormalTok{(.)}
\NormalTok{  res[}\KeywordTok{upper.tri}\NormalTok{(res)] <-}\StringTok{ }\OtherTok{NA}
  \KeywordTok{return}\NormalTok{(res)}
\NormalTok{\}}
\NormalTok{cor_dats <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\KeywordTok{cor_day}\NormalTok{(}\DecValTok{3}\NormalTok{), }\KeywordTok{cor_day}\NormalTok{(}\DecValTok{7}\NormalTok{), }\KeywordTok{cor_day}\NormalTok{(}\DecValTok{14}\NormalTok{),}
               \KeywordTok{cor_day}\NormalTok{(}\DecValTok{28}\NormalTok{), }\KeywordTok{cor_day}\NormalTok{(}\DecValTok{56}\NormalTok{), }\KeywordTok{cor_day}\NormalTok{(}\DecValTok{100}\NormalTok{))}
\NormalTok{melt_day <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(df, d)\{}
\NormalTok{  df }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{melt}\NormalTok{() }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{day =}\NormalTok{ d)}
\NormalTok{\}}
\NormalTok{melt_dats <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{1}\NormalTok{], }\DecValTok{3}\NormalTok{), }\KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{2}\NormalTok{], }\DecValTok{7}\NormalTok{),}
                  \KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{3}\NormalTok{], }\DecValTok{14}\NormalTok{), }\KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{4}\NormalTok{],}\DecValTok{28}\NormalTok{),}
                  \KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{5}\NormalTok{], }\DecValTok{56}\NormalTok{), }\KeywordTok{melt_day}\NormalTok{(cor_dats[}\DecValTok{6}\NormalTok{],}\DecValTok{100}\NormalTok{))}
\NormalTok{melt_dat_final <-}\StringTok{ }\NormalTok{melt_dats }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{reduce}\NormalTok{(rbind) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(value }\OperatorTok{!=}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{melt_dat_final}\OperatorTok{$}\NormalTok{day <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(melt_dat_final}\OperatorTok{$}\NormalTok{day)}
\KeywordTok{levels}\NormalTok{(melt_dat_final}\OperatorTok{$}\NormalTok{day) <-}\StringTok{ }\NormalTok{f_lvl}
\NormalTok{melt_dat_final }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(Var1, }\KeywordTok{desc}\NormalTok{(Var1)), }\DataTypeTok{y=}\NormalTok{Var2, }\DataTypeTok{fill=}\NormalTok{value)) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_tile}\NormalTok{(}\DataTypeTok{color =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{day, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient2}\NormalTok{(}\DataTypeTok{low =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"blue"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"white"}\NormalTok{, }
   \DataTypeTok{midpoint =} \DecValTok{0}\NormalTok{, }\DataTypeTok{limit =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{),}\DataTypeTok{name=}\NormalTok{name, }\DataTypeTok{na.value=}\StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---correlation-over-time}{%
\subsubsection{Figure - Correlation over
time}\label{figure---correlation-over-time}}

\label{show-correlation-mpa}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Correlation over time}
\NormalTok{cap <-}\StringTok{ "Correlation of variables with compressive strength over time"}
\NormalTok{name <-}\StringTok{ "Correlation"}
\NormalTok{melt_dat_final }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(Var1 }\OperatorTok{==}\StringTok{ "mpa"} \OperatorTok{|}\StringTok{ }\NormalTok{Var2 }\OperatorTok{==}\StringTok{ "mpa"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(Var1, }\KeywordTok{desc}\NormalTok{(Var1)), }\DataTypeTok{y=}\NormalTok{Var2, }\DataTypeTok{fill=}\NormalTok{value)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_tile}\NormalTok{(}\DataTypeTok{color =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{day, }\DataTypeTok{ncol=}\DecValTok{6}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient2}\NormalTok{(}\DataTypeTok{low =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"blue"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"white"}\NormalTok{,}
   \DataTypeTok{midpoint =} \DecValTok{0}\NormalTok{, }\DataTypeTok{limit =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{),}\DataTypeTok{name=}\NormalTok{name, }\DataTypeTok{na.value=}\StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =} \KeywordTok{round}\NormalTok{(value, }\DecValTok{2}\NormalTok{)), }\DataTypeTok{size =} \FloatTok{2.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---relationship-between-approximated-mix-water-mpa-and-age}{%
\subsubsection{Figure - Relationship between approximated mix, water,
MPa and
age}\label{figure---relationship-between-approximated-mix-water-mpa-and-age}}

\label{show-mix-app-mpa}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Relationship between approximated mix, water, MPa and age}
\NormalTok{cap <-}\StringTok{ "Relationship between approximated mix, water, MPa and age"}
\NormalTok{d <-}\StringTok{ " days"}
\NormalTok{xlabel <-}\StringTok{ "Approximated mix"} 
\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
\NormalTok{label <-}\StringTok{ "Water /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
\NormalTok{mix_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{c}\NormalTok{(day_}\DecValTok{3}\NormalTok{,day_}\DecValTok{7}\NormalTok{,day_}\DecValTok{14}\NormalTok{,day_}\DecValTok{28}\NormalTok{,day_}\DecValTok{56}\NormalTok{,day_}\DecValTok{100}\NormalTok{,}
\NormalTok{          mix_app, }\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{, }
          \StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{, }
          \StringTok{`}\DataTypeTok{coarse_aggregate_/_cement}\StringTok{`}\NormalTok{))}

\NormalTok{labs <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{), d, }\DataTypeTok{sep=}\StringTok{""}\NormalTok{)}
\NormalTok{mix_dat <-}\StringTok{ }\NormalTok{mix_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"day"}\NormalTok{, }\StringTok{"mpa"}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mix_app, }\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{,}
                          \StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{,}
                          \StringTok{`}\DataTypeTok{coarse_aggregate_/_cement}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{()}
\NormalTok{lvls <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\StringTok{"day_"}\NormalTok{,}\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{), }\DataTypeTok{sep=}\StringTok{""}\NormalTok{)}
\NormalTok{mix_dat}\OperatorTok{$}\NormalTok{day <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(mix_dat}\OperatorTok{$}\NormalTok{day, }\DataTypeTok{levels=}\NormalTok{lvls)}
\KeywordTok{levels}\NormalTok{(mix_dat}\OperatorTok{$}\NormalTok{day) <-}\StringTok{ }\NormalTok{labs}
\NormalTok{min_x <-}\StringTok{ }\KeywordTok{min}\NormalTok{(mix_dat}\OperatorTok{$}\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{)}
\NormalTok{max_x <-}\StringTok{ }\KeywordTok{max}\NormalTok{(mix_dat}\OperatorTok{$}\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{)}
\NormalTok{s_x <-}\StringTok{ }\NormalTok{max_x }\OperatorTok{-}\StringTok{ }\NormalTok{min_x}
\NormalTok{mix_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{mix_app, }\DataTypeTok{y=}\NormalTok{mpa, }\DataTypeTok{colour =} \StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{()  }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{day, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_shape_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{8}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_colour_gradient2}\NormalTok{(}\DataTypeTok{low =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"yellow"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"blue"}\NormalTok{,}
                         \DataTypeTok{midpoint =}\NormalTok{ s_x }\OperatorTok{/}\StringTok{ }\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{min_x ,}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(min_x, max_x),}
                         \DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\KeywordTok{round}\NormalTok{(min_x, }\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(s_x}\OperatorTok{*}\FloatTok{0.25} \OperatorTok{+}\StringTok{ }\NormalTok{min_x,}\DecValTok{2}\NormalTok{), }
                                    \KeywordTok{round}\NormalTok{(s_x}\OperatorTok{*}\FloatTok{0.5} \OperatorTok{+}\StringTok{ }\NormalTok{min_x,}\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(s_x }\OperatorTok{*}\StringTok{ }\FloatTok{0.75} \OperatorTok{+}\StringTok{ }\NormalTok{min_x,}\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(max_x, }\DecValTok{2}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{colour =}\NormalTok{ label) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(),}
      \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---relationship-between-concrete-main-features}{%
\subsubsection{Figure - Relationship between concrete main
features}\label{figure---relationship-between-concrete-main-features}}

\label{show-mix-mpa}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Relationship between concrete main features}
\NormalTok{cap <-}\StringTok{ "Relationship between concrete main features"}
\NormalTok{d <-}\StringTok{ " days"}
\NormalTok{xlabel <-}\StringTok{ "Water / Cement"}
\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
\NormalTok{colour <-}\StringTok{ "Fine Agg. /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
\NormalTok{size <-}\StringTok{ "Coarse Agg. /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
\NormalTok{min_x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{min}\NormalTok{(mix_dat}\OperatorTok{$}\StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{)}
\NormalTok{max_x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{max}\NormalTok{(mix_dat}\OperatorTok{$}\StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{)}
\NormalTok{s_x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{max_x_}\DecValTok{2} \OperatorTok{-}\StringTok{ }\NormalTok{min_x_}\DecValTok{2}
\NormalTok{mix_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\StringTok{`}\DataTypeTok{water_/_cement}\StringTok{`}\NormalTok{, }\DataTypeTok{y=}\NormalTok{mpa,}
             \DataTypeTok{colour =} \StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{,}
             \DataTypeTok{size =} \StringTok{`}\DataTypeTok{coarse_aggregate_/_cement}\StringTok{`}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{)  }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{day, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_colour_gradient2}\NormalTok{(}\DataTypeTok{low =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"yellow"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"blue"}\NormalTok{,}
                         \DataTypeTok{midpoint =}\NormalTok{ (s_x_}\DecValTok{2} \OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\NormalTok{min_x_}\DecValTok{2}\NormalTok{, }
                         \DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(min_x_}\DecValTok{2}\NormalTok{, max_x_}\DecValTok{2}\NormalTok{),}
                         \DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\KeywordTok{round}\NormalTok{(min_x_}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(s_x_}\DecValTok{2}\OperatorTok{*}\FloatTok{0.25} \OperatorTok{+}\StringTok{ }\NormalTok{min_x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(s_x_}\DecValTok{2}\OperatorTok{*}\FloatTok{0.5} \OperatorTok{+}\StringTok{ }\NormalTok{min_x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(s_x_}\DecValTok{2} \OperatorTok{*}\StringTok{ }\FloatTok{0.75} \OperatorTok{+}\StringTok{ }\NormalTok{min_x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{),}
                                    \KeywordTok{round}\NormalTok{(max_x_}\DecValTok{2} \OperatorTok{-}\StringTok{ }\FloatTok{0.01}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\NormalTok{                         ) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{colour =}\NormalTok{ colour, }\DataTypeTok{size =}\NormalTok{ size) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{5}\NormalTok{,}\DecValTok{85}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_radius}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(),}
      \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---variables-distribution}{%
\subsubsection{Figure - Variables
distribution}\label{figure---variables-distribution}}

\label{show-vars-distribution}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Variables distribution}

\NormalTok{cap <-}\StringTok{ "Variables distribution"}
\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
                  \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
                  \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
                  \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
                  \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
\NormalTok{dist_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, day_}\DecValTok{3}\NormalTok{, day_}\DecValTok{7}\NormalTok{, day_}\DecValTok{14}\NormalTok{, day_}\DecValTok{56}\NormalTok{, day_}\DecValTok{100}\NormalTok{, mix_app))}

\KeywordTok{colnames}\NormalTok{(dist_dat) <-}\StringTok{ }\NormalTok{colnames_dat}
\NormalTok{dist_dat <-}\StringTok{ }\NormalTok{dist_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"Var"}\NormalTok{, }\StringTok{"value"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{value =} \KeywordTok{as.numeric}\NormalTok{(value))}
\NormalTok{dist_dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(value)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ ..density..), }
                 \DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }
                 \DataTypeTok{fill =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{.5}\NormalTok{, }\DataTypeTok{fill =} \StringTok{"lightseagreen"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{Var, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{, }\DataTypeTok{scale =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---variables-distribution-grouped-by-age}{%
\subsubsection{Figure - Variables distribution grouped by
age}\label{figure---variables-distribution-grouped-by-age}}

\label{show-vars-distribution-time}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Variables distribution grouped by age}

\NormalTok{days_labs <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{)}
\NormalTok{cap <-}\StringTok{ "Variables distribution grouped by age"}
\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
                  \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
\NormalTok{                  days_labs,}
                  \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
                  \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
                  \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
\NormalTok{dist_dat_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, mix_app))}
\KeywordTok{colnames}\NormalTok{(dist_dat_}\DecValTok{2}\NormalTok{) <-}\StringTok{ }\NormalTok{colnames_dat}
\NormalTok{dist_dat_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dist_dat_}\DecValTok{2} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"day"}\NormalTok{, }\StringTok{"mpa"}\NormalTok{, days_labs) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"Var"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{day =} \KeywordTok{factor}\NormalTok{(day, }\DataTypeTok{levels =}\NormalTok{ days_labs))}
\NormalTok{dist_dat_}\DecValTok{2} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ day, }\DataTypeTok{y =}\NormalTok{ value)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_violin}\NormalTok{(}\DataTypeTok{color =} \OtherTok{NA}\NormalTok{,}
                \DataTypeTok{fill =} \StringTok{"lightseagreen"}\NormalTok{,}
                \DataTypeTok{alpha =} \FloatTok{.5}\NormalTok{,}
                \DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{,}
                \DataTypeTok{scale =} \StringTok{"count"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{Var, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{, }\DataTypeTok{scale =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---principal-component-analysis-on-ingredients}{%
\subsubsection{Figure - Principal component analysis on
ingredients}\label{figure---principal-component-analysis-on-ingredients}}

\label{show-pca}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Principal component analysis on ingredients}

\NormalTok{cap <-}\StringTok{ "Principal component analysis on ingredients"}
\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
  \StringTok{"Cem."}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"F.A."}\NormalTok{, }\StringTok{"Wat."}\NormalTok{, }\StringTok{"Sup."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{,}
  \StringTok{"F.Agg."}\NormalTok{,}\StringTok{"WxC"}\NormalTok{, }\StringTok{"F.Agg.xC"}\NormalTok{,}\StringTok{"C.Agg.xC"}\NormalTok{,}\StringTok{"F.Agg.xC"}\NormalTok{,}
  \StringTok{"WxC.Agg."}\NormalTok{,}\StringTok{"WxF.Agg."}
\NormalTok{)}
\NormalTok{class_list <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{low=}\StringTok{"0"}\NormalTok{, }\DataTypeTok{normal=}\StringTok{"20"}\NormalTok{, }\DataTypeTok{medium=}\StringTok{"40"}\NormalTok{, }\DataTypeTok{high=}\StringTok{"70"}\NormalTok{)}
\NormalTok{dat_pca <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, mix_app,}
            \StringTok{"day_3"}\NormalTok{, }\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{))}
\KeywordTok{colnames}\NormalTok{(dat_pca) <-}\StringTok{ }\NormalTok{colnames_dat}
\NormalTok{class_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{day_}\DecValTok{28}
\NormalTok{class_}\DecValTok{2}\NormalTok{[class_}\DecValTok{2} \OperatorTok{<}\StringTok{ }\DecValTok{20}\NormalTok{] =}\StringTok{ "0"}
\NormalTok{class_}\DecValTok{2}\NormalTok{[class_}\DecValTok{2} \OperatorTok{>=}\StringTok{ }\DecValTok{20} \OperatorTok{&}\StringTok{ }\NormalTok{class_}\DecValTok{2} \OperatorTok{<}\StringTok{ }\DecValTok{40}\NormalTok{] =}\StringTok{ "20"}
\NormalTok{class_}\DecValTok{2}\NormalTok{[class_}\DecValTok{2} \OperatorTok{>=}\StringTok{ }\DecValTok{40} \OperatorTok{&}\StringTok{ }\NormalTok{class_}\DecValTok{2} \OperatorTok{<}\StringTok{ }\DecValTok{70}\NormalTok{] =}\StringTok{ "40"}
\NormalTok{class_}\DecValTok{2}\NormalTok{[class_}\DecValTok{2} \OperatorTok{>=}\StringTok{ }\DecValTok{70}\NormalTok{] =}\StringTok{ "70"}
\NormalTok{class_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(class_}\DecValTok{2}\NormalTok{)}
\KeywordTok{levels}\NormalTok{(class_}\DecValTok{2}\NormalTok{) <-}\StringTok{ }\NormalTok{class_list}
\NormalTok{pca <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(dat_pca, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)}
\KeywordTok{fviz_pca_ind}\NormalTok{(}
\NormalTok{  pca,}
  \DataTypeTok{geom.ind =} \StringTok{"point"}\NormalTok{,}
  \DataTypeTok{habillage=}\NormalTok{class_}\DecValTok{2}\NormalTok{,}
  \DataTypeTok{addEllipses =}\NormalTok{ T,}
  \DataTypeTok{ellipse.level=}\FloatTok{0.95}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\hypertarget{machine-learning-models-1}{%
\subsection{Machine learning models}\label{machine-learning-models-1}}

\hypertarget{dummy-variables}{%
\subsubsection{Dummy variables}\label{dummy-variables}}

\label{show-dummy-var}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Dummy variables}
\NormalTok{dummies <-}\StringTok{ }\KeywordTok{dummyVars}\NormalTok{( }\OperatorTok{~}\StringTok{ }\NormalTok{mix_app, }\DataTypeTok{data =}\NormalTok{ dat)}
\NormalTok{dummyDat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(dummies, }\DataTypeTok{newdata =}\NormalTok{ dat))}
\NormalTok{dummyDat}\OperatorTok{$}\NormalTok{id <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{id}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mix_app)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{full_join}\NormalTok{(., dummyDat)}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-preparation-2}{%
\subsubsection{Data preparation}\label{data-preparation-2}}

\label{show-preparation}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Data preparation}
\KeywordTok{names}\NormalTok{(dat) <-}\StringTok{ }\KeywordTok{gsub}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{names}\NormalTok{(dat), }\DataTypeTok{pattern =} \StringTok{"/"}\NormalTok{, }\DataTypeTok{replacement =} \StringTok{"."}\NormalTok{)  }
\NormalTok{dat_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_3"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\NormalTok{dat_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_3"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_7"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\NormalTok{dat_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_3"}\NormalTok{, }\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_14"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\NormalTok{dat_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_3"}\NormalTok{, }\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_28"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\NormalTok{dat_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_3"}\NormalTok{, }\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_100"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_56"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\NormalTok{dat_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day_3"}\NormalTok{, }\StringTok{"day_7"}\NormalTok{, }\StringTok{"day_14"}\NormalTok{, }\StringTok{"day_28"}\NormalTok{, }\StringTok{"day_56"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename_at}\NormalTok{(}\StringTok{"day_100"}\NormalTok{,}\OperatorTok{~}\StringTok{"mpa"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---first-18-columns-of-the-6-first-samples-of-28-days}{%
\subsubsection{Table - First 18 columns of the 6 first samples of 28
days}\label{table---first-18-columns-of-the-6-first-samples-of-28-days}}

\label{show-table-preparated-samples}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - First 18 columns of the 6 first samples of 28 days}
\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"water"}\NormalTok{,}
              \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Agg."}\NormalTok{, }\StringTok{"Fine Agg."}\NormalTok{, }\StringTok{"MPa"}\NormalTok{, }\StringTok{"Class"}\NormalTok{,}
              \StringTok{"Wat./"}\NormalTok{, }\StringTok{"F.Agg./"}\NormalTok{, }\StringTok{"C.Agg./"}\NormalTok{,}
              \StringTok{"F.Agg./"}\NormalTok{,}\StringTok{"Wat./"}\NormalTok{, }\StringTok{"Wat./"}\NormalTok{, }\StringTok{"App Mix"}\NormalTok{ =}\StringTok{ }\DecValTok{2}\NormalTok{)}
\NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
              \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}\StringTok{""}\NormalTok{, }\StringTok{"Ci."}\NormalTok{,}
              \StringTok{"Ce."}\NormalTok{, }\StringTok{"Ce."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"F.Agg."}\NormalTok{, }\StringTok{"1:1:2"}\NormalTok{, }\StringTok{"1:2:2"}\NormalTok{)}
\NormalTok{caption <-}\StringTok{ "First 18 columns of the 6 first samples of 28 days"}
\NormalTok{dat_table <-}\StringTok{ }\NormalTok{dat_}\DecValTok{28}\NormalTok{[}\DecValTok{1}\OperatorTok{:}\DecValTok{18}\NormalTok{]}
\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{head}\NormalTok{(dat_table[}\KeywordTok{order}\NormalTok{(dat_table}\OperatorTok{$}\NormalTok{id),]),}
    \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{digits =} \DecValTok{2}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{removing-near-zero-variance-columns}{%
\subsubsection{Removing near zero variance
columns}\label{removing-near-zero-variance-columns}}

\label{show-nzv}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Removing near zero variance columns}
\NormalTok{nzv_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{3}\NormalTok{)}
\NormalTok{nzv_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{7}\NormalTok{)}
\NormalTok{nzv_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{14}\NormalTok{)}
\NormalTok{nzv_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{28}\NormalTok{)}
\NormalTok{nzv_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{56}\NormalTok{)}
\NormalTok{nzv_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{100}\NormalTok{)}
\NormalTok{dat_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{3}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{3}\NormalTok{]}
\NormalTok{dat_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{7}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{7}\NormalTok{]}
\NormalTok{dat_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{14}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{14}\NormalTok{]}
\NormalTok{dat_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{28}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{28}\NormalTok{]}
\NormalTok{dat_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{56}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{56}\NormalTok{]}
\NormalTok{dat_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\NormalTok{dat_}\DecValTok{100}\NormalTok{[,}\OperatorTok{-}\NormalTok{nzv_}\DecValTok{100}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\hypertarget{high-correlation-variables-verification}{%
\subsubsection{High correlation variables
verification}\label{high-correlation-variables-verification}}

\label{show-cors}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# High correlation variables verification}
\NormalTok{descr_cor_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{3}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{descr_cor_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{7}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{descr_cor_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{14}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{descr_cor_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{28}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{descr_cor_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{56}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{descr_cor_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{100}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
\NormalTok{high_cor_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{3}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{3}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\NormalTok{high_cor_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{7}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{7}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\NormalTok{high_cor_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{14}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{14}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\NormalTok{high_cor_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{28}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{28}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\NormalTok{high_cor_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{56}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{56}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\NormalTok{high_cor_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{abs}\NormalTok{(descr_cor_}\DecValTok{100}\NormalTok{[}\KeywordTok{upper.tri}\NormalTok{(descr_cor_}\DecValTok{100}\NormalTok{)]) }\OperatorTok{>}\StringTok{ }\FloatTok{.999}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{test-and-train-data-split}{%
\subsubsection{Test and train data
split}\label{test-and-train-data-split}}

\label{show-split}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Test and train data split}
\NormalTok{reg <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}
\NormalTok{  dat_}\DecValTok{3} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{3}\OperatorTok{$}\NormalTok{mpa),}
\NormalTok{  dat_}\DecValTok{7} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{7}\OperatorTok{$}\NormalTok{mpa),}
\NormalTok{  dat_}\DecValTok{14} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{14}\OperatorTok{$}\NormalTok{mpa),}
\NormalTok{  dat_}\DecValTok{28} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{28}\OperatorTok{$}\NormalTok{mpa),}
\NormalTok{  dat_}\DecValTok{56} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{56}\OperatorTok{$}\NormalTok{mpa),}
\NormalTok{  dat_}\DecValTok{100} \OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, class, id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ dat_}\DecValTok{100}\OperatorTok{$}\NormalTok{mpa)}
\NormalTok{)}
\NormalTok{reg_seed <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
  \DecValTok{1111}\NormalTok{, }\CommentTok{# 3}
  \DecValTok{1}\NormalTok{, }\CommentTok{# 7}
  \DecValTok{22}\NormalTok{, }\CommentTok{# 14}
  \DecValTok{11111}\NormalTok{, }\CommentTok{# 28}
  \DecValTok{111}\NormalTok{, }\CommentTok{# 56}
  \DecValTok{11} \CommentTok{# 100}
\NormalTok{  )}
\NormalTok{split_reg <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
  \KeywordTok{set.seed}\NormalTok{(reg_seed[[n]], }\DataTypeTok{sample.kind=}\StringTok{"Rounding"}\NormalTok{)}
  \KeywordTok{createDataPartition}\NormalTok{(reg[[n]]}\OperatorTok{$}\NormalTok{y, }\DataTypeTok{p =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{list =}\NormalTok{ F)}
\NormalTok{\}}
\NormalTok{trainIndex_reg <-}\StringTok{ }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{6}\NormalTok{), split_reg)}
\NormalTok{gen_dat <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
\NormalTok{  regIndex <-}\StringTok{ }\NormalTok{trainIndex_reg[[n]]}
  \KeywordTok{list}\NormalTok{(}\DataTypeTok{train =}\NormalTok{ reg[[n]][regIndex,], }\DataTypeTok{test =}\NormalTok{ reg[[n]][}\OperatorTok{-}\NormalTok{regIndex,])}
\NormalTok{\}}
\NormalTok{dats_reg <-}\StringTok{ }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{6}\NormalTok{), gen_dat)}
\KeywordTok{names}\NormalTok{(dats_reg) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"d3"}\NormalTok{, }\StringTok{"d7"}\NormalTok{, }\StringTok{"d14"}\NormalTok{, }\StringTok{"d28"}\NormalTok{, }\StringTok{"d56"}\NormalTok{, }\StringTok{"d100"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{distribution-of-test-and-train-data}{%
\subsubsection{Distribution of test and train
data}\label{distribution-of-test-and-train-data}}

\label{show-dist-split}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Distribution of test and train data}
\NormalTok{cap <-}\StringTok{ "Distribution of test and train data"}
\NormalTok{d_lab <-}\StringTok{ " days"}
\NormalTok{ylabel <-}\StringTok{ "Density"}
\NormalTok{g1 <-}\StringTok{ "Train"}
\NormalTok{g2 <-}\StringTok{ "Test"}
\NormalTok{dens <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d, n)\{}
\NormalTok{  dens <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(}
\NormalTok{      d}\OperatorTok{$}\NormalTok{train }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{select}\NormalTok{(y) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Group =}\NormalTok{ g1, }\DataTypeTok{day =} \KeywordTok{str_sub}\NormalTok{(n, }\DecValTok{2}\NormalTok{)),}
\NormalTok{      d}\OperatorTok{$}\NormalTok{test }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{select}\NormalTok{(y) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Group =}\NormalTok{ g2, }\DataTypeTok{day =} \KeywordTok{str_sub}\NormalTok{(n, }\DecValTok{2}\NormalTok{))}
\NormalTok{    )}
\NormalTok{  dens}
\NormalTok{\}}
\NormalTok{densities <-}\StringTok{ }\KeywordTok{imap}\NormalTok{(dats_reg, dens) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{reduce}\NormalTok{(rbind) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{day_f =} \KeywordTok{factor}\NormalTok{(day, }\DataTypeTok{levels=}\KeywordTok{c}\NormalTok{(}\StringTok{'3'}\NormalTok{,}\StringTok{'7'}\NormalTok{,}\StringTok{'14'}\NormalTok{,}\StringTok{'28'}\NormalTok{, }\StringTok{'56'}\NormalTok{, }\StringTok{'100'}\NormalTok{)))}
\NormalTok{labs <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{), d_lab, }\DataTypeTok{sep=}\StringTok{""}\NormalTok{)}
\KeywordTok{levels}\NormalTok{(densities}\OperatorTok{$}\NormalTok{day_f) <-}\StringTok{ }\NormalTok{labs}
\NormalTok{densities }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(y, }\DataTypeTok{fill =}\NormalTok{ Group)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{day_f, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"MPa"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{naive-model}{%
\subsubsection{Naive model}\label{naive-model}}

\label{show-naive-model-reg}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Naive model}
\NormalTok{res <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d, n)\{}
  \KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{day =} \KeywordTok{str_sub}\NormalTok{(n, }\DecValTok{2}\NormalTok{),}
    \DataTypeTok{mean_mpa =} \KeywordTok{mean}\NormalTok{(d}\OperatorTok{$}\NormalTok{train}\OperatorTok{$}\NormalTok{y)   }
\NormalTok{    )}
\NormalTok{\}}
\NormalTok{ing_model <-}\StringTok{ }\KeywordTok{imap}\NormalTok{(dats_reg, res)}
\NormalTok{get_rmse <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d, n)\{}
   \KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{rmse_train =} \KeywordTok{RMSE}\NormalTok{(d}\OperatorTok{$}\NormalTok{train}\OperatorTok{$}\NormalTok{y, ing_model[[n]]}\OperatorTok{$}\NormalTok{mean_mpa),}
    \DataTypeTok{rmse_test =}  \KeywordTok{RMSE}\NormalTok{(d}\OperatorTok{$}\NormalTok{test}\OperatorTok{$}\NormalTok{y, ing_model[[n]]}\OperatorTok{$}\NormalTok{mean_mpa),}
    \DataTypeTok{day =} \KeywordTok{str_sub}\NormalTok{(n, }\DecValTok{2}\NormalTok{) }
\NormalTok{    )}
\NormalTok{\}}
\NormalTok{rmses <-}\StringTok{ }\KeywordTok{imap}\NormalTok{(dats_reg, get_rmse) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{reduce}\NormalTok{(rbind)}
\NormalTok{ing_model_df <-}\StringTok{ }\NormalTok{ing_model }\OperatorTok{%>%}\StringTok{ }\KeywordTok{reduce}\NormalTok{(rbind)}
\NormalTok{df_performance_reg <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(ing_model_df, rmses)}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---naive-models}{%
\subsubsection{Table - Naive models}\label{table---naive-models}}

\label{show-table-naive-model-reg}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Naive models}
\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Age"}\NormalTok{, }\StringTok{"Mean $MPa$ (train)"}\NormalTok{, }\StringTok{"RMSE (train)"}\NormalTok{, }\StringTok{"RMSE (test)"}\NormalTok{)}
\NormalTok{caption <-}\StringTok{ "Naive models"}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{    df_performance_reg,}
    \DataTypeTok{col.names =}\NormalTok{ colNames,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    )  }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{features-selection}{%
\subsubsection{Features selection}\label{features-selection}}

\label{show-test-models-reg}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Features selection}
\NormalTok{data1 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train }\CommentTok{# full}
\NormalTok{data2 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# no dummy vars}
\NormalTok{data3 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{7}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# only original variables}
\NormalTok{data4 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{8}\OperatorTok{:}\DecValTok{21}\NormalTok{)] }\CommentTok{# only new variables}
\NormalTok{data5 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{8}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# only new varibels, no dummy vars}
\NormalTok{test1 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test }\CommentTok{# full}
\NormalTok{test2 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# no dummy vars}
\NormalTok{test3 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{7}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# only original variables}
\NormalTok{test4 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{8}\OperatorTok{:}\DecValTok{21}\NormalTok{)] }\CommentTok{# only new variables}
\NormalTok{test5 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{8}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)] }\CommentTok{# only new varibels, no dummy vars}
\CommentTok{# parRF}
\KeywordTok{modelLookup}\NormalTok{(}\StringTok{"parRF"}\NormalTok{)}
\NormalTok{control_parRF <-}\StringTok{ }\KeywordTok{trainControl}\NormalTok{(}\DataTypeTok{method=}\StringTok{'repeatedcv'}\NormalTok{,}\DataTypeTok{number=}\DecValTok{10}\NormalTok{,}\DataTypeTok{repeats=}\DecValTok{5}\NormalTok{,}\DataTypeTok{search=}\StringTok{'grid'}\NormalTok{)}
\NormalTok{fit_parRF<-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(data, n) \{}
  \KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{  tunegrid_parRF<-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}\DataTypeTok{mtry =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{length}\NormalTok{(data), n))}
  \KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
    \DataTypeTok{data =}\NormalTok{ data,}
    \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
    \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
    \DataTypeTok{tuneGrid=}\NormalTok{tunegrid_parRF,}
    \DataTypeTok{trControl=}\NormalTok{control_parRF)}
\NormalTok{\}}
\NormalTok{fit_parRF1 <-}\StringTok{ }\KeywordTok{fit_parRF}\NormalTok{(data1, }\DecValTok{3}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(fit_parRF1)}
\KeywordTok{min}\NormalTok{(fit_parRF1}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE) }\CommentTok{# 6.26108}
\NormalTok{p1_parRF1 <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_parRF1, }\DataTypeTok{newdata =}\NormalTok{ test1)}
\KeywordTok{RMSE}\NormalTok{(p1_parRF1, test1}\OperatorTok{$}\NormalTok{y) }\CommentTok{# 4.887264}
\NormalTok{fit_parRF2 <-}\StringTok{ }\KeywordTok{fit_parRF}\NormalTok{(data2, }\DecValTok{1}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(fit_parRF2)}
\KeywordTok{min}\NormalTok{(fit_parRF2}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE) }\CommentTok{# 6.268789}
\NormalTok{p_parRF2 <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_parRF2, }\DataTypeTok{newdata =}\NormalTok{ test2)}
\KeywordTok{RMSE}\NormalTok{(p_parRF2, test2}\OperatorTok{$}\NormalTok{y) }\CommentTok{# 4.812994}
\NormalTok{fit_parRF3 <-}\StringTok{ }\KeywordTok{fit_parRF}\NormalTok{(data3, }\DecValTok{1}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(fit_parRF3)}
\KeywordTok{min}\NormalTok{(fit_parRF3}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE) }\CommentTok{# 6.361905}
\NormalTok{p_parRF3 <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_parRF3, }\DataTypeTok{newdata =}\NormalTok{ test3)}
\KeywordTok{RMSE}\NormalTok{(p_parRF3, test3}\OperatorTok{$}\NormalTok{y) }\CommentTok{# 5.083467}
\NormalTok{fit_parRF4 <-}\StringTok{ }\KeywordTok{fit_parRF}\NormalTok{(data4, }\DecValTok{1}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(fit_parRF4)}
\KeywordTok{min}\NormalTok{(fit_parRF4}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE) }\CommentTok{# 8.705984}
\NormalTok{p_parRF4 <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_parRF4, }\DataTypeTok{newdata =}\NormalTok{ test4)}
\KeywordTok{RMSE}\NormalTok{(p_parRF4, test4}\OperatorTok{$}\NormalTok{y) }\CommentTok{# 8.083962}
\NormalTok{fit_parRF5 <-}\StringTok{ }\KeywordTok{fit_parRF}\NormalTok{(data5, }\DecValTok{1}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(fit_parRF5)}
\KeywordTok{min}\NormalTok{(fit_parRF5}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE) }\CommentTok{# 8.691755}
\NormalTok{p_parRF5 <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_parRF5, }\DataTypeTok{newdata =}\NormalTok{ test5)}
\KeywordTok{RMSE}\NormalTok{(p_parRF5, test5}\OperatorTok{$}\NormalTok{y) }\CommentTok{# 7.991537}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---first-6-samples-of-train-data-of-the-28-days-model}{%
\subsubsection{Table - First 6 samples of train data of the 28 days
model}\label{table---first-6-samples-of-train-data-of-the-28-days-model}}

\label{show-table-done-samples-28}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - First 6 samples of train data of the 28 days model}
\NormalTok{caption <-}\StringTok{ "First 6 samples of train data of the 28 days model"}
\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"water"}\NormalTok{,}
              \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Agg."}\NormalTok{, }\StringTok{"Fine Agg."}\NormalTok{,}
              \StringTok{"Wat./"}\NormalTok{, }\StringTok{"F.Agg./"}\NormalTok{, }\StringTok{"C.Agg./"}\NormalTok{,}
              \StringTok{"F.Agg./"}\NormalTok{,}\StringTok{"Wat./"}\NormalTok{, }\StringTok{"Wat./"}\NormalTok{, }\StringTok{"y"}\NormalTok{)}
\NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
              \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"Ce."}\NormalTok{,}
              \StringTok{"Ce."}\NormalTok{, }\StringTok{"Ce."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"F.Agg."}\NormalTok{, }\StringTok{"MPa"}\NormalTok{)}
\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Features"}\NormalTok{=}\DecValTok{13}\NormalTok{, }\StringTok{"Outcome"}\NormalTok{=}\DecValTok{1}\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{head}\NormalTok{(data2),}
    \DataTypeTok{col.names =}\NormalTok{ dfUnits,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{digits =} \DecValTok{2}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames2, }\DataTypeTok{line =}\NormalTok{ T, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{regression-models-1}{%
\subsubsection{Regression models}\label{regression-models-1}}

\label{show-reg-models}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Regression models}
\CommentTok{# no dummy vars:}
\NormalTok{data3 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d3}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{22}\NormalTok{)]}
\NormalTok{data7 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d7}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{, }\DecValTok{19}\NormalTok{)]}
\NormalTok{data14 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d14}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)]}
\NormalTok{data28 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)]}
\NormalTok{data56 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d56}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{22}\NormalTok{)]}
\NormalTok{data100 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d100}\OperatorTok{$}\NormalTok{train[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{23}\NormalTok{)]}
\NormalTok{test3 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d3}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{22}\NormalTok{)]}
\NormalTok{test7 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d7}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{, }\DecValTok{19}\NormalTok{)]}
\NormalTok{test14 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d14}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)]}
\NormalTok{test28 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{21}\NormalTok{)]}
\NormalTok{test56 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d56}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{22}\NormalTok{)]}
\NormalTok{test100 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d100}\OperatorTok{$}\NormalTok{test[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{13}\NormalTok{, }\DecValTok{23}\NormalTok{)]}
\NormalTok{get_trControl <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n, r)\{}
  \KeywordTok{trainControl}\NormalTok{(}\DataTypeTok{method=}\StringTok{'repeatedcv'}\NormalTok{,}\DataTypeTok{number=}\NormalTok{n,}\DataTypeTok{repeats=}\NormalTok{r,}\DataTypeTok{search=}\StringTok{'grid'}\NormalTok{)}
\NormalTok{\}}
\NormalTok{trControl3 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{30}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{trControl7 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{trControl14 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{30}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{trControl28 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{30}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{trControl56 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{30}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{trControl100 <-}\StringTok{ }\KeywordTok{get_trControl}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{get_tuneGrid <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(data)\{}
  \KeywordTok{expand.grid}\NormalTok{(}\DataTypeTok{mtry =} \KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{length}\NormalTok{(data), }\DecValTok{1}\NormalTok{))}
\NormalTok{\}}
\NormalTok{tuneGrid3 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data3)}
\NormalTok{tuneGrid7 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data7)}
\NormalTok{tuneGrid14 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data14)}
\NormalTok{tuneGrid28 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data28)}
\NormalTok{tuneGrid56 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data56)}
\NormalTok{tuneGrid100 <-}\StringTok{ }\KeywordTok{get_tuneGrid}\NormalTok{(data100)}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data3,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid3,}
            \DataTypeTok{trControl=}\NormalTok{trControl3)}
\NormalTok{p_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{3}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test3)}
\NormalTok{RMSE_test_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{3}\NormalTok{, test3}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{3} \CommentTok{# 3.31037}
\NormalTok{fit_}\DecValTok{3}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 6}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data7,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid7,}
            \DataTypeTok{trControl=}\NormalTok{trControl7)}
\NormalTok{p_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{7}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test7)}
\NormalTok{RMSE_test_}\DecValTok{7}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{7}\NormalTok{, test7}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{7} \CommentTok{# 4.361987}
\NormalTok{fit_}\DecValTok{7}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 2}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data14,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid14,}
            \DataTypeTok{trControl=}\NormalTok{trControl14)}
\NormalTok{p_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{14}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test14)}
\NormalTok{RMSE_test_}\DecValTok{14}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{14}\NormalTok{, test14}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{14} \CommentTok{# 4.620515}
\NormalTok{fit_}\DecValTok{14}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 13}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data28,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid28,}
            \DataTypeTok{trControl=}\NormalTok{trControl28)}
\NormalTok{p_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{28}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test28)}
\NormalTok{RMSE_test_}\DecValTok{28}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{28}\NormalTok{, test28}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{28} \CommentTok{# 4.716698}
\NormalTok{fit_}\DecValTok{28}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 11}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data56,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid56,}
            \DataTypeTok{trControl=}\NormalTok{trControl56)}
\NormalTok{p_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{56}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test56)}
\NormalTok{RMSE_test_}\DecValTok{56}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{56}\NormalTok{, test56}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{56} \CommentTok{# 5.939163}
\NormalTok{fit_}\DecValTok{56}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 8}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DataTypeTok{sample.kind =} \StringTok{"Rounding"}\NormalTok{)}
\NormalTok{fit_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{train}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{., }
            \DataTypeTok{data =}\NormalTok{ data100,}
            \DataTypeTok{preProcess =} \KeywordTok{c}\NormalTok{(}\StringTok{"center"}\NormalTok{,}\StringTok{"scale"}\NormalTok{),}
            \DataTypeTok{method=}\StringTok{'parRF'}\NormalTok{, }
            \DataTypeTok{tuneGrid=}\NormalTok{tuneGrid100,}
            \DataTypeTok{trControl=}\NormalTok{trControl100)}
\NormalTok{p_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_}\DecValTok{100}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ test100)}
\NormalTok{RMSE_test_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\KeywordTok{RMSE}\NormalTok{(p_}\DecValTok{100}\NormalTok{, test100}\OperatorTok{$}\NormalTok{y)}
\NormalTok{RMSE_test_}\DecValTok{100} \CommentTok{# 5.851088}
\NormalTok{fit_}\DecValTok{100}\OperatorTok{$}\NormalTok{bestTune }\CommentTok{# mtry = 8}
\end{Highlighting}
\end{Shaded}

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

\hypertarget{table---models-details}{%
\subsubsection{Table - Models details}\label{table---models-details}}

\label{show-table-reg-models}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Models details}
\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"mtry"}\NormalTok{, }\StringTok{"CV"}\NormalTok{, }\StringTok{"Repetitions"}\NormalTok{, }
              \StringTok{"RMSE (train)"}\NormalTok{, }\StringTok{"RMSE (test)"}\NormalTok{)}
\NormalTok{caption <-}\StringTok{ "Regression models results"}
\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
\NormalTok{dat_reg_models <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{dia =}\NormalTok{ day,}
  \DataTypeTok{mtry =} \KeywordTok{c}\NormalTok{(fit_}\DecValTok{3}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry, fit_}\DecValTok{7}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry, fit_}\DecValTok{14}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry,}
\NormalTok{           fit_}\DecValTok{28}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry, fit_}\DecValTok{56}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry, fit_}\DecValTok{100}\OperatorTok{$}\NormalTok{bestTune}\OperatorTok{$}\NormalTok{mtry),}
  \DataTypeTok{number =} \KeywordTok{c}\NormalTok{(trControl3}\OperatorTok{$}\NormalTok{number, trControl7}\OperatorTok{$}\NormalTok{number, trControl14}\OperatorTok{$}\NormalTok{number,}
\NormalTok{             trControl28}\OperatorTok{$}\NormalTok{number,trControl56}\OperatorTok{$}\NormalTok{number,trControl100}\OperatorTok{$}\NormalTok{number),}
  \DataTypeTok{repeats =} \KeywordTok{c}\NormalTok{(trControl3}\OperatorTok{$}\NormalTok{repeats, trControl7}\OperatorTok{$}\NormalTok{repeats, trControl14}\OperatorTok{$}\NormalTok{repeats,}
\NormalTok{             trControl28}\OperatorTok{$}\NormalTok{repeats,trControl56}\OperatorTok{$}\NormalTok{repeats,trControl100}\OperatorTok{$}\NormalTok{repeats),}
  \DataTypeTok{RMSE_train =} \KeywordTok{c}\NormalTok{(}\KeywordTok{min}\NormalTok{(fit_}\DecValTok{3}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE), }\KeywordTok{min}\NormalTok{(fit_}\DecValTok{7}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE),}
                  \KeywordTok{min}\NormalTok{(fit_}\DecValTok{14}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE), }\KeywordTok{min}\NormalTok{(fit_}\DecValTok{28}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE),}
                  \KeywordTok{min}\NormalTok{(fit_}\DecValTok{56}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE), }\KeywordTok{min}\NormalTok{(fit_}\DecValTok{100}\OperatorTok{$}\NormalTok{results}\OperatorTok{$}\NormalTok{RMSE)),}
  \DataTypeTok{RMSE_test =} \KeywordTok{c}\NormalTok{(RMSE_test_}\DecValTok{3}\NormalTok{, RMSE_test_}\DecValTok{7}\NormalTok{, RMSE_test_}\DecValTok{14}\NormalTok{,}
\NormalTok{                RMSE_test_}\DecValTok{28}\NormalTok{, RMSE_test_}\DecValTok{56}\NormalTok{, RMSE_test_}\DecValTok{100}\NormalTok{)}
\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{    dat_reg_models,}
    \DataTypeTok{col.names =}\NormalTok{ colNames,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    )  }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figure---models-comparison}{%
\subsubsection{Figure - Models
comparison}\label{figure---models-comparison}}

\label{show-results-comparison}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Figure - Models comparison}
\NormalTok{cap <-}\StringTok{ "Actual vs Predicted values in each model"}
\NormalTok{models <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
\NormalTok{xlabel <-}\StringTok{ "Actual (MPa)"}
\NormalTok{ylabel <-}\StringTok{ "Predicted (MPa)"}
\NormalTok{preds <-}\StringTok{ }\KeywordTok{list}\NormalTok{(p_}\DecValTok{3}\NormalTok{, p_}\DecValTok{7}\NormalTok{, p_}\DecValTok{14}\NormalTok{, p_}\DecValTok{28}\NormalTok{, p_}\DecValTok{56}\NormalTok{, p_}\DecValTok{100}\NormalTok{)}
\NormalTok{actuals <-}\StringTok{ }\KeywordTok{list}\NormalTok{(test3}\OperatorTok{$}\NormalTok{y, test7}\OperatorTok{$}\NormalTok{y, test14}\OperatorTok{$}\NormalTok{y, test28}\OperatorTok{$}\NormalTok{y, test56}\OperatorTok{$}\NormalTok{y, test100}\OperatorTok{$}\NormalTok{y)}
\NormalTok{gen_res_df <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(ind)\{}
  \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{actual =}\NormalTok{ actuals[[ind]], }\DataTypeTok{pred =}\NormalTok{ preds[[ind]], }\DataTypeTok{model =}\NormalTok{ models[[ind]])}
\NormalTok{\}}
\NormalTok{res_df <-}\StringTok{ }\KeywordTok{lapply}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{6}\NormalTok{, gen_res_df)}
\NormalTok{res_df <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(res_df)}
\NormalTok{res_df}\OperatorTok{$}\NormalTok{model <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(res_df}\OperatorTok{$}\NormalTok{model,}\DataTypeTok{levels=}\NormalTok{models)}
\NormalTok{res_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(actual, pred)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{model, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{slope=}\DecValTok{1}\NormalTok{, }\DataTypeTok{intercept=}\DecValTok{0}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel)}
\end{Highlighting}
\end{Shaded}

\hypertarget{results-tables-of-10-best-and-worst-results}{%
\subsubsection{Results tables of 10 best and worst
results}\label{results-tables-of-10-best-and-worst-results}}

\label{show-table-10}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Results tables of 10 best and worst results}
\NormalTok{colNames1 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Actual"}\NormalTok{, }\StringTok{"Predicted"}\NormalTok{, }\StringTok{"Error"}\NormalTok{, }\StringTok{""}\NormalTok{,}\StringTok{"Actual"}\NormalTok{, }\StringTok{"Predicted"}\NormalTok{, }\StringTok{"Error"}\NormalTok{) }
\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Best 10"}\NormalTok{=}\DecValTok{3}\NormalTok{,}\StringTok{""}\NormalTok{, }\StringTok{"Worst 10"}\NormalTok{=}\DecValTok{3}\NormalTok{) }
\NormalTok{caption_}\DecValTok{0}\NormalTok{ <-}\StringTok{ "Model of "}
\NormalTok{get_X <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(pred, actual, }\DataTypeTok{X=}\DecValTok{10}\NormalTok{)\{}
\NormalTok{  diff <-}\StringTok{ }\KeywordTok{abs}\NormalTok{(pred }\OperatorTok{-}\StringTok{ }\NormalTok{actual)}
\NormalTok{  diff_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{pred }\OperatorTok{-}\StringTok{ }\NormalTok{actual}
\NormalTok{  ind_min <-}\StringTok{ }\KeywordTok{which}\NormalTok{(diff }\OperatorTok{<=}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{sort}\NormalTok{(diff, }\DataTypeTok{decreasing =}\NormalTok{ F)[}\DecValTok{1}\OperatorTok{:}\NormalTok{X]), }\DataTypeTok{arr.ind =}\NormalTok{ T)}
\NormalTok{  ind_max <-}\StringTok{ }\KeywordTok{which}\NormalTok{(diff }\OperatorTok{>=}\StringTok{ }\KeywordTok{min}\NormalTok{(}\KeywordTok{sort}\NormalTok{(diff, }\DataTypeTok{decreasing =}\NormalTok{ T)[}\DecValTok{1}\OperatorTok{:}\NormalTok{X]), }\DataTypeTok{arr.ind =}\NormalTok{ T)}
\NormalTok{  df_min <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{actual_min=}\NormalTok{actual[ind_min],}
    \DataTypeTok{pred_min=}\NormalTok{pred[ind_min],}
    \DataTypeTok{diff_min=}\NormalTok{diff[ind_min],}
    \DataTypeTok{diff_min_2=}\NormalTok{diff_}\DecValTok{2}\NormalTok{[ind_min]}
\NormalTok{  )}
\NormalTok{  df_max <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{actual_max=}\NormalTok{actual[ind_max],}
    \DataTypeTok{pred_max=}\NormalTok{pred[ind_max],}
    \DataTypeTok{diff_max=}\NormalTok{diff[ind_max],}
    \DataTypeTok{diff_max_2=}\NormalTok{diff_}\DecValTok{2}\NormalTok{[ind_max]}
\NormalTok{  )}
\NormalTok{  df_max <-}\StringTok{ }\NormalTok{df_max[}\KeywordTok{order}\NormalTok{(}\OperatorTok{-}\NormalTok{df_max}\OperatorTok{$}\NormalTok{diff_max),]}
\NormalTok{  df_min <-}\StringTok{ }\NormalTok{df_min[}\KeywordTok{order}\NormalTok{(df_min}\OperatorTok{$}\NormalTok{diff_min),]}
\NormalTok{  df_null <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{null_col =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{), X))}
\NormalTok{  res <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(df_min,df_null, df_max)}
\NormalTok{  res <-}\StringTok{ }\NormalTok{res }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(diff_max, diff_min))}
  \KeywordTok{rownames}\NormalTok{(res) <-}\StringTok{ }\KeywordTok{c}\NormalTok{()}
\NormalTok{  res}
\NormalTok{\}}
\NormalTok{gen_kable <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(ind)\{}
\NormalTok{  df <-}\StringTok{ }\KeywordTok{get_X}\NormalTok{(preds[[ind]], actuals[[ind]])}
\NormalTok{  caption <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(caption_}\DecValTok{0}\NormalTok{,models[ind])}
  \KeywordTok{kable}\NormalTok{(}
\NormalTok{    df,}
    \DataTypeTok{col.names =}\NormalTok{ colNames1,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    )  }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DataTypeTok{width =} \StringTok{"1cm"}\NormalTok{,) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames2, }\DataTypeTok{line =}\NormalTok{ T, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
\NormalTok{\}}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{2}\NormalTok{)}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{4}\NormalTok{)}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\KeywordTok{gen_kable}\NormalTok{(}\DecValTok{6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{discussion-and-conclusion-1}{%
\subsection{Discussion and
conclusion}\label{discussion-and-conclusion-1}}

\hypertarget{table---comparison-of-other-works}{%
\subsubsection{Table - Comparison of other
works}\label{table---comparison-of-other-works}}

\label{show-works-comparison}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Comparison of other works}
\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Author"}\NormalTok{,}\StringTok{"Year"}\NormalTok{ ,}\StringTok{"Algorithm"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{) }
\NormalTok{caption <-}\StringTok{ "Comparison of other works"}
\NormalTok{works <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{autor =} \KeywordTok{c}\NormalTok{(}\StringTok{"Pierobon"}\NormalTok{, }\StringTok{"Hameed"}\NormalTok{, }\StringTok{"Raj"}\NormalTok{,}\StringTok{"Modukuru"}\NormalTok{ ,}\StringTok{"Alshamiri"}\NormalTok{, }\StringTok{"Abban"}\NormalTok{),}
  \DataTypeTok{ano =} \KeywordTok{c}\NormalTok{(}\StringTok{"2018"}\NormalTok{, }\StringTok{"2020"}\NormalTok{, }\StringTok{"2018"}\NormalTok{,}\StringTok{"2020"}\NormalTok{ ,}\StringTok{"2020"}\NormalTok{, }\StringTok{"2016"}\NormalTok{),}
  \DataTypeTok{modelo =} \KeywordTok{c}\NormalTok{(}\StringTok{"Ensemble com 5 algorítimos"}\NormalTok{, }\StringTok{"Artificial Neural Networks"}\NormalTok{, }
          \StringTok{"Gradient Boosting Regressor"}\NormalTok{,}\StringTok{"Random Forest Regressor"}\NormalTok{ ,}
          \StringTok{"Regularized Extreme Learning Machine"}\NormalTok{,}
          \StringTok{"Support Vector Machines with Radial Basis Function Kernel"}\NormalTok{),}
  \DataTypeTok{RMSE =} \KeywordTok{c}\NormalTok{(}\StringTok{"4.150"}\NormalTok{, }\StringTok{"4.736"}\NormalTok{, }\StringTok{"4.957"}\NormalTok{,}\StringTok{"5.080"}\NormalTok{,}\StringTok{"5.508"}\NormalTok{, }\StringTok{"6.105"}\NormalTok{)}
\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{  works,}
  \DataTypeTok{col.names =}\NormalTok{ colNames,}
  \DataTypeTok{escape =}\NormalTok{ F,}
  \DataTypeTok{booktabs =}\NormalTok{ T,}
  \DataTypeTok{caption =}\NormalTok{ caption,}
  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
  \DataTypeTok{align =} \StringTok{"l"}
\NormalTok{  )  }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{table---final-results}{%
\subsubsection{Table - Final results}\label{table---final-results}}

\label{show-results}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Table - Final results}
\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{)}
\NormalTok{caption <-}\StringTok{ "Final result"}
\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
\NormalTok{dat_reg_models <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{dia =}\NormalTok{ day,}
  \DataTypeTok{RMSE_test =} \KeywordTok{c}\NormalTok{(RMSE_test_}\DecValTok{3}\NormalTok{, RMSE_test_}\DecValTok{7}\NormalTok{, RMSE_test_}\DecValTok{14}\NormalTok{,}
\NormalTok{                RMSE_test_}\DecValTok{28}\NormalTok{, RMSE_test_}\DecValTok{56}\NormalTok{, RMSE_test_}\DecValTok{100}\NormalTok{)}
\NormalTok{)}
\KeywordTok{kable}\NormalTok{(}
\NormalTok{    dat_reg_models,}
    \DataTypeTok{col.names =}\NormalTok{ colNames,}
    \DataTypeTok{escape =}\NormalTok{ F,}
    \DataTypeTok{booktabs =}\NormalTok{ T,}
    \DataTypeTok{caption =}\NormalTok{ caption,}
    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
    \DataTypeTok{align =} \StringTok{"c"}
\NormalTok{    )  }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2020-08-14 _meta/generated/paper_EN_html.Rmd
refactor

---
title: "Concrete compressive strength prediction with machine learning"
author: "Pedro Bernardino Alves Moreira"
date: "April 22, 2020"
output:
  bookdown::html_document2
# header-includes:
# - \usepackage{booktabs}
# - \usepackage{longtable}
# - \usepackage{supertabular}
# - \usepackage{array}
# - \usepackage{multirow}
# - \usepackage[table]{xcolor}
# - \usepackage{wrapfig}
# - \usepackage{colortbl}
# - \usepackage{pdflscape}
# - \usepackage{tabu}
# - \usepackage{threeparttable}
# - \usepackage[normalem]{ulem}
# - \usepackage{caption}
# - \usepackage{floatrow}
# - \floatsetup[table]{capposition=top}
# - \floatsetup[figure]{capposition=top}
# - \captionsetup{options=chunk}
# - \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
# - \renewcommand{\thechunk}{\arabic{chunk}}
# - \usepackage{indentfirst}
# - \usepackage{sectsty} \sectionfont{\centering}
indent: yes
bibliography: references.bib
abstract: Compressive strength is the main characteristic of concrete. This work used
  a set of samples widely used to build 6 forecast models of this characteristic,
  each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients
  (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and
  superplasticizers) and 6 relationships between the main ingredients (water/cement,
  coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate,
  water / coarse aggregate and water/fine aggregate) to make the predictions. The
  algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*)
  . The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31)
  and the worst for the 100-day model (*RMSE* of 5.85). The 28-day model had *RMSE*
  of 4.72. The work indicated that the separation in a model for each age is promising
  and was able to generate compatible results or better than other studies, opening
  way for new models creation following this proposal, but using different algorithms
  or the combination of different algorithms.
---
```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=F, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```
&nbsp;

\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

# Introduction

see \@ref(slug)

Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens. Measurement at 28 days is mandatory and represents the grade of the concrete. Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.

This compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed. @Hasan2011 proposed a mathematical model to predict from the results of tests of 7 and 14 days, and @Kabir2012 from 7 days. However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.

Many previous studies use the same dataset used by @Yeh1998 to predict the compressive strength of concrete. @Alshamiri2020 got good results with the regularized extreme learning machine (RELM) technique, and @Hameed2020 got even better results with the Artificial Neural Networks and cross-validation technique. This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as @Abban2016, @Raj2018, @Modukuru2020 and @Pierobon2018. At the end of the work, the results found are compared to the works cited here.

Unlike previous studies with this dataset, this work does data preparation differently. The age of the concrete is the most unique feature that contributes to its compressive strength. For this reason, age is treated separately in the machine learning models, creating models for each age group.

# Materials and Methods

## Materials
The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore]. Throughout the process, all code executed was documented in the same order as its execution in [Appendix 3](#appendix3), and reference was always made to codes throughout the text. All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1). In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in [Appendix 2](#appendix2).

## Reproducibility
In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result. The *seeds* can be checked throughout [Appendix 3](#appendix3) or directly on *Github*.

## Obtaining the data
The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}). In total there are 1030 samples with 9 columns. The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}). The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}). The first samples can be viewed in the table \ref{tab: first-samples} \@ref(tab-first-samples) \@ref(tab:first-samples).

```{r download-data, echo=F, include=F}
# Data download
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```
```{r rename-dat-cols, echo=F}
# Ranaming the columns
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )
dat$id <- seq.int(nrow(dat))
```
```{r reorder-dat, echo=F}
# Reordering the data
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```
```{r col-names-and-units, echo=F}
# Defining column names and units
colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```
```{r first-samples, echo=F}
# Table First samples
caption <- "First 6 samples"
kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

## Data preparation
The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies. Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed. In addition, the relevant data has been transformed to better fit the studies in the next steps.

### Initial data cleaning
Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

```{r removing-duplicated-samples, echo=F}
# Removing duplicated samples
n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```
The data show the variables in the columns and samples in the rows. However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.

```{r similar-samples, echo=FALSE}
# Tabela - Samples with same composition
same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))
caption <- "Samples with same composition"
kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process. This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.

```{r similar-samples-2, echo=FALSE}
# Tabela - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.

```{r initial-data-cleaning, echo=FALSE}
# Data cleaning
dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()
dat <- dat[!duplicated(select(dat, -c(id))),]
dat$id<-factor(dat$id)
n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```
The result can be seen in the table \ref{tab:similar-samples-same-id}. All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
# Tabela - Previous samples after processing
same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)
caption <- "Previous samples after processing"
kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### Age selection
As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models. However, it is necessary to verify how relevant the data of these other ages are. Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}. \@ref(fig-boxplot) See Figure \@ref(fig:boxplot)

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
# Figure - Compressive strength (MPa) vs age (days)
cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
ylabel <- "Compressive strength (MPa)"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```
It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.

This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}). The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis. The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
# Figure - Principal component analysis - 90, 91 e 100 days
dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))
cap <- "Principal component analysis - 90, 91 e 100 days"
colnames(dat_90_91_100) <- c(
   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
habillage <- dat_90_91_100$day
fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```
Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value. The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Compressive strength through time
cap <- "Compressive strength through time"
ylabel <- "MPa"
xlabel <- "Age (days)"
dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)
dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```
For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise. They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}. As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.

```{r join-90-91-100, echo=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum_duplicated <- sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Ages frequency
cap <- "Ages frequency"
ylabel <- "Frequency"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r remove-ages-lower-50, echo=F}
# Removing ages with frequency lower than 50
dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

### Data reorganization
The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}). The result in the first samples after this processing is shown in the table \ref{tab:new-features}.

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
# Table - First 6 samples after reorganization
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}
# Total samples
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables
To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
```{r new-features-2, echo=F}
# Adding New features
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```
```{r new-features-table, echo=F}
# Table - New features
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

## Data visualization
In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.

### Descriptive statistics
The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}). The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data. As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age. The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

### Correlation between ingredients and compressive strength
The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}). The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.",
                "3", "7", "14", "28", "56", "100",
                "class", "mix_app", "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
# Figure - Correlation over time
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
mix_dat$day <- factor(mix_dat$day, levels=lvls)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

### Variables distribution
A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
# Figure - Variables distribution

cap <- "Variables distribution"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat <- dat %>%
  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))

colnames(dist_dat) <- colnames_dat
dist_dat <- dist_dat %>%
  gather("Var", "value") %>%
  mutate(value = as.numeric(value))
dist_dat %>%
  ggplot(aes(value)) +
  geom_histogram(aes(y = ..density..), 
                 colour = "black", 
                 fill = "white") +
  geom_density(alpha = .5, fill = "lightseagreen") +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  xlab("") +
  ylab("") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}. It shows that the resistance to compression gradually increases over time, as expected. Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
# Figure - Variables distribution grouped by age

days_labs <- c("3","7","14","28","56","100")
cap <- "Variables distribution grouped by age"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  days_labs,
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat_2 <- dat %>%
  select(-c(id, class, mix_app))
colnames(dist_dat_2) <- colnames_dat
dist_dat_2 <- dist_dat_2 %>%
  gather("day", "mpa", days_labs) %>%
  drop_na() %>%
  gather("Var", "value", -c("day")) %>%
  mutate(day = factor(day, levels = days_labs))
dist_dat_2 %>%
  ggplot(aes(x = day, y = value)) +
  geom_violin(color = NA,
                fill = "lightseagreen",
                alpha = .5,
                na.rm = TRUE,
                scale = "count") +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  xlab("") +
  ylab("")
```

### Principal component analysis
Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}). The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that. It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
```{r pca, echo=F, fig.height=3, fig.cap=cap}
# Figure - Principal component analysis on ingredients

cap <- "Principal component analysis on ingredients"
colnames_dat <- c(
  "Cem.", "B.F.S.", "F.A.", "Wat.", "Sup.", "C.Agg.",
  "F.Agg.","WxC", "F.Agg.xC","C.Agg.xC","F.Agg.xC",
  "WxC.Agg.","WxF.Agg."
)
class_list <- list(low="0", normal="20", medium="40", high="70")
dat_pca <- dat %>%
  select(-c(id, class, mix_app,
            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))
colnames(dat_pca) <- colnames_dat
class_2 <- dat$day_28
class_2[class_2 < 20] = "0"
class_2[class_2 >= 20 & class_2 < 40] = "20"
class_2[class_2 >= 40 & class_2 < 70] = "40"
class_2[class_2 >= 70] = "70"
class_2 <- factor(class_2)
levels(class_2) <- class_list
pca <- prcomp(dat_pca, scale = TRUE)
fviz_pca_ind(
  pca,
  geom.ind = "point",
  habillage=class_2,
  addEllipses = T,
  ellipse.level=0.95) +
  ggtitle("") +
  theme_bw()
```

## Machine learning models
The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.

### Pre-processing and data separation
Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado. 
As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.

```{r dummy-var, echo=F, message=F, warning=F}
# Dummy variables
dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))
dummyDat$id <- dat$id
dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```
The samples were separated based on age. One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}). For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.

```{r preparation, echo=F}
# Data preparation
names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```
```{r table-preparated-samples, echo=F}
# Table - First 18 columns of the 6 first samples of 28 days
colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
caption <- "First 18 columns of the 6 first samples of 28 days"
dat_tabela <- dat_28[1:18]
kable(
    head(dat_tabela[order(dat_tabela$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}). Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact. In addition to them, in the case of the 7-day set, the fly ash variable was also removed. Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}). After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.

```{r nzv, echo=F}
# Removing near zero variance columns
nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)
dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```
```{r cors, echo=F}
# High correlation variables verification
descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```
The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package. If performed at this time, it would be necessary to manually undo these transformations in the predictions. The *caret* allows you to transform before training the models and already transforms the results back.

Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}). The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
```{r split, echo=F, message=F, warning=F}
# Test and train data split
reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$mpa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$mpa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$mpa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$mpa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$mpa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$mpa)
)
reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )
split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}
trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}
dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```
```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
# Distribution of test and train data
cap <- "Distribution of test and train data"
d_lab <- " days"
ylabel <- "Density"
g1 <- "Train"
g2 <- "Test"
dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Group = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Group = g2, day = str_sub(n, 2))
    )
  dens
}
densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs
densities %>%
  ggplot(aes(y, fill = Group)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~day_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

### Performance measures
The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*). The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.

### Naive models
Before creating the real models, for comparison purposes, naive models were created. They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}). In other words, naive models are simply the best guess possible. The results can be checked in the table \ref{tab:table-naive-model-reg}.
```{r naive-model-reg, echo=F, message=F, warning=F}
# Naive model
res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}
ing_model <- imap(dats_reg, res)
get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}
rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)
ing_model_df <- ing_model %>% reduce(rbind)
df_performance_reg <- full_join(ing_model_df, rmses)
```
```{r table-naive-model-reg, echo=F}
# Table - Naive models
colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
caption <- "Naive models"
kable(
    df_performance_reg,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

### Choice of algorithms
The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models. The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result. According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).

# Literature cited
<div id="refs"></div>
\newpage

# Appendix 1 - Virtual environment {#appendix1}

## Operational system
```{r system, echo=FALSE}
v <- as.matrix(R.Version())
kable(v, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```

## Packages
```{r packages, echo=FALSE}
p <- matrix(
      c(
        "caret", as.character(packageVersion("caret")),
        "cowplot",as.character(packageVersion("cowplot")),
        "dplyr", as.character(packageVersion("dplyr")),
        "factoextra", as.character(packageVersion("factoextra")),
        "gdata", as.character(packageVersion("gdata")),
        "ggplot2", as.character(packageVersion("ggplot2")),
        "gridExtra",as.character(packageVersion("gridExtra")),
        "kableExtra",as.character(packageVersion("kableExtra")),
        "knitr",as.character(packageVersion("knitr")),
        "pastecs",as.character(packageVersion("pastecs")),
        "purrr", as.character(packageVersion("purrr")),
        "questionr", as.character(packageVersion("questionr")),
        "reshape2", as.character(packageVersion("reshape2")),
        "tidyr", as.character(packageVersion("tidyr")),
        "tidyverse", as.character(packageVersion("tidyverse"))
      ),
      ncol = 2,
      byrow=TRUE
    )
kable(p, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```
\newpage

# Appendix 2 - Online repository {#appendix2}
https://github.com/pedrobern/concrete-compressive-strength-prediction
\newpage

# Appendix 3 - Code {#appendix3}

## Obtaining the data
### Data download

```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
```

### Ranaming the columns

```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
```

### Reordering the data

```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
```

### Defining column names and units

```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
```

### Table First samples

```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
```

## Data preparation
### Removing duplicated samples

```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
```

### Tabela - Samples with same composition

```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
```

### Tabela - Same samples with different results

```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
```

### Data cleaning

```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
```

### Tabela - Previous samples after processing

```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
```

### Figure - Compressive strength (MPa) vs age (days)

```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
```

### Figure - Principal component analysis - 90, 91 e 100 days

```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
```

### Figure - Compressive strength through time

```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
```

### Joining 90, 91 and 100 days data

```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
```

### Figure - Ages frequency

```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
```

### Removing ages with frequency lower than 50

```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
```

### Data reorganization

```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
```

### Table - First 6 samples after reorganization

```{r show-new-features, ref.label='new-features', ref='show-new-features', eval=FALSE}
```

### Total samples

```{r show-total-samples-2, ref.label='total-samples-2', ref='show-total-samples-2', eval=FALSE}
```

### Adding New features

```{r show-new-features-2, ref.label='new-features-2', ref='show-new-features-2', eval=FALSE}
```

### Table - New features

```{r show-new-features-table, ref.label='new-features-table', ref='show-new-features-table', eval=FALSE}
```

## Data visualization
### Table - Descriptive statistics - continuous variables

```{r show-stat-summ, ref.label='stat-summ', ref='show-stat-summ', eval=FALSE}
```

### Figure - Descriptive statistics - categorical variables

```{r show-stat-summ-categorical, ref.label='stat-summ-categorical', ref='show-stat-summ-categorical', eval=FALSE}
```

### Figure - Correlation grouped by age

```{r show-correlation, ref.label='correlation', ref='show-correlation', eval=FALSE}
```

### Figure - Correlation over time

```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
```

### Figure - Relationship between approximated mix, water, MPa and age

```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
```

### Figure - Relationship between concrete main features

```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
```

### Figure - Variables distribution

```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
```

### Figure - Variables distribution grouped by age

```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
```

### Figure - Principal component analysis on ingredients

```{r show-pca, ref.label='pca', ref='show-pca', eval=FALSE}
```

## Machine learning models
### Dummy variables

```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
```

### Data preparation

```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
```

### Table - First 18 columns of the 6 first samples of 28 days

```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
```

### Removing near zero variance columns

```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
```

### High correlation variables verification

```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
```

### Test and train data split

```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
```

### Distribution of test and train data

```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
```

### Naive model

```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
```

### Table - Naive models

```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
```

### Features selection

```{r show-test-models-reg, ref.label='test-models-reg', ref='show-test-models-reg', eval=FALSE}
```

### Table - First 6 samples of train data of the 28 days model {#slug}

```{r show-table-done-samples-28, ref.label='table-done-samples-28', ref='show-table-done-samples-28', eval=FALSE}
```

### 2020-08-14 _meta/generated/paper_PT.Rmd
refactor

--- 2020-05-08 generated/paper_PT.Rmd
@@
 library(questionr)
+
 # Set global options
@@
 
+# ```{r param-optimization-plots, echo=F, warning=F, message=F, include=F}
+# 
+# ggplot(fit_3) + xlab("Características") + ylab("RMSE")
+# ggplot(fit_7) + xlab("Características") + ylab("RMSE")
+# ggplot(fit_14) + xlab("Características") + ylab("RMSE")
+# ggplot(fit_28) + xlab("Características") + ylab("RMSE")
+# ggplot(fit_56) + xlab("Características") + ylab("RMSE")
+# ggplot(fit_100) + xlab("Características") + ylab("RMSE")
+# ```
+
 # Resultados

### 2020-08-14 _meta/generated/paper_PT.tex
refactor

--- 2020-08-14 _meta/generated/paper_EN.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[
-]{article}
+\documentclass[]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
-% Use upquote if available, for straight quotes in verbatim environments
+% use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
+% use microtype if available
+\IfFileExists{microtype.sty}{%
 \usepackage[]{microtype}
@@
 }{}
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{hyperref}
 \hypersetup{
-  pdftitle={Concrete compressive strength prediction with machine learning},
+            pdftitle={Prevendo a resistência à compressão do concreto com técnicas de machine learning},
             pdfauthor={Pedro Bernardino Alves Moreira},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage[margin=1in]{geometry}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
 \setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\renewcommand{\abstractname}{Resumo}
 \usepackage{booktabs}
@@
 
-\title{Concrete compressive strength prediction with machine learning}
+\title{Prevendo a resistência à compressão do concreto com técnicas de machine
+learning}
 \author{Pedro Bernardino Alves Moreira}
-\date{April 22, 2020}
+\date{Abril 22, 2020}
 
@@
 \begin{abstract}
-Compressive strength is the main characteristic of concrete. The correct
-prediction of this parameter means cost and time reduction. This work
-built predictive models for 6 different ages of concrete samples (3, 7,
-14, 28, 56, and 100 days). A set of data obtained in previous studies
-was used, a total of 1030 samples, with 9 variables: compressive
-strength, age, and 7 ingredients (water, cement, fine aggregate, coarse
-aggregate, fly ash, blast furnace slag, and superplasticizers). Another
-6 variables were added to represent the proportions of the main
-ingredients in each sample (water/cement, fine aggregate/cement, coarse
-aggregate/cement, fine aggregate/coarse aggregate, water/coarse
-aggregate, and water/fine aggregate). The predictive models were
-developed in \emph{R} language, using the \emph{caret} package with the
-\emph{Parallel Random Forest} algorithm and repeated cross-validation
-technique to optimize the parameters. The results were satisfactory and
-compatible with other studies using the same data set. The most
-important model, 28 days old, obtained \emph{RMSE} of 4.717. The 3-day
-model obtained the best result, \emph{RMSE} of 3.310. The worst result
-was the 56-day model, with \emph{RMSE} of 5.939. The work showed that
-the compressive strength of concrete can be predicted. The choice of
-creating a model for each age, instead of using age as a predictor,
-allowed to get compatible results with the available data at each age.
-It was a promising alternative since good results were achieved by
-training with just one algorithm. This work facilitates exploration and
-new efforts to predict the compressive strength of concrete, it can be
-replicated using different algorithms or the combination of several.
+A resistência à compressão é a principal característica do concreto. A
+previsão correta desse parâmetro significa redução de custo e tempo.
+Esse trabalho construiu modelos de previsão em 6 idades diferentes de
+amostras de concreto (3, 7, 14, 28, 56, e 100 dias). Foi utilizado um
+conjunto de dados obtido em estudos anteriores, um total de 1030
+amostras, com 9 variáveis: resistência à compressão, idade e 7
+ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza
+volante, escória granulada de alto forno e superplastificantes). Outras
+6 variáveis foram adicionadas para representar as proporções dos
+principais ingredientes em cada amostra (água/cimento, agregado
+miúdo/cimento, agregado graúdo/cimento, agregado miúdo/agregado graúdo,
+água/agregado graúdo e água/agregado miúdo). Os modelos de previsão
+foram desenvolvidos em linguagem \emph{R}, utilizando o pacote
+\emph{caret} com o algoritmo \emph{Parallel Random Forest} e técnica de
+validação cruzada repetida para otimização dos parâmetros. Os resultados
+foram satisfatórios e compatíveis com outros estudos utilizando o mesmo
+conjunto de dados. O modelo mais importante, de 28 dias, obteve
+\emph{RMSE} de 4,717. O modelo de 3 dias obteve o melhor resultado,
+\emph{RMSE} de 3,310. O pior resultado foi o modelo de 56 dias, com
+\emph{RMSE} de 5,939. O trabalho mostrou que a resistência à compressão
+do concreto pode ser prevista. A escolha de criar um modelo para cada
+idade, ao invés de utilizar a idade como característica para previsão,
+permitiu chegar a resultados compatíveis com os dados disponíveis de
+cada idade. Foi uma alternativa promissora, visto que bons resultados
+foram atingidos treinando com apenas um algoritmo. Esse trabalho
+facilita a exploração e novos esforços para previsão da resistência à
+compressão do concreto, ele pode ser replicado utilizando diferentes
+algoritmos ou a combinação de diversos.
 \end{abstract}
 
+\renewcommand{\figurename}{Figura}
+\renewcommand{\tablename}{Tabela}
+\renewcommand{\contentsname}{Sumário}
+\renewcommand{\chunkname}{Código}
+\renewcommand{\listfigurename}{Lista de figuras}
+\renewcommand{\listtablename}{Lista de tabelas}
 \newpage
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Compressive strength is the main characteristic of concrete, measured by
-tests of international standards that consist of the breaking of
-specimens. Measurement at 28 days is mandatory and represents the grade
-of the concrete. Knowing in advance what the result will be obtained for
-a given age, based on the proportions of its ingredients, is of great
-interest to concrete manufacturers, construction companies, and civil
-engineers.
+\hypertarget{introduuxe7uxe3o}{%
+\section{Introdução}\label{introduuxe7uxe3o}}
+
+A resistência à compressão é a principal característica do concreto,
+medida por testes de padrões internacionais que consistem na quebra de
+corpos de prova. A medição aos 28 dias é obrigatória e representa a
+classe do concreto. Saber com antecêdencia qual o resultado será obtido
+para uma determinada idade, a partir das proporções de seus
+ingredientes, é de grande interesse para os fabricantes de concreto,
+construtoras e engenheiros civis.
 
@@
 
-This compressive strength is a nonlinear function of its ingredients and
-age, making it difficult to establish an analytical formula, although
-some formulas have already been proposed. Hasan (2011) proposed a
-mathematical model to predict from the results of tests of 7 and 14
-days, and Kabir (2012) from 7 days. However, machine learning techniques
-can be used to model this characteristic from real sample data, using
-only the ingredients.
+Essa resistência à compressão é uma função não linear de seus
+ingredientes e idade, tornando difícil o estabelecimento de uma fórmula
+analítica, apesar de algumas fórmulas já haveram sido propostas. Hasan
+(2011) propós um modelo matemático para prever a partir dos resultados
+de testes de 7 e 14 dias, e Kabir (2012) a partir de 7 dias. Porém
+técnicas de machine learning podem ser utilizadas para modelar essa
+característica a partir de dados reais de amostras, utilizando apenas os
+ingredientes.
 
@@
 
-Many previous studies use the same dataset used by Yeh (1998) to predict
-the compressive strength of concrete. Alshamiri (2020) got good results
-with the regularized extreme learning machine (RELM) technique, and
-Hameed (2020) got even better results with the Artificial Neural
-Networks and cross-validation technique. This set of samples is so well
-known that there are many pages on the internet of unpublished studies
-that use it and have good results, such as Abban (2016), Raj (2018),
-Modukuru (2020) and Pierobon (2018). At the end of the work, the results
-found are compared to the works cited here.
+Muitos estudos anteriores utilizam o mesmo conjunto de dados utilizado
+por Yeh (1998) para prever a resistência à compressão do concreto.
+Alshamiri (2020) obteve bons resultados com a técnica de regularized
+extreme learning machine (RELM), e Hameed (2020) obteve resultados ainda
+melhores com a técnica de Artificial Neural Networks e cross-validation.
+Esse conjunto de amostras é tão conhecido que há ainda páginas na
+internet de estudos não publicados que o utilizam e possuem bons
+resultados, como Abban (2016), Raj (2018), Modukuru (2020) e Pierobon
+(2018). Ao final do trabalho os resultados encontrados são comparados
+aos trabalhos citados aqui.
 
@@
 
-Unlike previous studies with this dataset, this work does data
-preparation differently. The age of the concrete is the most unique
-feature that contributes to its compressive strength. For this reason,
-age is treated separately in the machine learning models, creating
-models for each age group.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{materials}{%
-\subsection{Materials}\label{materials}}
-
-The methodology was carried out using RStudio software (RStudio Team
-2020), an integrated virtual environment for code development in
-\emph{R} (R Core Team 2020). Throughout the process, all code executed
-was documented in the same order as its execution in
-\protect\hyperlink{appendix2}{Appendix 2}, and reference was always made
-to codes throughout the text. All relevant information related to the
-operating system and installed packages has been presented in
-\protect\hyperlink{appendix1}{Appendix 1}.
-
-\hypertarget{reproducibility}{%
-\subsection{Reproducibility}\label{reproducibility}}
-
-In order to guarantee reproducibility, whenever there was some code that
-could use probabilistic operations, a \emph{seed} was defined before its
-execution, ensuring that when run on another machine, with the same
-version of \emph{R} and the same \emph{seed}, get the same result. The
-\emph{seeds} can be checked throughout
-\protect\hyperlink{appendix2}{Appendix 2}.
-
-\hypertarget{obtaining-the-data}{%
-\subsection{Obtaining the data}\label{obtaining-the-data}}
-
-The data was downloaded from the University of California Irvine website
-(``Concrete Compressive Strength Data Set'' 2008)
-(\ref{show-download-data}). In total there are 1030 samples with 9
-columns. The samples were renamed and an id column was added to
-facilitate data manipulation (\ref{show-rename-dat-cols}). The columns
-were reordered to put the new id column in the first position
-(\ref{show-reorder-dat}). The first samples can be viewed in the table
-\ref{tab:first-samples}.
+Diferente dos estudos anteriores com esse conjunto de amostras, este
+trabalho faz a preparação dos dados de forma diferente. A idade do
+concreto é a variável mais singular que contribui para sua resistência à
+compressão, por esse motivo, a idade é tratada separadamente nos modelos
+de machine learning, criando modelos para cada faixa de idade.
+
+\hypertarget{metodologia}{%
+\section{Metodologia}\label{metodologia}}
+
+\hypertarget{materiais-utilizados}{%
+\subsection{Materiais utilizados}\label{materiais-utilizados}}
+
+A metodologia foi realizada utilizando o software RStudio (RStudio Team
+2020), ambiente virtual integrado para desenvolvimento de código na
+linguagem \emph{R} (R Core Team 2020). Ao longo do processo, todo código
+executado foi documentado na mesma ordem de sua execução no
+\protect\hyperlink{appendix3}{Apêndice 3}, e foi sempre realizada
+referência aos códigos ao longo do texto. Todas as informações
+relevantes relacionados ao sistema operacional e pacotes instalados
+foram apresentados no \protect\hyperlink{appendix1}{Apêndice 1}. Além
+disso foi criado um repositorio online e de código aberto no
+\emph{Github}, abrigando todo o código utilizado para gerar esse
+trabalho, o link foi diponibilizado no
+\protect\hyperlink{appendix2}{Apêndice 2}.
+
+\hypertarget{reprodutibilidade}{%
+\subsection{Reprodutibilidade}\label{reprodutibilidade}}
+
+Para garantir a reprodutibilidade, sempre que houve algum código que
+podesse utilizar operações probabilísticas, foi definido um \emph{seed}
+antes da sua execução, garantindo que quando rodado em outra máquina,
+com a mesma versão de \emph{R} e o mesmo \emph{seed}, chegue ao mesmo
+resultado. Os \emph{seeds} podem ser conferidos ao longo do
+\protect\hyperlink{appendix}{apêndice 3} ou diretamente no
+\emph{Github}.
+
+\hypertarget{obtenuxe7uxe3o-dos-dados}{%
+\subsection{Obtenção dos dados}\label{obtenuxe7uxe3o-dos-dados}}
+
+O download dos dados foi realizado no website da Universidade da
+California Irvine (``Concrete Compressive Strength Data Set'' 2008)
+(\ref{show-download-data}). No total são 1030 amostras com 9 colunas. As
+amostras foram renomeadas e foi adicionado uma coluna de id para
+facilitar na manipulação dos dados (\ref{show-rename-dat-cols}). As
+colunas foram reordenadas para colocar a nova coluna id em primeira
+posição (\ref{show-reorder-dat}). As primeiras amostras podem ser
+visualizadas na tabela \ref{tab:first-samples}.
 
@@
 
+\hypertarget{preparauxe7uxe3o-dos-dados}{%
+\subsection{Preparação dos dados}\label{preparauxe7uxe3o-dos-dados}}
+
+A preparação dos dados consistiu em transformar o conjunto de amostra
+afim de manter apenas dados relevantes para os estudos subsequêntes.
+Foram retirandos dados considerados irrelevantes ou que com potencial de
+adicionar um ruído indesejavel nas análises. Além disso, os dados
+relevântes foram transformados para melhor se enquadrar para os estudos
+nas próximas etapas.
+
+\hypertarget{limpeza-inicial-dos-dados}{%
+\subsubsection{Limpeza inicial dos
+dados}\label{limpeza-inicial-dos-dados}}
+
+Inicialmente foram observadas a existência de 25 amostras duplicadas que
+foram retiradas, resultando em um novo total de 1005 amostras
+(\ref{show-removing-duplicated-samples}).
+
+~
+
+Os dados apresentam as variaveis nas colunas e amostras nas linhas.
+Porém foi verificado que algumas amostras são identicas em proporções de
+ingredientes, alterando apenas o valor da idade e resistência à
+compressão, por exemplo as amostras 653, 654, 678 e 681, mostradas na
+tabela \ref{tab:similar-samples}.
+
+~
+
 \begin{table}[H]
 
-\caption{\label{tab:first-samples}First 6 samples}
+\caption{\label{tab:similar-samples}Amostras com a mesma composição}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F} & \multicolumn{1}{c}{C.Volante} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{Dia} & \multicolumn{1}{c}{Comp.Str.} \\
  & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
 \midrule
-1 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1040.0 & 676.0 & 28 & 79.99\\
-\addlinespace
-2 & 540.0 & 0.0 & 0 & 162 & 2.5 & 1055.0 & 676.0 & 28 & 61.89\\
-\addlinespace
-3 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 270 & 40.27\\
-\addlinespace
-4 & 332.5 & 142.5 & 0 & 228 & 0.0 & 932.0 & 594.0 & 365 & 41.05\\
-\addlinespace
-5 & 198.6 & 132.4 & 0 & 192 & 0.0 & 978.4 & 825.5 & 360 & 44.30\\
-\addlinespace
-6 & 266.0 & 114.0 & 0 & 228 & 0.0 & 932.0 & 670.0 & 90 & 47.03\\
+653 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 3 & 4.57\\
+\addlinespace
+678 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 7 & 7.68\\
+\addlinespace
+681 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 28 & 17.28\\
+\addlinespace
+654 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 90 & 25.46\\
 \bottomrule
@@
 
-\hypertarget{data-preparation}{%
-\subsection{Data preparation}\label{data-preparation}}
-
-The preparation of the data consisted of transforming the sample set in
-order to maintain only relevant data for the subsequent studies. Data
-that were considered irrelevant or that had the potential to add
-undesirable noise to the analysis were removed. In addition, the
-relevant data has been transformed to better fit the studies in the next
-steps.
-
-\hypertarget{initial-data-cleaning}{%
-\subsubsection{Initial data cleaning}\label{initial-data-cleaning}}
-
-Initially, there were 25 duplicate samples that were removed, resulting
-in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).
+Além disso, também existem amostras com os mesmos valores e proporções
+de ingredientes, mas com resistência à compressão diferente,
+provavelmente devido a diferenças na execução. É o caso por exemplo das
+amostras 472, 473 e 474, mostradas na tabela
+\ref{tab:similar-samples-2}.
 
@@
 
-The data show the variables in the columns and samples in the rows.
-However it was found that some samples are identical in proportions of
-ingredients, changing only the value of age and compressive strength,
-for example, samples 653, 654, 678 and 681, shown in the table
-\ref{tab:similar-samples}.
-
-~
-
-\begin{table}[H]
-
-\caption{\label{tab:similar-samples}Samples with same composition}
+\begin{table}
+
+\caption{\label{tab:similar-samples-2}Amostras iguais com resultados diferentes}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F} & \multicolumn{1}{c}{C.Volante} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{Dia} & \multicolumn{1}{c}{Comp.Str.} \\
  & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
 \midrule
-653 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 3 & 4.57\\
-\addlinespace
-678 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 7 & 7.68\\
-\addlinespace
-681 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 28 & 17.28\\
-\addlinespace
-654 & 102 & 153 & 0 & 192 & 0 & 887 & 942 & 90 & 25.46\\
+472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 57.03\\
+\addlinespace
+473 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 44.42\\
+\addlinespace
+474 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 51.02\\
 \bottomrule
@@
 
-In addition, there are also samples with the same values and proportions
-of ingredients, but with different compressive strength, probably due to
-differences in the building process. This is the case, for example, of
-samples 472, 473 and 474, shown in the table
-\ref{tab:similar-samples-2}.
+Para facilitar a analise das amostras, todas as amostras que são iguais
+em relação aos ingredientes, foram atribuidos o mesmo \emph{id}. Além
+disso, como a resistência à compressão aos 28 dias é o parâmetro para
+determinar a classe do concreto, foi mantido apenas os elementos que
+contenham esse dia entre suas amostras. No caso das amostras iguais mas
+com resultado diferentes de resistência à compressão, foi calculado a
+média dos valores. Após todas essas alterações
+(\ref{show-initial-data-cleaning}), o novo total de amostras foi
+reduzido para 970, contendo 416 configurações diferentes das proporções
+de ingredientes.
 
@@
 
-\begin{table}
-
-\caption{\label{tab:similar-samples-2}Same samples with different results}
+O resultado pode ser conferido na tabela
+\ref{tab:similar-samples-same-id}. Todas as amostras com configurações
+iguais de ingredientes possuem o mesmo id, e quando possuiam resultados
+diferentes para os mesmos dias, foram transformadas em apenas uma
+amostra, com a média aritmética na resistência à compressão.
+
+~
+
+\begin{table}[H]
+
+\caption{\label{tab:similar-samples-same-id}Amostras anteriores após processamento}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
- & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
-\midrule
-472 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 57.03\\
-\addlinespace
-473 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 44.42\\
-\addlinespace
-474 & 446 & 24 & 79 & 162 & 11.6 & 967 & 712 & 28 & 51.02\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-To facilitate the analysis of the samples, all samples that are the same
-in relation to the ingredients, have been assigned the same \emph{id}.
-In addition, as the compressive strength at 28 days is the parameter to
-determine the grade of the concrete, only the elements containing that
-day among its samples were maintained. In the case of the same samples
-but with different results of compressive strength, the values were
-averaged. After all these changes (\ref{show-initial-data-cleaning}),
-the new total of samples was reduced to 970, containing 416 different
-settings for the proportions of ingredients.
-
-~
-
-The result can be seen in the table \ref{tab:similar-samples-same-id}.
-All samples with equal ingredient settings have the same id, and when
-they had different results for the same days, they were transformed into
-just one sample, with the arithmetic mean in the compressive strength.
-
-~
-
-\begin{table}[H]
-
-\caption{\label{tab:similar-samples-same-id}Previous samples after processing}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccccccc}
-\toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{C.Aggregate} & \multicolumn{1}{c}{F.Aggregate} & \multicolumn{1}{c}{Day} & \multicolumn{1}{c}{Comp.Str.} \\
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F} & \multicolumn{1}{c}{C.Volante} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{Dia} & \multicolumn{1}{c}{Comp.Str.} \\
  & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ &  & $MPa$\\
@@
 
-\hypertarget{age-selection}{%
-\subsubsection{Age selection}\label{age-selection}}
-
-As previously described, the main age for analysis of compressive
-strength is 28 days, but other ages can also be used to build predictive
-models. However, it is necessary to verify how relevant the data of
-these other ages are. Starting with the distribution of the samples in
-relation to each age (\ref{show-boxplot}) shown in the figure
-\ref{fig:boxplot}.
+\hypertarget{seleuxe7uxe3o-das-idades}{%
+\subsubsection{Seleção das idades}\label{seleuxe7uxe3o-das-idades}}
+
+Como descrito anteriormente, a principal idade para análise da
+resistência à compressão é aos 28 dias, mas as outras idades também
+podem ser utilizadas para construir modelos de previsão. Porém é
+necessário verificar o quanto relevante os dados dessas outras idades
+são. Iniciando pela distribuição das amostras em relação a cada idade
+(\ref{show-boxplot}) mostrado na figura \ref{fig:boxplot}.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/boxplot-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/boxplot-1}
 
@@
 
-\caption{Boxplot - Compressive strength (MPa) vs age (days)}\label{fig:boxplot}
+\caption{Boxplot - Resistência à compressão (MPa) vs idade (dias)}\label{fig:boxplot}
 \end{figure}
 
-It was observed that the ages of 90, 91 and 100 days probably represent
-extremes to each other in the ingredient configurations, since they are
-relatively close ages but with very different values, especially for 90
-and 91.
+Foi observado que as idades de 90, 91 e 100 dias provavelmente
+representam extremos entre si das configurações de ingredientes, uma vez
+que são idades relativamente próximas porém com valores muito
+diferentes, especialmente para 90 e 91.
 
@@
 
-This hypothesis was verified using the principal component analysis
-method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}).
-The figure \ref{fig:pca-90-91-100} shows how the samples relate to each
-other (which are similar or different) and revealed how each variable
-contributes to the analysis. The first two dimensions represent 37\% and
-24\% \$ respectively of the variance.
+Essa hipótese foi verificada utilizando o método de análise de
+componente principal, aplicado às amostras dessas 3 idades
+(\ref{show-pca-90-91-100}). A fígura \ref{fig:pca-90-91-100} mostra como
+as amostras se relacionam umas com as outras (quais são parecidas ou
+diferentes) e revelou como cada variável contribui para a análise. As
+duas primeiras dimensões representam 37\% e 24\% respectivamente da
+variância.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/pca-90-91-100-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/pca-90-91-100-1}
 
@@
 
-\caption{Principal component analysis - 90, 91 e 100 days}\label{fig:pca-90-91-100}
+\caption{Análise componente principal - 90, 91 e 100 dias}\label{fig:pca-90-91-100}
 \end{figure}
 
-Another important point considered, of the concrete nature itself, is
-the fact that the growth rate of its compressive strength decreases with
-time, reaching a certain stability value. The figure
-\ref{fig:mpa-on-time} shows the compressive strength over the days for
-samples with more than 5 data, that is, data available for at least 6
-different ages (\ref{show-mpa-on-time}).
+Outro ponto importante considerado, da própria natureza do concreto, é o
+fato da taxa de crescimento de sua resistência à compressão diminuir com
+o tempo, chegando a um certo valor de establidade. A figura
+\ref{fig:mpa-on-time} mostra a resistência à compressão ao longo dos
+dias para amostras com mais de 5 dados, ou seja, dados disponiveis para
+no mínimo 6 idades distintas (\ref{show-mpa-on-time}).
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/mpa-on-time-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/mpa-on-time-1}
 
@@
 
-\caption{Compressive strength through time}\label{fig:mpa-on-time}
+\caption{Resistência à compressão ao longo do tempo}\label{fig:mpa-on-time}
 \end{figure}
 
-For the reasons presented in the figures \ref{fig:boxplot},
-\ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered
-that the ages of 90, 91 and 100 days can be grouped to improve reading
-and decrease sample noise. They were converted to the same value, the
-age of 100 days was chosen (\ref{show-join-90-91-100}). As shown in the
-figure \ref{fig:mpa-on-time}, the resistance only increases, after 100
-days the resistance to compression will be greater than or equal to the
-value of 90 or 91 days.
+Pelos motivos apresentados nas figuras \ref{fig:boxplot},
+\ref{fig:pca-90-91-100} e \ref{fig:mpa-on-time}, foi considerado que as
+idades de 90, 91 e 100 dias podem ser agrupadas para melhorar a leitura
+e diminuir o ruído das amostras. Elas foram convertidas para o mesmo
+valor, que no caso foi escolhido a idade de 100 dias
+(\ref{show-join-90-91-100}), pois como mostrado na figura
+\ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a
+resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
 
@@
 
-Another topic analyzed in the selection of ages was the observed
-frequency of each age value after this transformation from 90, 91 in 100
-days, shown in the figure \ref{fig:freq-ages}. Some values of days have
-very low concentrations of samples, at the risk of damaging more than
-helping to create the models, so they were removed
-(\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain
-only ages with a frequency greater than 50, only the values of 3, 7, 14,
-28, 56 and 100 days.
+Mais um tópico analisado na seleção das idades foi a frequência
+observada de cada valor de idade após essa transformação dos 90, 91 em
+100 dias, mostrada na figura \ref{fig:freq-ages}. Alguns valores de dias
+apresentam concentrações muito baixas de amostras, correndo o risco de
+prejudicar mais do que ajudar na criação dos modelos, logo elas foram
+removidas (\ref{show-remove-ages-lower-50}). O critério adotado foi
+manter apenas idades com frequência maior que 50, ou seja, apenas os
+valores de 3, 7, 14, 28, 56 e 100 dias.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/freq-ages-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/freq-ages-1}
 
@@
 
-\caption{Ages frequency}\label{fig:freq-ages}
+\caption{Frequência das idades}\label{fig:freq-ages}
 \end{figure}
 
-\hypertarget{data-reorganization}{%
-\subsubsection{Data reorganization}\label{data-reorganization}}
-
-The samples were grouped to maintain only one sample from each set of
-configuration of the proportions of the ingredients, adding new
-variables/columns for resistance at each age
-(\ref{show-reorganizing-dat}). The result in the first samples after
-this processing is shown in the table \ref{tab:new-features}.
+\hypertarget{reorganizauxe7uxe3o-dos-dados}{%
+\subsubsection{Reorganização dos
+dados}\label{reorganizauxe7uxe3o-dos-dados}}
+
+As amostras foram agrupadas para manter apenas uma amostra distinta de
+cada conjunto de configuração das proporções dos ingredientes,
+adicionando novas variáveis/colunas para a resistência em cada idade
+(\ref{show-reorganizing-dat}). O resultado nas primeiras amostras após
+esse processamento é mostrado na tabela \ref{tab:new-features}.
 
@@
 
-\caption{\label{tab:new-features}First 6 samples after reorganization}
+\caption{\label{tab:new-features}Primeiras 6 amostras reorganizadas}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{Water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Ag.} & \multicolumn{1}{c}{Fine Ag.} & \multicolumn{1}{c}{3 days} & \multicolumn{1}{c}{7 days} & \multicolumn{1}{c}{14 days} & \multicolumn{1}{c}{28 days} & \multicolumn{1}{c}{56 days} & \multicolumn{1}{c}{100 days} \\
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F.} & \multicolumn{1}{c}{C.Volante} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{3 dias} & \multicolumn{1}{c}{7 dias} & \multicolumn{1}{c}{14 dias} & \multicolumn{1}{c}{28 dias} & \multicolumn{1}{c}{56 dias} & \multicolumn{1}{c}{100 dias} \\
  & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$ & $MPa$\\
@@
 
-The number of samples and distinct samples after all this manipulation
-remained the same, a total of 416 (\ref{show-total-samples-2}).
-
-\hypertarget{adding-new-variables}{%
-\subsubsection{Adding new variables}\label{adding-new-variables}}
-
-To finish the data preparation, new columns were added to the dataset
-(\ref{show-new-features-2}). Starting with the concrete class, for
-example if the compressive strength is between 25 and 30, it receives
-the class \emph{C25}. The inclusion of the class was important because
-the compressive strength in \emph{MPa} is a continuous variable, which
-will be used in the regression models, but the class as a discrete
-variable can provide another visualization of the data. The approximate
-mix of concrete was also added, which represents the proportions of
-aggregates (fine and coarse) for cement. Other proportions between the
-main ingredients were also added. The new variables are presented in the
-table \ref{tab:new-features-table}.
+O número de amostras e amostras distintas após toda essa manipulação
+permaneceu o mesmo, um total de 416 (\ref{show-total-samples-2}).
+
+\hypertarget{adicionando-novas-variuxe1veis}{%
+\subsubsection{Adicionando novas
+variáveis}\label{adicionando-novas-variuxe1veis}}
+
+Para finalizar a preparação dos dados, novas colunas foram adicionadas
+ao conjunto de amostras (\ref{show-new-features-2}). Iniciando pela
+classe do concreto, por exemplo se a resistência à compressão está entre
+25 e 30, recebe a classe \emph{C25}. A inclusão da classe foi importânte
+pois a resistência em \(MPa\) é uma variável contínua, que será
+utilizada nos modelos de regressão, mas a classe como variável discreta
+pode fornecer outro ângulo de visualização dos dados. Também foi
+adicionado o traço aproximado do concreto, que representa as proporções
+de agregados (miúdo e graúdo) para o cimento. Outras proporções entre os
+principais ingredientes também foram adicionadas. As novas variáveis são
+apresentadas na tabela \ref{tab:new-features-table}.
 
@@
 
-\caption{\label{tab:new-features-table}New features}
+\caption{\label{tab:new-features-table}Novas variáveis}
 \centering
@@
 \toprule
-ID & Class & Approximated Mix & Water / Cement & Fine Ag. / Cement & Coarse Ag. / Cement & Fine Ag. / Coarse Ag. & Water / Coarse Ag. & Water / Fine Ag.\\
+ID & Classe & Traço aproximado & Água / Cimento & A.Miúdo / Cimento & A.Graúdo / Cimento & A.Miúdo / A.Graúdo & Água / A.Graúdo & Água / A.Miúdo\\
 \midrule
@@
 
-\hypertarget{data-visualization}{%
-\subsection{Data visualization}\label{data-visualization}}
-
-In order to assess the need for further manipulation before building the
-models, in this step the 416 samples already processed were visualized
-and analyzed.
-
-\hypertarget{descriptive-statistics}{%
-\subsubsection{Descriptive statistics}\label{descriptive-statistics}}
-
-The table \ref{tab:stat-summ} presents the statistical data of the
-continuous variables (\ref {show-stat-summ}). The \emph{Null} line
-represents the number of zeroed values for the ingredients, and the
-\emph{NA} line represents the number of missing data. As the samples
-were filtered to maintain only sets of samples with known values of
-compressive strength at 28 days, the number of \emph{NAs} is zero for
-that age. The figure \ref{fig:stat-summ-categorical} presents the
-statistical data of the discrete variables
-(\ref{show-stat-summ-categorical}).
+\hypertarget{visualizauxe7uxe3o-dos-dados}{%
+\subsection{Visualização dos dados}\label{visualizauxe7uxe3o-dos-dados}}
+
+Para avaliar a necessidade de mais manipulações antes da construção dos
+modelos, nesta etapa as 416 amostras já processadas foram visualizadas e
+analisadas.
+
+\hypertarget{estatistica-descritiva}{%
+\subsubsection{Estatistica descritiva}\label{estatistica-descritiva}}
+
+A tabela \ref{tab:stat-summ} apresenta os dados estatísticos das
+variáveis contínuas (\ref{show-stat-summ}). A linha de \emph{Null}
+representa o número de valores zerados para os ingredientes, e a linha
+\emph{NA} representa o número de dados faltando. Como as amostras foram
+filtradas para manter apenas conjuntos de amostras com valores
+conhecidos da resistência à compressão aos 28 dias, o número de
+\emph{NAs} é zero para essa idade. A figura
+\ref{fig:stat-summ-categorical} apresenta os dados estatísticos das
+variáveis discretas (\ref{show-stat-summ-categorical}).
 
@@
 
-\caption{\label{tab:stat-summ}Descriptive statistics - continuous variables}
+\caption{\label{tab:stat-summ}Estatística descritiva - variáveis contínuas}
 \centering
@@
 \toprule
-  & Samples & Null & NA & Min & Max & Range & Sum & Median & Mean & SE mean & CI mean & Variance & Std.Dev. & Coef.Var\\
+  & Amostras & Null & NA & Min & Max & Intervalo & Soma & Mediana & Média & Erro padrão  da média & Intervalo de confiânça da média & Variância & Desvio Padrão & Coeficiênte de variação\\
 \midrule
-Cement & 416 & 0 & 0 & 102.00 & 540.00 & 438.00 & 109373.10 & 257.70 & 262.92 & 5.10 & 10.02 & 10817.50 & 104.01 & 0.40\\
-\addlinespace
-B.F.S. & 416 & 174 & 0 & 0.00 & 359.40 & 359.40 & 35824.60 & 94.25 & 86.12 & 4.32 & 8.49 & 7755.00 & 88.06 & 1.02\\
-\addlinespace
-Fly ash & 416 & 202 & 0 & 0.00 & 200.10 & 200.10 & 26389.00 & 71.25 & 63.44 & 3.26 & 6.40 & 4407.81 & 66.39 & 1.05\\
-\addlinespace
-Water & 416 & 0 & 0 & 121.80 & 247.00 & 125.20 & 76335.60 & 185.00 & 183.50 & 0.94 & 1.86 & 370.73 & 19.25 & 0.10\\
+Cimento & 416 & 0 & 0 & 102.00 & 540.00 & 438.00 & 109373.10 & 257.70 & 262.92 & 5.10 & 10.02 & 10817.50 & 104.01 & 0.40\\
+\addlinespace
+E.G.A.F. & 416 & 174 & 0 & 0.00 & 359.40 & 359.40 & 35824.60 & 94.25 & 86.12 & 4.32 & 8.49 & 7755.00 & 88.06 & 1.02\\
+\addlinespace
+Cinza Volante & 416 & 202 & 0 & 0.00 & 200.10 & 200.10 & 26389.00 & 71.25 & 63.44 & 3.26 & 6.40 & 4407.81 & 66.39 & 1.05\\
+\addlinespace
+Água & 416 & 0 & 0 & 121.80 & 247.00 & 125.20 & 76335.60 & 185.00 & 183.50 & 0.94 & 1.86 & 370.73 & 19.25 & 0.10\\
 \addlinespace
@@
 \addlinespace
-Coarse agg. & 416 & 0 & 0 & 801.00 & 1145.00 & 344.00 & 397799.90 & 953.35 & 956.25 & 4.12 & 8.10 & 7063.06 & 84.04 & 0.09\\
-\addlinespace
-Fine agg. & 416 & 0 & 0 & 594.00 & 992.60 & 398.60 & 317809.80 & 769.65 & 763.97 & 3.59 & 7.06 & 5371.89 & 73.29 & 0.10\\
-\addlinespace
-3 days & 121 & 0 & 295 & 2.33 & 41.64 & 39.31 & 2210.82 & 15.52 & 18.27 & 0.87 & 1.72 & 91.64 & 9.57 & 0.52\\
-\addlinespace
-7 days & 114 & 0 & 302 & 7.51 & 59.09 & 51.58 & 2845.52 & 21.06 & 24.96 & 1.29 & 2.55 & 188.81 & 13.74 & 0.55\\
-\addlinespace
-14 days & 62 & 0 & 354 & 12.84 & 59.76 & 46.92 & 1782.56 & 26.54 & 28.75 & 1.10 & 2.19 & 74.62 & 8.64 & 0.30\\
-\addlinespace
-28 days & 416 & 0 & 0 & 8.54 & 81.75 & 73.21 & 15101.13 & 33.72 & 36.30 & 0.70 & 1.38 & 206.30 & 14.36 & 0.40\\
-\addlinespace
-56 days & 83 & 0 & 333 & 23.25 & 80.20 & 56.95 & 4178.77 & 50.77 & 50.35 & 1.52 & 3.02 & 190.82 & 13.81 & 0.27\\
-\addlinespace
-100 days & 120 & 0 & 296 & 21.86 & 82.60 & 60.74 & 5701.90 & 45.61 & 47.52 & 1.17 & 2.31 & 163.40 & 12.78 & 0.27\\
-\addlinespace
-Water / Cement & 416 & 0 & 0 & 0.27 & 1.88 & 1.62 & 340.60 & 0.73 & 0.82 & 0.02 & 0.03 & 0.11 & 0.34 & 0.41\\
-\addlinespace
-Fine agg. / Cement & 416 & 0 & 0 & 1.14 & 9.24 & 8.10 & 1415.82 & 2.94 & 3.40 & 0.07 & 0.14 & 1.96 & 1.40 & 0.41\\
-\addlinespace
-Coarse agg. / Cement & 416 & 0 & 0 & 1.55 & 8.70 & 7.14 & 1761.06 & 3.67 & 4.23 & 0.08 & 0.16 & 2.70 & 1.64 & 0.39\\
-\addlinespace
-Fine agg. / Coarse agg. & 416 & 0 & 0 & 0.53 & 1.16 & 0.63 & 335.42 & 0.80 & 0.81 & 0.01 & 0.01 & 0.01 & 0.11 & 0.14\\
-\addlinespace
-Water / Coarse agg. & 416 & 0 & 0 & 0.12 & 0.29 & 0.17 & 80.66 & 0.19 & 0.19 & 0.00 & 0.00 & 0.00 & 0.03 & 0.16\\
-\addlinespace
-Water / Fine agg. & 416 & 0 & 0 & 0.13 & 0.38 & 0.26 & 101.26 & 0.24 & 0.24 & 0.00 & 0.00 & 0.00 & 0.04 & 0.17\\
+A.Graúdo & 416 & 0 & 0 & 801.00 & 1145.00 & 344.00 & 397799.90 & 953.35 & 956.25 & 4.12 & 8.10 & 7063.06 & 84.04 & 0.09\\
+\addlinespace
+A.Miúdo & 416 & 0 & 0 & 594.00 & 992.60 & 398.60 & 317809.80 & 769.65 & 763.97 & 3.59 & 7.06 & 5371.89 & 73.29 & 0.10\\
+\addlinespace
+3 dias & 121 & 0 & 295 & 2.33 & 41.64 & 39.31 & 2210.82 & 15.52 & 18.27 & 0.87 & 1.72 & 91.64 & 9.57 & 0.52\\
+\addlinespace
+7 dias & 114 & 0 & 302 & 7.51 & 59.09 & 51.58 & 2845.52 & 21.06 & 24.96 & 1.29 & 2.55 & 188.81 & 13.74 & 0.55\\
+\addlinespace
+14 dias & 62 & 0 & 354 & 12.84 & 59.76 & 46.92 & 1782.56 & 26.54 & 28.75 & 1.10 & 2.19 & 74.62 & 8.64 & 0.30\\
+\addlinespace
+28 dias & 416 & 0 & 0 & 8.54 & 81.75 & 73.21 & 15101.13 & 33.72 & 36.30 & 0.70 & 1.38 & 206.30 & 14.36 & 0.40\\
+\addlinespace
+56 dias & 83 & 0 & 333 & 23.25 & 80.20 & 56.95 & 4178.77 & 50.77 & 50.35 & 1.52 & 3.02 & 190.82 & 13.81 & 0.27\\
+\addlinespace
+100 dias & 120 & 0 & 296 & 21.86 & 82.60 & 60.74 & 5701.90 & 45.61 & 47.52 & 1.17 & 2.31 & 163.40 & 12.78 & 0.27\\
+\addlinespace
+Água/ Cimento & 416 & 0 & 0 & 0.27 & 1.88 & 1.62 & 340.60 & 0.73 & 0.82 & 0.02 & 0.03 & 0.11 & 0.34 & 0.41\\
+\addlinespace
+A.Miúdo/ Cimento & 416 & 0 & 0 & 1.14 & 9.24 & 8.10 & 1415.82 & 2.94 & 3.40 & 0.07 & 0.14 & 1.96 & 1.40 & 0.41\\
+\addlinespace
+A.Graúdo/ Cimento & 416 & 0 & 0 & 1.55 & 8.70 & 7.14 & 1761.06 & 3.67 & 4.23 & 0.08 & 0.16 & 2.70 & 1.64 & 0.39\\
+\addlinespace
+A.Miúdo / A.Graúdo & 416 & 0 & 0 & 0.53 & 1.16 & 0.63 & 335.42 & 0.80 & 0.81 & 0.01 & 0.01 & 0.01 & 0.11 & 0.14\\
+\addlinespace
+Água / A.Graúdo & 416 & 0 & 0 & 0.12 & 0.29 & 0.17 & 80.66 & 0.19 & 0.19 & 0.00 & 0.00 & 0.00 & 0.03 & 0.16\\
+\addlinespace
+Água / A.Miúdo & 416 & 0 & 0 & 0.13 & 0.38 & 0.26 & 101.26 & 0.24 & 0.24 & 0.00 & 0.00 & 0.00 & 0.04 & 0.17\\
 \bottomrule
@@
 \end{table}
-
 \begin{figure}
 
-{\centering \includegraphics{paper_EN_files/figure-latex/stat-summ-categorical-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/stat-summ-categorical-1}
 
@@
 
-\caption{Descriptive statistics - categorical variables}\label{fig:stat-summ-categorical}
+\caption{Estatística descritiva - variáveis discretas}\label{fig:stat-summ-categorical}
 \end{figure}
 
-\hypertarget{correlation-between-ingredients-and-compressive-strength}{%
-\subsubsection{Correlation between ingredients and compressive
-strength}\label{correlation-between-ingredients-and-compressive-strength}}
-
-The figure \ref{fig:correlation} shows the correlation of variables for
-each set of ages (\ref{show-correlation}). The figure
-\ref{fig:correlation-mpa} presents the same data, but instead of
-correlating them all, it only correlates with the compressive strength,
-showing the values in more detail (\ref{show-correlation-mpa}).
+\hypertarget{correlauxe7uxe3o-dos-ingredientes-e-resistuxeancia-uxe0-compressuxe3o}{%
+\subsubsection{Correlação dos ingredientes e resistência à
+compressão}\label{correlauxe7uxe3o-dos-ingredientes-e-resistuxeancia-uxe0-compressuxe3o}}
+
+A figura \ref{fig:correlation} apresenta a correlação das variáveis para
+cada conjunto de idades (\ref{show-correlation}). A figura
+\ref{fig:correlation-mpa} apresenta os mesmos dados, mas em vez de
+correlacionar todos, correlaciona apenas com a resistência à compressão,
+mostrando os valores mais detalhadamente (\ref{show-correlation-mpa}).
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/correlation-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/correlation-1}
 
@@
 
-\caption{Correlations at each age}\label{fig:correlation}
+\caption{Correlações em cada idade}\label{fig:correlation}
 \end{figure}
-
 \begin{figure}
 
-{\centering \includegraphics{paper_EN_files/figure-latex/correlation-mpa-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/correlation-mpa-1}
 
@@
 
-\caption{Correlation of variables with compressive strength over time}\label{fig:correlation-mpa}
+\caption{Correlação das variáveis com a resistência à compressão no tempo}\label{fig:correlation-mpa}
 \end{figure}
 
-The interpretation of the figure \ref{fig:correlation-mpa} suggests that
-the strength of the concrete is positively related mainly to the cement
-and superplasticizer ingredients and negatively to the water and fine
-aggregate. The smaller the amount of cement for aggregates and water,
-the more negatively they are correlated with compressive strength.
+A interpretação da figura \ref{fig:correlation-mpa} sugere que a
+resistência do concreto está relacionada positivamente principalmente
+com os ingredientes cimento e superplastificante e negativamente com a
+água e agregado miúdo. Quanto menor a quantidade de cimento para os
+agregados e para água, mais negativamente estão correlacionados com a
+resistência à compressão.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/mix-app-mpa-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/mix-app-mpa-1}
 
@@
 
-\caption{Relationship between approximated mix, water, MPa and age}\label{fig:mix-app-mpa}
+\caption{Relação entre o traço aproximado, água, resistência à compressão e idade}\label{fig:mix-app-mpa}
 \end{figure}
-
 \begin{figure}
 
-{\centering \includegraphics{paper_EN_files/figure-latex/mix-mpa-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/mix-mpa-1}
 
@@
 
-\caption{Relationship between concrete main features}\label{fig:mix-mpa}
+\caption{Relação das principais proporções do concreto}\label{fig:mix-mpa}
 \end{figure}
 
-The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the
-relationship between the main ingredients (known as mix) in relation to
-the compressive strength (\ref{show-mix-app-mpa} and
-\ref{show-mix-mpa}). The interpretation of these figures shows that the
-greater the amount of cement in relation to the other ingredients, the
-greater the resistance to compression.
-
-\hypertarget{variables-distribution}{%
-\subsubsection{Variables distribution}\label{variables-distribution}}
-
-The figure\ref{fig:vars-distribution} shows the distribution of
-variables in the samples (\ref{show-vars-distribution}). It was
-calculated using data only at 28 days.
+As figuras \ref{fig:mix-app-mpa} e \ref{fig:mix-mpa} mostram a relação
+entre os principais ingredientes (conhecida como traço) em relação à
+resistência à compressão (\ref{show-mix-app-mpa} e \ref{show-mix-mpa}).
+A interpretação dessas figuras mostra que quanto maior a quantidade de
+cimento em relação aos outros ingredientes, maior será a resistência à
+compressão.
+
+\hypertarget{distribuiuxe7uxe3o-das-variuxe1veis}{%
+\subsubsection{Distribuição das
+variáveis}\label{distribuiuxe7uxe3o-das-variuxe1veis}}
+
+A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis
+nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando
+apenas os dados aos 28 dias.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/vars-distribution-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/vars-distribution-1}
 
@@
 
-\caption{Variables distribution}\label{fig:vars-distribution}
+\caption{Distribuição das variáveis}\label{fig:vars-distribution}
 \end{figure}
 
-The figure \ref{fig:vars-distribution-time} shows the distribution of
-ingredients and compressive strength for each set of ages
-(\ref{show-vars-distribution-time}), in case of the 28 days it presents
-the same information as the figure \ref{fig:vars-distribution}. It shows
-that the resistance to compression gradually increases over time, as
-expected. Furthermore it is seen that the concentration of the
-ingredients can vary a lot when stratified by ages.
+De outra forma, a figura \ref{fig:vars-distribution-time} mostra a
+distribuição dos ingredientes e da resistência à compressão para cada
+conjunto de idades (\ref{show-vars-distribution-time}), ou seja, no caso
+dos 28 dias apresenta a mesma informação que a figura
+\ref{fig:vars-distribution}. Através dela é visualizado que como
+esperado, a resistência à compressão gradualmente aumenta ao longo do
+tempo. Além disso é visualizado que a concentração dos ingredientes
+podem variar muito quando estratíficado pelas idades.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/vars-distribution-time-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/vars-distribution-time-1}
 
@@
 
-\caption{Variables distribution grouped by age}\label{fig:vars-distribution-time}
+\caption{Distribuição das variáveis em relação a idade}\label{fig:vars-distribution-time}
 \end{figure}
 
-\hypertarget{principal-component-analysis}{%
-\subsubsection{Principal component
-analysis}\label{principal-component-analysis}}
-
-In the figure \ref{fig:pca}, using an alternative classification, the
-principal component analysis was performed on the ingredients
-(\ref{show-pca}). The classification separates concrete into 4 different
-compressive strength groups, low up to \emph{20 MPa}, normal up to
-\emph{40 MPa}, medium up to \emph{70 MPa} and high above that. It is
-possible to notice that the groups overlap, but there is a
-differentiation between the high and low group.
+\hypertarget{anuxe1lise-de-componente-principal}{%
+\subsubsection{Análise de componente
+principal}\label{anuxe1lise-de-componente-principal}}
+
+Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi
+realizada a análise de componente principal nos ingredientes das
+amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos
+de resistência a compressão diferentes, baixo até \emph{20 MPa}, normal
+até \emph{40 MPa}, médio até \emph{70 MPa} e alto acima disso. É
+possivel perceber que os grupos se sobrepõem, mas existem uma
+diferenciação entre o grupo alto e baixo.
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/pca-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/pca-1}
 
@@
 
-\caption{Principal component analysis on ingredients}\label{fig:pca}
+\caption{Análise componente principal nos ingredientes}\label{fig:pca}
 \end{figure}
 
-\hypertarget{machine-learning-models}{%
-\subsection{Machine learning models}\label{machine-learning-models}}
-
-The development of machine learning models was carried out with the
-\emph{caret} package (Kuhn 2020) and based on Irizarry (2019) and Kuhn
+\hypertarget{modelos-de-machine-learning}{%
+\subsection{Modelos de machine
+learning}\label{modelos-de-machine-learning}}
+
+O desenvolvimento dos modelos de machine learning foi realizado com o
+pacote \emph{caret} (Kuhn 2020) e baseado em Irizarry (2019) e Kuhn
 (2008).
 
-\hypertarget{pre-processing-and-data-separation}{%
-\subsubsection{Pre-processing and data
-separation}\label{pre-processing-and-data-separation}}
-
-As the approximate mix is a categorical variable, it was converted into
-dummy variables (\ref{show-dummy-var}), going from 22 columns (id,
-class, compressive strength and 19 more \emph{features}) to 45 columns,
-an addition of 23 variables, one for each approximate mix.
+\hypertarget{pre-processamento-e-separauxe7uxe3o-dos-dados}{%
+\subsubsection{Pre processamento e separação dos
+dados}\label{pre-processamento-e-separauxe7uxe3o-dos-dados}}
+
+Como existe a variável categórica para o traço aproximado do concreto,
+foi realizada a conversão dessa variável em variáveis fictiícias
+(\emph{dummy vars}) (\ref{show-dummy-var}), passando de 22 colunas (id,
+classe, resistência à compressão e mais 19 \emph{features}) para 45
+colunas, uma adição de 23 variáveis, uma para cada traço aproximado.
 
@@
 
-The samples were separated based on age. One dataset was created for
-each age value, totaling 6 different sets (\ref{show-preparation}). For
-illustrative purposes, the first 18 of 45 columns of the first 6 samples
-from the 28-day set are shown in the table
-\ref{tab:table-preparated-samples}.
+As amostras foram separadas baseado nas idades. Foram criados um
+conjunto de dados para cada valor de idade, totalizando 6 conjuntos
+diferentes (\ref{show-preparation}). Para fins ilustrativos, as
+primeiras 18 de 45 colunas das primeiras 6 amostras do conjunto de 28
+dias são mostradas na tabela \ref{tab:table-preparated-samples}.
 
@@
 
-\caption{\label{tab:table-preparated-samples}First 18 columns of the 6 first samples of 28 days}
+\caption{\label{tab:table-preparated-samples}Primeiras 18 colunas das primeiras 6 amostras de 28 dias}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Agg.} & \multicolumn{1}{c}{Fine Agg.} & \multicolumn{1}{c}{MPa} & \multicolumn{1}{c}{Class} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{C.Agg./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{Wat./} & \multicolumn{2}{c}{App Mix} \\
- & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ &  & Ci. & Ce. & Ce. & C.Agg. & C.Agg. & F.Agg. & 1:1:2 & 1:2:2\\
+\multicolumn{1}{c}{ID} & \multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F.} & \multicolumn{1}{c}{C.Vol.} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{MPa} & \multicolumn{1}{c}{Classe} & \multicolumn{1}{c}{Ág./} & \multicolumn{1}{c}{A.M./} & \multicolumn{1}{c}{A.G./} & \multicolumn{1}{c}{A.M./} & \multicolumn{1}{c}{Ág./} & \multicolumn{1}{c}{Ág./} & \multicolumn{2}{c}{Traço Apox.} \\
+ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $MPa$ &  & Ci. & Ci. & Ci. & A.G. & A.G. & Ag.M. & 1:1:2 & 1:2:2\\
 \midrule
@@
 
-For each of the 6 sets, the existence or not of variables with variance
-close to zero and their subsequent removal was verified
-(\ref{show-nzv}). Many of the 23 variables added referring to the
-approximate concrete mix were removed due to this fact. In addition to
-them, in the case of the 7-day set, the fly ash variable was also
-removed. Then it was verified that there are no variables with high
-correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}).
-After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25
-columns respectively for ages in increasing sequence.
+Para cada um dos 6 conjuntos foi verificado a existência ou não de
+variáveis com variância próxima a zero e sua subsequênte remoção
+(\ref{show-nzv}). Muitas das 23 variáveis adicionadas referentes ao
+traço aproximado do concreto foram removidas devido a esse fato. Além
+delas, no caso do conjunto de 7 dias, a variável cinza volante também
+foi removida. Depois foi verificado que não existem variáveis com alta
+correlação, acima de 0,999 em nenhum dos 6 conjuntos de dados
+(\ref{show-cors}). Após essas etapas, os conjuntos de amostras
+apresentaram 24, 21, 23, 23, 24 e 25 colunas respectivamente para as
+idades em sequência crescente.
 
@@
 
-The stage of centralization and normalization of the variables was
-carried out later, together with the application of the models, as it is
-simpler to do this with the \emph{caret} package. If performed at this
-time, it would be necessary to manually undo these transformations in
-the predictions. The \emph{caret} allows you to transform before
-training the models and already transforms the results back.
+A etapa de centralização e normalização das variáveis foi realizada mais
+a frente, junto com a aplicação dos modelos, pois é mais simples fazer
+dessa forma com o pacote \emph{caret}. Se fosse realizada nesse momento,
+seria nescessário manualmente desfazer essas transformações nas
+previsões. O \emph{caret} permite transformar antes do treino dos
+modelos e já transforma de volta os resultados.
 
@@
 
-Each data set was separated into test and training sets, \emph{20\%} and
-\emph{80\%} respectively (\ref{show-split}). The figure
-\ref{fig:dist-split} shows the distribution of data between the sets in
-relation to the compressive strength for each model
-(\ref{show-dist-split}).
+Cada um dos conjuntos de dados foi separado em conjuntos de teste e
+treino, \emph{20\%} e \emph{80\%} respectivamente (\ref{show-split}). A
+figura \ref{fig:dist-split} mostra como ficou a distribuição dos dados
+entre os conjuntos em relação a resistência à compressão para cada
+modelo (\ref{show-dist-split}).
 
@@
 
-{\centering \includegraphics{paper_EN_files/figure-latex/dist-split-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/dist-split-1}
 
@@
 
-\caption{Distribution of test and train data}\label{fig:dist-split}
+\caption{Distribuição dos conjuntos de teste e treino}\label{fig:dist-split}
 \end{figure}
 
-\hypertarget{performance-measures}{%
-\subsubsection{Performance measures}\label{performance-measures}}
-
-The performance evaluation of the models was performed by the Root Mean
-Square Error (\emph{RMSE}). The \emph{RMSE} is the measure used in all
-the works mentioned in the introduction and will allow the comparison of
-the models in the discussion.
-
-\hypertarget{naive-models}{%
-\subsubsection{Naive models}\label{naive-models}}
-
-Before creating the real models, for comparison purposes, naive models
-were created. They simply predict that the compressive strength of the
-test set is the average compressive strength of the training set
-(\ref{show-naive-model-reg}). In other words, naive models are simply
-the best guess possible. The results can be checked in the table
+\hypertarget{medidas-de-performance}{%
+\subsubsection{Medidas de performance}\label{medidas-de-performance}}
+
+A avaliação da performance dos modelos foi realizada pela Raiz do Erro
+Quadrático Médio (\emph{RMSE}). O \emph{RMSE} é a medida utilizada em
+todos os trabalhos citados na introdução e permitirá a comparação dos
+modelos na discussão.
+
+\hypertarget{modelos-inguxeanuos}{%
+\subsubsection{Modelos ingênuos}\label{modelos-inguxeanuos}}
+
+Antes de criar os modelos verdadeiros, para fins de comparação, foram
+criados modelos ingênuos. Eles simplesmente prevêem que a resistência à
+compressão do conjunto de teste, é a média da resistência à compressão
+do conjunto de treino (\ref{show-naive-model-reg}). Em outras palavras,
+os modelos ingênuos são simplesmente o melhor palpite possível. Os
+resultados podem ser conferidos na tabela
 \ref{tab:table-naive-model-reg}.
@@
 
-\caption{\label{tab:table-naive-model-reg}Naive models}
+\caption{\label{tab:table-naive-model-reg}Modelos ingênuos}
 \centering
@@
 \toprule
-Age & Mean $MPa$ (train) & RMSE (train) & RMSE (test)\\
+Idade & Média $MPa$ (treino) & RMSE (treino) & RMSE (teste)\\
 \midrule
@@
 
-\hypertarget{choice-of-algorithms}{%
-\subsubsection{Choice of algorithms}\label{choice-of-algorithms}}
-
-The \emph{caret} (Kuhn 2020) package exposes more than 200 different
-algorithms for building \emph{machine learning} models. The package
-documentation presents an initial code (``Models Clustered by Tag
-Similarity'' 2020) as a suggestion to select a portfolio of the most
-distinct algorithms possible in relation to some pre-selected algorithm,
-but for agility and due to technical limitations, it was chosen to use
-an algorithm with the highest probability to achieve the best possible
-result. According to Fernandez-Delgado et al. (2014), who compared 179
-algorithms in 121 different databases, the algorithm most likely to
-achieve the best possible results is the \emph{Parallel Random Forest}
-(called \emph{prRF } in \emph{caret}).
-
-\hypertarget{regression-models}{%
-\subsubsection{Regression models}\label{regression-models}}
-
-As new variables were added throughout the processing (the relationships
-between the ingredients and a few more \emph{dummy vars} for each age
-set), 5 possibilities for configuring the \emph{features} for the models
-were studied:
+\hypertarget{escolha-do-algoruxedtimo}{%
+\subsubsection{Escolha do algorítimo}\label{escolha-do-algoruxedtimo}}
+
+O pacote \emph{caret} (Kuhn 2020) expõe mais de 200 algorítimos
+diferentes para criar modelos de \emph{machine learning}. A documentação
+do pacote apresenta um código inicial (``Models Clustered by Tag
+Similarity'' 2020) como sugestão para selecionar um portifólio de
+algóritmos mais distintos possíveis em relação à algum algorítimo
+pré-selecionado, mas para agilidade e devido a limitações técnicas, foi
+escolhido utilizar um algorítimo com a maior probabilidade de atingir o
+melhor resultado possível. Segundo Fernandez-Delgado et al. (2014), que
+comparou 179 algorítimos em 121 banco de dados diferentes, o algorítimo
+mais provável de atingir os melhores resultados possíveis é o
+\emph{Parallel Random Forest} (denominado \emph{prRF} no \emph{caret}).
+
+\hypertarget{modelos-de-regressuxe3o}{%
+\subsubsection{Modelos de regressão}\label{modelos-de-regressuxe3o}}
+
+Como ao longo do processamento foram adicionadas novas variáveis (as
+relações entre os ingredientes e mais algumas \emph{dummy vars} para
+cada conjunto de idade), foram estudadas 5 possibilidades de
+configurações das \emph{features} para os modelos:
 
@@
 \item
-  All \emph{features};
+  Todas as \emph{features};
 \item
-  All \emph{features} without the \emph{dummy vars};
+  Sem as \emph{dummy vars};
 \item
-  Only the original \emph{features};
+  Apenas \emph{features} originais;
 \item
-  Only new \emph{features};
+  Apenas \emph{features} adicionadas;
 \item
-  Only new \emph{features}, without \emph{dummy vars};
+  Apenas \emph{features} adicionadas, sem as \emph{dummy vars};
 \end{enumerate}
@@
 
-Building a model for each of these configurations using the set at 28
-days (\ref{show-test-models-reg}), showed that the best option is
-configuration 2, the \emph{dummy vars} were completely discarded, but
-the other new variables were kept. For illustrative purposes, the
-\ref{tab:table-done-samples-28} table shows the first 6 samples from the
-28-day model training set. The samples of the other models, of the test
-and training sets are similar, the only difference being in the 7-day
-model, which excludes fly ash due to the variance close to zero,
-performed previously.
+Construindo um modelo para cada uma dessas configurações utilizando o
+conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou
+que a melhor opção é a configuração 2, ou seja, as \emph{dummy vars}
+foram completamente descartadas, porém foram mantidas as outras novas
+varáveis. Para fins ilustrativos, a tabela
+\ref{tab:table-done-samples-28} mostra as primeiras 6 amostras do
+conjunto de treino do modelo de 28 dias. As amostras dos outros modelos,
+dos conjuntos de teste e treino são similares, sendo a única diferença
+no modelo de 7 dias, que exlcui a cinza volante devido a variância
+próxima a zero, realizado anteriormente.
 
@@
 
-\caption{\label{tab:table-done-samples-28}First 6 samples of train data of the 28 days model}
+\caption{\label{tab:table-done-samples-28}Primeiras 6 amostras do conjunto de treino do modelo de 28 dias}
 \centering
@@
 \cmidrule(l{3pt}r{3pt}){1-13} \cmidrule(l{3pt}r{3pt}){14-14}
-\multicolumn{1}{c}{Cement} & \multicolumn{1}{c}{B.F.S.} & \multicolumn{1}{c}{Fly ash} & \multicolumn{1}{c}{water} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{Coarse Agg.} & \multicolumn{1}{c}{Fine Agg.} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{C.Agg./} & \multicolumn{1}{c}{F.Agg./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{Wat./} & \multicolumn{1}{c}{y} \\
-$kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & Ce. & Ce. & Ce. & C.Agg. & C.Agg. & F.Agg. & MPa\\
+\multicolumn{1}{c}{Cimento} & \multicolumn{1}{c}{E.G.A.F.} & \multicolumn{1}{c}{C.Vol.} & \multicolumn{1}{c}{Água} & \multicolumn{1}{c}{Superp.} & \multicolumn{1}{c}{A.Graúdo} & \multicolumn{1}{c}{A.Miúdo} & \multicolumn{1}{c}{Ág./} & \multicolumn{1}{c}{A.M./} & \multicolumn{1}{c}{A.G./} & \multicolumn{1}{c}{A.M./} & \multicolumn{1}{c}{Ág./} & \multicolumn{1}{c}{Ág./} & \multicolumn{1}{c}{y} \\
+$kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & $kg/m^3$ & Ci. & Ci. & Ci. & A.G. & A.G. & Ag.M. & $MPa$\\
 \midrule
@@
 
-For each age set, a model was created using the \emph{Parallel Random
-Forest} algorithm, previously defined (\ref{show-reg-models}). For each
-of the 6 models, the parameter \emph{mtry} was optimized, and
-\emph{repeated cross-validation } was performed, dividing into 10 or 30
-parts and repeating 10 times.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-The test \emph{RMSE} for each model in ascending order of age was 3.31,
-4.36, 4.62, 4.72, 5.94 and 5.85 respectively. The table
-\ref{tab:table-reg-models} presents the details and results of each
-model (\ref{show-table-reg-models}). The figure
-\ref{fig:results-comparison} compares the actual and predicted values
-(\ref{show-results-comparison}), and the following tables show the best
-and worst results for each model (\ref{show-table-10}).
+Para cada conjunto de idade foi criado um modelo utilizando o algorítimo
+\emph{Parallel Random Forest}, definido anteriormente
+(\ref{show-reg-models}). Para cada um dos 6 modelos, o parâmetro
+\emph{mtry} foi otimizado, e foi realizado \emph{repeated
+cross-validation}, dividindo em 10 ou 30 partes e repetindo 10 vezes.
+
+\hypertarget{r-param-optimization-plots-echof-warningf-messagef-includef-ggplotfit_3-xlabcaracteruxedsticas-ylabrmse-ggplotfit_7-xlabcaracteruxedsticas-ylabrmse-ggplotfit_14-xlabcaracteruxedsticas-ylabrmse-ggplotfit_28-xlabcaracteruxedsticas-ylabrmse-ggplotfit_56-xlabcaracteruxedsticas-ylabrmse-ggplotfit_100-xlabcaracteruxedsticas-ylabrmse}{%
+\section{\texorpdfstring{\texttt{\{r\ param-optimization-plots,\ echo=F,\ warning=F,\ message=F,\ include=F\}\ \#\ \ \#\ ggplot(fit\_3)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#\ ggplot(fit\_7)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#\ ggplot(fit\_14)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#\ ggplot(fit\_28)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#\ ggplot(fit\_56)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#\ ggplot(fit\_100)\ +\ xlab("Características")\ +\ ylab("RMSE")\ \#}}{\{r param-optimization-plots, echo=F, warning=F, message=F, include=F\} \#  \# ggplot(fit\_3) + xlab("Características") + ylab("RMSE") \# ggplot(fit\_7) + xlab("Características") + ylab("RMSE") \# ggplot(fit\_14) + xlab("Características") + ylab("RMSE") \# ggplot(fit\_28) + xlab("Características") + ylab("RMSE") \# ggplot(fit\_56) + xlab("Características") + ylab("RMSE") \# ggplot(fit\_100) + xlab("Características") + ylab("RMSE") \#}}\label{r-param-optimization-plots-echof-warningf-messagef-includef-ggplotfit_3-xlabcaracteruxedsticas-ylabrmse-ggplotfit_7-xlabcaracteruxedsticas-ylabrmse-ggplotfit_14-xlabcaracteruxedsticas-ylabrmse-ggplotfit_28-xlabcaracteruxedsticas-ylabrmse-ggplotfit_56-xlabcaracteruxedsticas-ylabrmse-ggplotfit_100-xlabcaracteruxedsticas-ylabrmse}}
+
+\hypertarget{resultados}{%
+\section{Resultados}\label{resultados}}
+
+O \emph{RMSE} de teste de cada modelo em ordem crescente de idade foi
+respectivamente 3.31, 4.36, 4.62, 4.72, 5.94 e 5.85. A tabela
+\ref{tab:table-reg-models} apresenta os detalhes e resultados de cada
+modelo (\ref{show-table-reg-models}). A figura
+\ref{fig:results-comparison} compara os valores reais e previstos
+(\ref{show-results-comparison}), e as tabelas seguintes mostram os
+melhores e piores resultados de cada modelo (\ref{show-table-10}).
 
@@
 
-\caption{\label{tab:table-reg-models}Regression models results}
+\caption{\label{tab:table-reg-models}Resultados dos modelos de regressão}
 \centering
@@
 \toprule
-Model & mtry & CV & Repetitions & RMSE (train) & RMSE (test)\\
+Modelo & mtry & CV & Repetições & RMSE (treino) & RMSE (teste)\\
 \midrule
-3 days & 6 & 30 & 10 & 3.905196 & 3.310370\\
-\addlinespace
-7 days & 2 & 10 & 10 & 4.475981 & 4.361987\\
-\addlinespace
-14 days & 13 & 30 & 10 & 5.136687 & 4.620515\\
-\addlinespace
-28 days & 11 & 30 & 10 & 5.847334 & 4.716698\\
-\addlinespace
-56 days & 8 & 30 & 10 & 6.702565 & 5.939163\\
-\addlinespace
-100 days & 8 & 10 & 10 & 6.381940 & 5.851088\\
+3 dias & 6 & 30 & 10 & 3.905196 & 3.310370\\
+\addlinespace
+7 dias & 2 & 10 & 10 & 4.475981 & 4.361987\\
+\addlinespace
+14 dias & 13 & 30 & 10 & 5.136687 & 4.620515\\
+\addlinespace
+28 dias & 11 & 30 & 10 & 5.847334 & 4.716698\\
+\addlinespace
+56 dias & 8 & 30 & 10 & 6.702565 & 5.939163\\
+\addlinespace
+100 dias & 8 & 10 & 10 & 6.381940 & 5.851088\\
 \bottomrule
@@
 \end{table}
-
 \begin{figure}
 
-{\centering \includegraphics{paper_EN_files/figure-latex/results-comparison-1}
+{\centering \includegraphics{paper_PT_files/figure-latex/results-comparison-1}
 
@@
 
-\caption{Actual vs Predicted values in each model}\label{fig:results-comparison}
+\caption{Comparação dos valores reais e previstos em cada modelo}\label{fig:results-comparison}
 \end{figure}
-
 \begin{table}[H]
 
-\caption{\label{tab:table-10}Model of 3 days}
+\caption{\label{tab:table-10}Modelo de 3 dias}
 \centering
@@
 \toprule
-\multicolumn{3}{c}{Best 10} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Worst 10} \\
+\multicolumn{3}{c}{10 melhores} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{10 piores} \\
 \cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){5-7}
-Actual & Predicted & Error &  & Actual & Predicted & Error\\
+Real & Previsto & Erro &  & Real & Previsto & Erro\\
 \midrule
@@
 
-\caption{\label{tab:table-10}Model of 7 days}
+\caption{\label{tab:table-10}Modelo de 7 dias}
 \centering
@@
 
-\caption{\label{tab:table-10}Model of 14 days}
+\caption{\label{tab:table-10}Modelo de 14 dias}
 \centering
@@
 
-\caption{\label{tab:table-10}Model of 28 days}
+\caption{\label{tab:table-10}Modelo de 28 dias}
 \centering
@@
 
-\caption{\label{tab:table-10}Model of 56 days}
+\caption{\label{tab:table-10}Modelo de 56 dias}
 \centering
@@
 
-\caption{\label{tab:table-10}Model of 100 days}
+\caption{\label{tab:table-10}Modelo de 100 dias}
 \centering
@@
 
-\hypertarget{discussion-and-conclusion}{%
-\section{Discussion and conclusion}\label{discussion-and-conclusion}}
-
-The models built present satisfactory results and prove that the
-compressive strength of concrete can be predicted relatively easily. The
-alternative adopted to create a model for each set of age proved to be a
-valid method, managing to stratify to obtain specific results for each
-set. The studies cited in the introduction using the same dataset have
-similar results, as expected. The \ref{tab:works-comparison} table
-presents the results of these works (\ref{show-works-comparison}), and
-the table \ref{tab:results} presents the values found for easy
-comparison (\ref{show-results}).
+\hypertarget{discussuxe3o}{%
+\section{Discussão}\label{discussuxe3o}}
+
+Os modelos construidos apresentam resultados satisfatórios e provam que
+a resistência à compressão do concreto pode ser prevista de forma
+relativamente fácil. A alternativa adotada de criar um modelo para cada
+conjunto de idade se mostrou como uma alternativa válida, conseguindo
+estratificar para obter resultados específicos de cada conjunto. Os
+estudos citados na introdução utilizando o mesmo conjunto de dados
+possuem resultados similares, como esperado. A tabela
+\ref{tab:works-comparison} apresenta os resultados desses trabalhos
+(\ref{show-works-comparison}), e a tabela \ref{tab:results} apresenta os
+valores encontrados (\ref{show-results}) para fácil comparação.
 
@@
 
-\caption{\label{tab:works-comparison}Comparison of other works}
+\caption{\label{tab:works-comparison}Comparação dos estudos de outros autores}
 \centering
@@
 \toprule
-Author & Year & Algorithm & RMSE\\
+Autor & Ano & Algorítimo & RMSE\\
 \midrule
@@
 \end{table}
-
 \begin{table}[H]
 
-\caption{\label{tab:results}Final result}
+\caption{\label{tab:results}Resultados finais}
 \centering
@@
 \toprule
-Model & RMSE\\
+Modelo & RMSE\\
 \midrule
-3 days & 3.310370\\
-\addlinespace
-7 days & 4.361987\\
-\addlinespace
-14 days & 4.620515\\
-\addlinespace
-28 days & 4.716698\\
-\addlinespace
-56 days & 5.939163\\
-\addlinespace
-100 days & 5.851088\\
+3 dias & 3.310370\\
+\addlinespace
+7 dias & 4.361987\\
+\addlinespace
+14 dias & 4.620515\\
+\addlinespace
+28 dias & 4.716698\\
+\addlinespace
+56 dias & 5.939163\\
+\addlinespace
+100 dias & 5.851088\\
 \bottomrule
@@
 
-Following the line of reasoning of this work, it can be performed with
-different algorithms, the results found here used only one
-(\emph{Parallel Random Forest}), even though it was theoretically the
-``best'' found, other algorithms can present even better results.
-Another option is to create an \emph{ensemble} of various algorithms,
-just like Pierobon (2018), but with the separation of age sets proposed
-here. In addition, it can be performed with a larger dataset, ideally
-with the same number of samples in each age set, a more homogeneous
-distribution of compressive strength, and less variance between samples.
-
-\hypertarget{literature-cited}{%
-\section{Literature cited}\label{literature-cited}}
+Seguindo a linha de raciocínio desse trabalho, ele pode ser realizado
+com diferentes algorítimos, os resultados aqui encontrados utilizaram
+apenas um único (\emph{Parallel Random Forest}), mesmo que tenha sido
+teoricamente o ``melhor'' encontrado, outros algorítmos podem apresentar
+resultados ainda melhores. Outra opção é criar um \emph{ensemble} com
+diversos algorítimos, como realizado por Pierobon (2018), mas com a
+separação de conjuntos de idade aqui proposto. Além disso, pode ser
+realizado com um conjunto maior de dados, idealmente com o mesmo número
+de amostras em cada conjunto de idade, distribuição mais homgênia da
+resistência à compressão e menor variância entre as amostras.
+
+\hypertarget{bibliografia}{%
+\section{Bibliografia}\label{bibliografia}}
 
@@
 \leavevmode\hypertarget{ref-irizarry2019}{}%
-Irizarry, R. A. 2019. \emph{Introduction to Data Science: Data Analysis
+Irizarry, R.A. 2019. \emph{Introduction to Data Science: Data Analysis
 and Prediction Algorithms with R}. Chapman \& Hall/Crc Data Science
@@
 \hypertarget{appendix1}{%
-\section{Appendix 1 - Virtual environment}\label{appendix1}}
-
-\hypertarget{operational-system}{%
-\subsection{Operational system}\label{operational-system}}
+\section{Appendix 1 - Ambiente virtual}\label{appendix1}}
+
+\hypertarget{sistema-operacional}{%
+\subsection{Sistema operacional}\label{sistema-operacional}}
 
@@
 
-\hypertarget{packages}{%
-\subsection{Packages}\label{packages}}
+\hypertarget{pacotes-utilizados}{%
+\subsection{Pacotes utilizados}\label{pacotes-utilizados}}
 
@@
 \hypertarget{appendix2}{%
-\section{Appendix 2 - Code}\label{appendix2}}
-
-\hypertarget{obtaining-the-data-1}{%
-\subsection{Obtaining the data}\label{obtaining-the-data-1}}
-
-\hypertarget{data-download}{%
-\subsubsection{Data download}\label{data-download}}
+\section{Appendix 2 - Repositório online}\label{appendix2}}
+
+\url{https://github.com/pedrobern/concrete-compressive-strength-prediction}
+\newpage
+
+\hypertarget{appendix3}{%
+\section{Appendix 3 - Código}\label{appendix3}}
+
+\hypertarget{obtenuxe7uxe3o-dos-dados-1}{%
+\subsection{Obtenção dos dados}\label{obtenuxe7uxe3o-dos-dados-1}}
+
+\hypertarget{download-dos-dados}{%
+\subsubsection{Download dos dados}\label{download-dos-dados}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Data download}
+\CommentTok{# Download dos dados}
 \NormalTok{url_base <-}\StringTok{ "https://archive.ics.uci.edu"}
@@
 
-\hypertarget{renaming-the-columns}{%
-\subsubsection{Renaming the columns}\label{renaming-the-columns}}
+\hypertarget{renomeando-as-colunas}{%
+\subsubsection{Renomeando as colunas}\label{renomeando-as-colunas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Renaming the columns}
+\CommentTok{# Renomeando as colunas}
 \KeywordTok{colnames}\NormalTok{(dat) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
@@
 
-\hypertarget{reordering-the-data}{%
-\subsubsection{Reordering the data}\label{reordering-the-data}}
+\hypertarget{reordenando-os-dados}{%
+\subsubsection{Reordenando os dados}\label{reordenando-os-dados}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Reordering the data}
+\CommentTok{# Reordenando os dados}
 \NormalTok{col_order <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
@@
 
-\hypertarget{defining-column-names-and-units}{%
-\subsubsection{Defining column names and
-units}\label{defining-column-names-and-units}}
+\hypertarget{definindo-nomes-e-unidades-das-colunas}{%
+\subsubsection{Definindo nomes e unidades das
+colunas}\label{definindo-nomes-e-unidades-das-colunas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Defining column names and units}
-\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-             \StringTok{"Superp."}\NormalTok{, }\StringTok{"C.Aggregate"}\NormalTok{, }\StringTok{"F.Aggregate"}\NormalTok{, }\StringTok{"Day"}\NormalTok{, }\StringTok{"Comp.Str."}\NormalTok{)}
+\CommentTok{# Definindo nomes e unidades das colunas}
+\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F"}\NormalTok{, }\StringTok{"C.Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+             \StringTok{"Superp."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }\StringTok{"Dia"}\NormalTok{, }\StringTok{"Comp.Str."}\NormalTok{)}
 \NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
@@
 
-\hypertarget{table---first-samples}{%
-\subsubsection{Table - First samples}\label{table---first-samples}}
+\hypertarget{tabela---primeiras-amostras}{%
+\subsubsection{Tabela - Primeiras
+amostras}\label{tabela---primeiras-amostras}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - First samples}
-\NormalTok{caption <-}\StringTok{ "First 6 samples"}
+\CommentTok{# Tabela - Primeiras amostras}
+\NormalTok{caption <-}\StringTok{ "Primeiras 6 amostras"}
 \KeywordTok{kable}\NormalTok{(}
@@
 \StringTok{    }\KeywordTok{add_header_above}\NormalTok{(}\DataTypeTok{header =}\NormalTok{ colNames, }\DataTypeTok{line =}\NormalTok{ F, }\DataTypeTok{align =} \StringTok{"c"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{data-preparation-1}{%
-\subsection{Data preparation}\label{data-preparation-1}}
-
-\hypertarget{removing-duplicated-samples}{%
-\subsubsection{Removing duplicated
-samples}\label{removing-duplicated-samples}}
+\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{save_kable}\NormalTok{(}\StringTok{"tables_PT/test.pdf"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{preparauxe7uxe3o-dos-dados-1}{%
+\subsection{Preparação dos dados}\label{preparauxe7uxe3o-dos-dados-1}}
+
+\hypertarget{removendo-amostras-duplicadas}{%
+\subsubsection{Removendo amostras
+duplicadas}\label{removendo-amostras-duplicadas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Removing duplicated samples}
+\CommentTok{# Removendo amostras duplicadas}
 \NormalTok{n_distinct_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{n_distinct}\NormalTok{()}
@@
 
-\hypertarget{table---samples-with-same-composition}{%
-\subsubsection{Table - Samples with same
-composition}\label{table---samples-with-same-composition}}
+\hypertarget{tabela---amostras-com-a-mesma-composiuxe7uxe3o}{%
+\subsubsection{Tabela - Amostras com a mesma
+composição}\label{tabela---amostras-com-a-mesma-composiuxe7uxe3o}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Samples with same composition}
+\CommentTok{# Tabela - Amostras com a mesma composição}
 \NormalTok{same_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{653}\NormalTok{, }\DecValTok{678}\NormalTok{, }\DecValTok{654}\NormalTok{, }\DecValTok{681}\NormalTok{))}
-\NormalTok{caption <-}\StringTok{ "Samples with same composition"}
+\NormalTok{caption <-}\StringTok{ "Amostras com a mesma composição"}
 \KeywordTok{kable}\NormalTok{(}
@@
 
-\hypertarget{table---same-samples-with-different-results}{%
-\subsubsection{Table - Same samples with different
-results}\label{table---same-samples-with-different-results}}
+\hypertarget{tabela---amostras-iguais-com-resultados-diferentes}{%
+\subsubsection{Tabela - Amostras iguais com resultados
+diferentes}\label{tabela---amostras-iguais-com-resultados-diferentes}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Same samples with different results}
+\CommentTok{# Tabela - Amostras iguais com resultados diferentes}
 \NormalTok{same_samples_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{472}\NormalTok{, }\DecValTok{473}\NormalTok{, }\DecValTok{474}\NormalTok{))}
-\NormalTok{caption <-}\StringTok{ "Same samples with different results"}
+\NormalTok{caption <-}\StringTok{ "Amostras iguais com resultados diferentes"}
 \KeywordTok{kable}\NormalTok{(}
@@
 
-\hypertarget{data-cleaning}{%
-\subsubsection{Data cleaning}\label{data-cleaning}}
+\hypertarget{limpeza-inicial-das-amotras}{%
+\subsubsection{Limpeza inicial das
+amotras}\label{limpeza-inicial-das-amotras}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Data cleaning}
+\CommentTok{# Limpeza inicial das amotras}
 \NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{table---previous-samples-after-processing}{%
-\subsubsection{Table - Previous samples after
-processing}\label{table---previous-samples-after-processing}}
+\hypertarget{tabela---amostras-anteriores-apuxf3s-processamento}{%
+\subsubsection{Tabela - Amostras anteriores após
+processamento}\label{tabela---amostras-anteriores-apuxf3s-processamento}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Previous samples after processing}
+\CommentTok{# Tabela - Amostras anteriores após processamento}
 \NormalTok{same_samples <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{==}\StringTok{ }\DecValTok{653} \OperatorTok{|}\StringTok{ }\NormalTok{id }\OperatorTok{==}\StringTok{ }\DecValTok{472} \OperatorTok{&}\StringTok{ }\NormalTok{day }\OperatorTok{==}\StringTok{ }\DecValTok{28}\NormalTok{)}
-\NormalTok{caption <-}\StringTok{ "Previous samples after processing"}
+\NormalTok{caption <-}\StringTok{ "Amostras anteriores após processamento"}
 \KeywordTok{kable}\NormalTok{(}
@@
 
-\hypertarget{figure---compressive-strength-mpa-vs-age-days}{%
-\subsubsection{Figure - Compressive strength (MPa) vs age
-(days)}\label{figure---compressive-strength-mpa-vs-age-days}}
+\hypertarget{figura---resistuxeancia-uxe0-compressuxe3o-mpa-vs-idade-dias}{%
+\subsubsection{Figura - Resistência à compressão (MPa) vs idade
+(dias)}\label{figura---resistuxeancia-uxe0-compressuxe3o-mpa-vs-idade-dias}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Compressive strength (MPa) vs age (days)}
-\NormalTok{cap <-}\StringTok{ "Boxplot - Compressive strength (MPa) vs age (days)"}
-\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
-\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
+\CommentTok{# Figura - Resistência à compressão (MPa) vs idade (dias)}
+\NormalTok{cap <-}\StringTok{ "Boxplot - Resistência à compressão (MPa) vs idade (dias)"}
+\NormalTok{ylabel <-}\StringTok{ "Resistência à compressão (MPa)"}
+\NormalTok{xlabel <-}\StringTok{ "Idade (dias)"}
 \NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{figure---principal-component-analysis---90-91-e-100-days}{%
-\subsubsection{Figure - Principal component analysis - 90, 91 e 100
-days}\label{figure---principal-component-analysis---90-91-e-100-days}}
+\hypertarget{figura---anuxe1lise-componente-principal---90-91-e-100-dias}{%
+\subsubsection{Figura - Análise componente principal - 90, 91 e 100
+dias}\label{figura---anuxe1lise-componente-principal---90-91-e-100-dias}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Principal component analysis - 90, 91 e 100 days}
+\CommentTok{# Figura - Análise componente principal - 90, 91 e 100 dias}
 \NormalTok{dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 \StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, mpa))}
-\NormalTok{cap <-}\StringTok{ "Principal component analysis - 90, 91 e 100 days"}
+\NormalTok{cap <-}\StringTok{ "Análise componente principal - 90, 91 e 100 dias"}
 \KeywordTok{colnames}\NormalTok{(dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
-   \StringTok{"Cem."}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly.A."}\NormalTok{,}\StringTok{"Water"}\NormalTok{,}\StringTok{"Sup."}\NormalTok{,}\StringTok{"C.Ag."}\NormalTok{,}\StringTok{"F.Ag."}\NormalTok{,}\StringTok{"day"}\NormalTok{)}
-\NormalTok{pca <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(day)), }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{habillage <-}\StringTok{ }\NormalTok{dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\OperatorTok{$}\NormalTok{day}
+   \StringTok{"Cim."}\NormalTok{,}\StringTok{"E.G.A.F."}\NormalTok{,}\StringTok{"Ci.Vo."}\NormalTok{,}\StringTok{"Agu."}\NormalTok{,}\StringTok{"Sup."}\NormalTok{,}\StringTok{"Ag.G."}\NormalTok{,}\StringTok{"Ag.M."}\NormalTok{,}\StringTok{"dia"}
+\NormalTok{)}
+\NormalTok{pca <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(dia)), }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{habillage <-}\StringTok{ }\NormalTok{dat_}\DecValTok{90}\NormalTok{_}\DecValTok{91}\NormalTok{_}\DecValTok{100}\OperatorTok{$}\NormalTok{dia}
 \KeywordTok{fviz_pca_biplot}\NormalTok{(}
@@
 
-\hypertarget{figure---compressive-strength-through-time}{%
-\subsubsection{Figure - Compressive strength through
-time}\label{figure---compressive-strength-through-time}}
+\hypertarget{figura---resistuxeancia-uxe0-compressuxe3o-ao-longo-do-tempo}{%
+\subsubsection{Figura - Resistência à compressão ao longo do
+tempo}\label{figura---resistuxeancia-uxe0-compressuxe3o-ao-longo-do-tempo}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Compressive strength through time}
-\NormalTok{cap <-}\StringTok{ "Compressive strength through time"}
+\CommentTok{# Figura - Resistência à compressão ao longo do tempo}
+\NormalTok{cap <-}\StringTok{ "Resistência à compressão ao longo do tempo"}
 \NormalTok{ylabel <-}\StringTok{ "MPa"}
-\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
+\NormalTok{xlabel <-}\StringTok{ "Idade (dias)"}
 \NormalTok{dat_duplicated_only <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{joining-90-91-and-100-days-data}{%
-\subsubsection{Joining 90, 91 and 100 days
-data}\label{joining-90-91-and-100-days-data}}
+\hypertarget{juntando-amostras-de-90-91-e-100-dias}{%
+\subsubsection{Juntando amostras de 90, 91 e 100
+dias}\label{juntando-amostras-de-90-91-e-100-dias}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Joining 90, 91 and 100 days data}
+\CommentTok{# Juntando amostras de 90, 91 e 100 dias}
 \NormalTok{ind_}\DecValTok{90}\NormalTok{ <-}\StringTok{ }\NormalTok{dat}\OperatorTok{$}\NormalTok{id[}\KeywordTok{which}\NormalTok{(dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{==}\StringTok{ "90"}\NormalTok{)]}
@@
 
-\hypertarget{figure---ages-frequency}{%
-\subsubsection{Figure - Ages frequency}\label{figure---ages-frequency}}
+\hypertarget{figura---frequuxeancia-das-idades}{%
+\subsubsection{Figura - Frequência das
+idades}\label{figura---frequuxeancia-das-idades}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Ages frequency}
-\NormalTok{cap <-}\StringTok{ "Ages frequency"}
-\NormalTok{ylabel <-}\StringTok{ "Frequency"}
-\NormalTok{xlabel <-}\StringTok{ "Age (days)"}
-\NormalTok{dat }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{factor}\NormalTok{(day))) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_bar}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(xlabel) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{ylab}\NormalTok{(ylabel)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{removing-ages-with-frequency-lower-than-50}{%
-\subsubsection{Removing ages with frequency lower than
-50}\label{removing-ages-with-frequency-lower-than-50}}
+\CommentTok{# Figura - Frequência das idades}
+\NormalTok{cap <-}\StringTok{ "Frequência das idades"}
+\NormalTok{ylabel <-}\StringTok{ "Frequência"}
+\StringTok{xlabel <- "}\KeywordTok{Idade}\NormalTok{ (dias)}\StringTok{"}
+\StringTok{dat %>%}
+\StringTok{  ggplot(aes(x = factor(day))) +}
+\StringTok{  geom_bar() +}
+\StringTok{  theme_bw() +}
+\StringTok{  xlab(xlabel) +}
+\StringTok{  ylab(ylabel)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{removendo-idades-com-frequuxeancia-menor-que-50}{%
+\subsubsection{Removendo idades com frequência menor que
+50}\label{removendo-idades-com-frequuxeancia-menor-que-50}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Removing ages with frequency lower than 50}
+\CommentTok{# Removendo idades com frequência menor que 50}
 \NormalTok{dat <-}\StringTok{ }\NormalTok{dat[dat}\OperatorTok{$}\NormalTok{day }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{28}\NormalTok{, }\DecValTok{56}\NormalTok{, }\DecValTok{100}\NormalTok{),]}
@@
 
-\hypertarget{data-reorganization-1}{%
-\subsubsection{Data reorganization}\label{data-reorganization-1}}
+\hypertarget{reorganizauxe7uxe3o-das-amotras}{%
+\subsubsection{Reorganização das
+amotras}\label{reorganizauxe7uxe3o-das-amotras}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Data reorganization}
+\CommentTok{# Reorganização das amotras}
 \NormalTok{dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{table---first-6-samples-after-reorganization}{%
-\subsubsection{Table - First 6 samples after
-reorganization}\label{table---first-6-samples-after-reorganization}}
+\hypertarget{tabela---primeiras-6-amostras-reorganizadas}{%
+\subsubsection{Tabela - Primeiras 6 amostras
+reorganizadas}\label{tabela---primeiras-6-amostras-reorganizadas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - First 6 samples after reorganization}
-\NormalTok{caption <-}\StringTok{ "First 6 samples after reorganization"}
-\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S"}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-             \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Ag."}\NormalTok{, }\StringTok{"Fine Ag."}\NormalTok{, }\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{,}
-             \StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
+\CommentTok{# Tabela - Primeiras 6 amostras reorganizadas}
+\NormalTok{caption <-}\StringTok{ "Primeiras 6 amostras reorganizadas"}
+\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+             \StringTok{"Superp."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{,}
+             \StringTok{"14 dias"}\NormalTok{, }\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{)}
 \NormalTok{dfUnits2 <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
@@
 
-\hypertarget{total-samples}{%
-\subsubsection{Total samples}\label{total-samples}}
+\hypertarget{total-de-amostras}{%
+\subsubsection{Total de amostras}\label{total-de-amostras}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Total samples}
+\CommentTok{# Total de amostras}
 \NormalTok{n_samples <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(dat) }\CommentTok{# 416}
@@
 
-\hypertarget{adding-new-features}{%
-\subsubsection{Adding New features}\label{adding-new-features}}
+\hypertarget{adicionando-novas-variuxe1veis-1}{%
+\subsubsection{Adicionando novas
+variáveis}\label{adicionando-novas-variuxe1veis-1}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Adding New features}
+\CommentTok{# Adicionando novas variáveis}
 \NormalTok{concrete_class <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mpa)\{}
@@
 
-\hypertarget{table---new-features}{%
-\subsubsection{Table - New features}\label{table---new-features}}
+\hypertarget{tabela---novas-variuxe1veis}{%
+\subsubsection{Tabela - Novas
+variáveis}\label{tabela---novas-variuxe1veis}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - New features}
-\NormalTok{caption <-}\StringTok{ "New features"}
-\NormalTok{colNames7 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Class"}\NormalTok{,}\StringTok{"Approximated Mix"}\NormalTok{,}
-             \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine Ag. / Cement"}\NormalTok{, }
-             \StringTok{"Coarse Ag. / Cement"}\NormalTok{, }\StringTok{"Fine Ag. / Coarse Ag."}\NormalTok{,}
-             \StringTok{"Water / Coarse Ag."}\NormalTok{, }\StringTok{"Water / Fine Ag."}\NormalTok{)}
-\KeywordTok{kable}\NormalTok{(}
-    \KeywordTok{head}\NormalTok{(dat[}\KeywordTok{order}\NormalTok{(dat}\OperatorTok{$}\NormalTok{id),][,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{15}\OperatorTok{:}\DecValTok{22}\NormalTok{)]),}
-    \DataTypeTok{col.names =}\NormalTok{ colNames7,}
-    \DataTypeTok{escape =}\NormalTok{ F,}
-    \DataTypeTok{booktabs =}\NormalTok{ T,}
-    \DataTypeTok{caption =}\NormalTok{ caption,}
-    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
-    \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
-    \DataTypeTok{digits =} \DecValTok{4}
-\NormalTok{    ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DataTypeTok{width =} \StringTok{"1.5cm"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DataTypeTok{width =} \StringTok{"2cm"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\DecValTok{4}\OperatorTok{:}\DecValTok{9}\NormalTok{, }\DataTypeTok{width =} \StringTok{"1.7cm"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{data-visualization-1}{%
-\subsection{Data visualization}\label{data-visualization-1}}
-
-\hypertarget{table---descriptive-statistics---continuous-variables}{%
-\subsubsection{Table - Descriptive statistics - continuous
-variables}\label{table---descriptive-statistics---continuous-variables}}
+\CommentTok{# Tabela - Novas variáveis}
+\NormalTok{caption <-}\StringTok{ "Novas variáveis"}
+\NormalTok{colNames7 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Classe"}\NormalTok{,}\StringTok{"Traço aproximado"}\NormalTok{,}
+             \StringTok{"Água / Cimento"}\NormalTok{, }\StringTok{"A.Miúdo / Cimento"}\NormalTok{, }
+             \StringTok{"A.Graúdo / Cimento"}\NormalTok{, }\StringTok{"A.Miúdo / A.Graúdo",}
+\StringTok{             "}\NormalTok{Água }\OperatorTok{/}\StringTok{ }\NormalTok{A.Graúdo", }\StringTok{"Água / A.Miúdo")}
+\StringTok{kable(}
+\StringTok{    head(dat[order(dat$id),][,c(1,15:22)]),}
+\StringTok{    col.names = colNames7,}
+\StringTok{    escape = F,}
+\StringTok{    booktabs = T,}
+\StringTok{    caption = caption,}
+\StringTok{    linesep = "}\NormalTok{\textbackslash{}\textbackslash{}addlinespace}\StringTok{",}
+\StringTok{    align = "}\NormalTok{c}\StringTok{",}
+\StringTok{    digits = 4}
+\StringTok{    ) %>%}
+\StringTok{  kable_styling(latex_options = c("}\NormalTok{HOLD_position}\StringTok{", "}\NormalTok{scale_down}\StringTok{")) %>%}
+\StringTok{  column_spec(2, width = "}\FloatTok{1.5}\NormalTok{cm}\StringTok{") %>%}
+\StringTok{  column_spec(3, width = "}\NormalTok{2cm}\StringTok{") %>%}
+\StringTok{  column_spec(4:9, width = "}\FloatTok{1.7}\NormalTok{cm}\StringTok{")}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{visualizauxe7uxe3o-dos-dados-1}{%
+\subsection{Visualização dos
+dados}\label{visualizauxe7uxe3o-dos-dados-1}}
+
+\hypertarget{tabela---estatuxedstica-descritiva---variuxe1veis-contuxednuas}{%
+\subsubsection{Tabela - Estatística descritiva - variáveis
+contínuas}\label{tabela---estatuxedstica-descritiva---variuxe1veis-contuxednuas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Descriptive statistics - continuous variables}
+\CommentTok{# Tabela - Estatística descritiva - variáveis contínuas}
 \NormalTok{summ <-}\StringTok{ }\KeywordTok{t}\NormalTok{(}
   \KeywordTok{stat.desc}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, mix_app))))}
-\NormalTok{caption <-}\StringTok{ "Descriptive statistics - continuous variables"}
-\KeywordTok{colnames}\NormalTok{(summ) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Samples"}\NormalTok{, }\StringTok{"Null"}\NormalTok{, }\StringTok{"NA"}\NormalTok{, }\StringTok{"Min"}\NormalTok{, }\StringTok{"Max"}\NormalTok{, }\StringTok{"Range"}\NormalTok{,}
-              \StringTok{"Sum"}\NormalTok{, }\StringTok{"Median"}\NormalTok{, }\StringTok{"Mean"}\NormalTok{, }\StringTok{"SE mean"}\NormalTok{, }
-              \StringTok{"CI mean"}\NormalTok{,}\StringTok{"Variance"}\NormalTok{, }\StringTok{"Std.Dev."}\NormalTok{, }\StringTok{"Coef.Var"}\NormalTok{)}
-\KeywordTok{rownames}\NormalTok{(summ) =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-              \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{, }\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{,}
-              \StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{,}
-              \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
-              \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
-              \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
-\KeywordTok{kable}\NormalTok{(}
-\NormalTok{  summ,}
-  \DataTypeTok{escape =}\NormalTok{ F,}
-  \DataTypeTok{booktabs =}\NormalTok{ T,}
-  \DataTypeTok{caption =}\NormalTok{ caption,}
-  \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
-  \DataTypeTok{align =} \StringTok{"c"}\NormalTok{,}
-  \DataTypeTok{digits =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{)}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{column_spec}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{10}\OperatorTok{:}\DecValTok{15}\NormalTok{), }\DataTypeTok{width =} \StringTok{"1.5cm"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{figure---descriptive-statistics---categorical-variables}{%
-\subsubsection{Figure - Descriptive statistics - categorical
-variables}\label{figure---descriptive-statistics---categorical-variables}}
+\NormalTok{caption <-}\StringTok{ "Estatística descritiva - variáveis contínuas"}
+\KeywordTok{colnames}\NormalTok{(summ) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Amostras"}\NormalTok{, }\StringTok{"Null"}\NormalTok{, }\StringTok{"NA"}\NormalTok{, }\StringTok{"Min"}\NormalTok{, }\StringTok{"Max"}\NormalTok{, }\StringTok{"Intervalo"}\NormalTok{,}
+              \StringTok{"Soma"}\NormalTok{, }\StringTok{"Mediana"}\NormalTok{, }\StringTok{"Média"}\NormalTok{, }\StringTok{"Erro padrão  da média"}\NormalTok{, }
+              \StringTok{"Intervalo de confiânça da média"}\NormalTok{,}
+              \StringTok{"Variância"}\NormalTok{, }\StringTok{"Desvio Padrão"}\NormalTok{, }\StringTok{"Coeficiênte de variação"}\NormalTok{)}
+\KeywordTok{rownames}\NormalTok{(summ) =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"Cinza Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+              \StringTok{"Superplast."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{,}
+              \StringTok{"14 dias"}\NormalTok{, }\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{,}
+              \StringTok{"Água/ Cimento"}\NormalTok{, }\StringTok{"A.Miúdo/ Cimento"}\NormalTok{, }
+              \StringTok{"A.Graúdo/ Cimento"}\NormalTok{, }\StringTok{"A.Miúdo / A.Graúdo",}
+\StringTok{              "}\NormalTok{Água }\OperatorTok{/}\StringTok{ }\NormalTok{A.Graúdo", }\StringTok{"Água / A.Miúdo")}
+\StringTok{kable(}
+\StringTok{  summ,}
+\StringTok{  escape = F,}
+\StringTok{  booktabs = T,}
+\StringTok{  caption = caption,}
+\StringTok{  linesep = "}\NormalTok{\textbackslash{}\textbackslash{}addlinespace}\StringTok{",}
+\StringTok{  align = "}\NormalTok{c}\StringTok{",}
+\StringTok{  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)}
+\StringTok{  ) %>%}
+\StringTok{  kable_styling(latex_options = c("}\NormalTok{HOLD_position}\StringTok{", "}\NormalTok{scale_down}\StringTok{")) %>%}
+\StringTok{  column_spec(c(1,10:15), width = "}\FloatTok{1.5}\NormalTok{cm}\StringTok{")}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{figura---estatuxedstica-descritiva---variuxe1veis-discretas}{%
+\subsubsection{Figura - Estatística descritiva - variáveis
+discretas}\label{figura---estatuxedstica-descritiva---variuxe1veis-discretas}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Descriptive statistics - categorical variables}
-\NormalTok{cap <-}\StringTok{ "Descriptive statistics - categorical variables"}
-\NormalTok{name1 <-}\StringTok{ "Percentage"}
-\NormalTok{name2 <-}\StringTok{ "Accumulated percentage"}
-\NormalTok{ylabel <-}\StringTok{ "Frequency"}
-\NormalTok{xlabel1 <-}\StringTok{ "Class"}
-\NormalTok{xlabel2 <-}\StringTok{ "Approximate Mix"}
+\CommentTok{# Figura - Estatística descritiva - variáveis discretas}
+\NormalTok{cap <-}\StringTok{ "Estatística descritiva - variáveis discretas"}
+\NormalTok{name1 <-}\StringTok{ "Porcentagem"}
+\NormalTok{name2 <-}\StringTok{ "Porcentagem acumulada"}
+\NormalTok{ylabel <-}\StringTok{ "Frequencia"}
+\NormalTok{xlabel1 <-}\StringTok{ "Classe"}
+\NormalTok{xlabel2 <-}\StringTok{ "Traço aproximado"}
 \NormalTok{format_percent =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(n)\{}
@@
 
-\hypertarget{figure---correlation-grouped-by-age}{%
-\subsubsection{Figure - Correlation grouped by
-age}\label{figure---correlation-grouped-by-age}}
+\hypertarget{figura---correlauxe7uxf5es-em-cada-idade}{%
+\subsubsection{Figura - Correlações em cada
+idade}\label{figura---correlauxe7uxf5es-em-cada-idade}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Correlation grouped by age}
+\CommentTok{# Figura - Correlações em cada idade}
 \NormalTok{cor_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id))}
-\NormalTok{cap <-}\StringTok{ "Correlations at each age"}
-\NormalTok{f_lvl <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{,}\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
-\NormalTok{name <-}\StringTok{ "Correlation"}
-\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-                \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
-                \StringTok{"3"}\NormalTok{, }\StringTok{"7"}\NormalTok{, }\StringTok{"14"}\NormalTok{, }\StringTok{"28"}\NormalTok{, }\StringTok{"56"}\NormalTok{, }\StringTok{"100"}\NormalTok{,}
-                \StringTok{"class"}\NormalTok{, }\StringTok{"mix_app"}\NormalTok{, }\StringTok{"W./C."}\NormalTok{, }\StringTok{"F.A./C."}\NormalTok{, }
-                \StringTok{"C.A./C."}\NormalTok{, }\StringTok{"F.A./C.A."}\NormalTok{,}
-                \StringTok{"W./C.A."}\NormalTok{, }\StringTok{"W./F.A."}\NormalTok{)}
+\NormalTok{cap <-}\StringTok{ "Correlações em cada idade"}
+\NormalTok{f_lvl <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{, }\StringTok{"14 dias"}\NormalTok{,}\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{)}
+\NormalTok{name <-}\StringTok{ "Correlação"}
+\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+                  \StringTok{"Superp."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo",}
+                  \StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{,}
+                  \StringTok{"class"}\NormalTok{,}\StringTok{"mix_app"}\NormalTok{ ,}\StringTok{"Ág./Ci."}\NormalTok{,}
+                  \StringTok{"A.M./Ci."}\NormalTok{, }\StringTok{"A.G./Ci."}\NormalTok{, }\StringTok{"A.M./A.G."}\NormalTok{,}
+                  \StringTok{"Ág./A.G."}\NormalTok{, }\StringTok{"Ág./A.M."}\NormalTok{)}
 \KeywordTok{colnames}\NormalTok{(cor_dat) <-}\StringTok{ }\NormalTok{colnames_dat}
@@
 
-\hypertarget{figure---correlation-over-time}{%
-\subsubsection{Figure - Correlation over
-time}\label{figure---correlation-over-time}}
+\hypertarget{figura---correlauxe7uxf5es-no-tempo}{%
+\subsubsection{Figura - Correlações no
+tempo}\label{figura---correlauxe7uxf5es-no-tempo}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Correlation over time}
-\NormalTok{cap <-}\StringTok{ "Correlation of variables with compressive strength over time"}
-\NormalTok{name <-}\StringTok{ "Correlation"}
+\CommentTok{# Figura - Correlações no tempo}
+\NormalTok{cap <-}\StringTok{ "Correlação das variáveis com a resistência à compressão no tempo"}
+\NormalTok{name <-}\StringTok{ "Correlação"}
 \NormalTok{melt_dat_final }\OperatorTok{%>%}
@@
 
-\hypertarget{figure---relationship-between-approximated-mix-water-mpa-and-age}{%
-\subsubsection{Figure - Relationship between approximated mix, water,
-MPa and
-age}\label{figure---relationship-between-approximated-mix-water-mpa-and-age}}
+\hypertarget{figura---relauxe7uxe3o-entre-o-trauxe7o-aproximado-uxe1gua-mpa-e-idade}{%
+\subsubsection{Figura - Relação entre o traço aproximado, água, MPa e
+idade}\label{figura---relauxe7uxe3o-entre-o-trauxe7o-aproximado-uxe1gua-mpa-e-idade}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Relationship between approximated mix, water, MPa and age}
-\NormalTok{cap <-}\StringTok{ "Relationship between approximated mix, water, MPa and age"}
-\NormalTok{d <-}\StringTok{ " days"}
-\NormalTok{xlabel <-}\StringTok{ "Approximated mix"} 
-\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
-\NormalTok{label <-}\StringTok{ "Water /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
+\CommentTok{# Figura - Relação entre o traço aproximado, água, MPa e idade}
+\NormalTok{cap <-}\StringTok{ "Relação entre o traço aproximado, água, resistência à compressão e idade"}
+\NormalTok{d <-}\StringTok{ " dias"}
+\NormalTok{xlabel <-}\StringTok{ "Traço aproximado"} 
+\NormalTok{ylabel <-}\StringTok{ "Resistência à compressão (MPa)"}
+\NormalTok{label <-}\StringTok{ "Água /}\CharTok{\textbackslash{}n}\StringTok{Cimento"}
 \NormalTok{mix_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{figure---relationship-between-concrete-main-features}{%
-\subsubsection{Figure - Relationship between concrete main
-features}\label{figure---relationship-between-concrete-main-features}}
+\hypertarget{figura---relauxe7uxe3o-das-principais-caracteruxedsticas-do-concreto}{%
+\subsubsection{Figura - Relação das principais características do
+concreto}\label{figura---relauxe7uxe3o-das-principais-caracteruxedsticas-do-concreto}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Relationship between concrete main features}
-\NormalTok{cap <-}\StringTok{ "Relationship between concrete main features"}
-\NormalTok{d <-}\StringTok{ " days"}
-\NormalTok{xlabel <-}\StringTok{ "Water / Cement"}
-\NormalTok{ylabel <-}\StringTok{ "Compressive strength (MPa)"}
-\NormalTok{colour <-}\StringTok{ "Fine Agg. /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
-\NormalTok{size <-}\StringTok{ "Coarse Agg. /}\CharTok{\textbackslash{}n}\StringTok{Cement"}
+\CommentTok{# Figura - Relação das principais características do concreto}
+\NormalTok{cap <-}\StringTok{ "Relação das principais proporções do concreto"}
+\NormalTok{d <-}\StringTok{ " dias"}
+\NormalTok{xlabel <-}\StringTok{ "Água / Cimento"}
+\NormalTok{ylabel <-}\StringTok{ "Resistência à compressão (MPa)"}
+\NormalTok{colour <-}\StringTok{ "Ag. Miúdo /}\CharTok{\textbackslash{}n}\StringTok{Cimento"}
+\NormalTok{size <-}\StringTok{ "Ag. Graúdo /}\CharTok{\textbackslash{}n}\StringTok{Cimento"}
 \NormalTok{min_x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{min}\NormalTok{(mix_dat}\OperatorTok{$}\StringTok{`}\DataTypeTok{fine_aggregate_/_cement}\StringTok{`}\NormalTok{)}
@@
 
-\hypertarget{figure---variables-distribution}{%
-\subsubsection{Figure - Variables
-distribution}\label{figure---variables-distribution}}
+\hypertarget{figura---distribuiuxe7uxe3o-das-variuxe1veis}{%
+\subsubsection{Figura - Distribuição das
+variáveis}\label{figura---distribuiuxe7uxe3o-das-variuxe1veis}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Variables distribution}
-
-\NormalTok{cap <-}\StringTok{ "Variables distribution"}
-\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-                  \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
-                  \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
-                  \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
-                  \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
-\NormalTok{dist_dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, day_}\DecValTok{3}\NormalTok{, day_}\DecValTok{7}\NormalTok{, day_}\DecValTok{14}\NormalTok{, day_}\DecValTok{56}\NormalTok{, day_}\DecValTok{100}\NormalTok{, mix_app))}
-
-\KeywordTok{colnames}\NormalTok{(dist_dat) <-}\StringTok{ }\NormalTok{colnames_dat}
-\NormalTok{dist_dat <-}\StringTok{ }\NormalTok{dist_dat }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"Var"}\NormalTok{, }\StringTok{"value"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{value =} \KeywordTok{as.numeric}\NormalTok{(value))}
-\NormalTok{dist_dat }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(value)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ ..density..), }
-                 \DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }
-                 \DataTypeTok{fill =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{.5}\NormalTok{, }\DataTypeTok{fill =} \StringTok{"lightseagreen"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{Var, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{, }\DataTypeTok{scale =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
-        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
-        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{figure---variables-distribution-grouped-by-age}{%
-\subsubsection{Figure - Variables distribution grouped by
-age}\label{figure---variables-distribution-grouped-by-age}}
+\CommentTok{# Figura - Distribuição das variáveis  }
+\NormalTok{cap <-}\StringTok{ "Distribuição das variáveis"}
+\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+                 \StringTok{"Superplastificante"}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }\StringTok{"MPa"}\NormalTok{,}
+                 \StringTok{"Água / Cimento"}\NormalTok{, }\StringTok{"A.Miúdo / Cimento"}\NormalTok{, }\StringTok{"A.Graúdo / Cimento"}\NormalTok{,}
+                 \StringTok{"A.Miúdo / A.Graúdo","}\NormalTok{Água }\OperatorTok{/}\StringTok{ }\NormalTok{A.Graúdo", }\StringTok{"Água / A.Miúdo")}
+\StringTok{dist_dat <- dat %>%}
+\StringTok{  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))}
+\StringTok{  }
+\StringTok{colnames(dist_dat) <- colnames_dat}
+\StringTok{dist_dat <- dist_dat %>%}
+\StringTok{  gather("}\NormalTok{Var}\StringTok{", "}\NormalTok{value}\StringTok{") %>%}
+\StringTok{  mutate(value = as.numeric(value))}
+\StringTok{dist_dat %>%}
+\StringTok{  ggplot(aes(value)) +}
+\StringTok{  geom_histogram(aes(y = ..density..), }
+\StringTok{                 colour = "}\NormalTok{black}\StringTok{", }
+\StringTok{                 fill = "}\NormalTok{white}\StringTok{") +}
+\StringTok{  geom_density(alpha = .5, fill = "}\NormalTok{lightseagreen}\StringTok{") +}
+\StringTok{  facet_wrap(~ Var, ncol=3, scale = "}\NormalTok{free}\StringTok{") +}
+\StringTok{  theme_minimal() +}
+\StringTok{  xlab("") +}
+\StringTok{  ylab("") +}
+\StringTok{  theme(panel.border = element_blank(), panel.grid.major = element_blank(),}
+\StringTok{        panel.grid.minor = element_blank(), }
+\StringTok{        axis.line = element_line(colour = "}\NormalTok{black}\StringTok{"))}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{figura---distribuiuxe7uxe3o-das-variuxe1veis-agrupadas-por-idade}{%
+\subsubsection{Figura - Distribuição das variáveis agrupadas por
+idade}\label{figura---distribuiuxe7uxe3o-das-variuxe1veis-agrupadas-por-idade}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Variables distribution grouped by age}
-
+\CommentTok{# Figura - Distribuição das variáveis agrupadas por idade  }
 \NormalTok{days_labs <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3"}\NormalTok{,}\StringTok{"7"}\NormalTok{,}\StringTok{"14"}\NormalTok{,}\StringTok{"28"}\NormalTok{,}\StringTok{"56"}\NormalTok{,}\StringTok{"100"}\NormalTok{)}
-\NormalTok{cap <-}\StringTok{ "Variables distribution grouped by age"}
-\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"Water"}\NormalTok{,}
-                  \StringTok{"Superplast."}\NormalTok{, }\StringTok{"Coarse agg."}\NormalTok{, }\StringTok{"Fine agg."}\NormalTok{,}
+\NormalTok{cap <-}\StringTok{ "Distribuição das variáveis em relação a idade"}
+\NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Volante"}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+                       \StringTok{"Superplastificante"}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }
 \NormalTok{                       days_labs,}
-                  \StringTok{"Water / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Cement"}\NormalTok{, }
-                  \StringTok{"Coarse agg. / Cement"}\NormalTok{, }\StringTok{"Fine agg. / Coarse agg."}\NormalTok{,}
-                  \StringTok{"Water / Coarse agg."}\NormalTok{, }\StringTok{"Water / Fine agg."}\NormalTok{)}
-\NormalTok{dist_dat_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(id, class, mix_app))}
-\KeywordTok{colnames}\NormalTok{(dist_dat_}\DecValTok{2}\NormalTok{) <-}\StringTok{ }\NormalTok{colnames_dat}
-\NormalTok{dist_dat_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{dist_dat_}\DecValTok{2} \OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"day"}\NormalTok{, }\StringTok{"mpa"}\NormalTok{, days_labs) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{drop_na}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(}\StringTok{"Var"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{"day"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{day =} \KeywordTok{factor}\NormalTok{(day, }\DataTypeTok{levels =}\NormalTok{ days_labs))}
-\NormalTok{dist_dat_}\DecValTok{2} \OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ day, }\DataTypeTok{y =}\NormalTok{ value)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_violin}\NormalTok{(}\DataTypeTok{color =} \OtherTok{NA}\NormalTok{,}
-                \DataTypeTok{fill =} \StringTok{"lightseagreen"}\NormalTok{,}
-                \DataTypeTok{alpha =} \FloatTok{.5}\NormalTok{,}
-                \DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{,}
-                \DataTypeTok{scale =} \StringTok{"count"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\StringTok{ }\NormalTok{Var, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{, }\DataTypeTok{scale =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{panel.border =} \KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{panel.grid.major =} \KeywordTok{element_blank}\NormalTok{(),}
-        \DataTypeTok{panel.grid.minor =} \KeywordTok{element_blank}\NormalTok{(), }
-        \DataTypeTok{axis.line =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{""}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{""}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{figure---principal-component-analysis-on-ingredients}{%
-\subsubsection{Figure - Principal component analysis on
-ingredients}\label{figure---principal-component-analysis-on-ingredients}}
+                       \StringTok{"Água / Cimento"}\NormalTok{, }\StringTok{"A.Miúdo / Cimento"}\NormalTok{, }\StringTok{"A.Graúdo / Cimento"}\NormalTok{,}
+                       \StringTok{"A.Miúdo / A.Graúdo","}\NormalTok{Água }\OperatorTok{/}\StringTok{ }\NormalTok{A.Graúdo", }\StringTok{"Água / A.Miúdo")}
+\StringTok{dist_dat_2 <- dat %>%}
+\StringTok{  select(-c(id, class, mix_app))}
+\StringTok{colnames(dist_dat_2) <- colnames_dat}
+\StringTok{dist_dat_2 <- dist_dat_2 %>%}
+\StringTok{  gather("}\NormalTok{day}\StringTok{", "}\NormalTok{mpa}\StringTok{", days_labs) %>%}
+\StringTok{  drop_na() %>%}
+\StringTok{  gather("}\NormalTok{Var}\StringTok{", "}\NormalTok{value}\StringTok{", -c("}\NormalTok{day}\StringTok{")) %>%}
+\StringTok{  mutate(day = factor(day, levels = days_labs))}
+\StringTok{dist_dat_2 %>%}
+\StringTok{  ggplot(aes(x = day, y = value)) +}
+\StringTok{  geom_violin(color = NA,}
+\StringTok{                fill = "}\NormalTok{lightseagreen}\StringTok{",}
+\StringTok{                alpha = .5,}
+\StringTok{                na.rm = TRUE,}
+\StringTok{                scale = "}\NormalTok{count}\StringTok{") +}
+\StringTok{  geom_boxplot(alpha = 0.2) +}
+\StringTok{  facet_wrap(~ Var, ncol=3, scale = "}\NormalTok{free}\StringTok{") +}
+\StringTok{  theme_minimal() +}
+\StringTok{  theme(panel.border = element_blank(), panel.grid.major = element_blank(),}
+\StringTok{        panel.grid.minor = element_blank(), }
+\StringTok{        axis.line = element_line(colour = "}\NormalTok{black}\StringTok{")) +}
+\StringTok{  xlab("") +}
+\StringTok{  ylab("")}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{figura---anuxe1lise-componente-principal-nos-ingredientes}{%
+\subsubsection{Figura - Análise componente principal nos
+ingredientes}\label{figura---anuxe1lise-componente-principal-nos-ingredientes}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Principal component analysis on ingredients}
-
-\NormalTok{cap <-}\StringTok{ "Principal component analysis on ingredients"}
+\CommentTok{# Figura - Análise componente principal nos ingredientes  }
+\NormalTok{cap <-}\StringTok{ "Análise componente principal nos ingredientes"}
 \NormalTok{colnames_dat <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}
-  \StringTok{"Cem."}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"F.A."}\NormalTok{, }\StringTok{"Wat."}\NormalTok{, }\StringTok{"Sup."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{,}
-  \StringTok{"F.Agg."}\NormalTok{,}\StringTok{"WxC"}\NormalTok{, }\StringTok{"F.Agg.xC"}\NormalTok{,}\StringTok{"C.Agg.xC"}\NormalTok{,}\StringTok{"F.Agg.xC"}\NormalTok{,}
-  \StringTok{"WxC.Agg."}\NormalTok{,}\StringTok{"WxF.Agg."}
+  \StringTok{"Cim."}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"Ci.Vo."}\NormalTok{, }\StringTok{"Águ."}\NormalTok{, }\StringTok{"Sup."}\NormalTok{, }\StringTok{"Ag.G."}\NormalTok{,}
+  \StringTok{"Ag.M."}\NormalTok{,}\StringTok{"AxC"}\NormalTok{, }\StringTok{"MxC"}\NormalTok{,}\StringTok{"GxC"}\NormalTok{,}\StringTok{"MxC"}\NormalTok{,}\StringTok{"AxG"}\NormalTok{,}\StringTok{"AxM"}
 \NormalTok{)}
-\NormalTok{class_list <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{low=}\StringTok{"0"}\NormalTok{, }\DataTypeTok{normal=}\StringTok{"20"}\NormalTok{, }\DataTypeTok{medium=}\StringTok{"40"}\NormalTok{, }\DataTypeTok{high=}\StringTok{"70"}\NormalTok{)}
+\NormalTok{class_list <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{baixo=}\StringTok{"0"}\NormalTok{, }\DataTypeTok{normal=}\StringTok{"20"}\NormalTok{, mé}\DataTypeTok{dio=}\StringTok{"40"}\NormalTok{, }\DataTypeTok{alto=}\StringTok{"70"}\NormalTok{)}
 \NormalTok{dat_pca <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
@@
 
-\hypertarget{machine-learning-models-1}{%
-\subsection{Machine learning models}\label{machine-learning-models-1}}
-
-\hypertarget{dummy-variables}{%
-\subsubsection{Dummy variables}\label{dummy-variables}}
+\hypertarget{modelos-de-machine-learning-1}{%
+\subsection{Modelos de machine
+learning}\label{modelos-de-machine-learning-1}}
+
+\hypertarget{variuxe1veis-fictuxedcias---dummy-vars}{%
+\subsubsection{Variáveis fictícias - dummy
+vars}\label{variuxe1veis-fictuxedcias---dummy-vars}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Dummy variables}
+\CommentTok{# Variáveis fictícias - dummy vars}
 \NormalTok{dummies <-}\StringTok{ }\KeywordTok{dummyVars}\NormalTok{( }\OperatorTok{~}\StringTok{ }\NormalTok{mix_app, }\DataTypeTok{data =}\NormalTok{ dat)}
@@
 
-\hypertarget{data-preparation-2}{%
-\subsubsection{Data preparation}\label{data-preparation-2}}
+\hypertarget{preparauxe7uxe3o-dos-dados-2}{%
+\subsubsection{Preparação dos
+dados}\label{preparauxe7uxe3o-dos-dados-2}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Data preparation}
+\CommentTok{# Preparação dos dados}
 \KeywordTok{names}\NormalTok{(dat) <-}\StringTok{ }\KeywordTok{gsub}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{names}\NormalTok{(dat), }\DataTypeTok{pattern =} \StringTok{"/"}\NormalTok{, }\DataTypeTok{replacement =} \StringTok{"."}\NormalTok{)  }
@@
 
-\hypertarget{table---first-18-columns-of-the-6-first-samples-of-28-days}{%
-\subsubsection{Table - First 18 columns of the 6 first samples of 28
-days}\label{table---first-18-columns-of-the-6-first-samples-of-28-days}}
+\hypertarget{tabela---primeiras-18-colunas-das-primeiras-6-amostras-de-28-dias}{%
+\subsubsection{Tabela - Primeiras 18 colunas das primeiras 6 amostras de
+28
+dias}\label{tabela---primeiras-18-colunas-das-primeiras-6-amostras-de-28-dias}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - First 18 columns of the 6 first samples of 28 days}
-\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"water"}\NormalTok{,}
-              \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Agg."}\NormalTok{, }\StringTok{"Fine Agg."}\NormalTok{, }\StringTok{"MPa"}\NormalTok{, }\StringTok{"Class"}\NormalTok{,}
-              \StringTok{"Wat./"}\NormalTok{, }\StringTok{"F.Agg./"}\NormalTok{, }\StringTok{"C.Agg./"}\NormalTok{,}
-              \StringTok{"F.Agg./"}\NormalTok{,}\StringTok{"Wat./"}\NormalTok{, }\StringTok{"Wat./"}\NormalTok{, }\StringTok{"App Mix"}\NormalTok{ =}\StringTok{ }\DecValTok{2}\NormalTok{)}
+\CommentTok{# Tabela - Primeiras 18 colunas das primeiras 6 amostras de 28 dias}
+\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Vol."}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+              \StringTok{"Superp."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo", }\StringTok{"MPa"}\NormalTok{, }\StringTok{"Classe"}\NormalTok{,}
+              \StringTok{"Ág./"}\NormalTok{, }\StringTok{"A.M./"}\NormalTok{, }\StringTok{"A.G./"}\NormalTok{,}
+              \StringTok{"A.M./"}\NormalTok{,}\StringTok{"Ág./"}\NormalTok{, }\StringTok{"Ág./"}\NormalTok{, }\StringTok{"Traço Apox."}\NormalTok{ =}\StringTok{ }\DecValTok{2}\NormalTok{)}
 \NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{""}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
               \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$MPa$"}\NormalTok{,}\StringTok{""}\NormalTok{, }\StringTok{"Ci."}\NormalTok{,}
-              \StringTok{"Ce."}\NormalTok{, }\StringTok{"Ce."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"F.Agg."}\NormalTok{, }\StringTok{"1:1:2"}\NormalTok{, }\StringTok{"1:2:2"}\NormalTok{)}
-\NormalTok{caption <-}\StringTok{ "First 18 columns of the 6 first samples of 28 days"}
+              \StringTok{"Ci."}\NormalTok{, }\StringTok{"Ci."}\NormalTok{, }\StringTok{"A.G."}\NormalTok{, }\StringTok{"A.G."}\NormalTok{, }\StringTok{"Ag.M."}\NormalTok{, }\StringTok{"1:1:2"}\NormalTok{, }\StringTok{"1:2:2"}\NormalTok{)}
+\NormalTok{caption <-}\StringTok{ "Primeiras 18 colunas das primeiras 6 amostras de 28 dias"}
 \NormalTok{dat_table <-}\StringTok{ }\NormalTok{dat_}\DecValTok{28}\NormalTok{[}\DecValTok{1}\OperatorTok{:}\DecValTok{18}\NormalTok{]}
@@
 
-\hypertarget{removing-near-zero-variance-columns}{%
-\subsubsection{Removing near zero variance
-columns}\label{removing-near-zero-variance-columns}}
+\hypertarget{removendo-colunas-com-variuxe2ncia-pruxf3xima-a-zero}{%
+\subsubsection{Removendo colunas com variância próxima a
+zero}\label{removendo-colunas-com-variuxe2ncia-pruxf3xima-a-zero}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Removing near zero variance columns}
+\CommentTok{# Removendo colunas com variância próxima a zero}
 \NormalTok{nzv_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{nearZeroVar}\NormalTok{(dat_}\DecValTok{3}\NormalTok{)}
@@
 
-\hypertarget{high-correlation-variables-verification}{%
-\subsubsection{High correlation variables
-verification}\label{high-correlation-variables-verification}}
+\hypertarget{verificauxe7uxe3o-de-variuxe1veis-com-alta-correlauxe7uxe3o}{%
+\subsubsection{Verificação de variáveis com alta
+correlação}\label{verificauxe7uxe3o-de-variuxe1veis-com-alta-correlauxe7uxe3o}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# High correlation variables verification}
+\CommentTok{# Verificação de variáveis com alta correlação}
 \NormalTok{descr_cor_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(}\KeywordTok{select}\NormalTok{(dat_}\DecValTok{3}\NormalTok{, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(mpa, id, class)))}
@@
 
-\hypertarget{test-and-train-data-split}{%
-\subsubsection{Test and train data
-split}\label{test-and-train-data-split}}
+\hypertarget{separauxe7uxe3o-em-conjunto-de-teste-e-treino}{%
+\subsubsection{Separação em conjunto de teste e
+treino}\label{separauxe7uxe3o-em-conjunto-de-teste-e-treino}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Test and train data split}
+\CommentTok{# Separação em conjunto de teste e treino}
 \NormalTok{reg <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}
@@
 
-\hypertarget{distribution-of-test-and-train-data}{%
-\subsubsection{Distribution of test and train
-data}\label{distribution-of-test-and-train-data}}
+\hypertarget{distribuiuxe7uxe3o-dos-conjuntos-de-teste-e-treino}{%
+\subsubsection{Distribuição dos conjuntos de teste e
+treino}\label{distribuiuxe7uxe3o-dos-conjuntos-de-teste-e-treino}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Distribution of test and train data}
-\NormalTok{cap <-}\StringTok{ "Distribution of test and train data"}
-\NormalTok{d_lab <-}\StringTok{ " days"}
-\NormalTok{ylabel <-}\StringTok{ "Density"}
-\NormalTok{g1 <-}\StringTok{ "Train"}
-\NormalTok{g2 <-}\StringTok{ "Test"}
+\CommentTok{# Distribuição dos conjuntos de teste e treino}
+\NormalTok{cap <-}\StringTok{ "Distribuição dos conjuntos de teste e treino"}
+\NormalTok{d_lab <-}\StringTok{ " dias"}
+\NormalTok{ylabel <-}\StringTok{ "Densidade"}
+\NormalTok{g1 <-}\StringTok{ "Treino"}
+\NormalTok{g2 <-}\StringTok{ "Teste"}
 \NormalTok{dens <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d, n)\{}
@@
 
-\hypertarget{naive-model}{%
-\subsubsection{Naive model}\label{naive-model}}
+\hypertarget{modelo-inguxeanuo}{%
+\subsubsection{Modelo ingênuo}\label{modelo-inguxeanuo}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Naive model}
+\CommentTok{# Modelo ingênuo}
 \NormalTok{res <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(d, n)\{}
@@
 
-\hypertarget{table---naive-models}{%
-\subsubsection{Table - Naive models}\label{table---naive-models}}
+\hypertarget{tabela---modelos-inguxeanuo}{%
+\subsubsection{Tabela - Modelos
+ingênuo}\label{tabela---modelos-inguxeanuo}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Naive models}
-\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Age"}\NormalTok{, }\StringTok{"Mean $MPa$ (train)"}\NormalTok{, }\StringTok{"RMSE (train)"}\NormalTok{, }\StringTok{"RMSE (test)"}\NormalTok{)}
-\NormalTok{caption <-}\StringTok{ "Naive models"}
-\KeywordTok{kable}\NormalTok{(}
-\NormalTok{    df_performance_reg,}
-    \DataTypeTok{col.names =}\NormalTok{ colNames,}
-    \DataTypeTok{escape =}\NormalTok{ F,}
-    \DataTypeTok{booktabs =}\NormalTok{ T,}
-    \DataTypeTok{caption =}\NormalTok{ caption,}
-    \DataTypeTok{linesep =} \StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{addlinespace"}\NormalTok{,}
-    \DataTypeTok{align =} \StringTok{"c"}
-\NormalTok{    )  }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"HOLD_position"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{features-selection}{%
-\subsubsection{Features selection}\label{features-selection}}
+\CommentTok{# Tabela - Modelos ingênuo}
+\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Idade"}\NormalTok{, }\StringTok{"Média $MPa$ (treino)"}\NormalTok{, }\StringTok{"RMSE (treino)"}\NormalTok{, }\StringTok{"RMSE (teste)"}\NormalTok{)}
+\NormalTok{caption <-}\StringTok{ "Modelos ingênuos"}
+\StringTok{kable(}
+\StringTok{    df_performance_reg,}
+\StringTok{    col.names = colNames,}
+\StringTok{    escape = F,}
+\StringTok{    booktabs = T,}
+\StringTok{    caption = caption,}
+\StringTok{    linesep = "}\NormalTok{\textbackslash{}\textbackslash{}addlinespace}\StringTok{",}
+\StringTok{    align = "}\NormalTok{c}\StringTok{"}
+\StringTok{    )  %>%}
+\StringTok{  kable_styling(latex_options = c("}\NormalTok{HOLD_position}\StringTok{"))}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{escolha-das-caracteruxedsitcas-features}{%
+\subsubsection{Escolha das caracterísitcas
+(features)}\label{escolha-das-caracteruxedsitcas-features}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Features selection}
+\CommentTok{# Escolha das caracterísitcas (features)}
 \NormalTok{data1 <-}\StringTok{ }\NormalTok{dats_reg}\OperatorTok{$}\NormalTok{d28}\OperatorTok{$}\NormalTok{train }\CommentTok{# full}
@@
 
-\hypertarget{table---first-6-samples-of-train-data-of-the-28-days-model}{%
-\subsubsection{Table - First 6 samples of train data of the 28 days
-model}\label{table---first-6-samples-of-train-data-of-the-28-days-model}}
+\hypertarget{tabela---primeiras-6-amostras-do-conjunto-de-treino-do-modelo-de-28-dias}{%
+\subsubsection{Tabela - Primeiras 6 amostras do conjunto de treino do
+modelo de 28
+dias}\label{tabela---primeiras-6-amostras-do-conjunto-de-treino-do-modelo-de-28-dias}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - First 6 samples of train data of the 28 days model}
-\NormalTok{caption <-}\StringTok{ "First 6 samples of train data of the 28 days model"}
-\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cement"}\NormalTok{, }\StringTok{"B.F.S."}\NormalTok{, }\StringTok{"Fly ash"}\NormalTok{, }\StringTok{"water"}\NormalTok{,}
-              \StringTok{"Superp."}\NormalTok{, }\StringTok{"Coarse Agg."}\NormalTok{, }\StringTok{"Fine Agg."}\NormalTok{,}
-              \StringTok{"Wat./"}\NormalTok{, }\StringTok{"F.Agg./"}\NormalTok{, }\StringTok{"C.Agg./"}\NormalTok{,}
-              \StringTok{"F.Agg./"}\NormalTok{,}\StringTok{"Wat./"}\NormalTok{, }\StringTok{"Wat./"}\NormalTok{, }\StringTok{"y"}\NormalTok{)}
+\CommentTok{# Tabela - Primeiras 6 amostras do conjunto de treino do modelo de 28 dias}
+\NormalTok{caption <-}\StringTok{ "Primeiras 6 amostras do conjunto de treino do modelo de 28 dias"}
+\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Cimento"}\NormalTok{, }\StringTok{"E.G.A.F."}\NormalTok{, }\StringTok{"C.Vol."}\NormalTok{, }\StringTok{"Água"}\NormalTok{,}
+              \StringTok{"Superp."}\NormalTok{, }\StringTok{"A.Graúdo", "}\NormalTok{A.Miúdo",}
+              \StringTok{"Ág./"}\NormalTok{, }\StringTok{"A.M./"}\NormalTok{, }\StringTok{"A.G./"}\NormalTok{,}
+              \StringTok{"A.M./"}\NormalTok{,}\StringTok{"Ág./"}\NormalTok{, }\StringTok{"Ág./"}\NormalTok{, }\StringTok{"y"}\NormalTok{)}
 \NormalTok{dfUnits <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}
-              \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"Ce."}\NormalTok{,}
-              \StringTok{"Ce."}\NormalTok{, }\StringTok{"Ce."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"C.Agg."}\NormalTok{, }\StringTok{"F.Agg."}\NormalTok{, }\StringTok{"MPa"}\NormalTok{)}
+              \StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"$kg/m^3$"}\NormalTok{,}\StringTok{"$kg/m^3$"}\NormalTok{, }\StringTok{"Ci."}\NormalTok{,}
+              \StringTok{"Ci."}\NormalTok{, }\StringTok{"Ci."}\NormalTok{, }\StringTok{"A.G."}\NormalTok{, }\StringTok{"A.G."}\NormalTok{, }\StringTok{"Ag.M."}\NormalTok{, }\StringTok{"$MPa$"}\NormalTok{)}
 \NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Features"}\NormalTok{=}\DecValTok{13}\NormalTok{, }\StringTok{"Outcome"}\NormalTok{=}\DecValTok{1}\NormalTok{)}
@@
 
-\hypertarget{regression-models-1}{%
-\subsubsection{Regression models}\label{regression-models-1}}
+\hypertarget{modelos-de-regressuxe3o-1}{%
+\subsubsection{Modelos de regressão}\label{modelos-de-regressuxe3o-1}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Regression models}
+\CommentTok{# Modelos de regressão}
 \CommentTok{# no dummy vars:}
@@
 
-\hypertarget{results-1}{%
-\subsection{Results}\label{results-1}}
-
-\hypertarget{table---models-details}{%
-\subsubsection{Table - Models details}\label{table---models-details}}
+\hypertarget{resultados-1}{%
+\subsection{Resultados}\label{resultados-1}}
+
+\hypertarget{tabela---detalhes-dos-modelos}{%
+\subsubsection{Tabela - Detalhes dos
+modelos}\label{tabela---detalhes-dos-modelos}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Models details}
-\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"mtry"}\NormalTok{, }\StringTok{"CV"}\NormalTok{, }\StringTok{"Repetitions"}\NormalTok{, }
-              \StringTok{"RMSE (train)"}\NormalTok{, }\StringTok{"RMSE (test)"}\NormalTok{)}
-\NormalTok{caption <-}\StringTok{ "Regression models results"}
-\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
+\CommentTok{# Tabela - Detalhes dos modelos}
+\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Modelo"}\NormalTok{, }\StringTok{"mtry"}\NormalTok{, }\StringTok{"CV"}\NormalTok{, }\StringTok{"Repetições"}\NormalTok{, }
+              \StringTok{"RMSE (treino)"}\NormalTok{, }\StringTok{"RMSE (teste)"}\NormalTok{)}
+\NormalTok{caption <-}\StringTok{ "Resultados dos modelos de regressão"}
+\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{, }\StringTok{"14 dias"}\NormalTok{, }\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{)}
 \NormalTok{dat_reg_models <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
@@
 
-\hypertarget{figure---models-comparison}{%
-\subsubsection{Figure - Models
-comparison}\label{figure---models-comparison}}
+\hypertarget{figura---comparauxe7uxe3o-dos-modelos}{%
+\subsubsection{Figura - Comparação dos
+modelos}\label{figura---comparauxe7uxe3o-dos-modelos}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Figure - Models comparison}
-\NormalTok{cap <-}\StringTok{ "Actual vs Predicted values in each model"}
-\NormalTok{models <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
-\NormalTok{xlabel <-}\StringTok{ "Actual (MPa)"}
-\NormalTok{ylabel <-}\StringTok{ "Predicted (MPa)"}
+\CommentTok{# Figura - Comparação dos modelos}
+\NormalTok{cap <-}\StringTok{ "Comparação dos valores reais e previstos em cada modelo"}
+\NormalTok{models <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{, }\StringTok{"14 dias"}\NormalTok{, }\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{)}
+\NormalTok{xlabel <-}\StringTok{ "Real (MPa)"}
+\NormalTok{ylabel <-}\StringTok{ "Previsto (MPa)"}
 \NormalTok{preds <-}\StringTok{ }\KeywordTok{list}\NormalTok{(p_}\DecValTok{3}\NormalTok{, p_}\DecValTok{7}\NormalTok{, p_}\DecValTok{14}\NormalTok{, p_}\DecValTok{28}\NormalTok{, p_}\DecValTok{56}\NormalTok{, p_}\DecValTok{100}\NormalTok{)}
@@
 
-\hypertarget{results-tables-of-10-best-and-worst-results}{%
-\subsubsection{Results tables of 10 best and worst
-results}\label{results-tables-of-10-best-and-worst-results}}
+\hypertarget{tabelas-dos-10-melhores-e-piores-resultados}{%
+\subsubsection{Tabelas dos 10 melhores e piores
+resultados}\label{tabelas-dos-10-melhores-e-piores-resultados}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Results tables of 10 best and worst results}
-\NormalTok{colNames1 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Actual"}\NormalTok{, }\StringTok{"Predicted"}\NormalTok{, }\StringTok{"Error"}\NormalTok{, }\StringTok{""}\NormalTok{,}\StringTok{"Actual"}\NormalTok{, }\StringTok{"Predicted"}\NormalTok{, }\StringTok{"Error"}\NormalTok{) }
-\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Best 10"}\NormalTok{=}\DecValTok{3}\NormalTok{,}\StringTok{""}\NormalTok{, }\StringTok{"Worst 10"}\NormalTok{=}\DecValTok{3}\NormalTok{) }
-\NormalTok{caption_}\DecValTok{0}\NormalTok{ <-}\StringTok{ "Model of "}
+\CommentTok{# Tabelas dos 10 melhores e piores resultados}
+\NormalTok{colNames1 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Real"}\NormalTok{, }\StringTok{"Previsto"}\NormalTok{, }\StringTok{"Erro"}\NormalTok{, }\StringTok{""}\NormalTok{,}\StringTok{"Real"}\NormalTok{, }\StringTok{"Previsto"}\NormalTok{, }\StringTok{"Erro"}\NormalTok{) }
+\NormalTok{colNames2 =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"10 melhores"}\NormalTok{=}\DecValTok{3}\NormalTok{,}\StringTok{""}\NormalTok{, }\StringTok{"10 piores"}\NormalTok{=}\DecValTok{3}\NormalTok{)}
+\NormalTok{caption_}\DecValTok{0}\NormalTok{ <-}\StringTok{ "Modelo de "}
 \NormalTok{get_X <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(pred, actual, }\DataTypeTok{X=}\DecValTok{10}\NormalTok{)\{}
@@
 
-\hypertarget{discussion-and-conclusion-1}{%
-\subsection{Discussion and
-conclusion}\label{discussion-and-conclusion-1}}
-
-\hypertarget{table---comparison-of-other-works}{%
-\subsubsection{Table - Comparison of other
-works}\label{table---comparison-of-other-works}}
+\hypertarget{discussuxe3o-1}{%
+\subsection{Discussão}\label{discussuxe3o-1}}
+
+\hypertarget{tabela---comparauxe7uxe3o-dos-estudos-de-outros-autores}{%
+\subsubsection{Tabela - ``Comparação dos estudos de outros
+autores''}\label{tabela---comparauxe7uxe3o-dos-estudos-de-outros-autores}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Comparison of other works}
-\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Author"}\NormalTok{,}\StringTok{"Year"}\NormalTok{ ,}\StringTok{"Algorithm"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{) }
-\NormalTok{caption <-}\StringTok{ "Comparison of other works"}
+\CommentTok{# Tabela - "Comparação dos estudos de outros autores"}
+\NormalTok{colNames =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Autor"}\NormalTok{,}\StringTok{"Ano"}\NormalTok{ ,}\StringTok{"Algorítimo"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{) }
+\NormalTok{caption <-}\StringTok{ "Comparação dos estudos de outros autores"}
 \NormalTok{works <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
@@
 
-\hypertarget{table---final-results}{%
-\subsubsection{Table - Final results}\label{table---final-results}}
+\hypertarget{tabela---resultados-finais}{%
+\subsubsection{Tabela - Resultados
+finais}\label{tabela---resultados-finais}}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Table - Final results}
-\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{)}
-\NormalTok{caption <-}\StringTok{ "Final result"}
-\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 days"}\NormalTok{, }\StringTok{"7 days"}\NormalTok{, }\StringTok{"14 days"}\NormalTok{, }\StringTok{"28 days"}\NormalTok{, }\StringTok{"56 days"}\NormalTok{, }\StringTok{"100 days"}\NormalTok{)}
+\CommentTok{# Tabela - Resultados finais}
+\NormalTok{colNames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Modelo"}\NormalTok{, }\StringTok{"RMSE"}\NormalTok{)}
+\NormalTok{caption <-}\StringTok{ "Resultados finais"}
+\NormalTok{day <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"3 dias"}\NormalTok{, }\StringTok{"7 dias"}\NormalTok{, }\StringTok{"14 dias"}\NormalTok{, }\StringTok{"28 dias"}\NormalTok{, }\StringTok{"56 dias"}\NormalTok{, }\StringTok{"100 dias"}\NormalTok{)}
 \NormalTok{dat_reg_models <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}

### 2020-08-14 _meta/templates/Abstract_&_Setup.Rmd
refactor

--- 2020-04-29 templates/Abstract_&_Setup.Rmd
@@
 indent: true
-<EN>abstract:
- Compressive strength is the main characteristic of concrete.
+<EN>abstract: "Compressive strength is the main characteristic of concrete.
 The correct prediction of this parameter means cost and time reduction.
 This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
-A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: resistance to compression, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
+A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
 Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
@@
 It was a promising alternative since good results were achieved by training with just one algorithm.
-This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several.</EN>
-<PT>abstract:
-  A resistência à compressão é a principal característica do concreto.
+This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several."</EN>
+<PT>abstract: "A resistência à compressão é a principal característica do concreto.
 A previsão correta desse parâmetro significa redução de custo e tempo.
@@
 Foi uma alternativa promissora, visto que bons resultados foram atingidos treinando com apenas um algoritmo.
-Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos.</PT>
+Esse trabalho facilita a exploração e novos esforços para previsão da resistência à compressão do concreto, ele pode ser replicado utilizando diferentes algoritmos ou a combinação de diversos."</PT>
 output:

### 2020-08-14 _meta/templates/Introduction.Rmd
refactor

--- 2020-07-13 HarvardX_Data_Science_Capstone/CopyOfcapstone.Rmd
@@
----
-title:
-  Concrete compressive strength prediction with machine learning
-date: "April 22, 2020"
-author: "Pedro Bernardino Alves Moreira"
-indent: true
-abstract: "Compressive strength is the main characteristic of concrete.
-The correct prediction of this parameter means cost and time reduction.
-This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days).
-A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers).
-Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate).
-The predictive models were developed in *R* language, using the *caret* package with the *Parallel Random Forest* algorithm and repeated cross-validation technique to optimize the parameters.
-The results were satisfactory and compatible with other studies using the same data set.
-The most important model, 28 days old, obtained *RMSE* of 4.717.
-The 3-day model obtained the best result, *RMSE* of 3.310.
-The worst result was the 56-day model, with *RMSE* of 5.939.
-The work showed that the compressive strength of concrete can be predicted.
-The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age.
-It was a promising alternative since good results were achieved by training with just one algorithm.
-This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several."
-output:
-  pdf_document:
-    keep_tex: yes
-    number_sections: yes
-    toc_depth: 2
-# bibliography: references.bib
-references:
-- id: Yeh1998
-  author:
-  - family: Yeh
-    given: I-Cheng
-  year: 1998
-  month: 12
-  pages: 1797-1808
-  title: Modeling of Strength of High-Performance Concrete Using Artificial Neural Networks
-  volume: 28
-  journal: Cement and Concrete Research
-  DOI: 10.1016/S0008-8846(98)00165-3
-- id: Hameed2020
-  author:
-  - family: Hameed
-    given: Mohammed
-  - family: Khalid
-    given: Mohamed
-   year: 2020
-   month: 01
-   pages: 323-335
-   title: "Prediction of Compressive Strength of High-Performance Concrete: Hybrid Artificial Intelligence Technique"
-   # isbn: 978-3-030-38751-8
-   DOI: 10.1007/978-3-030-38752-5_26
-header-includes:
-- \usepackage{booktabs}
-- \usepackage{longtable}
-- \usepackage{supertabular}
-- \usepackage{array}
-- \usepackage{multirow}
-- \usepackage{wrapfig}
-- \usepackage{colortbl}
-- \usepackage{pdflscape}
-- \usepackage{tabu}
-- \usepackage{threeparttable}
-- \usepackage[normalem]{ulem}
-- \usepackage{caption}
-- \usepackage{floatrow}
-- \floatsetup[table]{capposition=top}
-- \floatsetup[figure]{capposition=top}
-- \captionsetup{options=chunk}
-- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
-- \renewcommand{\thechunk}{\arabic{chunk}}
-- \usepackage{indentfirst}
-- \usepackage{sectsty} \sectionfont{\centering}
----
-```{r setup, include=FALSE, cache=F}
-# Install libraries, if not already installed
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(ggplot2)) install.packages("ggplot2")
-if(!require(caret)) install.packages("caret")
-if(!require(knitr)) install.packages("knitr")
-if(!require(kableExtra)) install.packages("kableExtra")
-if(!require(gdata)) install.packages("gdata")
-if(!require(dplyr)) install.packages("dplyr")
-if(!require(tidyr)) install.packages("tidyr")
-if(!require(cowplot)) install.packages("cowplot")
-if(!require(pastecs)) install.packages("pastecs")
-if(!require(factoextra)) install.packages("factoextra")
-if(!require(reshape2)) install.packages("reshape2")
-if(!require(purrr)) install.packages("purrr")
-if(!require(gridExtra)) install.packages("gridExtra")
-if(!require(questionr)) install.packages("questionr")
-# Load libraries
-library(tidyverse)
-library(caret)
-library(ggplot2)
-library(knitr)
-library(kableExtra)
-library(dplyr)
-library(tidyr)
-library(gdata)
-library(reshape2)
-library(cowplot)
-library(pastecs)
-library(factoextra)
-library(purrr)
-library(gridExtra)
-library(questionr)
-# Set global options
-options(pillar.sigfig = 5)
-options(knitr.kable.NA = '')
-opts_knit$set(eval.after = "fig.cap")
-opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
-oldSource <- knit_hooks$get("source")
-knit_hooks$set(source = function(x, options) {
-  x <- oldSource(x, options)
-  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
-  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
-})
-```
-
-\newpage
-\tableofcontents
-\newpage
-\listoffigures
-\listoftables
-\newpage
+<EN>
 
@@
 
-# Materials and Methods
+</EN>
 
-## Materials
+<PT>
 
-The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore].
-Throughout the process, all code executed was documented in the same order as its execution in [Appendix 2](#appendix2), and reference was always made to codes throughout the text.
-All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1).
+# Introdução
 
-## Reproducibility
-
-In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result.
-The *seeds* can be checked throughout [Appendix 2](#appendix2).
-
-## Obtaining the data
-
-The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}).
-In total there are 1030 samples with 9 columns.
-The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}).
-The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}).
-The first samples can be viewed in the table \ref{tab:first-samples}.
+A resistência à compressão é a principal característica do concreto, medida por testes de padrões internacionais que consistem na quebra de corpos de prova.
+A medição aos 28 dias é obrigatória e representa a classe do concreto.
+Saber com antecêdencia qual o resultado será obtido para uma determinada idade, a partir das proporções de seus ingredientes, é de grande interesse para os fabricantes de concreto, construtoras e engenheiros civis.
 
@@
 
-```{r download-data, echo=F, include=F}
-# Data download
-# url_base <- "https://archive.ics.uci.edu"
-# url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
-# download.file(paste0(url_base, url), "data.xls")
-dat <- read.xls("data.xls")
-n_inicial_samples <- nrow(dat)
-colnames(dat)
-```
-
-```{r rename-dat-cols, echo=F}
-# Renaming the columns
-colnames(dat) <- c(
-  "cement",
-  "blast_furnace_slag",
-  "fly_ash",
-  "water",
-  "superplasticizers",
-  "coarse_aggregate",
-  "fine_aggregate",
-  "day",
-  "mpa"
-  )
-dat$id <- seq.int(nrow(dat))
-```
-
-```{r reorder-dat, echo=F}
-# Reordering the data
-col_order <- c(
-  "id",
-  "cement",
-  "blast_furnace_slag",
-  "fly_ash",
-  "water",
-  "superplasticizers",
-  "coarse_aggregate",
-  "fine_aggregate",
-  "day",
-  "mpa"
-)
-dat <- dat[, col_order]
-```
-
-```{r col-names-and-units, echo=F}
-# Defining column names and units
-colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
-             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
-dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
-             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
-```
-
-```{r first-samples, echo=F}
-# Table - First samples
-caption <- "First 6 samples"
-kable(
-    dat[1:6,],
-    col.names = dfUnits,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    ) %>%
-    add_header_above(header = colNames, line = F, align = "c") %>%
-    kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-
-## Data preparation
-
-The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies.
-Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed.
-In addition, the relevant data has been transformed to better fit the studies in the next steps.
-
-### Initial data cleaning
-
-Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).
+Essa resistência à compressão é uma função não linear de seus ingredientes e idade, tornando difícil o estabelecimento de uma fórmula analítica, apesar de algumas fórmulas já haveram sido propostas.
+@Hasan2011 propós um modelo matemático para prever a partir dos resultados de testes de 7 e 14 dias, e @Kabir2012 a partir de 7 dias.
+Porém técnicas de machine learning podem ser utilizadas para modelar essa característica a partir de dados reais de amostras, utilizando apenas os ingredientes.
 
@@
 
-```{r removing-duplicated-samples, echo=F}
-# Removing duplicated samples
-n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
-n_duplicated_samples <- n_inicial_samples - n_distinct_samples
-dat <- dat[!duplicated(select(dat, -c(id))),]
-n_samples <- nrow(dat)
-```
-
-The data show the variables in the columns and samples in the rows.
-However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
+Muitos estudos anteriores utilizam o mesmo conjunto de dados utilizado por @Yeh1998 para prever a resistência à compressão do concreto.
+@Alshamiri2020 obteve bons resultados com a técnica de regularized extreme learning machine (RELM), e @Hameed2020 obteve resultados ainda melhores com a técnica de Artificial Neural Networks e cross-validation.
+Esse conjunto de amostras é tão conhecido que há ainda páginas na internet de estudos não publicados que o utilizam e possuem bons resultados, como @Abban2016, @Raj2018, @Modukuru2020 e @Pierobon2018.
+Ao final do trabalho os resultados encontrados são comparados aos trabalhos citados aqui.
 
@@
 
-```{r similar-samples, echo=FALSE}
-# Table - Samples with same composition
-same_samples <- dat %>%
-  filter(id %in% c(653, 678, 654, 681))
-caption <- "Samples with same composition"
-kable(
-  same_samples[order(same_samples$day),],
-  col.names = dfUnits,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "c",
-  row.names = FALSE
-  ) %>%
-  add_header_above(header = colNames, line = F, align = "c") %>%
-  kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
+Diferente dos estudos anteriores com esse conjunto de amostras, este trabalho faz a preparação dos dados de forma diferente.
+A idade do concreto é a variável mais singular que contribui para sua resistência à compressão, por esse motivo, a idade é tratada separadamente nos modelos de machine learning, criando modelos para cada faixa de idade.
 
-In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process.
-This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.
-
-&nbsp;
-
-```{r similar-samples-2, echo=FALSE}
-# Table - Same samples with different results
-same_samples_2 <- dat %>%
-  filter(id %in% c(472, 473, 474))
-caption <- "Same samples with different results"
-kable(
-  same_samples_2[order(same_samples_2$day),],
-  col.names = dfUnits,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "c",
-  row.names = FALSE
-  ) %>%
-  add_header_above(header = colNames, line = F, align = "c") %>%
-  kable_styling(latex_options = c("scale_down"))
-```
-
-To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*.
-In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained.
-In the case of the same samples but with different results of compressive strength, the values were averaged.
-After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.
-
-&nbsp;
-
-```{r initial-data-cleaning, echo=FALSE}
-# Data cleaning
-dat <- dat %>%
-  group_by(
-    cement,
-    blast_furnace_slag,
-    fly_ash,
-    water,
-    superplasticizers,
-    coarse_aggregate,
-    fine_aggregate,
-  ) %>%
-  filter("28" %in% day) %>%
-  mutate(id = id[which.min(id)]) %>%
-  ungroup() %>%
-  group_by(id, day) %>%
-  mutate(mpa = mean(mpa)) %>%
-  ungroup()
-dat <- dat[!duplicated(select(dat, -c(id))),]
-dat$id<-factor(dat$id)
-n_samples <- nrow(dat)
-n_distinct_samples <- n_distinct(dat$id)
-```
-
-The result can be seen in the table \ref{tab:similar-samples-same-id}.
-All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.
-
-&nbsp;
-
-```{r similar-samples-same-id, echo=FALSE, include=TRUE}
-# Table - Previous samples after processing
-same_samples <- dat %>%
-  filter(id == 653 | id == 472 & day == 28)
-caption <- "Previous samples after processing"
-kable(
-  same_samples[order(same_samples$id, same_samples$day),],
-  col.names = dfUnits,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "c",
-  row.names = FALSE,
-  digits = 2
-  ) %>%
-  add_header_above(header = colNames, line = F, align = "c") %>%
-  kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-
-### Age selection
-
-As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models.
-However, it is necessary to verify how relevant the data of these other ages are.
-Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}.
-
-&nbsp;
-
-```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
-# Figure - Compressive strength (MPa) vs age (days)
-cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
-ylabel <- "Compressive strength (MPa)"
-xlabel <- "Age (days)"
-dat %>%
-  ggplot(aes(x=factor(day), y=mpa)) +
-  geom_boxplot() +
-  geom_jitter(alpha=0.2)  +
-  theme_bw() +
-  ylab(ylabel) +
-  xlab(xlabel)
-```
-
-It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.
-
-&nbsp;
-
-This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}).
-The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis.
-The first two dimensions represent 37% and 24% $ respectively of the variance.
-
-&nbsp;
-
-```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
-# Figure - Principal component analysis - 90, 91 e 100 days
-dat_90_91_100 <- dat %>%
-  ungroup() %>%
-  filter(day %in% c(90, 91, 100)) %>%
-  select(-c(id, mpa))
-cap <- "Principal component analysis - 90, 91 e 100 days"
-colnames(dat_90_91_100) <- c(
-   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
-pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
-habillage <- dat_90_91_100$day
-fviz_pca_biplot(
-  pca,
-  geom.ind = "point",
-  habillage=habillage,
-  addEllipses = TRUE,
-  ellipse.level=0.75) +
-  ggtitle("") +
-  theme_bw() +
-  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
-```
-
-Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value.
-The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-time}).
-
-&nbsp;
-
-```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
-# Figure - Compressive strength through time
-cap <- "Compressive strength through time"
-ylabel <- "MPa"
-xlabel <- "Age (days)"
-dat_duplicated_only <- dat %>%
-  group_by(id) %>%
-  filter(n()>5) %>%
-  select(id, mpa, day)
-dat_duplicated_only %>%
-  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
-  geom_line() +
-  xlab(xlabel) +
-  ylab(ylabel) +
-  theme_bw() +
-  theme(legend.position = "none")
-```
-
-For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise.
-They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}).
-As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.
-
-&nbsp;
-
-```{r join-90-91-100, echo=F}
-# Joining 90, 91 and 100 days data
-ind_90 <- dat$id[which(dat$day == "90")]
-ind_91 <- dat$id[which(dat$day == "91")]
-ind_100 <- dat$id[which(dat$day == "100")]
-sum_duplicated <- sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
-dat <- dat %>%
-  ungroup() %>%
-  mutate(day = ifelse(day %in% c(91, 90), 100, day))
-```
-
-Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}.
-Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}).
-The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.
-
-&nbsp;
-
-```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
-# Figure - Ages frequency
-cap <- "Ages frequency"
-ylabel <- "Frequency"
-xlabel <- "Age (days)"
-dat %>%
-  ggplot(aes(x = factor(day))) +
-  geom_bar() +
-  theme_bw() +
-  xlab(xlabel) +
-  ylab(ylabel)
-```
-
-```{r remove-ages-lower-50, echo=F}
-# Removing ages with frequency lower than 50
-dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
-```
-
-### Data reorganization
-
-The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
-The result in the first samples after this processing is shown in the table \ref{tab:new-features}.
-
-&nbsp;
-
-```{r reorganizing-dat, echo=F}
-# Data reorganization
-dat <- dat %>%
-  group_by_at(vars(-mpa)) %>%
-  mutate(row_id = 1:n()) %>% ungroup() %>%
-  spread(day, mpa, sep = "_") %>%
-  select(-row_id)
-```
-
-```{r new-features, echo=F}
-# Table - First 6 samples after reorganization
-caption <- "First 6 samples after reorganization"
-colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
-             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
-             "14 days", "28 days", "56 days", "100 days")
-dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
-              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
-              "$MPa$","$MPa$","$MPa$","$MPa$")
-kable(
-    head(dat[order(dat$id),]),
-    col.names = dfUnits2,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c"
-    ) %>%
-    add_header_above(header = colNames2, line = F, align = "c") %>%
-    kable_styling(latex_options = c("HOLD_position", "scale_down"))
-```
-
-The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
-
-```{r total-samples-2, echo=F}
-# Total samples
-n_samples <- nrow(dat) # 416
-n_distinct_samples <- n_distinct(dat$id) # 416
-```
-
-### Adding new variables
-
-To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}).
-Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*.
-The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data.
-The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement.
-Other proportions between the main ingredients were also added.
-The new variables are presented in the table \ref{tab:new-features-table}.
-
-```{r new-features-2, echo=F}
-# Adding New features
-concrete_class <- function(mpa){
-  if (mpa >= 10) {
-    s <- as.character(mpa)
-    first <- substr(s, start = 1, stop = 1)
-    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
-  }
-  else {
-    first <- ""
-    second <- "5"
-  }
-  paste("C", first, second, sep = "")
-}
-mix <- function(c, f_ag, c_ag){
-  paste(
-    1,
-    round(f_ag/c, 0),
-    round(c_ag/c, 0),
-    sep = ":")
-}
-dat <- dat %>%
-  mutate(class = sapply(day_28, concrete_class)) %>%
-  mutate(class = as.factor(class)) %>%
-  mutate(mix_app = factor(
-    mix(cement, fine_aggregate, coarse_aggregate))) %>%
-  mutate(`water_/_cement` = water / cement) %>%
-  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
-  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
-  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
-  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
-  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
-lvl <- levels(dat$class)
-dat$class <- factor(
-  dat$class, 
-  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
-```
-
-```{r new-features-table, echo=F}
-# Table - New features
-caption <- "New features"
-colNames7 = c("ID", "Class","Approximated Mix",
-             "Water / Cement", "Fine Ag. / Cement", 
-             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
-             "Water / Coarse Ag.", "Water / Fine Ag.")
-kable(
-    head(dat[order(dat$id),][,c(1,15:22)]),
-    col.names = colNames7,
-    escape = F,
-    booktabs = T,
-    caption = caption,
-    linesep = "\\addlinespace",
-    align = "c",
-    digits = 4
-    ) %>%
-  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
-  column_spec(2, width = "1.5cm") %>%
-  column_spec(3, width = "2cm") %>%
-  column_spec(4:9, width = "1.7cm")
-```
-
-## Data visualization
-
-In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.
-
-### Descriptive statistics
-
-The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}).
-The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data.
-As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age.
-The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).
-
-&nbsp;
-
-```{r stat-summ, echo=FALSE}
-# Table - Descriptive statistics - continuous variables
-summ <- t(
-  stat.desc(select(dat, -c(id, class, mix_app))))
-caption <- "Descriptive statistics - continuous variables"
-colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
-              "Sum", "Median", "Mean", "SE mean", 
-              "CI mean","Variance", "Std.Dev.", "Coef.Var")
-rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
-              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
-              "14 days", "28 days", "56 days", "100 days",
-              "Water / Cement", "Fine agg. / Cement", 
-              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
-              "Water / Coarse agg.", "Water / Fine agg.")
-kable(
-  summ,
-  escape = F,
-  booktabs = T,
-  caption = caption,
-  linesep = "\\addlinespace",
-  align = "c",
-  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
-  ) %>%
-  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
-  column_spec(c(1,10:15), width = "1.5cm")
-```
-
-```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
-# Figure - Descriptive statistics - categorical variables
-cap <- "Descriptive statistics - categorical variables"
-name1 <- "Percentage"
-name2 <- "Accumulated percentage"
-ylabel <- "Frequency"
-xlabel1 <- "Class"
-xlabel2 <- "Approximate Mix"
-format_percent = function(n){
-  paste(n, "%", sep = "")
-}
-format_class <- function(cls){
-  str_remove_all(cls, "C")
-}
-f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
-  select(n, "%", "%cum") %>%
-  mutate(class = row.names(.)) %>%
-  mutate(class_n = as.numeric(format_class(class)))
-f_cls_labels = function(n) {
-  f_class$class[n]
-}
-f_cls_acc_labels = function(n){
-  paste(f_class$`%cum`[n], "%", sep = "")
-}
-p1 <- f_class %>%
-  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
-  geom_bar(stat = 'identity') +
-  scale_y_continuous(
-      sec.axis = sec_axis(~./length(dat$class) * 100,
-                          name = name1,
-                          labels = format_percent)) +
-  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
-                     sec.axis = sec_axis(~., breaks = 1:16,
-                              name = name2,
-                              labels = f_cls_acc_labels)) +
-  theme_bw() +
-  theme(panel.grid.minor.y = element_blank()) +
-  xlab(xlabel1) +
-  ylab(ylabel) +
-  coord_flip()
-format_mix <- function(mix){
-  str_remove_all(mix, ":")
-}
-f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
-  select(n, "%", "%cum") %>%
-  mutate(mix = row.names(.)) %>%
-  mutate(mix_n = as.numeric(format_mix(mix)))
-f_mix_labels = function(n) {
-  f_mix$mix[n]
-}
-f_mix_acc_labels = function(n){
-  paste(f_mix$`%cum`[n], "%", sep = "")
-}
-p2 <- f_mix %>%
-  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
-  geom_bar(stat = 'identity') +
-  scale_y_continuous(
-      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
-                          name = name1,
-                          labels = format_percent)) +
-  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
-                     sec.axis = sec_axis(~., breaks = 1:24,
-                              name = name2,
-                              labels = f_mix_acc_labels)) +
-  theme_bw() +
-  theme(panel.grid.minor.y = element_blank()) +
-  xlab(xlabel2) +
-  ylab(ylabel) +
-  coord_flip()
-grid.arrange(p1, p2, ncol=2)
-```
-
-### Correlation between ingredients and compressive strength
-
-The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}).
-The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).
-
-&nbsp;
-
-```{r correlation, echo=F, fig.height=4, fig.cap=cap}
-# Figure - Correlation grouped by age
-cor_dat <- dat %>% select(-c(id))
-cap <- "Correlations at each age"
-f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
-name <- "Correlation"
-colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
-                "Superplast.", "Coarse agg.", "Fine agg.",
-                "3", "7", "14", "28", "56", "100",
-                "class", "mix_app", "W./C.", "F.A./C.", 
-                "C.A./C.", "F.A./C.A.",
-                "W./C.A.", "W./F.A.")
-colnames(cor_dat) <- colnames_dat
-cor_dat <- cor_dat %>%
-  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
-  drop_na()
-cor_day <- function(d){
-  res <- cor_dat %>%
-    filter(day == d) %>%
-    select(-c(day, class, mix_app)) %>%
-    cor(.)
-  res[upper.tri(res)] <- NA
-  return(res)
-}
-cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
-               cor_day(28), cor_day(56), cor_day(100))
-melt_day <- function(df, d){
-  df %>%
-    melt() %>%
-    mutate(day = d)
-}
-melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
-                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
-                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
-melt_dat_final <- melt_dats %>%
-  reduce(rbind) %>%
-  filter(value != 1)
-melt_dat_final$day <- factor(melt_dat_final$day)
-levels(melt_dat_final$day) <- f_lvl
-melt_dat_final %>%
-  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
-  geom_tile(color = "white") +
-  facet_wrap(~day, ncol=3) +
-  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
-   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
-  xlab("") +
-  ylab("") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(), 
-        axis.line = element_line(colour = "black"))
-```
-
-```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
-# Figure - Correlation over time
-cap <- "Correlation of variables with compressive strength over time"
-name <- "Correlation"
-melt_dat_final %>%
-  filter(Var1 == "mpa" | Var2 == "mpa") %>%
-  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
-  geom_tile(color = "white") +
-  facet_wrap(~day, ncol=6) +
-  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
-   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
-  xlab("") +
-  ylab("") +
-  geom_text(aes(label = round(value, 2)), size = 2.5) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.line = element_line(colour = "black"))
-```
-
-The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate.
-The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.
-
-&nbsp;
-
-```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
-# Figure - Relationship between approximated mix, water, MPa and age
-cap <- "Relationship between approximated mix, water, MPa and age"
-d <- " days"
-xlabel <- "Approximated mix" 
-ylabel <- "Compressive strength (MPa)"
-label <- "Water /\nCement"
-mix_dat <- dat %>%
-  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
-          mix_app, `water_/_cement`, 
-          `fine_aggregate_/_cement`, 
-          `coarse_aggregate_/_cement`))
-
-labs <- paste(c("3","7","14","28","56","100"), d, sep="")
-mix_dat <- mix_dat %>%
-  gather("day", "mpa", -c(mix_app, `water_/_cement`,
-                          `fine_aggregate_/_cement`,
-                          `coarse_aggregate_/_cement`)) %>%
-  drop_na()
-lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
-mix_dat$day <- factor(mix_dat$day, levels=lvls)
-levels(mix_dat$day) <- labs
-min_x <- min(mix_dat$`water_/_cement`)
-max_x <- max(mix_dat$`water_/_cement`)
-s_x <- max_x - min_x
-mix_dat %>%
-  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
-  geom_point()  +
-  facet_wrap(~ day, ncol=2) +
-  theme_bw() +
-  ylab(ylabel) +
-  xlab(xlabel) +
-  scale_shape_manual(values=c(16, 2, 8)) +
-  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
-                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
-                         breaks = c(round(min_x, 2),
-                                    round(s_x*0.25 + min_x,2), 
-                                    round(s_x*0.5 + min_x,2),
-                                    round(s_x * 0.75 + min_x,2),
-                                    round(max_x, 2))) +
-  labs(colour = label) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  theme(panel.grid.minor = element_blank(),
-      axis.line = element_line(colour = "black"))
-```
-
-```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
-# Figure - Relationship between concrete main features
-cap <- "Relationship between concrete main features"
-d <- " days"
-xlabel <- "Water / Cement"
-ylabel <- "Compressive strength (MPa)"
-colour <- "Fine Agg. /\nCement"
-size <- "Coarse Agg. /\nCement"
-min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
-max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
-s_x_2 <- max_x_2 - min_x_2
-mix_dat %>%
-  ggplot(aes(x=`water_/_cement`, y=mpa,
-             colour = `fine_aggregate_/_cement`,
-             size = `coarse_aggregate_/_cement`)) +
-  geom_point(alpha = 0.5)  +
-  facet_wrap(~ day, ncol=2) +
-  theme_bw() +
-  ylab(ylabel) +
-  xlab(xlabel) +
-  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
-                         midpoint = (s_x_2 / 2) + min_x_2, 
-                         limits = c(min_x_2, max_x_2),
-                         breaks = c(round(min_x_2, 2),
-                                    round(s_x_2*0.25 + min_x_2,2),
-                                    round(s_x_2*0.5 + min_x_2,2),
-                                    round(s_x_2 * 0.75 + min_x_2,2),
-                                    round(max_x_2 - 0.01, 2))
-                         ) +
-  labs(colour = colour, size = size) +
-  ylim(c(-5,85)) +
-  scale_radius() +
-  theme_minimal() +
-  theme(panel.grid.minor = element_blank(),
-      axis.line = element_line(colour = "black"))
-```
-
-The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}).
-The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.
-
-### Variables distribution
-
-The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.
-
-&nbsp;
-
-```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
-# Figure - Variables distribution
-
-cap <- "Variables distribution"
-colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
-                  "Superplast.", "Coarse agg.", "Fine agg.",
-                  "Water / Cement", "Fine agg. / Cement", 
-                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
-                  "Water / Coarse agg.", "Water / Fine agg.")
-dist_dat <- dat %>%
-  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))
-
-colnames(dist_dat) <- colnames_dat
-dist_dat <- dist_dat %>%
-  gather("Var", "value") %>%
-  mutate(value = as.numeric(value))
-dist_dat %>%
-  ggplot(aes(value)) +
-  geom_histogram(aes(y = ..density..), 
-                 colour = "black", 
-                 fill = "white") +
-  geom_density(alpha = .5, fill = "lightseagreen") +
-  facet_wrap(~ Var, ncol=3, scale = "free") +
-  theme_minimal() +
-  xlab("") +
-  ylab("") +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(), 
-        axis.line = element_line(colour = "black"))
-```
-
-The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}.
-It shows that the resistance to compression gradually increases over time, as expected.
-Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
-
-```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
-# Figure - Variables distribution grouped by age
-
-days_labs <- c("3","7","14","28","56","100")
-cap <- "Variables distribution grouped by age"
-colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
-                  "Superplast.", "Coarse agg.", "Fine agg.",
-                  days_labs,
-                  "Water / Cement", "Fine agg. / Cement", 
-                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
-                  "Water / Coarse agg.", "Water / Fine agg.")
-dist_dat_2 <- dat %>%
-  select(-c(id, class, mix_app))
-colnames(dist_dat_2) <- colnames_dat
-dist_dat_2 <- dist_dat_2 %>%
-  gather("day", "mpa", days_labs) %>%
-  drop_na() %>%
-  gather("Var", "value", -c("day")) %>%
-  mutate(day = factor(day, levels = days_labs))
-dist_dat_2 %>%
-  ggplot(aes(x = day, y = value)) +
-  geom_violin(color = NA,
-                fill = "lightseagreen",
-                alpha = .5,
-                na.rm = TRUE,
-                scale = "count") +
-  geom_boxplot(alpha = 0.2) +
-  facet_wrap(~ Var, ncol=3, scale = "free") +
-  theme_minimal() +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(), 
-        axis.line = element_line(colour = "black")) +
-  xlab("") +
-  ylab("")
-```
-
-### Principal component analysis
-
-In the figure \ref{fig:pca}, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}).
-The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that.
-It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
-
-```{r pca, echo=F, fig.height=3, fig.cap=cap}
-# Figure - Principal component analysis on ingredients
-
-cap <- "Principal component analysis on ingredients"
-colnames_dat <- c(
-  "Cem.", "B.F.S.", "F.A.", "Wat.", "Sup.", "C.Agg.",
-  "F.Agg.","WxC", "F.Agg.xC","C.Agg.xC","F.Agg.xC",
-  "WxC.Agg.","WxF.Agg."
-)
-class_list <- list(low="0", normal="20", medium="40", high="70")
-dat_pca <- dat %>%
-  select(-c(id, class, mix_app,
-            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))
-colnames(dat_pca) <- colnames_dat
-class_2 <- dat$day_28
-class_2[class_2 < 20] = "0"
-class_2[class_2 >= 20 & class_2 < 40] = "20"
-class_2[class_2 >= 40 & class_2 < 70] = "40"
-class_2[class_2 >= 70] = "70"
-class_2 <- factor(class_2)
-levels(class_2) <- class_list
-pca <- prcomp(dat_pca, scale = TRUE)
-fviz_pca_ind(
-  pca,
-  geom.ind = "point",
-  habillage=class_2,
-  addEllipses = T,
-  ellipse.level=0.95) +
-  ggtitle("") +
-  theme_bw()
-```
-
-# Literature cited
-
-<div id="refs"></div>
+</PT>

### 2020-08-14 _meta/templates/Materials_and_methods__Data_preparation.Rmd
refactor

--- 2020-04-25 templates/Materials_and_methods__Data_preparation.Rmd
@@
 The data show the variables in the columns and samples in the rows.
-However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
+However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.
 </EN>
@@
 ```{r similar-samples, echo=FALSE}
-#<EN># Tabela - Samples with same composition</EN>
+#<EN># Table - Samples with same composition</EN>
 #<PT># Tabela - Amostras com a mesma composição</PT>
@@
 ```{r similar-samples-2, echo=FALSE}
-#<EN># Tabela - Same samples with different results</EN>
+#<EN># Table - Same samples with different results</EN>
 #<PT># Tabela - Amostras iguais com resultados diferentes</PT>
@@
 ```{r similar-samples-same-id, echo=FALSE, include=TRUE}
-#<EN># Tabela - Previous samples after processing</EN>
+#<EN># Table - Previous samples after processing</EN>
 #<PT># Tabela - Amostras anteriores após processamento</PT>
@@
 
-As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models.
+As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models.
 However, it is necessary to verify how relevant the data of these other ages are.
@@
 The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis.
-The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.
+The first two dimensions represent 37% and 24% $ respectively of the variance.
 </EN>
@@
 A fígura \ref{fig:pca-90-91-100} mostra como as amostras se relacionam umas com as outras (quais são parecidas ou diferentes) e revelou como cada variável contribui para a análise.
-As duas primeiras dimensões representam $37\%$ e $24\%$ respectivamente da variância.
+As duas primeiras dimensões representam 37% e 24% respectivamente da variância.
 </PT>
@@
 Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value.
-The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).
+The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-time}).
 </EN>
@@
 Pelos motivos apresentados nas figuras \ref{fig:boxplot}, \ref{fig:pca-90-91-100} e \ref{fig:mpa-on-time}, foi considerado que as idades de 90, 91 e 100 dias podem ser agrupadas para melhorar a leitura e diminuir o ruído das amostras.
-Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}, pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
+Elas foram convertidas para o mesmo valor, que no caso foi escolhido a idade de 100 dias (\ref{show-join-90-91-100}), pois como mostrado na figura \ref{fig:mpa-on-time}, a resistência apenas aumenta, logo aos 100 dias a resistência à compressão será maior ou igual ao valor de 90 ou 91 dias.
 </PT>
@@
 
-sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
+sum_duplicated <- sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
 
@@
 
-The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
+The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}).
 The result in the first samples after this processing is shown in the table \ref{tab:new-features}.
@@
 Other proportions between the main ingredients were also added.
-The new variables are presented in the table ref{tab:new-features-table}.
+The new variables are presented in the table \ref{tab:new-features-table}.
 </EN>

### 2020-08-14 _meta/templates/Materials_and_methods__Machine_learning_models.Rmd
refactor

--- 2020-04-25 templates/Materials_and_methods__Machine_learning_models.Rmd
@@
 
-The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.
+The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.
 
 ### Pre-processing and data separation
-
-Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado.
 
@@
 One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}).
-For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.
+For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-preparated-samples}.
 </EN>
@@
 
-dat_tabela <- dat_28[1:18]
+dat_table <- dat_28[1:18]
 
 kable(
-    head(dat_tabela[order(dat_tabela$id),]),
+    head(dat_table[order(dat_table$id),]),
     col.names = dfUnits,
@@
 As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
-
+&nbsp;
 1. All *features*;
@@
 5. Only new *features*, without *dummy vars*;
-
+&nbsp;
 Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
 
-For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set.
+For illustrative purposes, the table \ref{tab:table-done-samples-28} shows the first 6 samples from the 28-day model training set.
 The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
@@
 Como ao longo do processamento foram adicionadas novas variáveis (as relações entre os ingredientes e mais algumas *dummy vars* para cada conjunto de idade), foram estudadas 5 possibilidades de configurações das *features* para os modelos:
-
+&nbsp;
 1. Todas as *features*;
@@
 5. Apenas *features* adicionadas, sem as *dummy vars*;
-
+&nbsp;
 Construindo um modelo para cada uma dessas configurações utilizando o conjunto de amostras aos 28 dias (\ref{show-test-models-reg}), mostrou que a melhor opção é a configuração 2, ou seja, as *dummy vars* foram completamente descartadas, porém foram mantidas as outras novas varáveis.

### 2020-08-14 _meta/templates/Materials_and_methods__Replicability.Rmd
refactor

<EN>
## Reproducibility

In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result. The *seeds* can be checked throughout [Appendix 3](#appendix3) or directly on *Github*.
</EN>

<PT>
## Reprodutibilidade

Para garantir a reprodutibilidade, sempre que houve algum código que podesse utilizar operações probabilísticas, foi definido um *seed* antes da sua execução, garantindo que quando rodado em outra máquina, com a mesma versão de *R* e o mesmo *seed*, chegue ao mesmo resultado. Os *seeds* podem ser conferidos ao longo do [apêndice 3](#appendix) ou diretamente no *Github*.
</PT>

### 2020-08-14 _meta/templates/Results.Rmd
refactor

--- 2020-04-25 templates/Results.Rmd
@@
 <PT>
-# Results
+# Resultados
 
@@
 #<EN># Table - Models details</EN>
-#<PT># Tabela - Detalhes dos modelso</PT>
+#<PT># Tabela - Detalhes dos modelos</PT>
 
@@
 res_df <- bind_rows(res_df)
-res_df$model <- factor(res_df$modelo,levels=modelos)
+res_df$model <- factor(res_df$model,levels=models)
 
@@
   ggplot(aes(actual, pred)) +
-  facet_wrap(~ modelo, ncol=3) +
+  facet_wrap(~ model, ncol=3) +
   geom_point(alpha=0.5) +

### 2020-08-14 HarvardX_Data_Science_Capstone/capstone.Rmd
Save checkpoint xls file

@@
 
+```{r dat-preparated-2, echo=F}
+dat_preparated_2 <- dat
+```
+
+```{r dat-preparated-to-xls, echo=F}
+dat_preparated_2 <- dat
+names(dat_preparated_2) <-  gsub("day_", "", names(dat_preparated_2))
+dat_preparated_2 <- dat_preparated_2 %>%
+  gather("age", "Mpa", "3":"100", na.rm = TRUE, convert = TRUE)
+
+write.xlsx(dat_preparated_2, "checkpoint.xls", sheetName = "Sheet1", 
+  col.names = TRUE, row.names = TRUE, append = FALSE)
+
+```
+
 ## Data visualization

### 2020-08-14 HarvardX_Data_Science_Capstone/capstone.R

@@ -352,6 +352,16 @@ kable(
 	column_spec(4:9, width = "1.7cm")
 
 
+# Save the data to a xls file
+dat_preparated_2 <- dat
+names(dat_preparated_2) <-  gsub("day_", "", names(dat_preparated_2))
+dat_preparated_2 <- dat_preparated_2 %>%
+	gather("age", "Mpa", "3":"100", na.rm = TRUE, convert = TRUE)
+
+write.xlsx(dat_preparated_2, "checkpoint.xls", sheetName = "Sheet1", 
+											col.names = TRUE, row.names = TRUE, append = FALSE)
+
+
 # Table - Descriptive statistics - continuous variables
 summ <- t(
 	stat.desc(select(dat, -c(id, class, mix_app))))

### 2020-09-07 CILAMCE_2020/PT/paper.tex
add PT version

%% --------------------------------------------------------------------------
% LaTeX template for the XLI CILAMCE.
%
% This latex document tries to copy the Microsoft Word template.
% --------------------------------------------------------------------------
\documentclass[a4paper,10pt]{book}

% PACKAGES USED - packages that need to be previously installed on your computer
\usepackage[lmargin=2.5cm, rmargin=2.5cm, tmargin=2.5cm, bmargin=2.5cm ]{geometry}
\usepackage{graphicx}
\usepackage{times}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[english]{babel}
\usepackage{parskip} 
\usepackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% My Additional Packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
%\usepackage{amssymb} %Mathematics
%\usepackage{amsfonts}%Mathematics
%\usepackage{amsmath,amscd}%Mathematics
%\usepackage{amsthm}%Mathematics
%\usepackage{mathrsfs}%Mathematics font
%\usepackage{xspace}
%\usepackage{booktabs}
%\usepackage{stmaryrd}%Particular Brackets
%\usepackage{graphicx} %Tables and Figures
%\usepackage{subfigure}
%\usepackage{url}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{./pkg-crefNames}
\usepackage[labelsep=period]{caption}

%BibTeX compatible with the CILAMCE format
\usepackage[numbers,sort&compress]{natbib}

\setlength{\bibsep}{0pt plus 0.3ex}

\renewcommand*{\bibfont}{\small}

\makeatletter
\renewcommand\bibsection
{
  \section*{References}
}

\renewenvironment{thebibliography}[1]
      {\section*{Referêcias}%
       \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
       \list{\@biblabel{\@arabic\c@enumiv}}%
            {\settowidth\labelwidth{\@biblabel{#1}}%
             \leftmargin\labelwidth
             \advance\leftmargin-10pt% change 20 pt according to your needs
             \advance\leftmargin\labelsep
             \setlength\itemindent{10pt}% change using the inverse of the length used before
             \@openbib@code
             \usecounter{enumiv}%
             \let\p@enumiv\@empty
             \renewcommand\theenumiv{\@arabic\c@enumiv}}%
       \sloppy
       \clubpenalty4000
       \@clubpenalty \clubpenalty
       \widowpenalty4000%
       \sfcode`\.\@m}
      {\def\@noitemerr
        {\@latex@warning{Empty `thebibliography' environment}}%
       \endlist}
\renewcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother

\makeatother
\bibliographystyle{bib-cilamce}
%\bibliographystyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% CONFIGURATION
\renewcommand*\arraystretch{1.5}
\renewcommand*\thesection{\arabic{section}}
%\hyphenpenalty=10000 % You can uncomment this to avoid hyphenization
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\bfseries}
\titlespacing\section{0pt}{20pt plus 2pt minus 2pt}{12pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{20pt plus 0pt minus 0pt}{12pt plus 0pt minus 0pt}
\setlength{\parskip}{0pt} % Spacing between paragraphs
\setlength{\parindent}{0.75cm} % Paragraph identation
\setlength\abovecaptionskip{6pt}

% --------------------------------------------------------------------------
% DO NOT EDIT - SPECIAL HEADINGS OF XLI CILAMCE
% --------------------------------------------------------------------------
\fancypagestyle{first}
{
\fancyhf{}
\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}
\renewcommand{\headrulewidth}{.0pt}
\renewcommand{\footrulewidth}{.5pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[RO,LE]{}

\fancyfoot[RO]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\fancyfoot[LE]{\footnotesize \textit{CILAMCE 2020 \\
Proceedings of the XLI Ibero-Latin-American Congress on Computational Methods in Engineering, ABMEC.\\
Foz do Iguaçu/PR, Brazil, November 16-19, 2020}}

\renewcommand{\headrulewidth}{.5pt}
\renewcommand{\footrulewidth}{.5pt}

% --------------------------------------------------------------------------
% PLEASE, EDIT THIS!
\fancyhead[LE]{\footnotesize \textit{CILAMCE 2020 Previsão da resistência à compressão do concreto com técnicas de aprendizado de máquina}}
\fancyhead[RO]{\footnotesize \textit{P. Moreira, V. Silva}}
% --------------------------------------------------------------------------

\begin{document}\thispagestyle{first}

% --------------------------------------------------------------------------
% DO NOT EDIT - LOGO OF XLI CILAMCE
% --------------------------------------------------------------------------

\begin{figure}[ht!]
\vspace{-30pt}
\flushright
\includegraphics[width=5.5cm]{Figures/logoCILAMCE2020.png}
%scale=0.25
\end{figure}

% --------------------------------------------------------------------------
% TITLE OF PAPER
% --------------------------------------------------------------------------

\noindent
\textbf{\Large
Previsão da resistência à compressão do concreto com técnicas de aprendizado de máquina} 
\vspace{14pt} % <- keep this vertical space!

% --------------------------------------------------------------------------
% AUTHORS
% --------------------------------------------------------------------------

\noindent Pedro B. A. Moreira$^1$, Victor M. Silva$^2$

\vspace{14pt} % <- keep this vertical space!

\noindent $^1$\textit{Estudante, Engenharia Civil, Universidade Veiga de Almeida}

\noindent \textit{Avenida das Américas, 22631-004, Rio de Janeiro, Brazil}

\noindent \textit{pedrobermoreira@gmail.com}

\noindent $^2$\textit{Professor, Engenharias, IBMEC/RJ}

\noindent \textit{Avenida Armando Lombardi 940, 22640-000, Rio de Janeiro, Brazil}

\noindent \textit{victor.silva@professores.ibmec.edu.br}

\vspace{18pt} % <- keep this vertical space!

% --------------------------------------------------------------------------
% ABSTRACT
% --------------------------------------------------------------------------

\noindent \textbf{Resumo.}
A resistência à compressão é a principal característica do concreto. A previsão correta deste parâmetro significa redução de custo e tempo. Este trabalho construiu modelos preditivos para 6 diferentes idades de amostras de concreto. Foi utilizado um conjunto de 1.030 amostras de estudos anteriores, com 9 variáveis. Outras 6 variáveis ​​foram adicionadas para representar as proporções dos ingredientes principais em cada amostra. Os modelos preditivos foram desenvolvidos em linguagem {\it R}, utilizando o algoritmo {\it Parallel Random Forest} e a técnica de validação cruzada repetida para otimizar os parâmetros. Os resultados foram compatíveis com outros estudos usando o mesmo conjunto de dados. O modelo mais importante, de 28 dias, obteve um erro quadrático médio ({\it RMSE}) de 4,717. O modelo de 3 dias obteve o melhor resultado, {\it RMSE} de 3,310. O trabalho mostrou que a resistência à compressão do concreto pode ser prevista. Criar um modelo para cada idade permitiu obter resultados compatíveis com os dados disponíveis em cada idade. Foi uma alternativa promissora, pois bons resultados foram alcançados treinando com apenas um algoritmo. Este trabalho facilita a exploração e novos esforços para prever a resistência à compressão do concreto, pode ser usado como linha de base para diferentes algoritmos ou a combinação de vários.

\vspace{18pt} % <- keep this vertical space!

\noindent \textbf{Palavras-chave:} Concreto, Resistência à compressão, Aprendizado de Máquina, Previsão, {\it Parallel Random Forest}

% --------------------------------------------------------------------------
\section{Introdução}\label{sec:introduction}
% --------------------------------------------------------------------------

A resistência à compressão é a principal característica do concreto, medida por ensaios de padrões internacionais que consistem na quebra de corpos de prova \cite{Gambhir1990}. A medição aos 28 dias é obrigatória e representa a classe do concreto. Saber com antecedência o resultado em uma determinada idade, com base nas proporções de seus ingredientes, é de grande interesse para fabricantes de concreto, construtoras e engenheiros civis.

A resistência à compressão é uma função não linear de seus ingredientes e idade, dificultando o estabelecimento de uma fórmula analítica, embora algumas já tenham sido propostas por \citet{Hasan2011} e  \citet{Kabir2012}. Porém, a maioria dos estudos construiu modelos incluindo a idade como característica junto com os ingredientes, mas devido à não linearidade entre a resistência à compressão e a idade, constatamos a necessidade de maiores investigações de modelos que separem a idade e analisem apenas os ingredientes.

Portanto, o presente estudo visa a construção de modelos preditivos da resistência à compressão do concreto em diferentes idades utilizando apenas seus ingredientes como características.

% --------------------------------------------------------------------------
\section{Trabalhos Relacionados}\label{sec:RelatedWork}
% --------------------------------------------------------------------------

\citet{Yeh1998} demonstrou a possibilidade de usar Redes Neurais Artificiais para prever a resistência à compressão do concreto, concluindo que é um método mais preciso do que os modelos de regressão. Neste estudo, mais de 1.000 amostras de concreto foram coletadas de 17 fontes diferentes. Este conjunto de dados foi posteriormente utilizado em vários estudos sobre concreto, alguns dos quais são mencionados a seguir.

\citet{Alshamiri2020} propôs um novo método, {\it Regularized Extreme Learning Machine} ({\it RELM}) para treinar modelos de Redes Neurais Artificiais para predizer a resistência à compressão. Os resultados foram comparados com vários algoritmos conhecidos rodando no mesmo conjunto de dados, incluindo individual e {\it ensembles}, e o modelo proposto teve o melhor resultado de longe.

\citet{Hameed2020} compara modelos de Rede Neural Artificial com Regressão Linear Múltipla para prever a força de resistência à compressão e descobriu que os modelos de Rede Neural Artificial obtêm muito mais precisão do que a Regressão Linear Múltipla.

Além desses estudos publicados, agora é muito comum publicar projetos pessoais em páginas da web. Por fácil acesso a este banco de dados e o crescente interesse em ciência de dados e aprendizado de máquina, alguns estudos não publicados usando esse mesmo banco de dados incluem \citet{Modukuru2020}, \citet{Raj2018}, \citet{Abban2016} e  \citet{Pierobon2018}. No geral, todos seguiram as etapas padrão no desenvolvimento de modelos de aprendizado de máquina, os dois primeiros usando o pacote {\it scikit-learn} em linguagem {\it python} desenvolvido por  \citet{pedregosa2011scikit} e os dois últimos usaram o pacote {\it caret} desenvolvido por \citet{caret} em linguagem {\it R} \cite{RCore}.

Ao final deste trabalho, na seção de discussão e conclusão, os resultados encontrados neste trabalho são comparados com todos esses estudos relacionados.

% --------------------------------------------------------------------------
\section{Metodologia}\label{sec:MaterialsAndMethods}
% --------------------------------------------------------------------------

% --------------------------------------------------------------------------
\subsection{Materiais e reprodutibilidade} \label{subsec:materialsandreproducibility}
% --------------------------------------------------------------------------

A metodologia foi realizada no software {\it RStudio} \cite{RStudio}, um ambiente virtual integrado para desenvolvimento de código em linguagem {\it R} \cite{RCore}. Ao longo do processo, todo o código executado foi documentado na mesma ordem de sua execução e enviado para o repositório online hospedado no {\it github} \cite{github}. O repositório contém uma versão estendida deste trabalho, incluindo todo o código, pacotes necessários e versões. Para garantir a reprodutibilidade, sempre que um código utilizou operações probabilísticas, um {\it seed} foi definido antes de sua execução, garantindo a consistência dos resultados ao rodar em outra máquina.

% --------------------------------------------------------------------------
\subsection{Conjunto de dados} \label{subsec:dataset}
% --------------------------------------------------------------------------
Os dados foram baixados do site da Universidade da Califórnia em Irvine  \cite{downloadData}. No total, são 1.030 linhas com 9 colunas. Cada linha representa uma amostra com as variáveis: resistência à compressão, idade e 7 ingredientes (água, cimento, agregado miúdo, agregado graúdo, cinza volante, escória de alto-forno e superplastificantes).

% --------------------------------------------------------------------------
\subsection{Preparação dos dados} \label{subsec:datapreparation}
% --------------------------------------------------------------------------

Os trabalhos relacionados utilizaram o conjunto de dados em sua totalidade ou realizaram um mínimo de preparação. De forma diferente, neste trabalho uma etapa específica foi dedicada apenas a limpar as amostras e prepará-las para as próximas etapas. As principais etapas executadas nesta seção estão listadas abaixo:

\begin{enumerate}
\item Amostras duplicadas foram removidas;

\item As amostras foram agregadas e identificadas (com um nova coluna de {\it ID}) pelas configurações de proporção dos ingredientes, independente de sua idade;

\item Idades de 90, 91 e 100 dias foram fundidas em uma única categoria de 100 dias. Essa etapa foi realizada com o seguinte desenvolvimento: primeiro, traçando e analisando o {\it boxplot} agrupado por idade na Fig. \ref{fig:boxplot} mostrou que as amostras de 90, 91 e 100 dias têm concentrações distintas de valores de resistência à compressão. Em seguida, uma análise de componentes principais ({\it PCA}) dos ingredientes foi feita na Fig. \ref{fig:pca}, mostrando que as combinações de ingredientes dessas idades são muito distintas. Como são idades muito próximas, é razoável que possamos juntar essas idades sem qualquer prejuízo nas previsões. Por fim, a marca dos 100 dias foi escolhida porque a análise de amostras que continham dados em pelo menos cinco idades diferentes mostrou que a resistência à compressão tende a aumentar com o tempo. Portanto, uma idade mais avançada tende a fornecer resultados mais conservadores.

\item Após juntar as idades de 90, 91 e 100 dias, foram retiradas as idades com frequência menor que 50, restando apenas as idades de 3, 7, 14, 28, 56 e 100 dias;

\item Para amostras com o mesmo {\it ID} e a mesma idade, mas com diferentes valores de resistência à compressão, os dados foram combinados em uma única linha para cada combinação de {\it ID} e idade, contendo uma média das resistências à compressão;

\item Foram mantidos apenas os {\it ID}s que possuem dados na marca de 28 dias;

\item Adição de 6 novas variáveis contínuas que foram usados nos modelos de predição, representando as proporções entre os ingredientes principais (água / cimento, agregado miúdo / cimento, agregado graúdo / cimento, agregado miúdo / agregado graúdo, água / agregado graúdo e água / agregado miúdo);

\item Acréscimo de 2 novas variáveis ​​categóricas utilizadas para visualizar a distribuição das amostras (classe do concreto e traço aproximado).
\end{enumerate}

\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.8]{Figures/boxplot_PT.jpg}
\vspace{12pt}
\caption{{\it Boxplot} - resistência à compressão agrupada por idade}
\label{fig:boxplot}
\end{center}
\end{figure}
\vspace{8pt}

\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.8]{Figures/pca-90-91-100_PT.jpg}
\vspace{12pt}
\caption{Análise de componente principal - 90, 91 e 100 dias}
\label{fig:pca}
\end{center}
\end{figure}
\vspace{8pt}

Após essas manipulações, o número final de amostras foi reduzido de 1.030 linhas para 916 linhas, com 416 configurações de ingredientes diferentes ({\it IDs}). Um arquivo {\it xls} dos dados neste momento está disponível para download no repositório no github  \cite{github}.

% --------------------------------------------------------------------------
\subsection{Visualização dos dados} \label{subsec:datavisualization}
% --------------------------------------------------------------------------

Além dos gráficos já apresentados, vários outros gráficos e tabelas foram construídos para realizar a exploração e visualização das amostras preparadas na etapa anterior, dentre elas:

\begin{itemize}

\item Estatística descritiva das variáveis contínuas e categóricas;

\item Distribuição das variáveis em relação à resistência à compressão;

\item Correlação entre as variáveis agrupadas por idade;

\item Relação entre o traço aproximado e a resistência à compressão;

\item Relação entre os principais ingredientes do concreto e a resistência à compressão;

\item Análise de componentes principais ({\it PCA}) dos ingredientes.

\end{itemize}

Os gráficos para da estatística descritiva das variáveis categóricas são apresentados, na Fig. \ref{fig:categorical}.

\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.8]{Figures/stat-summ-categorical_PT.jpg}
\vspace{12pt}
\caption{Estatística descritiva das variáveis categóricas}
\label{fig:categorical}
\end{center}
\end{figure}
\vspace{8pt}

% --------------------------------------------------------------------------
\subsection{Pré-processamento e separação dos dados} \label{subsec:preprocessing}
% --------------------------------------------------------------------------

 O principal pacote usado para construir os modelos de aprendizado de máquina foi o pacote {\it Caret}  \cite{caret}. Ele fornece todas as funcionalidades e utilitários para construir modelos de previsão para qualquer conjunto de dados, tem uma documentação direta e clara que orienta o processo e fornece cerca de 200 algoritmos diferentes para construir modelos. Neste trabalho, foram realizados os principais passos descritos por \citet{irizarry2019} e \citet{Kuhn2008}. Começando por algumas etapas de pré-processamento descritas abaixo:

\begin{enumerate}

\item Remoção das variáveis categóricas;

\item Separação do conjunto de dados por idade, resultando em 6 conjuntos de dados menores;

\item Remoção de variáveis com variância próxima de zero (apenas a cinza volante do conjunto de 7 dias foi removida);

\item Verificação de variáveis com correlação acima de 0,999, o que não ocorreu;

\item Cada conjunto de dados foi dividido em conjuntos de teste e treino, 20\% e 80\% respectivamente, mostrado na Tabela \ref{tab:split-config} e a distribuição na Fig. \ref{fig:split-dist}.

\end{enumerate}

\vspace{8pt}
\begin{table}[!ht]
\centering
\caption{Separação das amostras}
\label{tab:split-config}
\vspace{6pt}
\begin{tabular}{cccc}
\hline
Modelo & Amostras & Treino (80\%) & Teste (20\%)\\ \hline
3 dias & 121 & 97 & 24\\
7 dias & 114 & 94 & 20\\
14 dias & 62 & 50 & 12\\
28 dias & 416 & 335 & 81\\
56 dias & 83 & 67 & 16\\
100 dias & 120 & 96 & 24\\ \hline
\end{tabular}
\end{table}
\vspace{8pt}

\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.9]{Figures/dist-split_PT.jpg}
\vspace{12pt}
\caption{Distribuição dos conjuntos de teste e treino em relação à resistência à compressão}
\label{fig:split-dist}
\end{center}
\end{figure}
\vspace{8pt}

% --------------------------------------------------------------------------
\subsection{Modelos ingênuos} \label{subsec:naivemodels}
% --------------------------------------------------------------------------

Antes de construir os modelos reais, para fins de comparação, foram criados modelos ingênuos. Eles simplesmente prevêem que a resistência à compressão do conjunto de teste é a resistência à compressão média do conjunto de treino. Em outras palavras, os modelos ingênuos são simplesmente a melhor estimativa possível para avaliar o quão próximo / distante o modelo real está de um palpite aleatório.

% --------------------------------------------------------------------------
\subsection{Modelos de aprendizado de máquina} \label{subsec:machinelearningmodels}
% --------------------------------------------------------------------------

Utilizamos apenas um algoritmo neste trabalho, escolhido pela maior probabilidade para atingir o melhor resultado possível, de acordo com \citet{Fernandez2014}, que comparou 179 algoritmos em 121 bancos de dados diferentes, e concluiu que o mais provável de alcançar o melhor resultado possível é o {\it Parallel Random Forest}, chamado {\it prRF} no pacote {\it Caret} \cite{caret}.

Seis modelos diferentes foram construídos, um para cada faixa de idade, e as seguintes etapas foram feitas para cada modelo:

\begin{enumerate}

\item Definir o esquema de reamostragem, com método de validação cruzada repetida;

\item Definir a {\it tuning grid} para o parâmetro {\it mtry}, que é uma sequência de 1 ao número de colunas de cada conjunto de dados. Todos, exceto o de 7 dias, são iguais, visto que apenas o conjunto de 7 dias teve uma coluna removida no pré-processamento;

\item Definir o {\it seed} igual a 1, escolhido arbitrariamente para garantir reprodutibilidade. Esse {\it seed} pode ser manipulado para obter resultados mais satisfatórios, mas foi optado por não fazê-lo.

\item Fazer a transformação do pré-processamento dos dados com os métodos {\it center} e {\it scale};

\item Executar a função {\it train} do {\it Caret} com as configurações acima e modelo {\it parRF};

\end{enumerate}

% --------------------------------------------------------------------------
\section{Resultados}\label{sec:Results}
% --------------------------------------------------------------------------
A avaliação de desempenho dos modelos foi realizada pelo erro de raiz quadrada média ({\it RMSE}). O {\it RMSE} é a medida utilizada em todos os trabalhos citados na introdução permitindo a comparação dos modelos na discussão.

O {\it RMSE} de teste para cada modelo em ordem crescente de idade foi 3,31, 4,36, 4,62, 4,72, 5,94 e 5,85, respectivamente. A tabela  \ref{tab:results} apresenta os detalhes e resultados de cada modelo, inclusive o ingênuo. A Fig.  \ref{fig:results} compara os valores reais e previstos para os modelos finais.

\vspace{8pt}
\begin{table}[!ht]
\centering
\caption{Resultados dos modelos finais}
\label{tab:results}
\vspace{6pt}
\begin{tabular}{ccccccc}
\hline
Modelo & {\it mtry} & {\it CV} &  Repetições & {\it RMSE} ingênuo (teste) &  {\it RMSE} final (treino) & {\it RMSE} final (teste)\\ \hline
3 dias & 6 & 30 & 10 & 9.303229 & 3.905196 & 3.310370\\
7 dias & 2 & 10 & 10 & 13.443646 & 4.475981 & 4.361987\\
14 dias & 13 & 30 & 10 & 7.593319 & 5.136687 & 4.620515\\
28 dias & 11 & 30 & 10 & 14.283824 & 5.847334 & 4.716698\\
56 dias & 8 & 30 & 10 & 12.702112 & 6.702565 & 5.939163\\
100 dias & 8 & 10 & 10 & 12.614652 & 6.381940 & 5.851088\\ \hline
\end{tabular}
\end{table}
\vspace{8pt}

\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.9]{Figures/results-comparison_PT.jpg}
\vspace{12pt}
\caption{Real {\it vs} previsto para cada modelo}
\label{fig:results}
\end{center}
\end{figure}
\vspace{8pt}

\pagebreak

% --------------------------------------------------------------------------
\section{Discussão e conclusão}\label{sec:DiscussionAndConclusion}
% --------------------------------------------------------------------------

Os modelos construídos apresentam resultados satisfatórios e comprovam que a resistência à compressão do concreto pode ser prevista com relativa facilidade. A alternativa adotada de criar um modelo para cada conjunto de idades mostrou-se um método válido, ao invés de usar a idade como preditor junto com os ingredientes, como realizado nos estudos relacionados com o mesmo conjunto de dados. A adoção dessa estratificação obteve resultados diferenciados para cada faixa de idade. O {\it RMSE} calculado em nosso trabalho e o obtido nos trabalhos relacionados foram próximos. A tabela \ref{tab:comparison} mostra a comparação entre esses estudos e o modelo de 28 dias aqui desenvolvido.

Seguindo a linha de raciocínio deste trabalho, a mesma hipótese pode ser avaliada utilizando outros algoritmos além do aqui utilizado  ({\it Parallel Random Forest}), pois podem apresentar resultados melhores. Outra opção é criar um {\it ensemble} de vários algoritmos, como realizado por \citet{Pierobon2018}, mas com a separação dos conjuntos de idades aqui proposta. Além disso, este estudo pode ser reproduzido com um conjunto de dados maior, idealmente com um número semelhante de amostras em cada faixa de idade e uma distribuição mais homogênea de resistência à compressão e classe de concreto, visto que na Fig. \ref{fig:categorical} mostra que esse conjunto de dados é muito inclinado para a classe de concreto entre C25 e C35.

\vspace{8pt}
\begin{table}[!ht]
\centering
\caption{Comparação com outros trabalhos com o mesmo conjunto de dados}
\label{tab:comparison}
\vspace{6pt}
\begin{tabular}{lclcc}
\hline
Autor & Ano & Algoritmo & {\it RMSE} & Diferença ({\%}) \\ \hline
\citet{Pierobon2018} & 2018 & {\it Ensemble} de 5 algoritmos & 4.150 & -12\\
\textbf{Esse trabalho (28 dias)} & \textbf{2020} & \textbf{{\it Parallel Random Forest}} & \textbf{4.717}  & \textbf{0}\\
\citet{Hameed2020} & 2020 & Redes neurais artificiais & 4.736 & 0\\
\citet{Raj2018} & 2018 & {\it Gradient Boosting Regressor} & 4.957  & +5\\
\citet{Modukuru2020} & 2020 & {\it Random Forest Regressor} & 5.080  & +8\\
\citet{Alshamiri2020} & 2020 & {\it Regularized Extreme Learning Machine} & 5.508 & +17\\
\citet{Abban2016} & 2016 & {\it SVM} com {\it Radial Basis Function Kernel} & 6.105 & +29\\ \hline
\end{tabular}
\end{table}
\vspace{8pt}

\pagebreak

%--------------------------------------------------------------------------
\vspace{20pt}
\noindent \textbf{Declaração de autoria.} Os autores confirmam que são os únicos responsáveis pela autoria deste trabalho, e que todo o material aqui incluído como parte do presente trabalho é propriedade (e autoria) dos autores, ou tem a permissão dos proprietários a serem incluídos aqui.

\bibliography{bibliography}

\end{document}
% --------------------------------------------------------------------------
